---
title: "time_to_size_clean"
author: "Denver Link"
date: "2025-10-08"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Library
```{r}
library(tidyverse)
library(arrow)
```

# Data
```{r}
mw_data <- open_dataset(sources = file.path("G:", 
                                            "Shared drives", 
                                            "Hansen Lab", 
                                            "RESEARCH PROJECTS", 
                                            "Fish Survey Data", 
                                            "Parquet files", 
                                            "hive_update"), partitioning = c("state"))
```

# Data prep for Analysis 
```{r}
# first we filter for where the alk level is lake-year, filter for species, and select columns of interest
aged_fish <- mw_data %>% 
  filter(alk == "year" &
           species_1 == "walleye") %>% 
    select(state,
         lake_id,
         lake_name,
         nhdhr_id,
         latitude_lake_centroid,
         longitude_lake_centroid,
         date_survey,
         date_total_effort_ident,
         date_sub_effort_ident,
         date_sample,
         species_1,
         length_1,
         length_unit_1,
         weight_1,
         sampling_method_simple,
         sampling_method,
         sampling_method_2,
         age,
         aging_structure_1,
         est_age,
         alk,
         alk_age_str,
         alk_length,
         alk_n) %>% 
  collect() 

#what is the best date for aging?
aged_fish %>% 
  group_by(state) %>% 
  summarise(has_survey_date = sum(!is.na(date_survey)),
            has_tei_date = sum(!is.na(date_total_effort_ident)),
            has_sub_date = sum(!is.na(date_sub_effort_ident)),
            has_sample_date = sum(!is.na(date_sample)),
            n = n(),
            prop_sample = has_sample_date /n,
            prop_sub = has_sub_date / n,
            prop_tei = has_tei_date/n)

#creating an aged date for fish,
#we typically can use the sample date (corresponding to the date the fish were sampled)
#SD and IA we use sub effort ident, which corresponds to when the net was pulled up
#MI our best date is total effort ident 
aged_fish <- aged_fish %>% 
  mutate(date_age = case_when(state == "Illinois" ~ date_sample,
                              state == "Iowa" ~ date_sub_effort_ident,
                              state == "Michigan" ~ date_total_effort_ident,
                              state == "Minnesota" ~ date_sample,
                              state == "South_Dakota" ~ date_sub_effort_ident,
                              state == "Wisconsin" ~ date_sample)) 
  

aged_fish %>% 
  group_by(state) %>% 
  summarise(n_na = sum(is.na(date_age)),
            n = n(),
            prop = n_na/n)

#have to remove fish without date (Wisconsin fish that we can't fill with any date)
aged_fish <- aged_fish %>% 
  filter(!is.na(date_age))

#adding decimal age
aged_fish %>% 
  group_by(state, length_unit_1) %>% 
  count()

aged_fish <- aged_fish %>% 
  mutate(year = year(date_age),
         doy = yday(date_age),
         days_in_year = case_when(leap_year(year) ~ 366,
                                  TRUE ~ 365),
         decimal_age = est_age + (doy/days_in_year),
         alk_length = case_when(length_unit_1 == "millimeters" ~ alk_length/25.4,
                                TRUE ~ alk_length))

aged_fish %>% 
  group_by(state) %>% 
  slice_sample(n = 1) %>% 
  select(state, lake_id, length_1, length_unit_1, alk_length) %>% 
  print(n = nrow(.))
```

We now have over 1.1 million fish with estimated ages at the lake-year level with a created "decimal age" to account for the growth that may happen within the summer prior to when the fish was sampled. Next up:
1.) filter for fish with otolith and spine ALKs and fish of 15 +/- 0.5 inches
2.) do some quick summary stats of the data distribution
3.) calculate a lake average decimal age for 15 inch walleye
4.) calculate lake-year specific values and make plots for a specific lake
5.) make ALK plot of a specific lake-year example
6.) explore the extremes to see how it is holding up


# quick summary stats of fish for this analysis 
```{r}
#how many 15 inch walleye do we have across states, how many lakes, how many lake years?
aged_fish %>% 
  filter(alk_length >= 14.5 & alk_length <= 15.5) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  group_by(state) %>% 
  summarise(mean_dec_age = mean(decimal_age),
            n_lakes = n_distinct(lake_id),
            n_lake_years = n_distinct(lake_id, year),
            n_fish = n()) 

#in the original data, how many fish from Wisconsin are otoliths?
mw_data %>% 
  filter(state == "Wisconsin") %>% 
  filter(!is.na(aging_structure_1)) %>% 
  summarise(sum_o = sum(aging_structure_1 == "otolith"),
            sum_s = sum(aging_structure_1 %in% c("spine", "scale")),
            n = n(),
            prop_o = sum_o/n,
            prop_s = sum_s/n) %>% 
  collect()

#how many fish per lake do we get?
aged_fish %>% 
  filter(alk_length >= 14.5 & alk_length <= 15.5) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>%
  group_by(state, lake_id, year) %>% 
  summarise(n_fish = n()) %>% 
  ggplot() +
  geom_histogram(aes(n_fish)) +
  facet_wrap(~state, scales = "free")
#over 1000 walleye of 15ish inches seems pretty crazy in one year

mw_data %>% 
  filter(state == "Wisconsin" & lake_id == "2294900" & year == 2009) %>% 
  filter(species_1 == "walleye") %>% 
  summarise(min = min(date_sample, na.rm = T),
            max = max(date_sample, na.rm = T),
            n_tei = n_distinct(total_effort_ident),
            n_gears = n_distinct(sampling_method),
            n_fish_15 = sum(alk_length >= 14.5 & alk_length <= 15.5, na.rm = T),
            n_fish = n()) %>% 
  collect()

mw_data %>% 
  filter(state == "Wisconsin" & lake_id == "2294900" & year == 2009) %>% 
  filter(species_1 == "walleye") %>% 
  collect() %>% 
  ggplot() +
  geom_histogram(aes(length_1, fill = as.factor(est_age)))
#appears to be just a crazy population estimate that they do in wisconsin
```

# mean decimal age by lake-year (with map)
```{r}
bin_summary <- aged_fish %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  filter(alk_length >= 14.5 & alk_length <= 15.5) %>% 
  group_by(state, lake_id, year) %>% 
  summarise(mean_dec_age = mean(decimal_age)) %>% 
  group_by(state, lake_id) %>%
  #now summarize across time
  summarise(mean_dec_age= mean(mean_dec_age))

#make map
lake_ids <- aged_fish %>% 
  distinct(state, lake_id, latitude_lake_centroid, longitude_lake_centroid)

bin_summary <- bin_summary %>% 
  left_join(lake_ids)
rm(lake_ids)

library(sf)
library(maps)

states <- st_as_sf(map("state", plot = FALSE, fill = TRUE)) %>%
  filter(ID %in% c("minnesota", 
                   "wisconsin", 
                   "south dakota",
                   "iowa", 
                   "michigan"
                   #"illinois"
                   ))

bin_summary_sf <- bin_summary %>%
  filter(!is.na(latitude_lake_centroid)) %>% 
  st_as_sf(coords = c("longitude_lake_centroid", "latitude_lake_centroid"), 
           crs = 4326) 

ggplot() +
  geom_sf(data = states, fill = "white", color = "black") +
  geom_sf(data = bin_summary_sf, aes(color = mean_dec_age), size = 2) +
  scale_color_viridis_c(option = "plasma", name = "Decimal Age", direction = -1) +
  theme_minimal() +
  theme(legend.position = c(.1,.75),
        legend.title.align =  0.5,
        legend.key.height = unit(0.3, "cm"),
        legend.key.width  = unit(0.3, "cm"),
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 8))
#ggsave("map_dec_age_otolith.jpg", height = 4, width = 6, dpi = 600)
rm(states, bin_summary_sf)
```

# mean decimal age over time in one lake
```{r}
bin_summary <- aged_fish %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  filter(alk_length >= 14.5 & alk_length <= 15.5) %>% 
  group_by(state, lake_id, year) %>% 
  summarise(mean_dec_age = mean(decimal_age))

bin_summary %>% 
  group_by(lake_id) %>% 
  count() %>% 
  arrange(desc(n))

#make plot showing the number of fish each year colored by the age class
#sticking with Winni in MN
aged_fish %>% 
  filter(lake_id == "69037800",
         alk_length >= 14.5, alk_length <= 15.5) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>%
  group_by(year) %>% 
  summarise(
    mean_dec_age = mean(decimal_age, na.rm = TRUE),
    sd_dec_age = sd(decimal_age, na.rm = TRUE),
    n = n(),
    se_dec_age = sd_dec_age / sqrt(n)
  ) %>% 
  ggplot(aes(x = as.factor(year), y = mean_dec_age)) +
  geom_point(aes(size = n), color = "darkblue") +
  scale_size_continuous(name = "Number of fish") +
  geom_errorbar(aes(ymin = mean_dec_age - se_dec_age,
                    ymax = mean_dec_age + se_dec_age),
                width = 0.2) +
  theme_minimal() +
  labs(
    y = "Decimal Age at 15 inches",
    x = element_blank()
  ) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 30, hjust = 1)
        )
#error bars show 1 standard error
#ggsave("winni_dec_age_summary.jpg", height = 5, width = 6, dpi = 600)
```

# Winni Most Recent Year by Gear?
```{r}
aged_fish %>% 
  filter(lake_id == "69037800",
         alk_length >= 14.5, alk_length <= 15.5) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>%
  filter(year == "2022") %>% 
  group_by(sampling_method_simple) %>% 
  summarise(
    mean_dec_age = mean(decimal_age, na.rm = TRUE),
    sd_dec_age = sd(decimal_age, na.rm = TRUE),
    n = n(),
    se_dec_age = sd_dec_age / sqrt(n)
  )

aged_fish %>% 
  filter(lake_id == "69037800",
         alk_length >= 14.5, alk_length <= 15.5) %>% 
  group_by(sampling_method_simple, year) %>% 
  count() %>% 
  arrange(year) %>% 
  print(n = nrow(.))
```


# Age-Length key of the most recent year (2022) in Winni
```{r}
#plot showing the ALK
aged_fish %>% 
  filter(
    lake_id == "69037800",
    year == 2022 ,
    !is.na(age)
  ) %>%
  mutate(age = fct_rev(as.factor(age))) %>% 
  ggplot(aes(x = alk_length, fill = age)) +
  geom_density(alpha = 0.7, color = "black", adjust = 2) +
  scale_fill_viridis_d(
    option = "plasma",
    name = "Age",
    guide = guide_legend(reverse = TRUE, ncol = 2)  # 2-column legend
  ) +
  theme_minimal() +
  labs(
    x = "Fish Length (in)",
    y = "Density"
  ) +
  theme(
    legend.position = c(0.75, 0.75),   # place legend inside plot
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )
#ggsave("winni_2022_ALK.jpg", height = 5, width = 6, dpi = 600)
```

# check the extremes
Lets look at the min and max to see if they look correct
```{r}
bin_summary %>%
  ungroup() %>%
  filter(mean_dec_age == min(mean_dec_age) | mean_dec_age == max(mean_dec_age))

############################################################################
# Minnesota min
aged_fish %>% 
  filter((lake_id == "66000800" & year == 2020)) %>% 
  filter(!is.na(age)) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
   mutate(age = fct_rev(as.factor(age))) %>% 
  ggplot() +
  geom_density(aes(x = alk_length, fill = age), alpha = 0.7, color = NA, adjust = 2) +
  geom_vline(xintercept = 15) +
  facet_wrap(~lake_id, scales = "free") +
  scale_fill_viridis_d(
    option = "plasma",
    name = "Age",
    guide = guide_legend(reverse = TRUE, ncol = 2)  # 2-column legend
  ) +
  theme_minimal() +
  labs(
    x = "Fish Length (in)",
    y = "Density"
  ) +
  theme(
    legend.position = c(0.75, 0.75),   # place legend inside plot
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

aged_fish %>% 
  filter((lake_id == "66000800" & year == 2020)) %>% 
  filter(alk_length >= 14.5, alk_length <= 15.5) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  summarise(mean_dec_age = mean(decimal_age),
            mean_length = mean(alk_length),
            est_age = est_age,
            age = mean(age),
            aging_structure_1,
            n = n())
#we also see at this extreme an estimated age-1 walleye was observed aged at 1 
#makes sense

##############################################################################
#Wisconsin max
aged_fish %>% 
  filter(lake_id == "439800" & year == 2008) %>%  
  filter(!is.na(age)) %>% 
   mutate(age = fct_rev(as.factor(age))) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  ggplot() +
  geom_density(aes(x = alk_length, fill = age), alpha = 0.7, color = NA, adjust = 2) +
  geom_vline(xintercept = 15) +
  facet_wrap(~lake_id, scales = "free") +
  scale_fill_viridis_d(
    option = "plasma",
    name = "Age",
    guide = guide_legend(reverse = TRUE, ncol = 2)  # 2-column legend
  ) +
  theme_minimal() +
  labs(
    x = "Fish Length (in)",
    y = "Density"
  ) +
  theme(
    legend.position = c(0.75, 0.75),   # place legend inside plot
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

aged_fish %>% 
  filter((lake_id == "439800" & year == 2008)) %>% 
  filter(alk_length >= 14.5, alk_length <= 15.5) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  summarise(mean_dec_age = mean(decimal_age),
            min = min(decimal_age),
            max = max(decimal_age),
            mean_length = mean(alk_length),
            n = n())

aged_fish %>% 
  filter((lake_id == "439800" & year == 2008)) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  glimpse()
#this example in Wisconsin also appears legit

###########################################################################
#extremes in SD?
bin_summary %>%
  ungroup() %>% 
  filter(state == "South_Dakota") %>% 
  filter(mean_dec_age == min(mean_dec_age) | mean_dec_age == max(mean_dec_age))

aged_fish %>% 
  filter(lake_id == "UJA-Lake-866-000" & year == 2018)  %>% 
  filter(!is.na(age)) %>% 
   mutate(age = fct_rev(as.factor(age))) %>% 
  filter(alk_age_str %in% c("otolith", "spine")) %>% 
  ggplot() +
  geom_density(aes(x = alk_length, fill = age), alpha = 0.7, color = NA, adjust = 2) +
  geom_vline(xintercept = 15) +
  facet_wrap(~lake_id, scales = "free") +
  scale_fill_viridis_d(
    option = "plasma",
    name = "Age",
    guide = guide_legend(reverse = TRUE, ncol = 2)  # 2-column legend
  ) +
  theme_minimal() +
  labs(
    x = "Fish Length (in)",
    y = "Density"
  ) +
  theme(
    legend.position = c(0.75, 0.75),   # place legend inside plot
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )
```

