---
title: "CASC_Data_Explainer_Work"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
---

#To-do: 
1. Can I add the suffixes to these field names in a file specific way? This would avoid the new data added with multiple [state] suddennly adding suffixes to all and breaking everyone else's code code.

The answer is yes, and the idea was that within a state, the suffixes could be dropped where the col name is otherwise not duplicated (this will avoid breaking working code in all the other states)

2. Build in somethign to strip ; or, after the last word in a string from the data explainer file. 











3/2/23
Preps data explainer
## Load libraries
```{r}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(arrow)
library(data.table)
  # update_dev_pkg()
library(googledrive)
library(janitor)
library(tidyr)
 # install.packages("tidyr")

```

## CASC Data Explainer
### Read in
```{r}

cde <- fread("https://docs.google.com/spreadsheets/d/e/2PACX-1vSE1oFKS5o65kqIkeeY5ISKx7oOHk5tBROE_g95JCaqcsLuxnmwwHvDhOPnFO0Okw/pub?gid=1011323178&single=true&output=csv", na.strings="")
# 
# cde <- data.table(read_excel("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/CASC_Data_Explainer_for_R_Data_Manipulation.xlsx")) #set as DT on the fly

```

### Prep 
```{r}
#tidy up those column names
# colnames(cde)
  cde <- clean_names(cde) #all lwr case, only _ and letters & numbers


# even we can't get our data straight! r/r commas with semicolons 
  cols = colnames(cde)[colnames(cde) != "source_agency" ] # can choose which cols to do action on here skipping source agency b/c "South Dakota Game, Fish and Parks"
  
  #r/r commas to semicolons
  cde[, (cols) := #rebuild each column ()
        lapply(.SD, function(x){str_replace_all(x , pattern = ",", replacement = ";")}), #by looking through each cell and r/r commas with semicolons
      .SDcols = cols ]  # do the rebuild on cols selected in the cols object

#expand each col to the maxN of col names in any row that column
  cols = colnames(cde)#what are current colnames
  
  #get a vector of how many columns named in each cde column:
  maxn <- unlist(cde[, 
        lapply(.SD,function(x){max(lengths(str_split(x, pattern = ";")))}), # how many units are in here? each unit is sep by a semicolon, count col names, then use the max
        .SD ][1], use.names = F)
    
  #split each multi-content column into many:
    for (i in 1:length(colnames(cde)[maxn > 1])) {
      # i =3
      cde[, paste(colnames(cde)[maxn > 1][i],
                  1:c(maxn[maxn > 1][i]),
                  sep = ".") :=
            tstrsplit(get(colnames(cde)[maxn > 1][i]),
                      ";",
                      fixed=TRUE
              )
      ]
      
      cde[ ,
           paste(colnames(cde)[maxn > 1][i], 1:c(maxn[maxn > 1][i]), sep = ".") := 
             lapply( .SD, trimws) ,
           .SDcols = paste(colnames(cde)[maxn > 1][i], 1:c(maxn[maxn > 1][i]), sep = ".") ]
      
      
      
    }
  
# now clean up all of the leading and trailing whitespace we created in that split
  
 
    
    # colnames(cde)
    # 
    # cbind(cols,maxn)
    
    cde[ , cols[maxn > 1] := NULL ,  ] #remove the original columns (now the first case within a col is colname.1) 
    
    # colnames(cde)


```

