---
title: "IL_Flat_File_Aggregation"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Preamble

Things to keep in mind:
-There was a mismatch between some surveys in the biodata and effort data - Denver sent email was sent to vic 11/16 for fixing
   -catch per effort would not be calculated correctly otherwise 
-Need to explore subsample time 
    -are we able to link fish obs with if it was captured in the sub sample?
    -are they capturing all fish before they sub sample (including sub sample species)?
-make sure I am retaining important columns that are listed before the effort section

things to do:
-make sure the effort ident is at the correct level (pretty sure Denver already took care of this)
     -"nets" includes gill nets and fyke nets
     -in the current situation we could have more than one sub-gear type in a survey
     -in short, make sure there is sampling method and finer scale sampling method in id creation
-the surveys with nothing caught need to be cleaned so that the survey info is retained for filtering (also think this is done)
      -right now the way I join the nothing caught surveys in they only contain the distinct columns
      -requires just the .keep_all = true in the tidy
-check targeted surveys if they have a targeted species - does this align with subsample?
     -do they only record that species in the targeted survey? - need to checked species reported in targeted surveys


##Libraries
```{r}
library(arrow)
library(readr)
library(dplyr)
library(stringr)
library(data.table)
library(janitor)
library(tidyr)
library(lubridate)
library(bit64)

options(scipen = 999)
```


##Data
This could readily be changed into a function that takes a filepath and returns files into environment.
* note Holly has to change file paths to "D" 
```{r}
#generate a file list to import
files_list <- list.files(path = "D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IL_Data/il_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files
files_list



#object for use in loop (simple length of file list)
n <- length(files_list)

for(i in 1:n) {
  #i = 1
  filei <- word(gsub(".csv","", files_list[i]), start = -1, sep = fixed("/"))
  #this does those two steps in one package
  assign(filei ,
          fread(paste0("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IL_Data/il_raw_disaggregated_data/",
                                          files_list[i])))
  
  # if the file is a crosswalk, do not rename anything, just loop to the confirm import line
  if(str_detect(filei, "crosswalk")) {  #confirm import of files:  
    print(paste(filei ,"added to workspace" ))  
    #confirm import of files:  
    print(paste(i ,"files added to workspace" )) ; next}
  
  #if the file is not in the data explainer, don't try to rename it:
  if(filei %in% cde$new_file_name) {
    print("renaming with data explainer")
  } else {next}
  
  
  
  
  
  # note we want to review a sorted list of column names to check misspelling etc.
  # we still need to use the columns with names like col_name_length_in, or known_units
  
  
  cde %>% # call data explainer file
    filter(`new_file_name`== filei)%>% #keep only the row relevant to this file
    select_if(~ !any(is.na(.))) %>% 
    transpose(keep.names = "newname") %>% 
    rename("oldname" = V1) %>% 
    assign("names", ., envir = .GlobalEnv)
  
  #see if any column names will not have a match! 
  # IF any pop FALSE, force stop and revist of data explainer ()
  # - e.g., named something "total catch" when actual column name was "total_catch"
  print(
    cbind(colnames(get(filei)),
          colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]
    )
  )
  
  
  # break the loop if the current file has column names not in the data explainer
  # if (all(cbind(colnames(get(filei)),  colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ])[,2]) == FALSE ) break
  if (all(colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]) == FALSE ) break
  
  
  # append old col names into new "notes" columns:
  get(filei)[ , (names[ str_detect(newname, "notes") , oldname   ,  ]) := Map(paste, colnames(.SD), .SD, sep = ':') , .SDcols =  names[ str_detect(newname, "notes") , oldname   ,  ] ]
  
  #now rename that file's colnames
  setnames(get(filei), colnames(get(filei)), names[!str_detect(newname,"unique_row_key")] [match(names(get(filei)),names[!str_detect(newname,"unique_row_key"),oldname]), newname] )
  
  #append all other data from data explainer
  unusedbits <- 
    data.table(
      matrix(
        rep(names[ !newname %in% colnames(get(filei)) , oldname , ],
            each = nrow(get(filei))
        ),
        nrow = nrow(get(filei)),
        dimnames = list(rep(NA,nrow(get(filei))),
                        names[ !newname %in% colnames(get(filei)) , newname , ])
        )
      )
  
  #add all not yet used columns from data explainer:
  get(filei)[ , (names[ !newname %in% colnames(get(filei)) , newname , ]) := unusedbits[] ]

  #confirm import of files:  
  print(paste(filei ,"added to workspace" ))  
  #confirm import of files:  
  print(paste(i ,"files added to workspace" )) 

  
} 
  #confirm import of files:  
  print(paste(i ,"files added to workspace" ))
  #confirm import of files:  
  print(paste(n-i ,"remaining to be added" )) 



```


#Data Review
```{r}

# review each dataset that we have, strategizing about how you'll use them to develop a obs-level file
# consider things like species scope (do I need to restrict all input data to just the 8ish game species?), file organization (is this file already in a obs-level format or is it a count of each species/size that I should uncount()?), linking keys (is there a fish obs ID in the age data that I can use to link to the fish observations data?), and what things (posisbly whole datasets) are unneeded for our work, here. I have left the Michigan work in here to give you an idea of what I did:


#aged fish
glimpse(il_aged_fish_surveys_28Dec2022)

il_aged_fish_surveys_28Dec2022 %>% 
  group_by(lake_name.1) %>% 
  count() 

#any na lengths? 
il_aged_fish_surveys_28Dec2022 %>% 
  filter(is.na(length.1)) %>% 
  group_by(survey_id) %>% 
  count()

#do the catch and aged data match for the surveys?
il_aged_fish_surveys_28Dec2022 %>% 
  group_by(year, survey_id, lake_name.1) %>% 
  count() %>% 
  print(n = nrow(.))
#only clinton and rend lake from 1985 - 2003

il_catch_age_effort_17Jan22 %>% 
  filter(lake_name.1 == "Clinton Lake" | lake_name.1 == "Rend Lake") %>% 
  group_by(survey_id, year) %>% 
  count() %>% 
  print(n = nrow(.))
#they aren't going to jive

#catch age effort
glimpse(il_catch_age_effort_17Jan22)

il_catch_age_effort_17Jan22 %>% 
  filter(!is.na(age)) %>% 
  glimpse()

il_catch_age_effort_17Jan22 %>% 
  filter(total_count.1 < 1) %>% 
  group_by(survey_id) %>% 
  count()

il_catch_age_effort_17Jan22 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_2)) 
#variable for sub-survey level effort ID

il_catch_age_effort_17Jan22 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_1))
  
#effort
glimpse(il_effort_17Jan22)
#efforts associated with lake/survey/date
#two efforts
#survey id matches with the catch file but not the aged fish surveys 

il_effort_17Jan22 %>% 
  filter(survey_id == "Survey_15") %>% 
  glimpse()

multieffort.effort <- il_effort_17Jan22 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_2)) %>%  
  filter(n >1)
#total effort 2 is a sub-survey level id for effort

multieffort.effort <- il_effort_17Jan22 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_1)) %>%  
  filter(n >1)

#lake info 
glimpse(il_lake_info_17Jan22)
#only has lake info (location, size, id)

#do survey ids match?
#do survey IDs match effort?
il_aged_fish_surveys_28Dec2022 %>% 
  group_by(lake_name.1, survey_id) %>% 
  summarise(n = n())
il_effort_17Jan22 %>% 
  group_by(lake_name.1, survey_id) %>% 
  summarise() %>% 
  print(n = 20)
il_catch_age_effort_17Jan22 %>% 
  group_by(lake_name.1, survey_id) %>% 
  summarise() %>% 
  print(n = 20)
il_catch_age_effort_17Jan22 %>%
  anti_join(il_effort_17Jan22, by = c("lake_name.1", "survey_id")) %>% 
   group_by(lake_name.1, survey_id) %>% 
   summarise(n = n())


#from this exploration, aged fish, lake info are of no use
#in order to use the effort from the catch file we must:
##1. find a survey level grouping that lumps all sub-survey levels (survey_id, year, sampling method)
##2.sum the sub-survey level effort (EF runs/number of nets)
##3. create a survey level identifier that lumps the sub-survey levels appropriately 

#exploring mismatching surveys that I found while merging fish level effort and survey level effort
effort.catch <- il_catch_age_effort_17Jan22 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1)

effort.effort <- il_effort_17Jan22 %>% 
    distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1)

mismatched.catch <- anti_join(effort.catch, effort.effort) %>% 
  rename(total_effort_2.catch = total_effort_2, 
         total_effort_1.catch =  total_effort_1)
mismatched.effort <- anti_join(effort.effort, effort.catch) %>% 
  rename(total_effort_2.effort = total_effort_2, 
         total_effort_1.effort =  total_effort_1)

final_mismatched_surveys <- bind_rows(
  mismatched.catch %>% select(survey_id, lake_id, year, date.1, sampling_method, gear_data_notes.1, total_effort_2.catch, total_effort_1.catch),
  mismatched.effort %>% select(survey_id, lake_id, year, date.1, sampling_method, gear_data_notes.1, total_effort_2.effort, total_effort_1.effort)
)

final_combined_surveys <- final_mismatched_surveys %>%
  group_by(
    survey_id, lake_id, year, date.1, sampling_method, gear_data_notes.1
  ) %>%
  summarize(
    total_effort_2.catch = sum(total_effort_2.catch, na.rm = TRUE),
    total_effort_1.catch = sum(total_effort_1.catch, na.rm = TRUE),
    total_effort_2.effort = sum(total_effort_2.effort, na.rm = TRUE),
    total_effort_1.effort = sum(total_effort_1.effort, na.rm = TRUE)
  ) %>% 
  mutate(total_effort_1.catch = case_when(total_effort_1.catch == 0 & total_effort_2.catch == 0 ~ NA,
                                          TRUE ~ total_effort_1.catch),
         total_effort_2.catch = case_when(is.na(total_effort_1.catch) ~ NA,
                                          TRUE ~ total_effort_2.catch),
         total_effort_1.effort = case_when(total_effort_1.effort == 0 & total_effort_2.effort == 0 ~ NA,
                                          TRUE ~ total_effort_1.effort),
         total_effort_2.effort = case_when(is.na(total_effort_1.effort) ~ NA,
                                          TRUE ~ total_effort_2.effort))
mismatched <- final_combined_surveys %>% 
  filter(!is.na(total_effort_1.catch) & !is.na(total_effort_2.effort))
nomatch <- final_combined_surveys %>% 
  filter(is.na(total_effort_1.catch) | is.na(total_effort_2.effort))


sur.503.catch <- effort.catch %>% 
  filter(survey_id == "Survey_503")

sur.503.effort <- effort.effort %>% 
  filter(survey_id == "Survey_503")
#survey503 has different effort in the catch and effort files

lake40.catch <- effort.catch %>% 
  filter(lake_id == "40")

lake40.effort <- effort.effort %>% 
  filter(lake_id == "40")
#several surveys that are mismatched within the same lake here

#sub effort exploration
il_catch_age_effort_17Jan22 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1,
           .keep_all = T) %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.1 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>% 
  group_by(subsample.time, sampling_method) %>% 
  count()%>% 
  print(n = nrow(.))
#only in efishing but 17% of them

il_catch_age_effort_17Jan22 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1,
           .keep_all = T) %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.1 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>% 
  group_by(garbage_bin_notes.1, sampling_method) %>% 
  count()%>% 
  print(n = nrow(.))

il_catch_age_effort_17Jan22 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1) %>% 
  count()

#effort
il_effort_17Jan22 %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.3 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>%
  group_by(subsample.time) %>% 
  count() %>% 
  print(n = nrow(.))

il_effort_17Jan22 %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.3 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>%
  group_by(subsample.time, sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))

il_effort_17Jan22 %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.3 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>%
  filter(subsample.time == "Y") %>% 
  group_by(garbage_bin_notes.4) %>% 
  count() %>% 
  print(n = nrow(.))
#most of the sub are for bluegill

il_effort_17Jan22 %>% 
  group_by(garbage_bin_notes.2, sampling_method) %>% 
  count()
```

A few columns to make sure I am keeping:
1. sub sample time and sub sample species (garbage bin notes)
2. community vs targeted sampling (garbage bin notes)

#Effort Merge
```{r}
#uncounting batch fish 
#makes fish per row from batch counted fish
il_catch_age_effort_17Jan22 <- il_catch_age_effort_17Jan22 %>% 
  uncount(total_count.1)

#collapsing effort from the obs level file 
survey.effort <- il_effort_17Jan22 %>% 
  #groups all sub-survey level effort into the larger survey
  group_by(survey_id, lake_id, year, date.1, sampling_method, gear_data_notes.1, total_effort_2, total_effort_1, original_file_name.1) %>% 
  #count of obs in each survey grouping
  count() %>% 
  #grouping by the survey id, year, and method to get survey level effort from sub-survey level 
  group_by(survey_id, lake_id, year, sampling_method, gear_data_notes.1, original_file_name.1) %>% 
  mutate(survey_effort = sum(total_effort_1),
         survey_effort_id = cur_group_id()) %>%  #gives a unique ident to this grouping
  rename(original_file_name.1_effort = original_file_name.1)

#quick gut check on survey-effort aggregation 
#are there any surveys that have more than one survey level effort?
survey.effort %>% 
  group_by(survey_id, sampling_method, gear_data_notes.1, survey_effort) %>% 
  count() %>% 
  group_by(survey_id, sampling_method, gear_data_notes.1) %>% 
  count() %>% 
  filter(n >1)

survey.effort %>% 
  filter(survey_id %in% c("Survey_1390", "Survey_684"))
#all of the duplicates have multiple years associated with the same effort so it wont be a problem
#surveys have two efforts for the survey id but they are in different years
#this does not hinder the join as date/year will be included

#joining the survey level effort with the individual fish obs
#this join adds the survey level effort and survey effort ident for each grouping
il_data <- left_join(survey.effort, il_catch_age_effort_17Jan22, by = c("survey_id", 
                                                                        "lake_id",
                                                                        "year", 
                                                                        "date.1", 
                                                                        "sampling_method", 
                                                                        "gear_data_notes.1",
                                                                        "total_effort_2", 
                                                                        "total_effort_1")) %>% 
  select(-n) %>% 
  rename(original_file_name.1_indvifish = original_file_name.1)
#this works, but all of the surveys that did not catch fish are missing important survey information 
#we want to know the surveys that did not catch fish, extract all information from the original file, put the information back in


##########saw an issue with loosing fish obs -checking that below#########
#taking a look at efforts without associated fish obs
no_matches <- anti_join(survey.effort, il_catch_age_effort_17Jan22, by = c("survey_id", 
                                                                        "lake_id",
                                                                              "year", 
                                                                              "date.1", 
                                                                              "sampling_method", 
                                                                              "gear_data_notes.1",
                                                                              "total_effort_2", 
                                                                              "total_effort_1"))
#this is an object that thas all of the survey that did not catch fish with all of the information
no_fish <- right_join(il_effort_17Jan22, no_matches, by = c("survey_id", 
                                                                        "lake_id",
                                                                              "year", 
                                                                              "date.1", 
                                                                              "sampling_method", 
                                                                              "gear_data_notes.1",
                                                                              "total_effort_2", 
                                                                              "total_effort_1"))
#lake id 4222 in 2015 has multiple surveys of the same effort at the same location
#returning no fish surveys back into the data
il_data <- il_data %>% 
  filter(!is.na(original_file_name.1_indvifish)) 
il_data <- bind_rows(il_data, no_fish)

#quick glance to see how net efforts are calculating to survey efforts
effort.check <- il_data %>% 
  distinct(survey_id, lake_id, year, date.1, sampling_method, total_effort_2, total_effort_1, .keep_all = T) %>% 
  select(survey_id, lake_id, lake_name.1, year, date.1, sampling_method, total_effort_2, total_effort_1, survey_effort)

#Are my original file names tracking?
il_data %>% 
  group_by(original_file_name.1_indvifish) %>% 
  count() %>% 
  print(n = nrow(.))
#this shows us that there are 227 records that have effort but are not associated with fish

il_data %>% 
  group_by(original_file_name.1_effort) %>% 
  count() %>% 
  print(n = nrow(.))
#all obs have effort associated with them

#fish that are in the catch but are not in the effort?
excluded <- anti_join(il_catch_age_effort_17Jan22, il_data, by = c("survey_id", 
                                                                   "lake_id",
                                                                   "year", 
                                                                   "date.1", 
                                                                   "sampling_method", 
                                                                   "gear_data_notes.1",
                                                                   "total_effort_2", 
                                                                   "total_effort_1"))
surveys.excluded <- excluded %>% 
  group_by(lake_id, lake_name.1, survey_id, year, sampling_method, gear_data_notes.1, total_effort_2, total_effort_1, species.1) %>% 
  count() 

il_effort_17Jan22 %>% 
  filter(survey_id == "Survey_1364") %>% 
  glimpse()

diff <- anti_join(il_catch_age_effort_17Jan22, il_data) %>% 
  distinct(survey_id, lake_id, year, sampling_method, gear_data_notes.1, original_file_name.1, total_effort_2, total_effort_1)
``` 
          
          
# Data tidying                        
```{r}                       
glimpse(il_data)

il_data %>% 
  group_by(state) %>% 
  count()

il_data %>% 
  group_by(county) %>% 
  count() %>% 
  print(n = nrow(.))
#might have to fix St.? do we spell it out?

il_data %>% 
  group_by(lake_name.1, lake_id) %>% 
  count() %>% 
  print(n = nrow(.))
#seems to be a problem with Pistakee Lake - several repeated rows
###most likely it is 4225 - the id with the most obs

il_data %>% 
  filter(lake_name.1 == "Pistakee Lake") %>% 
  group_by(lake_name.1, lake_id, survey_id, date.1, total_effort_1, survey_effort) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(date.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(survey_type.1) %>% 
  count() %>% 
  print(n = nrow(.))
#they are all impoundment surveys?

il_data %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))
#sampling method seems to be more sampling method agg

il_data %>% 
  group_by(gear_data_notes.1) %>% 
  count() %>% 
  print(n = nrow(.))
#gear notes 1 is more of the sampling method

il_data %>% 
  group_by(total_effort_2) %>% 
  count() %>% 
  print(n = nrow(.))
#this column needs to be renamed to sub-survey level effort id (net/shocking period)

il_data %>% 
  group_by(gear_data_notes.2) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(garbage_bin_notes.4) %>% 
  count() %>% 
  print(n = nrow(.))
#garbage bin 4 gives us a general survey purpose 

il_data %>% 
  group_by(garbage_bin_notes.1) %>% 
  count() %>% 
  print(n = nrow(.))
#gives subsample times - not clear to me what this is

il_data %>% 
  mutate(species.1 = str_to_lower(species.1),
         species.1 = str_replace_all(species.1, " ", "_")) %>% 
  group_by(species.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(species.1) %>% 
  filter(!is.na(length.1)) %>% 
  summarise(min = min(length.1), mean = mean(length.1), max = max(length.1))

il_data %>% 
  group_by(weight_unit.1) %>% 
  filter(!is.na(weight.1)) %>% 
  summarise(min = min(weight.1), mean = mean(weight.1), max = max(weight.1))

il_data %>% 
   mutate(weight_unit.1 = case_when(weight_unit.1 == "col_name_Wt(g)" ~ "g",
                                   TRUE ~ NA)) %>% 
  group_by(weight_unit.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(sex) %>% 
  count() %>% 
  print(n = nrow(.))
#no sex data

il_data %>% 
  group_by(reproductive_condition_notes) %>% 
  count() %>% 
  print(n = nrow(.))
#no maturity - useless column

il_data %>% 
  group_by(aging_structure.1) %>% 
  count() %>% 
  print(n = nrow(.))
#no aging structure data

il_data %>% 
  group_by(age) %>% 
  count() %>% 
  print(n = nrow(.))
#hardly any age data

il_data %>% 
  group_by(survey_id, total_effort_2, total_effort_1) %>% 
  count(survey_effort) %>% 
  print(n = nrow(.))

#####################final tidy#########################################
il_data <- il_data %>% 
  #lower case with no spaces for species name
  mutate(species.1 = str_to_lower(species.1),
         species.1 = str_replace_all(species.1, " ", "_")) %>% 
  #shorten unit to just g
  mutate(weight_unit.1 = case_when(weight_unit.1 == "col_name_Wt(g)" ~ "g",
                                   TRUE ~ NA)) %>%
  #effort units
  mutate(total_effort_1_units = "minutes",
         sub_effort_1_units = "minutes") %>% 
  #drop survey type and replace with garbage bin notes 4 for survey type
  select(-survey_type.1) %>% 
  rename(survey_type.1 = garbage_bin_notes.4,
         sub_sample_effort = garbage_bin_notes.1,
         sub_sample_species = ) %>% 
  mutate(survey_type.1 = gsub("Sample Type:", "", survey_type.1)) %>%  #cleans up column name
  #adding in length units
  #no units were specified within data but general assumption is mm
  mutate(length_unit.1 = "mm") %>% 
  #obvious typo of lake id Pistakee
  mutate(lake_id = case_when(lake_name.1 == "Pistakee Lake" ~ 4225,
                             TRUE ~ lake_id)) %>% 
  #renaming sampling hierarchy
  rename(sampling_method.1 = sampling_method) %>% 
  rename(sampling_method.2 = gear_data_notes.1) %>% 
  mutate(sampling_method.2 = gsub("Gear Used:", "", sampling_method.2)) %>% 
  #renaming effort hierarchy
  #this renames the survey level effort as total_effort_1 with the effort grouping noted by survey_effort_1_id
  #sub-survey level effort is named sub_effort_1 with sub-survey ids as sub_effort_1_id
  rename(sub_effort_1_id = total_effort_2) %>% 
  rename(total_effort_1_id = survey_effort_id) %>% 
  rename(sub_effort_1 = total_effort_1) %>% 
  rename(total_effort_1 = survey_effort) %>% 
  #making a caught nothing column to flag when no fish species were caught in a net/survey
  mutate(caught_nothing = case_when(is.na(species.1) ~ TRUE,
                                    TRUE ~ FALSE)) %>% 
  select(state, 
         county, 
         lake_name.1,
         lake_id,
         date.1,
         year,
         survey_id,
         lat_unspec,
         lon_unspec,
         survey_type.1,
         sampling_method.1,
         sampling_method.2,
         total_effort_1_id,
         total_effort_1,
         sub_effort_1_id,
         sub_effort_1,
         caught_nothing,
         total_effort_1_units,
         sub_effort_1_units,
         species.1,
         length.1,
         length_unit.1,
         weight.1,
         weight_unit.1,
         age,
         aging_structure.1,
         sex,
         original_file_name.1_indvifish,
         original_file_name.1_effort)
#no crosswalk for lake id exists for nhdids
#lat and long is retained for future cross walking

#all the columns appear to be present that we need
#need to clean up column naming (underscores for periods)
il_data <- il_data %>% 
  clean_names() 
```                        
                        
 
# Review & QC datasets
```{r}
#here I do some very basic checks on what the data structure and general outputs look like (i.e., s this thing behaving like the obs-level file I think it is?). MI work left here as an idea of a previous state's work. In my opinion, it is not our job to QC the actual observations at this point (like, is a WAE really going to be 500mm at age zero), but instead to use this QC as a check on the operations performed in this script.  

glimpse(il_data)

#how does the caught nothign column look?
il_data %>% 
  filter(caught_nothing == "TRUE") %>% 
  group_by(caught_nothing, lake_name_1, lake_id, survey_id, sub_effort_1_id) %>% 
  count() %>% 
  print(n = nrow(.))
#the only columns with an na lake name do not have a catch - this tell us the merge was done correctly

#how do the original file names track? - we should have 227 records of effort data without fish
il_data %>% 
  filter(is.na(original_file_name_1_indvifish)) %>% 
  glimpse()
#looks good

cpue <- il_data %>% 
  group_by(total_effort_1_id, species_1) %>% 
  mutate(n = n(),
         cpue = n/total_effort_1) %>% 
  distinct(survey_id, sampling_method_1, total_effort_1_id, total_effort_1, n, species_1, cpue)

#subeffort?
glimpse(il_data)
```


# Import/Export files
```{r}
str(il_data)

il_data <- as_arrow_table(il_data)

write_dataset(dataset = il_data, path = "Data_and_Scripts/Data/output/il_file_arrow")

il_data <- open_dataset("Data_and_Scripts/Data/output/il_file_arrow")

glimpse(il_data)
```

