---
title: "IL_Flat_File_Aggregation"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Preamble


Things to keep in mind:
-Need to explore subsample time 
    -are we able to link fish obs with if it was captured in the sub sample?
    -are they capturing all fish before they sub sample (including sub sample species)?
-make sure units on total effort make sense (seines might be hauls)


##Libraries
```{r}
library(arrow)
library(readr)
library(dplyr)
library(stringr)
library(data.table)
library(janitor)
library(tidyr)
library(lubridate)
library(bit64)

options(scipen = 999)
```


##Data
This could readily be changed into a function that takes a filepath and returns files into environment.
* note Holly has to change file paths to "D" 
```{r}
#generate a file list to import
files_list <- list.files(path = "D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IL_Data/il_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files
files_list



#object for use in loop (simple length of file list)
n <- length(files_list)

for(i in 1:n) {
  #i = 3
  filei <- word(gsub(".csv","", files_list[i]), start = -1, sep = fixed("/"))
  #this does those two steps in one package
  assign(filei ,
          fread(paste0("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IL_Data/il_raw_disaggregated_data/",
                                          files_list[i])))
  
  # if the file is a crosswalk, do not rename anything, just loop to the confirm import line
  if(str_detect(filei, "crosswalk")) {  #confirm import of files:  
    print(paste(filei ,"added to workspace" ))  
    #confirm import of files:  
    print(paste(i ,"files added to workspace" )) ; next}
  
  #if the file is not in the data explainer, don't try to rename it:
  if(filei %in% cde$new_file_name) {
    print("renaming with data explainer")
  } else {next}
  
  
  
  
  
  # note we want to review a sorted list of column names to check misspelling etc.
  # we still need to use the columns with names like col_name_length_in, or known_units
  
  
  cde %>% # call data explainer file
    filter(`new_file_name`== filei)%>% #keep only the row relevant to this file
    select_if(~ !any(is.na(.))) %>% 
    transpose(keep.names = "newname") %>% 
    rename("oldname" = V1) %>% 
    assign("names", ., envir = .GlobalEnv)
  
  #see if any column names will not have a match! 
  # IF any pop FALSE, force stop and revist of data explainer ()
  # - e.g., named something "total catch" when actual column name was "total_catch"
  print(
    cbind(colnames(get(filei)),
          colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]
    )
  )
  
  
  # break the loop if the current file has column names not in the data explainer
  # if (all(cbind(colnames(get(filei)),  colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ])[,2]) == FALSE ) break
  if (all(colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]) == FALSE ) break
  
  
  # append old col names into new "notes" columns:
  get(filei)[ , (names[ str_detect(newname, "notes") , oldname   ,  ]) := Map(paste, colnames(.SD), .SD, sep = ':') , .SDcols =  names[ str_detect(newname, "notes") , oldname   ,  ] ]
  
  #now rename that file's colnames
  setnames(get(filei), colnames(get(filei)), names[!str_detect(newname,"unique_row_key")] [match(names(get(filei)),names[!str_detect(newname,"unique_row_key"),oldname]), newname] )
  
  #append all other data from data explainer
  unusedbits <- 
    data.table(
      matrix(
        rep(names[ !newname %in% colnames(get(filei)) , oldname , ],
            each = nrow(get(filei))
        ),
        nrow = nrow(get(filei)),
        dimnames = list(rep(NA,nrow(get(filei))),
                        names[ !newname %in% colnames(get(filei)) , newname , ])
        )
      )
  
  #add all not yet used columns from data explainer:
  get(filei)[ , (names[ !newname %in% colnames(get(filei)) , newname , ]) := unusedbits[] ]

  #confirm import of files:  
  print(paste(filei ,"added to workspace" ))  
  #confirm import of files:  
  print(paste(i ,"files added to workspace" )) 

  
} 
  #confirm import of files:  
  print(paste(i ,"files added to workspace" ))
  #confirm import of files:  
  print(paste(n-i ,"remaining to be added" )) 



```


#Data Review
```{r}
# review each dataset that we have, strategizing about how you'll use them to develop a obs-level file
# consider things like species scope (do I need to restrict all input data to just the 8ish game species?), file organization (is this file already in a obs-level format or is it a count of each species/size that I should uncount()?), linking keys (is there a fish obs ID in the age data that I can use to link to the fish observations data?), and what things (posisbly whole datasets) are unneeded for our work, here. I have left the Michigan work in here to give you an idea of what I did:


#aged fish
glimpse(il_aged_fish_surveys_28Dec2022)

il_aged_fish_surveys_28Dec2022 %>% 
  group_by(lake_name.1) %>% 
  count() 

#any na lengths? 
il_aged_fish_surveys_28Dec2022 %>% 
  filter(is.na(length.1)) %>% 
  group_by(survey_id) %>% 
  count()

#do the catch and aged data match for the surveys?
il_aged_fish_surveys_28Dec2022 %>% 
  group_by(year, survey_id, lake_name.1) %>% 
  count() %>% 
  print(n = nrow(.))
#only clinton and rend lake from 1985 - 2003

il_catch_age_effort_30Nov23 %>% 
  filter(lake_name.1 == "Clinton Lake" | lake_name.1 == "Rend Lake") %>% 
  group_by(survey_id, year) %>% 
  count() %>% 
  print(n = nrow(.))
#they aren't going to jive

#catch age effort
glimpse(il_catch_age_effort_30Nov23)

il_catch_age_effort_30Nov23 %>% 
  filter(!is.na(age)) %>% 
  glimpse()

il_catch_age_effort_30Nov23 %>% 
  filter(total_count.1 < 1) %>% 
  group_by(survey_id) %>% 
  count()

il_catch_age_effort_30Nov23 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_2)) 
#variable for sub-survey level effort ID

il_catch_age_effort_30Nov23 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_1))
  
#effort
glimpse(il_effort_30Nov23)
#efforts associated with lake/survey/date
#two efforts
#survey id matches with the catch file but not the aged fish surveys 

il_effort_30Nov23 %>% 
  filter(survey_id == "Survey_15") %>% 
  glimpse()

multieffort.effort <- il_effort_30Nov23 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_2)) %>%  
  filter(n >1)
#total effort 2 is a sub-survey level id for effort

multieffort.effort <- il_effort_30Nov23 %>% 
  group_by(survey_id) %>% 
  summarise(n = n(), unique(total_effort_1)) %>%  
  filter(n >1)

#lake info 
glimpse(il_lake_info_17Jan22)
#only has lake info (location, size, id)

#do survey ids match?
#do survey IDs match effort?
il_aged_fish_surveys_28Dec2022 %>% 
  group_by(lake_name.1, survey_id) %>% 
  summarise(n = n())
il_effort_17Jan22 %>% 
  group_by(lake_name.1, survey_id) %>% 
  summarise() %>% 
  print(n = 20)



#from this exploration, aged fish, lake info are of no use
#in order to use the effort from the catch file we must:
##1. find a survey level grouping that lumps all sub-survey levels (survey_id, year, sampling method)
##2.sum the sub-survey level effort (EF runs/number of nets)
##3. create a survey level identifier that lumps the sub-survey levels appropriately 

#exploring mismatching surveys that I found while merging fish level effort and survey level effort
effort.catch <- il_catch_age_effort_30Nov23 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method.1, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1)

effort.effort <- il_effort_30Nov23 %>% 
    distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method.1, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1)

mismatched.catch <- anti_join(effort.catch, effort.effort) %>% 
  rename(total_effort_2.catch = total_effort_2, 
         total_effort_1.catch =  total_effort_1)
mismatched.effort <- anti_join(effort.effort, effort.catch) %>% 
  rename(total_effort_2.effort = total_effort_2, 
         total_effort_1.effort =  total_effort_1)

final_mismatched_surveys <- bind_rows(
  mismatched.catch %>% select(survey_id, lake_id, year, date.1, sampling_method.1, gear_data_notes.1, total_effort_2.catch, total_effort_1.catch),
  mismatched.effort %>% select(survey_id, lake_id, year, date.1, sampling_method.1, gear_data_notes.1, total_effort_2.effort, total_effort_1.effort)
)

final_combined_surveys <- final_mismatched_surveys %>%
  group_by(
    survey_id, lake_id, year, date.1, sampling_method.1, gear_data_notes.1
  ) %>%
  summarize(
    total_effort_2.catch = sum(total_effort_2.catch, na.rm = TRUE),
    total_effort_1.catch = sum(total_effort_1.catch, na.rm = TRUE),
    total_effort_2.effort = sum(total_effort_2.effort, na.rm = TRUE),
    total_effort_1.effort = sum(total_effort_1.effort, na.rm = TRUE)
  ) %>% 
  mutate(total_effort_1.catch = case_when(total_effort_1.catch == 0 & total_effort_2.catch == 0 ~ NA,
                                          TRUE ~ total_effort_1.catch),
         total_effort_2.catch = case_when(is.na(total_effort_1.catch) ~ NA,
                                          TRUE ~ total_effort_2.catch),
         total_effort_1.effort = case_when(total_effort_1.effort == 0 & total_effort_2.effort == 0 ~ NA,
                                          TRUE ~ total_effort_1.effort),
         total_effort_2.effort = case_when(is.na(total_effort_1.effort) ~ NA,
                                          TRUE ~ total_effort_2.effort))
mismatched <- final_combined_surveys %>% 
  filter(!is.na(total_effort_1.catch) & !is.na(total_effort_2.effort))
nomatch <- final_combined_surveys %>% 
  filter(is.na(total_effort_1.catch) | is.na(total_effort_2.effort))


sur.503.catch <- effort.catch %>% 
  filter(survey_id == "Survey_503")

sur.503.effort <- effort.effort %>% 
  filter(survey_id == "Survey_503")
#survey503 has different effort in the catch and effort files

lake40.catch <- effort.catch %>% 
  filter(lake_id == "40")

lake40.effort <- effort.effort %>% 
  filter(lake_id == "40")
#several surveys that are mismatched within the same lake here

#sub effort exploration
il_catch_age_effort_30Nov23 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method.1, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1,
           .keep_all = T) %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.1 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>% 
  group_by(subsample.time, sampling_method.1) %>% 
  count()%>% 
  print(n = nrow(.))
#only in efishing but 17% of them

il_catch_age_effort_30Nov23 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method.1, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1,
           .keep_all = T) %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.1 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>% 
  group_by(garbage_bin_notes.1, sampling_method.1) %>% 
  count()%>% 
  print(n = nrow(.))

il_catch_age_effort_30Nov23 %>% 
  distinct(survey_id, 
            lake_id,
             year, 
               date.1, 
                sampling_method.1, 
                gear_data_notes.1,
                total_effort_2, 
                 total_effort_1) %>% 
  count()

#effort
il_effort_30Nov23 %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.3 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>%
  group_by(subsample.time) %>% 
  count() %>% 
  print(n = nrow(.))

il_effort_30Nov23 %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.3 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>%
  group_by(subsample.time, sampling_method.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_effort_30Nov23 %>% 
  mutate(subsample.time = case_when(garbage_bin_notes.3 == "Subsample Time:0" ~ "N",
                                    TRUE ~ "Y")) %>%
  filter(subsample.time == "Y") %>% 
  group_by(garbage_bin_notes.4) %>% 
  count() %>% 
  print(n = nrow(.))
#most of the sub are for bluegill

il_effort_30Nov23 %>% 
  group_by(garbage_bin_notes.2, sampling_method.1) %>% 
  count()
```



Process to merge goes as follows:
1. Uncount fish that were binned in the fish obs file
2. Use the effort file to cacluate a total effort for a survey while still retaining sub effort 
3. Attach total effort to the fish as obs file
4. Retain surveys that caught no fish in the data
5. Review to make sure sub effort make up total efforts in a survey
6. Review surveys that might still have issues (ex. fish that are linking to a survey effort)

#Effort Merge
```{r}
#uncounting batch fish 
#makes fish per row from batch counted fish
il_catch_age_effort_30Nov23_uncount <- il_catch_age_effort_30Nov23 %>% 
  filter(survey_id != "Survey_355") %>% #date difference between effort and fish data that needs to be verified with Vic
  uncount(total_count.1)

#collapsing effort from the fish obs level file 
#this file is checked for conflicts between the effort file in a later step - keeping effort in this file makes joining cleaner
#I also retain surveys that didn't catch fish by finding surveys not present in the fish file that are present in the effort file and retaining them
survey.effort <- il_effort_30Nov23 %>% 
  filter(survey_id != "Survey_355") %>% 
  #groups all sub-survey level effort into the larger survey
  group_by(survey_id, lake_id, year, date.1, gear_data_notes.1, total_effort_2, total_effort_1, original_file_name.1) %>% 
  #count of obs in each survey grouping
  count() %>% #this count gives an idea of how many sub-survey go into a total survey but is dropped later
  #grouping by the survey id, year, and method to get survey level effort from sub-survey level 
  group_by(survey_id, lake_id, year, gear_data_notes.1, original_file_name.1) %>% 
  mutate(survey_effort = sum(total_effort_1),
         survey_effort_id = cur_group_id()) %>%  #gives a unique ident for total effort 
  rename(original_file_name.1_effort = original_file_name.1) %>% 
  select(-n)

#quick gut check on survey-effort aggregation 
#are there any surveys that have more than one survey level effort?
survey.effort %>% 
  group_by(survey_effort_id, survey_effort) %>% 
  count() %>% 
  group_by(survey_effort_id) %>% 
  count() %>% 
  filter(n >1)
#there are not two survey effort idents present for a given survey

#joining the survey level effort with the individual fish obs
#this join adds the survey level effort and survey effort ident for each grouping
il_data_merged <- left_join(survey.effort, il_catch_age_effort_30Nov23_uncount, by = c("survey_id",
                                                                                       "lake_id",
                                                                                       "year",
                                                                                       "date.1",
                                                                                       "gear_data_notes.1",
                                                                                       "total_effort_2",
                                                                                       "total_effort_1")) %>% 
  rename(original_file_name.1_indvifish = original_file_name.1)
#this works, but all of the surveys that did not catch fish are missing important survey information 
#we want to know the surveys that did not catch fish, extract all information from the original file, put the information back in

#finds surveys that are in the effort file but do not have associated fish
no_matches <- anti_join(survey.effort, il_catch_age_effort_30Nov23_uncount, by = c("survey_id",
                                                                                   "lake_id",
                                                                                   "year",
                                                                                   "date.1",
                                                                                   "gear_data_notes.1",
                                                                                   "total_effort_2",
                                                                                   "total_effort_1"))
#this is an object that has all of the survey that did not catch fish with all of the information
#I will bind this information to the the fish as rows object to ensure they are retained for cpe calculations
no_fish <- right_join(il_effort_30Nov23, no_matches, by = c("survey_id",
                                                            "lake_id",
                                                            "year", 
                                                            "date.1", 
                                                            "gear_data_notes.1",
                                                            "total_effort_2", 
                                                            "total_effort_1"))

#returning no fish surveys back into the data
il_data_all <- il_data_merged %>% 
  filter(!is.na(original_file_name.1_indvifish)) 
#this first step takes out surveys with no fish in the merged data so they can be returned with the object that returns all the survey information associated with surveys that did not catch fish

#this new object contains all fish data with associated effort and the surveys that did not catch fish 
il_data <- bind_rows(il_data_all, no_fish) %>% 
  ungroup() 

#quick glance to see how net efforts are calculating to survey efforts
effort.check <- il_data %>% 
  distinct(lake_id, year, date.1, sampling_method.1, total_effort_2, total_effort_1, .keep_all = T) %>% 
  select(lake_id, lake_name.1, year, date.1, sampling_method.1, total_effort_2, total_effort_1, survey_effort)

#Are my original file names tracking?
il_data %>% 
  group_by(original_file_name.1_indvifish) %>% 
  count() %>% 
  print(n = nrow(.))
#shows all fish records come from the original file for fish

na <- il_data %>% 
  filter(is.na(original_file_name.1_indvifish))
#all nas here are surveys that did not catch fish

il_data %>% 
  group_by(original_file_name.1_effort) %>% 
  count() %>% 
  print(n = nrow(.))
#all obs have effort associated with them from the effort file

#fish that are in the catch but are not in the effort?
excluded <- anti_join(il_catch_age_effort_30Nov23, il_data, by = c("lake_id",
                                                                   "year", 
                                                                   "date.1",
                                                                   "gear_data_notes.1",
                                                                   "total_effort_2", 
                                                                   "total_effort_1"))

surveys.excluded <- excluded %>% 
  group_by(lake_id, lake_name.1, survey_id, year, gear_data_notes.1, total_effort_2, total_effort_1, species.1) %>% 
  count() 
#one survey that has file records without any effort in the file 
#it looks like there is a difference in the date associated between the effort record and the fish records
``` 
          
          
# Data tidying                        
```{r}                       
glimpse(il_data)

il_data %>% 
  group_by(state) %>% 
  count()

il_data %>% 
  group_by(county) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(lake_name.1, lake_id) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  filter(lake_name.1 == "Pistakee Lake") %>% 
  group_by(lake_name.1, lake_id, survey_id, date.1, total_effort_1, survey_effort) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(date.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(survey_type.1) %>% 
  count() %>% 
  print(n = nrow(.))
#they are all impoundment surveys?

il_data %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))
#sampling method seems to be more sampling method agg

il_data %>% 
  group_by(gear_data_notes.1) %>% 
  count() %>% 
  print(n = nrow(.))
#gear notes 1 is more of the sampling method

il_data %>% 
  group_by(total_effort_2) %>% 
  count() %>% 
  print(n = nrow(.))
#this column needs to be renamed to sub-survey level effort id (net/shocking period)

il_data %>% 
  group_by(garbage_bin_notes.4) %>% 
  count() %>% 
  print(n = nrow(.))
#garbage bin 4 gives us a general survey purpose 

il_data %>% 
  group_by(garbage_bin_notes.1) %>% 
  count() %>% 
  print(n = nrow(.))
#gives subsample times 

il_data %>% 
  mutate(species.1 = str_to_lower(species.1),
         species.1 = str_replace_all(species.1, " ", "_")) %>% 
  group_by(species.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(species.1) %>% 
  filter(!is.na(length.1)) %>% 
  summarise(min = min(length.1), mean = mean(length.1), max = max(length.1))

il_data %>% 
  group_by(weight_unit.1) %>% 
  filter(!is.na(weight.1)) %>% 
  summarise(min = min(weight.1), mean = mean(weight.1), max = max(weight.1))

il_data %>% 
   mutate(weight_unit.1 = case_when(weight_unit.1 == "col_name_Wt(g)" ~ "g",
                                   TRUE ~ NA)) %>% 
  group_by(weight_unit.1) %>% 
  count() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(sex) %>% 
  count() %>% 
  print(n = nrow(.))
#not much sex data

il_data %>% 
  group_by(reproductive_condition_notes) %>% 
  count() %>% 
  print(n = nrow(.))
#no maturity - useless column

il_data %>% 
  group_by(aging_structure.1) %>% 
  count() %>% 
  print(n = nrow(.))
#limited aging data

il_data %>% 
  group_by(age) %>% 
  count() %>% 
  print(n = nrow(.))
#hardly any age data

il_data %>% 
  group_by(survey_id, total_effort_2, total_effort_1) %>% 
  count(survey_effort) %>% 
  print(n = nrow(.))
#how are sub effort s and total efforts looking?

#####################final tidy#########################################
il_data <- il_data %>% 
  #lower case with no spaces for species name
  mutate(species.1 = str_to_lower(species.1),
         species.1 = str_replace_all(species.1, " ", "_")) %>% 
  #shorten unit to just g
  mutate(weight_unit.1 = case_when(weight_unit.1 == "col_name_Wt(g)" ~ "g",
                                   TRUE ~ NA)) %>%
  #effort units
  mutate(total_effort_1_units = "minutes",
         sub_effort_1_units = "minutes") %>% 
  #drop survey type and replace with garbage bin notes 4 for survey type
  select(-survey_type.1) %>% 
  rename(survey_type.1 = garbage_bin_notes.4,
         sub_sample_effort = garbage_bin_notes.1) %>% 
  mutate(survey_type.1 = gsub("Sample Type:", "", survey_type.1)) %>%  #cleans up column name
  #adding in length units
  #no units were specified within data but general assumption is mm
  mutate(length_unit.1 = "mm") %>% 
  #renaming sampling hierarchy
  rename(sampling_method.1 = sampling_method.1) %>% 
  rename(sampling_method.2 = gear_data_notes.1) %>% 
  mutate(sampling_method.2 = gsub("Gear Used:", "", sampling_method.2)) %>% 
  #renaming effort hierarchy
  #this renames the survey level effort as total_effort_1 with the effort grouping noted by survey_effort_1_id
  #sub-survey level effort is named sub_effort_1 with sub-survey ids as sub_effort_1_id
  rename(sub_effort_ident = total_effort_2) %>% 
  rename(total_effort_ident = survey_effort_id) %>% 
  rename(sub_effort_1 = total_effort_1) %>% 
  rename(total_effort_1 = survey_effort) %>% 
  #making a caught nothing column to flag when no fish species were caught in a net/survey
  mutate(nothing_caught = case_when(is.na(original_file_name.1_indvifish) ~ TRUE,
                                    TRUE ~ FALSE)) %>%
  #retaining lake area in acres - can be removed if desired
  rename(lake_area_acres = garbage_bin_notes.3) %>% 
  mutate(lake_area_acres = gsub("Acres:", "", lake_area_acres)) %>% 
  #cleaning date
  mutate(date.1 = mdy(date.1)) %>% 
  #units of total effort - making net nights and hauls for nets/seines
  mutate(total_effort_1 = case_when(sampling_method.1 == "Nets" ~ total_effort_1/1440,
                                    TRUE ~ total_effort_1)) %>% 
  mutate(total_effort_1_units = case_when(sampling_method.1 == "Nets" ~ "net_nights",
                                          sampling_method.1 == "Seine Hauls" ~ "hauls",
                                          TRUE ~ total_effort_1_units)) %>% 
  #selecting columns to retain
  select(state, 
         county, 
         lake_name.1,
         lake_id,
         date.1,
         survey_id,
         lat_unspec,
         lon_unspec,
         survey_type.1,
         sampling_method.1,
         sampling_method.2,
         total_effort_ident,
         total_effort_1,
         sub_effort_ident,
         sub_effort_1,
         sub_sample_effort,
         nothing_caught,
         total_effort_1_units,
         sub_effort_1_units,
         species.1,
         length.1,
         length_unit.1,
         weight.1,
         weight_unit.1,
         age,
         aging_structure.1,
         sex,
         lake_area_acres,
         original_file_name.1_indvifish,
         original_file_name.1_effort)
#no crosswalk for lake id exists for nhdids
#lat and long is retained for future cross walking

#all the columns appear to be present that we need
#need to clean up column naming (underscores for periods)
il_data <- il_data %>% 
  clean_names() 
```                        
                        
 
# Review & QC datasets
```{r}
#here I do some very basic checks on what the data structure and general outputs look like (i.e., s this thing behaving like the obs-level file I think it is?). MI work left here as an idea of a previous state's work. In my opinion, it is not our job to QC the actual observations at this point (like, is a WAE really going to be 500mm at age zero), but instead to use this QC as a check on the operations performed in this script.  

glimpse(il_data)

#how does the caught nothing column look?
il_data %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(survey_id, total_effort_ident, sub_effort_ident, date_1) %>% 
  count() %>% 
  print(n = nrow(.))
#the only columns with an na lake name do not have a catch - this tell us the merge was done correctly
#important to note here - sub_effort_ident is taken from the agency itself (net number) - date is important to include because they could have sampled the same net or location several times in a total effort ident (within the same year)

#how do the original file names track? - we should have 174 records of effort data without fish
il_data %>% 
  filter(is.na(original_file_name_1_indvifish)) %>% 
  glimpse()
#looks good

cpue <- il_data %>% 
  group_by(total_effort_ident, species_1) %>% 
  mutate(n = n(),
         cpue = n/total_effort_1) %>% 
  distinct(sampling_method_2, total_effort_ident, total_effort_1, n, species_1, cpue)
```


# Import/Export files
```{r}
str(il_data)

il_data <- as_arrow_table(il_data)

write_dataset(dataset = il_data, path = "Data_and_Scripts/Data/output/il_file_arrow")

il_data <- open_dataset("Data_and_Scripts/Data/output/il_file_arrow")

glimpse(il_data)
```


