---
title: "MN_Flat_File_Aggregation"
author: "Holly Kundel & Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Preamble
updated 11/29
To Do:

DONE 4. Line 229 might offer a key to the large lakes missing data issue. Had a "total.count" val for those-- doubtful, but recheck this. 

DONE 1. where gear notes marked with dippers, bring back dipper # into end of gear desc.
DONE 2. Consider grabbing the mesh var?
DONE 3. Daylight EFing (gear data notes 2 - keep it around for filtering)
DONE 9. Move DL's hiving code up into the actual data agg scripts

DONE 11.We have numerous TEIs with multiple total efforts. That's not how total effort works. Fix.
DONE 10.Chris Rounds is finding Fall EF dates that are actually reporting the lakefinder date for this survey rather than the actual observation date.
        How are we currently handling date, are we using the “survey_date” and dumping the “observation_date” and losing info?
        See hive date setups
DONE 13. MOve lake area data over to the agg script (trims lots of NAs in other states)
DONE 15. Clean up sex column Male, Female, Unknown, NA (not reported, not sexed, data lost, all those things)
DONE 16. QC on age column
DONE 17. negative vals on sample ids (whats that mean and should they be removed?)
DONE 18. Add sampling method simple column to hive
DONE 24. Add flags for 0 efforts and high efforts
DONE 19. Fix units to match hive structure, make sure anything with na values are also na for the unit (e.g., units like net-nights -> net_nights)
DONE 25. Auto Flagging: (build a flagging template from the MW Hive then apply it state-by-state)
          0 and negative efforts
          10SDs above mean effort? (by gear/state?)
          Lengths (prob group by age/survey/something) - big and small
          Weights (prob group by age/survey/something)
          Counts (by TEI)
DONE 20. Where values are NA, units should also be NA (efforts, lengths, weights, others?)
DONE 22. Check N species in a survey (is 44 spp a reasonable finding?)
DONE 23. Check that our flagging is spelling out implications
DONE 7. Lake 28003200|1996|Trap has 2 idents one TEInothing caught=T and TEInothingcaught=F
        
        
NOT COMPLETELY RESOLVED:        
12.We have many TEIs tagged with nothing caught)
          Large Lakes data R/R with Kristen [somethings]'s data that she agg'd 
          Built a script in the project that lays out a few other cases of weirdness (white bear lake 2022 trap nets, Willard Lake, Cougar Lake) 
          Corey sees the same un-entered data we do in their data entry dashboard
ACTION    Send all Nothing caught = True total efforts off to Jon Hansen
CHECK:    Historical WAE survey duplication
          - check on outcome of fixes (section name: sleuth out the MN duplicated data issues)

5. In 13 Aug cpue export, State ==NA, total effort ident = NA, Note that in the filtered cpe data file these records are NA-laden
    Pulled example: feather lake, 06/10/1992, gill net, total effort ident = 44170. Only one observation, species is “no particular species”, survey type is historical

14. Grace found whole surveys-lake-year that were excluded from chris custer cpe output but were present in lake finder 
    Examples sent to denver to explore what might be happening
      LOW 2013, 1997, and 1996 not showing up in CPE
        2013 has a flag
        96 and 97 have NA dates - all historical surveys?
      Artichoke 2021 gillnet not showing up in CPE
        Total effort of 3 with an area of 1501+ (effort is confirmed on lake finder)





libraries
```{r}
library(arrow)
library(readr)
library(tidyverse)
library(stringr)
library(data.table)
library(janitor)
library(tidyr)
library(lubridate)
library(bit64)
library(ggplot2)
library(mwlaxeref)

options(scipen = 999)
```




# Load Data

* note depending on your local settings, file paths need correction to match local GDrive letter

```{r}
##build functions  
remove_unnecessary_suffixes <- function(dt) {
  # Identify columns with numeric suffixes
  numeric_suffix_cols <- grep("\\.\\d+$", names(dt), value = TRUE)
  
  # Create a list to store prefixes
  prefixes <- unique(sub("\\.\\d+$", "", numeric_suffix_cols))
  
  # Remove unnecessary numeric suffixes
  for (prefix in prefixes) {
    cols <- grep(paste0("^", prefix, "\\.\\d+$"), names(dt), value = TRUE)
    if (length(cols) == 1) {
      setnames(dt, cols, sub("\\.\\d+$", "", cols))
    }
  }
  
  return(dt)
}

remove_unnecessary_suffixes_2 <- function(dt) {
  # Identify columns with numeric suffixes
  numeric_suffix_cols <- grep("\\_\\d+$", names(dt), value = TRUE)
  
  # Create a list to store prefixes
  prefixes <- unique(sub("\\_\\d+$", "", numeric_suffix_cols))
  
  # Remove unnecessary numeric suffixes
  for (prefix in prefixes) {
    cols <- grep(paste0("^", prefix, "\\_\\d+$"), names(dt), value = TRUE)
    if (length(cols) == 1) {
      setnames(dt, cols, sub("\\_\\d+$", "", cols))
    }
  }
  
  return(dt)
}

#generate a file list to import
files_list <- list.files(path = "D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files
files_list

#object for use in loop (simple length of file list)
n <- length(files_list)

for(i in 1:n) {
  #i = 3
  outer_break <- FALSE# clean up so that break tool will refresh
  
  filei <- word(gsub(".csv","", files_list[i]), start = -1, sep = fixed("/"))
  #this does those two steps in one package
  assign(filei ,
          fread(paste0("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/",
                                          files_list[i])))
  
  # if the file is a crosswalk, do not rename anything, just loop to the confirm import line
  if(str_detect(filei, "crosswalk")) {  #confirm import of files:  
    print(paste(filei ,"added to workspace" ))  
    #confirm import of files:  
    print(paste(i ,"files added to workspace" )) ; next}
  
  
  # note we want to review a sorted list of column names to check misspelling etc.
  # we still need to use the columns with names like col_name_length_in, or known_units
  
  
  cde %>% # call data explainer file
    filter(`new_file_name`== filei)%>% #keep only the row relevant to this file
    select_if(~ !any(is.na(.))) %>% 
    data.table::transpose(keep.names = "newname") %>% 
    rename("oldname" = V1) %>% 
    assign("names", ., envir = .GlobalEnv)
  
  #see if any column names will not have a match! 
  # IF any pop FALSE, force stop and revist of data explainer ()
  # - e.g., named something "total catch" when actual column name was "total_catch"
  print(
    cbind(colnames(get(filei)),
          colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]
    )
  )
  
  
  # break the loop if the current file has column names not in the data explainer
  # if (all(cbind(colnames(get(filei)),  colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ])[,2]) == FALSE ) break
  if (all(colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]) == FALSE ) {
    warning(paste("Warning: Import renaming failed at file", filei))
    outer_break <-  TRUE
    break}
  
  
  # append old col names into new "notes" columns:
  get(filei)[ , (names[ str_detect(newname, "notes") , oldname   ,  ]) := Map(paste, colnames(.SD), .SD, sep = ':') , .SDcols =  names[ str_detect(newname, "notes") , oldname   ,  ] ]
  
  #now rename that file's colnames
  setnames(get(filei), colnames(get(filei)), names[!str_detect(newname,"unique_row_key")] [match(names(get(filei)),names[!str_detect(newname,"unique_row_key"),oldname]), newname] )
  
  #append all other data from data explainer
  unusedbits <- 
    data.table(
      matrix(
        rep(names[ !newname %in% colnames(get(filei)) , oldname , ],
            each = nrow(get(filei))
        ),
        nrow = nrow(get(filei)),
        dimnames = list(rep(NA,nrow(get(filei))),
                        names[ !newname %in% colnames(get(filei)) , newname , ])
        )
      )
  
  #add all not yet used columns from data explainer:
  get(filei)[ , (names[ !newname %in% colnames(get(filei)) , newname , ]) := unusedbits[] ]

  rm(names, unusedbits)
  
  
  #remove the unneeded suffixes from data
  assign(filei, remove_unnecessary_suffixes_2(remove_unnecessary_suffixes(get(filei))))
  
  #confirm import of files:  
  print(paste(filei ,"added to workspace" ))  
  #confirm import of files:  
  print(paste(i ,"files added to workspace" )) 

  
} 
if (outer_break) {print("Data import issue");rm(outer_break)
  } else {
  #confirm import of files:  
  print(paste(i ,"files added to workspace" ))
  #confirm import of files:  
  print(paste(n-i ,"remaining to be added" )) 
}
  
  
```


# Effort

```{r}

# effort dataset has a unique key in the combo of (lake_id, date.2(survey date), gear, survey type)
mn_effort_18aug2023[          , .N , .(lake_id, date.1, sampling_method, survey_type) ]

#note that there is a component count column in here that seems like it's related to effort, but not exactly clear to me:
mn_effort_18aug2023[ , summary(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2))) ,]
mn_effort_18aug2023[ , hist(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), breaks = 100) ,]
#actually, these appear to be like total.count
mn_effort_18aug2023[ ,plot(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), total_count.1) , ]#CPUE included count
mn_effort_18aug2023[ ,plot(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), total_count.2) , ]#cpue excluded count
mn_effort_18aug2023[ ,plot(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), total_count.1+total_count.2) , ] #boom! there is is, component count = the sum of catch between fish marked to exclude form CPUE and those marked to include
mn_effort_18aug2023[ !(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)) == total_count.1+total_count.2) ,, ]


# we have two total effort cols, one is EF seconds, which I think is redundant to the main effort column:
mn_effort_18aug2023[ , total_effort_2 := gsub("EF_SECONDS:", "" , total_effort_2), ]
mn_effort_18aug2023[ total_effort_2 == "NA" , total_effort_2 := NA  ,]
mn_effort_18aug2023[  , total_effort_2 := as.numeric(total_effort_2) ,] #one weird one we'll ignore (retain total.count, drop component count implicity by process)


mn_effort_18aug2023[ , .N , .("te2NA" = is.na(total_effort_2), "te1NA" = is.na(total_effort_1))] # any cases where all dat in total.effort.2
#NOPE

#here we can see that ef time is covered by the total effort 1 column
plot(mn_effort_18aug2023[ !is.na(total_effort_2) , total_effort_2,]/3600~ mn_effort_18aug2023[ !is.na(total_effort_2) , total_effort_1 ,])
#what about the one's where they don't jive?
mn_effort_18aug2023[ round(total_effort_2/3600, 2) != total_effort_1, plot(total_effort_2/3600~total_effort_1), ]
abline(0,1)

#other weirdness?
mn_effort_18aug2023[ , summary(total_effort_1) , ]
mn_effort_18aug2023[total_effort_1 == 0, ]#review of these records indicates they oughtta stay in, effort moved over to tot eff 1
mn_effort_18aug2023[total_effort_1 == 0, total_effort_1 := total_effort_2]

mn_effort_18aug2023[is.na(total_effort_1)]


#we don't need that extra total eff column
mn_effort_18aug2023[ , total_effort_2 := NULL , ]

#the DNR sent along two values for counts, one where they count only cases where CPUE was marked yes in the indiv fish data and one count including only where cpue was marked no in the indiv fish data
#in the effort data thius is the number of units of gear that were used (and thus number of units marked us/don't use for cpues)

mn_effort_18aug2023[total_count.2>0 , .(total_count.1,total_count.2) , ]
mn_effort_18aug2023[ , plot(total_count.1~total_count.2) , ]

#are there any effort data with NO associated count (i.e., nothing caught)? 
mn_effort_18aug2023[   , .N , .("tc1_0" = total_count.1 == 0, "tc2_0" = total_count.2 == 0) ]
#NOPE

# Crazy that there are "no zeros". However, because these data are gear level, this is agg'd across species so relatively useless.

#drop these columns
mn_effort_18aug2023[ , `:=` ("total_count.1" = NULL,"total_count.2" = NULL) , ]


#fix up dates
mn_effort_18aug2023[ , date_survey :=  as.IDate(word(date.1, 1, sep = fixed(" ")), format = "%m/%d/%Y"),]
mn_effort_18aug2023[ is.na(date_survey), ]

mn_effort_18aug2023[ , hist(yday(date_survey)) ,]
mn_effort_18aug2023[ , hist(year(date_survey)) ,]

#check other fields
names(mn_effort_18aug2023)


#tidy sampling methods
mn_effort_18aug2023[ , .N , sampling_method ]
mn_effort_18aug2023[ , .N , .(sampling_method, sampling_method_abbrev)]

#keep the abbreviation/code column because that's the only thing in the CPUE file
mn_effort_18aug2023[ , .N , sampling_method_abbrev , ]
mn_cpue_21aug2023[ , .N , sampling_method_abbrev]

any(is.na(match(mn_effort_18aug2023[ , unique(sampling_method_abbrev),], mn_cpue_21aug2023[ , unique(sampling_method_abbrev) , ])))
#all abbreviations are covered (we could come back to that to add full gears names field to the cpues)

sort(names(mn_effort_18aug2023))

#year
mn_effort_18aug2023[ , unique(year) , ]
mn_effort_18aug2023[ is.na(year),]
 sum(mn_effort_18aug2023[ ,year != year(date_survey)])
#non-issue, delete year column
mn_effort_18aug2023[ , year := NULL] 

#effort units
mn_effort_18aug2023[ , .N , sampling_method ]


mn_effort_18aug2023[ , summary(total_effort_1) ,]

mn_effort_18aug2023[ total_effort_1 <0 , , ]

ggplot( mn_effort_18aug2023[!total_effort_1 < 0 ], aes(total_effort_1, group = sampling_method))+
  geom_density()+
  facet_wrap(~sampling_method, scales = "free")

# we can get units for these effort data from here (p45ish):
# https://files.dnr.state.mn.us/publications/fisheries/special_reports/180.pdf




```



# CPUE 2(3) ways
We have CPUE directly form the DNR, we can generate our own cpue using the individual fish data (both including and excluding the fish marked for CPUE inclusion)
Update 15 Sep 2023: We recieved a new data rip from MN DNR in three pieces: effort, catch, and cpue. Why catch and CPUE you ask? Because the DNR tells me that there are historical surveys (pre ~1985) that have CPUEs but no data for the indiv fish caught in those surveys. 

```{r}

  #cpue data
  mn_cpue_21aug2023[  , .N , .(lake_id, date, sampling_method_abbrev, survey_type, species) ]
  sum(duplicated(mn_cpue_21aug2023))
  #not really sure why there would be duplicated cpues in there... 
  mn_cpue_21aug2023[duplicated(mn_cpue_21aug2023) , .N, .(lake_name, lake_id, date) ]
  #if these cpues are calc'd per net or instance, that could cause it. how many of each gear in each survey?
  mn_cpue_21aug2023[ , .N, .(lake_name, lake_id, date, species, sampling_method_abbrev) ][, summary(N)]
  mn_cpue_21aug2023[ ,mean(total_effort), .(sampling_method_abbrev) ]
  mn_cpue_21aug2023[ , .N, .(lake_name, lake_id, date, species, sampling_method_abbrev) ][N>1, ]

  #not likely a problem. I think sometimes multiple rows are reported for a given gear on a survey, when the catch is == between two of them you get an aparrent "duplicate". This would be mitigated by a sub_effort_ident coming from the db. 
  
  #we'll be circling back to this dataset as a way to get the historical un-lengthed fish data into our fish-obs-as-rows format
  
  
  #fix up dates
  mn_cpue_21aug2023[ , date_survey :=  as.IDate(date),]
  mn_cpue_21aug2023[ is.na(date_survey), ]

  mn_cpue_21aug2023[ , hist(yday(date_survey)) ,]
  mn_cpue_21aug2023[ , .N ,.(lake_name,lake_id,year(date_survey))][ , hist(year) ,]
  
  #compared to the effort file, it's obvious there is overlap in these and that the CPUE stuff has more old data
  mn_effort_18aug2023[ ,.N ,.(lake_name,lake_id,year(date_survey))][ , hist(year) ,]
  
  
  #check overlap betweek effort and CPUE:
  mn_cpue_21aug2023[            , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]
  mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]

   mn_cpue_21aug2023[            , .("cpueN" = .N) , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ][
     mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , effortN := "N" 
   ][ , .N , .("missingCPUE" = is.na(cpueN), "missing_effort" = is.na(effortN))] #
   

   
  mn_cpue_21aug2023[            , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]
  mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]

   mn_cpue_21aug2023[            , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ][
     mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , 
   ]
  
  # Awesome-- the effort data are entirely contained in the cpue file. So here's a plan: 
   # 1. we use the fish obs and effort to set up our obs level file
   #  2a. Identify CPUE rows unmatched in the obs level surveys (for which we have nothing in effort or indiv fish). 
   #  2b. expand those rows into obs level format (but no wieghts or lengths)
   #  2c. bring them back to that main obs level file
  
```


# Prep & Scope Indiv Fish Data

```{r}

#inspect data
colnames(mn_indiv_fish_23nov2023)
sort(colnames(mn_indiv_fish_23nov2023))

setcolorder(mn_indiv_fish_23nov2023, c("lake_id", "date.2", "date.1", "sampling_method_abbrev", "survey_type")) 

mn_indiv_fish_23nov2023[ , .N , .(sampling_method_abbrev, species)  ]
mn_indiv_fish_23nov2023[ species == "WAE", .N , .(sampling_method_abbrev)  ] 
mn_indiv_fish_23nov2023[ species == "WAE", .N , .(sampling_method_abbrev)  ][ , sum(N) , ]

#to-do:
# make these clean: "lake_id", "date_clean", "sampling_method", "survey_type.1"

#dates (from data explainer we know that date.2 is survey date, we prob want to join on that one to connect to the effort data)
mn_indiv_fish_23nov2023[ , .(date.2, date.1), ]
mn_indiv_fish_23nov2023[ , .N , .(date1=is.na(date.1), date2=is.na(date.2)) ]


#change dates to Idate:
mn_indiv_fish_23nov2023[ , date_survey := as.IDate(date.2, format = "%m/%d/%Y") ,]
mn_indiv_fish_23nov2023[ , date_sample := as.IDate(date.1, format = "%m/%d/%Y") ,]


mn_indiv_fish_23nov2023[  , summary(date_sample - date_survey)   ,  ]
mn_indiv_fish_23nov2023[ (abs(date_sample - date_survey) < 400) == TRUE , hist(date_sample - date_survey)   ,  ]

mn_indiv_fish_23nov2023[ , summary(date_survey) , ]
mn_indiv_fish_23nov2023[is.na(date_survey) , .N ]
#no missingness!

mn_indiv_fish_23nov2023[ , summary(date_sample) , ]
mn_indiv_fish_23nov2023[is.na(date_sample) , .N ]
#no missingness!


#lake_id
mn_indiv_fish_23nov2023[ , summary(lake_id) ,]
mn_indiv_fish_23nov2023[ , str(lake_id), ]
mn_effort_18aug2023[ , str(lake_id) ,]
#no issues here


#tidy sampling methods
mn_indiv_fish_23nov2023[ , .N , .(sampling_method_abbrev)]
mn_effort_18aug2023[ , .N , .(sampling_method, sampling_method_abbrev)]


any(is.na(match(mn_effort_18aug2023[ , unique(sampling_method_abbrev),], mn_indiv_fish_23nov2023[ , unique(sampling_method_abbrev) , ])))
#all abbreviations are covered (we could come back to that to add full gears to the indiv fish, but that'll happen in the merge, so skipped here. )


#survey.type
mn_indiv_fish_23nov2023[ , .N , survey_type ]
mn_effort_18aug2023[ , .N , survey_type]

mn_indiv_fish_23nov2023[ , .N , survey_type ][ ,survey_type, ]


# generate a type table (these codes can be found in MN_Data folder)
# Assuming you have the data in a vector or a data frame column named 'codes'
codes <- c("IS", "RS", "PA", "SA", "CR", "WK", "DO", "GM", "CP", "LL", "SS", "NR", "OT", "RE", "XM", "XR", "SD", "TS")
full_names <- c("Initial Survey", "Re-Survey", "Population Assessment", "Special Assessment", "Creel Survey", "Winter Kill Assessment", 
                "Dissolved Oxygen Check", "Game Survey", "Collectors Permit", "Large Lake Survey", "Subsurvey", "Natural Reproduction Check", 
                "Investigational Survey", "Research Survey", "External Management Survey", "External Research Survey", "Standard Survey", "Targeted Survey")

# Create a data.table
survey_type_key <- data.table(code = codes, fullname = full_names)

# Print the resulting data.table
print(survey_type_key)


#execute
any(is.na(match(mn_indiv_fish_23nov2023[ , survey_type], survey_type_key[,code]))) #add codes are in key
# survey_type_key[match(mn_indiv_fish_23nov2023[ , survey_type], survey_type_key[,code]), fullname]

mn_indiv_fish_23nov2023[ , survey_type := survey_type_key[match(mn_indiv_fish_23nov2023[ , survey_type], survey_type_key[,code]), fullname]   ,]
rm(survey_type_key)

#uncount indiv_fish
mn_indiv_fish_23nov2023[, summary(total_count) , ]
mn_indiv_fish_23nov2023[total_count == 0 , .N , ]

#expected result size:
mn_indiv_fish_23nov2023[ , sum(total_count) , ] + mn_indiv_fish_23nov2023[total_count == 0 , .N , ]


#execute dataset expansion/uncount
mn_indiv_fish_23nov2023 <- 
rbind(
  mn_indiv_fish_23nov2023[total_count == 0 , , ], 
  uncount(mn_indiv_fish_23nov2023, weights = total_count, .remove = F)
)



#summarize ages
mn_indiv_fish_23nov2023[ , .N , age ]
mn_indiv_fish_23nov2023[ , .N , .(age=!is.na(age))]


```



# Merge effort and indiv fish
```{r}
#merge the effort and indiv fish data


#scope the merge:
#99% of the wb in the effort data are represneted in indiv fish data
sum(mn_indiv_fish_23nov2023[ , unique(lake_id) , ] %in% mn_effort_18aug2023[ , unique(lake_id)])/mn_effort_18aug2023[, length(unique(lake_id)) ,] 

#100% of the wb in the indiv fish data are represneted in  effort data
sum(mn_indiv_fish_23nov2023[ , unique(lake_id) , ] %in% mn_effort_18aug2023[ , unique(lake_id)])/mn_indiv_fish_23nov2023[, length(unique(lake_id)) ,] 

#merge the effort and indiv fish data

intersect(names(mn_effort_18aug2023), names(mn_indiv_fish_23nov2023))

setdiff(names(mn_effort_18aug2023), names(mn_indiv_fish_23nov2023))

# setnames(mn_effort_18aug2023, old = "date_clean", new = "date_survey")

#MEMORY HOG PINCH POINT
#we know that dates are a problem, can we schmooze that join? 
#we leave many column names out of this merge key to avoid disconnecting things where one dataset has empty info
flatfish <- merge(mn_effort_18aug2023,mn_indiv_fish_23nov2023, by = c("lake_id", "date_survey" , "sampling_method_abbrev", "survey_type"), suffixes = c("_effort", "_indivfish"), all = T)

rm(mn_indiv_fish_23nov2023,mn_effort_18aug2023)

flatfish[ , .N , sampling_method]
flatfish[ , .N, species]


#check on the annual #s of aged, lengthed, both, and neither numbers
a <- setnames(
  dcast(flatfish[ (species == "WAE" & sampling_method %in% c("Standard gill net sets")), .N, .(year = year(date_survey), age = !is.na(age), length = !is.na(length))][order(year, -N)],
        formula = year ~ age + length, fill = 0 ),
  old = c("FALSE_FALSE", "FALSE_TRUE", "TRUE_TRUE"), new = c("neither", "length_only", "both" ))


names(flatfish)

#how many unique surveys exist for each category of has effort and has indiv fish data?
flatfish[ , length(unique(paste(lake_id, date_survey , sampling_method, survey_type))) , .(effort = !is.na(state_effort), indiv_fish_dat = !is.na(state_indivfish)) ]
  
#sum of all
  flatfish[ , length(unique(paste(lake_id, date_survey , sampling_method, survey_type))) , .(effort = !is.na(state_effort), indiv_fish_dat = !is.na(state_indivfish)) ][ , sum(V1) , ]

  
  
#now roll the catch CPUE data back to this bigger file:
  
##Context for this next operation:
#   3. Historical WAE survey duplications show same gear/lake/date cathing many (from CPUE) and zero (from indivfish)
# - Consider dropping the effort table info where we have a HISTORICAL cpue record (or at least scope this strategy)


     
  #We've got a bad survey type thingy going on here where cpue is historical:
   mn_cpue_21aug2023[ , .N , survey_type ]
   flatfish[, .N , survey_type]
   
   # consider dropping records from the effort table where we have a cpue in a historical?
   #send this off to Corey Geving to review 
   
   # setnames(mn_cpue_21aug2023, old = "date_clean", new = "date_survey")
   
  flatfish[mn_cpue_21aug2023[survey_type == "HISTORICAL", , ], on = .(lake_id, date_survey, sampling_method_abbrev) , nomatch = NULL ][ , .(neffcatch = .N, n_cpue = sum(i.total_count), nlen = sum(!is.na(length))) , .(lake_id, date_survey, sampling_method_abbrev, survey_type, i.survey_type) ][order(lake_id,date_survey)] 
  
  
  #for these records we want to push the flatfish record out completely. 
  badflatfish <- flatfish[mn_cpue_21aug2023[survey_type == "HISTORICAL", , ], on = .(lake_id, date_survey, sampling_method_abbrev) , nomatch = NULL ][ , .(neffcatch = .N, n_cpue = sum(i.total_count), nlen = sum(!is.na(length))) , .(lake_id, date_survey, sampling_method_abbrev, survey_type, i.survey_type) ][order(lake_id,date_survey)][nlen %in% c(0), ]
  
  
  flatfish <- flatfish[!badflatfish, on = .(lake_id, date_survey, sampling_method_abbrev,survey_type) ]
  
  rm(badflatfish)

  #subset of MN_cpue where lake/date/method are not matched:
   glimpse(mn_cpue_21aug2023[ 
     !flatfish[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , 
   ])
  
  #trim to only the not already covered data:
  mn_cpue_21aug2023 <- mn_cpue_21aug2023[ 
     !flatfish[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , 
   ]
  
  mn_cpue_21aug2023[ , summary(total_count) , ]
  mn_cpue_21aug2023[total_count == 9999 ]
  
  
  
  #execute dataset expansion/uncount
mn_cpue_21aug2023 <- 
rbind(
  mn_cpue_21aug2023[total_count == 0 , , ], 
  uncount(mn_cpue_21aug2023, weights = total_count, .remove = F)
)

names(flatfish)


  #MEMORY HOG STEP
  #now bring that together with the flatfish data:
  flatfish <- rbindlist(list(flatfish, mn_cpue_21aug2023 ),
                               fill = TRUE,
                               use.names = TRUE)
  
  rm(mn_cpue_21aug2023)
  
  #for each of those with both effort and fish data, summarize catch by species (bring effort along too!)   
flatfish[ !is.na(state_effort)&!is.na(state_indivfish) , .N , .(lake_id, date_survey , sampling_method, survey_type, species, total_effort_1) ]


  
```


# Save an rds to work from 
```{r}





# saveRDS(flatfish, file = "Data_and_Scripts\\Data\\output\\mn_flat_effort_indivfish_merge.rds")


#this flatfish file is big AF and will slow down Mike's Hansen Lab Computer (32gb RAM)
# flatfish <- readRDS("Data_and_Scripts/Data/output/mn_flat_effort_indivfish_merge.rds")






```


#Gear codes, species codes, state, county, length age units, target spp
```{r}
#we need to
# 1. revises the data in a few columns: sampling method, nothing caught, species names, state, units, target species, effort,  
# 2. export notes for the workshop. (delete, flag, clarify, keep for workshop) then dump those columns
# 3. match to schema (https://docs.google.com/spreadsheets/d/1zfevASMxRMxMYNWm3Hq1yR2Qk0zcJV_JnO0LBeODaPQ/edit#gid=0)

mn_dat <- flatfish
rm(flatfish)
names(mn_dat)

# dump some known trash

mn_dat[, c("targeted_or_standard_effort", "targeted_or_standard_indivfish", "targeted_or_standard",
           "new_file_name_effort", "new_file_name_indivfish", "new_file_name",
           "source_contact_effort", "source_agency", "source_agency_effort", "source_contact_indivfish", "source_agency_indivfish", "source_contact",
           "file_status_effort", "file_type_effort", "file_status_indivfish", "file_type_indivfish", "file_status", "file_type",
           "jan2023_inventoried_effort", "jan2023_inventoried_indivfish", "jan2023_inventoried",
           "data_type_effort", "data_type_indivfish", "data_type",
           "date_recieved_effort", "date_recieved_indivfish", "date_recieved", 
           "unique_row_key_effort", "unique_row_key_indivfish", "unique_row_key",
           "total_count") := NULL, ]


#gear full names need filling out
    mn_dat[ , .N , .(sampling_method_abbrev, sampling_method)]
    
    #borrow codes as key from other rows
    gear_table_mn <- mn_dat[!is.na(sampling_method), .N , .(sampling_method_abbrev,sampling_method) ]
    
    
    #apply these to all rows
    #drop old col
    mn_dat[ , sampling_method := NULL ,]
    #rejoin to complete gear table
    mn_dat[gear_table_mn,
      on = .(sampling_method_abbrev), sampling_method := sampling_method
          ]
    #checkwork
    mn_dat[ , .N , .(sampling_method_abbrev, sampling_method)]
    
 
#add nothing_caught
    mn_dat[is.na(species), nothing_caught := T , ] #The only species == NA are places where lake/date/gear/surveytype found no species
     
    
#species names are mixed codes and names
    #what cols are have species in name?
    mn_dat %>%
      select(contains("species"))
    #show problem
    mn_dat[ , .N , species ]
    
    # Review "codes" as defined by 3 letter species names
    print(mn_dat[nchar(species) == 3, .N , species ],
          nrow(mn_dat[nchar(species) == 3, .N , species ]))
    
    # choose only species.1 length of 3, assign these to a species_code
    mn_dat[nchar(species) == 3, `:=` (species_abbrev = species, species = NA) , ]
    
    a <- mn_dat[ ,.N , species_abbrev ]
    
    #generate a species code table (will also use later to fix target species cols)    
      species_codes_mn <-  mn_dat[!is.na(species_abbrev), .(species_abbrev = unique(species_abbrev)) ,  ] 
    
    #adds names to that table (uses key provided by Jon Hansen, MNDNR on 10 Oct 2023: https://docs.google.com/spreadsheets/d/1ZakE4S2zfSdYNEWHfikFqILNnD1gYpNf?rtpof=true&usp=drive_fs )    
        #This chunk cleans codes and names within the species column, references? for codes?
      species_codes_mn[ , species := 
                          case_when(# first fish codes are converted,
                            species_abbrev == "LMB" ~ "largemouth_bass",
                            species_abbrev == "SMB" ~ "smallmouth_bass",
                            species_abbrev == "WAE" ~ "walleye", 
                            species_abbrev == "WTS" ~ "white_sucker",
                            species_abbrev == "BLC" ~ "black_crappie",
                            species_abbrev == "YEP" ~ "yellow_perch",
                            species_abbrev == "BLG" ~ "bluegill",
                            species_abbrev == "NOP" ~ "northern_pike",
                            species_abbrev == "PMK" ~ "pumpkinseed", 
                            species_abbrev == "RKB" ~ "rock_bass",
                            species_abbrev == "BKF" ~ "banded_killifish",
                            species_abbrev == "IOD" ~ "iowa_darter",
                            species_abbrev == "BCS" ~ "blackchin_shiner",
                            species_abbrev == "BLB" ~ "black_bullhead",
                            species_abbrev == "JND" ~ "johnny_darter",
                            species_abbrev == "GOS" ~ "golden_shiner",
                            species_abbrev == "YEB" ~ "yellow_bullhead",
                            species_abbrev == "BRB" ~ "brown_bullhead",
                            species_abbrev == "CNM" ~ "central_mudminnow",
                            species_abbrev == "NRD" ~ "northern_redbelly_dace",
                            species_abbrev == "CSH" ~ "chinook_salmon",
                            species_abbrev == "MTS" ~ "mottled_sculpin",
                            species_abbrev == "TPM" ~ "tadpole_madtom",
                            species_abbrev == "SHR" ~ "shorthead_redhorse",
                            species_abbrev == "HSF" ~ "hybrid_sunfish",
                            species_abbrev == "BOF" ~ "bowfin",
                            species_abbrev == "WDS" ~ "weed_shiner",
                            species_abbrev == "BNM" ~ "bluntnose_minnow",
                            species_abbrev == "BND" ~ "blacknose_dace",
                            species_abbrev == "OTM" ~ "minnows",
                            species_abbrev == "FHM" ~ "fathead_minnow",
                            species_abbrev == "FND" ~ "finescale_dace",
                            species_abbrev == "SDM" ~ "slender_madtom",
                            species_abbrev == "SPO" ~ "spottail_shiner",
                            species_abbrev == "BNS" ~ "blacknose_shiner",
                            species_abbrev == "BKT" ~ "brook_trout",
                            species_abbrev == "SPT" ~ "splake",
                            species_abbrev == "RBT" ~ "rainbow_trout",
                            species_abbrev == "BNT" ~ "brown_trout",
                            species_abbrev == "BUB" ~ "burbot",
                            species_abbrev == "BIB" ~ "bigmouth_buffalo",
                            species_abbrev == "BLH" ~ "bullheads",
                            species_abbrev == "TLC" ~ "cisco",
                            species_abbrev == "CCF" ~ "channel_catfish",
                            species_abbrev == "BSD" ~ "blackside_darter",
                            species_abbrev == "BKS" ~ "brook_silverside",
                            species_abbrev == "LED" ~ "least_darter",
                            species_abbrev == "LGP" ~ "logperch",
                            species_abbrev == "SLR" ~ "silver_redhorse",
                            species_abbrev == "HHC" ~ "hornyhead_chub",
                            species_abbrev == "MUE" ~ "muskellunge",
                            species_abbrev == "TRP" ~ "trout_perch",
                            species_abbrev == "RHS" ~ "redhorse",
                            species_abbrev == "CAP" ~ "common_carp",
                            species_abbrev == "MMS" ~ "mimic_shiner",
                            species_abbrev == "EMS" ~ "emerald_shiner",
                            species_abbrev == "SUN" ~ "sunfish",
                            species_abbrev == "CRC" ~ "creek_chub",
                            species_abbrev == "GSF" ~ "green_sunfish",
                            species_abbrev == "SFS" ~ "spotfin_shiner",
                            species_abbrev == "SHI" ~ "shiners",
                            species_abbrev == "BST" ~ "brook_stickleback",
                            species_abbrev == "CSR" ~ "central_stoneroller",
                            species_abbrev == "SDS" ~ "sand_shiner",
                            species_abbrev == "GRR" ~ "greater_redhorse",
                            species_abbrev == "PGS" ~ "pugnose_shiner",
                            species_abbrev == "SMS" ~ "slimy_sculpin",
                            species_abbrev == "LND" ~ "longnose_dace",
                            species_abbrev == "FRD" ~ "freshwater_drum",
                            species_abbrev == "WHC" ~ "white_crappie",
                            species_abbrev == "DAR" ~ "darters",
                            species_abbrev == "SAR" ~ "sauger",
                            species_abbrev == "FCF" ~ "flathead_catfish",
                            species_abbrev == "BMS" ~ "bigmouth_shiner",
                            species_abbrev == "LKS" ~ "lake_sturgeon",
                            species_abbrev == "LKW" ~ "lake_whitefish",
                            species_abbrev == "BRM" ~ "brassy_minnow",
                            species_abbrev == "SLM" ~ "mississippi_silvery_minnow",
                            species_abbrev == "GLR" ~ "golden_redhorse",
                            species_abbrev == "OTS" ~ "suckers",
                            species_abbrev == "NST" ~ "ninespine_stickleback",
                            species_abbrev == "GOE" ~ "goldeye",
                            species_abbrev == "QBS" ~ "quillback",
                            species_abbrev == "RVS" ~ "river_shiner",
                            species_abbrev == "LES" ~ "longear_sunfish",
                            species_abbrev == "CHL" ~ "chestnut_lamprey",
                            species_abbrev == "RRH" ~ "river_redhorse",
                            species_abbrev == "WHB" ~ "white_bass",
                            species_abbrev == "OSS" ~ "orangespotted_sunfish",
                            species_abbrev == "STC" ~ "stonecat",
                            species_abbrev == "SNG" ~ "shortnose_gar",
                            species_abbrev == "RFS" ~ "rosyface_shiner",
                            species_abbrev == "CMS" ~ "carmine_shiner",
                            species_abbrev == "SAB" ~ "smallmouth_buffalo",
                            species_abbrev == "GIS" ~ "gizzard_shad",
                            species_abbrev == "LNG" ~ "longnose_gar",
                            species_abbrev == "RCS" ~ "river_carpsucker",
                            species_abbrev == "SLS" ~ "shovelnose_sturgeon",
                            species_abbrev == "LAT" ~ "lake_trout",
                            species_abbrev == "TME" ~ "tiger_muskellunge",
                            species_abbrev == "PRD" ~ "pearl_dace",
                            species_abbrev == "PLS" ~ "pallid_shiner",
                            species_abbrev == "RBS" ~ "rainbow_smelt",
                            species_abbrev == "RBD" ~ "rainbow_darter",
                            species_abbrev == "LNS" ~ "longnose_sucker",
                            species_abbrev == "SJC" ~ "shortjaw_cisco",
                            species_abbrev == "DWS" ~ "deepwater_sculpin",
                            species_abbrev == "SHS" ~ "spoonhead_sculpin",
                            species_abbrev == "BDD" ~ "banded_darter",
                            species_abbrev == "MOE" ~ "mooneye",
                            species_abbrev == "BLS" ~ "blue_sucker",
                            species_abbrev == "SLC" ~ "silver_chub",
                            species_abbrev == "HFS" ~ "highfin_carpsucker",
                            species_abbrev == "MCP" ~ "mirror_carp",
                            species_abbrev == "WAS" ~ "walleye/sauger",
                            species_abbrev == "AME" ~ "american_eel",
                            species_abbrev == "GOF" ~ "goldfish",
                            species_abbrev == "SPS" ~ "spotted_sucker",
                            species_abbrev == "PAH" ~ "paddlefish",
                            species_abbrev == "RVD" ~ "river_darter",
                            species_abbrev == "NHS" ~ "northern_hog_sucker",
                            species_abbrev == "BHM" ~ "bullhead_minnow",
                            species_abbrev == "WSD" ~ "western_sand_darter",
                            species_abbrev == "MDD" ~ "mud_darter",
                            species_abbrev == "SIL" ~ "silver_lamprey",
                            species_abbrev == "SKJ" ~ "skipjack_herring",
                            species_abbrev == "PGM" ~ "pugnose_minnow",
                            species_abbrev == "BKB" ~ "black_buffalo",
                            species_abbrev == "FTD" ~ "fantail_darter",
                            species_abbrev == "YLB" ~ "yellow_bass",
                            species_abbrev == "PRP" ~ "pirate_perch",
                            species_abbrev == "LKC" ~ "lake_chub",
                            species_abbrev == "RIR" ~ "river_ruffe",
                            species_abbrev == "WHP" ~ "white_perch",
                            species_abbrev == "ALW" ~ "alewife",
                            species_abbrev == "SEL" ~ "sea_lamprey",
                            species_abbrev == "BHC" ~ "bighead_carp",
                            species_abbrev == "GLD" ~ "gilt_darter",
                            species_abbrev == "CRD" ~ "crystal_darter",
                            species_abbrev == "SMT" ~ "speckled_madtom",
                            species_abbrev == "RDS" ~ "red_shiner",
                            species_abbrev == "WAM" ~ "warmouth",
                            species_abbrev == "CIS" ~ "cisco",
                            species_abbrev == "CPS" ~ "carpsucker",
                            species_abbrev == "CRP" ~ "crappie",
                            species_abbrev == "HCR" ~ "hybrid_crappie",
                            species_abbrev == "SCU" ~ "sculpin",
                            species_abbrev == "SHD" ~ "slenderhead_darter",
                            species_abbrev == "SIP" ~ "northern_pike_silver_phase",
                            species_abbrev == "UK1" ~ "unknown_fish",
                            species_abbrev == "UK2" ~ "unknown_fish",
                            # below unknown codes specified as such
                            species_abbrev == "LXB" ~ "see_ambiguous_species_code",
                            TRUE ~ species_abbrev
                          )
      ]
      
      #some dulicated common names (two codes for one thing)
      species_codes_mn[duplicated(species_codes_mn$species), , ]
      
      setnames(species_codes_mn, "species" , "common.name" )
      
        #fill species names
      mn_dat[species_codes_mn,
             on = .(species_abbrev=species_abbrev),
             species.fix := common.name 
      ]
        #fill into species.1 col, drop fix column
        mn_dat[is.na(species), species := species.fix , ]
        mn_dat[ , species.fix := NULL ,]
       
        
        #clean up common names
        mn_dat[ , species := str_replace_all( str_replace_all( str_replace_all(tolower(species),
                                                                                 " ", "_") ,
                                                                 "\\)" , "") ,
                                                "\\(" , "")]
        
        mn_dat[ , species := 
                  case_when(
                    species == "tullibee_cisco" ~ "cisco",
                    species == "trout-perch" ~ "trout_perch",
                    species == "bowfin dogfish" ~ "bowfin",
                    # unknown species
                    species == "no_particular_species" ~ "no_particular_species",
                    species == "unknown_fish_one" ~ "unknown_fish",
                    species == "walleye/sauger" ~ "walleye_x_sauger",
                    TRUE ~ species
                  )
        ]
        
        #check work
        print(mn_dat[ , .N , species][order(species)], nrow(mn_dat[ , .N , species])) 
        
        #still 1876 NAs here (nothing caught surveyXgears)
        mn_dat[is.na(species) , .N , .(nothing_caught)  ]
        
        
        #drop species codes
        species_codes_mn
        mn_dat[, .N , .(species_abbrev,species) ]
        mn_dat[ , `:=` (species_abbrev = NULL)]

        
# fix state
       #what cols are have species in name?
    mn_dat[ , .N ,  .(state, state_effort, state_indivfish) ]
    
    #collapse all
    mn_dat[ , `:=` (state = "Minnesota", state_effort = NULL, state_indivfish = NULL) , ]
    
#fix county
    mn_dat[ , .N ,  .(county_effort, county_indivfish) ]
    mn_dat[is.na(county_effort), county_effort := county_indivfish ]
              mn_dat[ , c("county_indivfish") := NULL]
    

# fix unit names for length and age
    #formats the length unit for clarity
    mn_dat[ , .N , length_unit ]
      mn_dat[ !is.na(length) ,.N  , length_unit ]
    
    mn_dat[ , .N , weight_unit]
      mn_dat[ !is.na(weight) ,.N  , weight_unit ]
    
    mn_dat[!is.na(length_unit), length_unit := "mm" ,]
    mn_dat[!is.na(weight_unit), weight_unit := "g" ,]

# target species
    #targeted species - needs cleaning, has species codes (clean up with the species code table generated above)
    mn_dat[ , .N , target_species ]
    mn_dat[ , c("T1","T2", "T3", "T4", "T5")  :=   tstrsplit(target_species, split = ", "), ]
    
    mn_dat[species_codes_mn, on = .(T1=species_abbrev) , target_species_1 := common.name ]
    mn_dat[species_codes_mn, on = .(T2=species_abbrev) , target_species_2 := common.name ]
    mn_dat[species_codes_mn, on = .(T3=species_abbrev) , target_species_3 := common.name ]
    mn_dat[species_codes_mn, on = .(T4=species_abbrev) , target_species_4 := common.name ]
    mn_dat[species_codes_mn, on = .(T5=species_abbrev) , target_species_5 := common.name ]
    
    mn_dat[ !is.na(target_species), unique(paste(target_species_1,target_species_2, target_species_3, target_species_4,target_species_5))]
    
    mn_dat[ , target_species := str_remove_all(paste(target_species_1,target_species_2, target_species_3, target_species_4,target_species_5, sep = ","),
                                               ",NA"),]
    
    mn_dat[ , c("T1","T2", "T3", "T4", "T5",
                "target_species_1","target_species_2", "target_species_3", "target_species_4", "target_species_5") := NULL]
    
#fix sexes
mn_dat[,.N, sex]

mn_dat[ sex == "U", sex := "Unknown"]
mn_dat[ sex == "M", sex := "Male"]
mn_dat[ sex == "F", sex := "Female"]
mn_dat[ sex == "N" | sex == "" | is.na(sex), sex := NA]
    
    
```

# Efforts
```{r}
# effort needs attention. Currently we have a negative effort and no units (nab code to add units from Holly based on gear type)
    #effort
    mn_dat %>% 
      select(contains("effort")) %>% 
      colnames()
    #total_effort_1.1
    
    gear_table_mn[order(sampling_method_abbrev)]
    
    #we got one total eff from the cpues, one from the fish/effort
    mn_dat[ ,.N  , .(is.na(total_effort_1), is.na( total_effort))]
    
    
    mn_dat[ !is.na(total_effort), .(mean = mean(total_effort), min =min(total_effort), max = max(total_effort)) , ]
    mn_dat[ !is.na(total_effort_1), .(mean = mean(total_effort_1), min =min(total_effort_1), max = max(total_effort_1)) , ]
    
    mn_dat[ !is.na(total_effort), .(mean = mean(total_effort), min =min(total_effort), max = max(total_effort)) , .(sampling_method_abbrev)]
    mn_dat[ !is.na(total_effort_1), .(mean = mean(total_effort_1), min =min(total_effort_1), max = max(total_effort_1)) , .(sampling_method_abbrev) ]
    
    #consolidate cols
    mn_dat[is.na(total_effort_1), total_effort_1 := total_effort]
    mn_dat[ ,total_effort := NULL]
    
    #no effort units, we have a negative effort? 
    
    # we can get units for these effort data from here (p45ish):
    # https://files.dnr.state.mn.us/publications/fisheries/special_reports/180.pdf
    #manually added the units for AOW, TMF, and GFM -- assuming same as similar gear categories
    # effort_table <- data.table( CODE = c(c("AIF", "TOU", "ATL", "STL", "EXT", "XXX", "MT", "SA", "ZO", "AOW"),#NONE
    #                              c("TR" , "STR", "SEF", "EFB", "EW", "EF"), #HOUR 
    #                              c("SSE", "SE", "S18", "S58"), #HAUL
    #                              c("GN", "GDE", "GSH", "GSM", "GST", "GSU", "GNP", "FCI", "TML", "SGN", "GNA", "VGN", "TN", "LTN", "DTN", "TQU", "T38", "THA", "T34", "T1", "STN", "GFM", "TMF")),#SET  
    #                     effort_unit = c(rep("no_unit", 10), rep("hours", 6), rep("haul", 4), rep("set", 23)))
    
    effort_table <- data.table(
  Code = c("GN", "GDE", "GSH", "GSM", "GST", "GSU", "GNP", "FCI", "TML", "SGN", "GNA", "VGN", "TN", "LTN", "DTN", "TQU", "T38", "THA", "T34", "T1", "STN", "EF", "EW", "EFB", "SEF", "S58", "S18", "SE", "SSE", "AIF", "AOW", "TOU", "ATL", "STL", "TR", "STR", "EXT", "XXX", "MT", "SA", "ZO"),
  Method = c("Gill net Standard graduated-mesh gill net", "Gill net Deep GN net, in waters colder than 12 o C", "Gill net Shallow GN net, in waters warmer than 12 o C", "Gill net Small-mesh (0.375 and 0.5-in-bar) net", "Gill net Short-term sets, monofilament net", "Gill net Suspended set, any gill net type", "Gill net Large-mesh (1.5, 2.0, 2.5-in bar) net", "Gill net Fish Community Index nets (Ontario)", "Gill net Trammel net", "Gill net Other (special) gill net", "Gill net Standard North American gill net", "Gill net Vertical gill net", "Trap net Standard 0.75-in-mesh double-frame net", "Trap net Large-frame (5x6 ft) net", "Trap net Pair of TN nets, set in double-pot configuration", "Trap net 0.25-in-mesh net", "Trap net 0.375-in-mesh net", "Trap net 0.5-in-mesh net", "Trap net 0.75-in-mesh single-frame net", "Trap net 1.0-in-mesh net", "Trap net Other (special) trap net", "Electrofishing Standard Boat electrofishing", "Electrofishing Fall electrofishing targeting Walleye", "Electrofishing Backpack electrofishing", "Electrofishing Other (special) electrofishing", "Seine 50-ft bag seine, 0.125-in mesh", "Seine 15-ft seine, 0.125-in mesh", "Seine 50-ft bag seine, 0.25-in mesh", "Seine Other (special) seining", "Angling Ice fishing", "Angling Open-water angling", "Angling Tournament angling", "Trot line Normal trot-line configuration", "Trot line Other (special) trot line", "Trawl Large lake trawl", "Trawl Other (special) trawl", "Miscellaneous Fish collected from or by an external source", "Miscellaneous Fish taken by miscellaneous means", "Miscellaneous 0.25-in-mesh minnow trap", "Miscellaneous Fish spawning trap", "Zooplankton Zooplankton sampling gear"),
  effort_unit = c("net_night", "net_night", "net_night", "net_night", "sets", "net_night", "net_night", "net_night", "net_night", "sets", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night","sets", "hours", "hours", "hours", "hours", "hauls", "hauls", "hauls", "hauls", "none", "none", "none", "none", "none", "hours", "hours", "none", "none", "none", "none", "none")
)
    
    
    #add effort units to gear tbale       
    gear_table_mn[effort_table, on = .(sampling_method_abbrev=Code), `:=` (total_effort_unit = effort_unit, sampling_method_desc = Method)]
    
    mn_dat[ , .N , .(is.na(sampling_method_abbrev), is.na(sampling_method)) ]
    
    #and push those over to the main df
    mn_dat[gear_table_mn, on = .(sampling_method_abbrev), `:=` (total_effort_unit = total_effort_unit, sampling_method_desc = sampling_method_desc)]
    
    #cleanup and checkwork
    rm(effort_table)
    mn_dat[ , .N , .(sampling_method,sampling_method_desc,  total_effort_unit)]
    
    mn_dat[ total_effort_1 <= 0 ][order(site_id)] #weird, but not diagnosable/fixable IMO.
    mn_dat[ total_effort_1 <= 0, flag := "effort value error" ]
    
    
#survey_id 
    mn_dat %>%
      select(contains("survey_")) %>% 
      colnames()
    
    mn_dat[ ,  .N, .(is.na(survey_id), is.na(survey_id_effort))]
    
    #something looks wonky in these survey Ids but the MNDNR says they're not useful anyways... see Minnnesota_README (https://docs.google.com/document/d/1sXag9RDvbKN0pUB6uRlADz36-kFNqRPwHVGlkJgyWTM/edit)
    
    mn_dat[ , `:=` (survey_id = NULL, survey_id_effort = NULL, survey_id_indivfish = NULL) , ]
    
    
#lake_name
              mn_dat[, .N, .(catchisNA = is.na(lake_name),
                                                                effortisNA = is.na(lake_name_effort),
                                                                fishisNA = is.na(lake_name_indivfish))
                 ]
            mn_dat %>% 
            mutate(same_survID_eff_fish = ifelse(lake_name_effort == lake_name_indivfish, "Same", "Different")) %>% 
            mutate(same_survID_eff_catch = ifelse(lake_name_effort == lake_name, "Same", "Different")) %>% 
            mutate(same_survID_catch_fish = ifelse(lake_name == lake_name_indivfish, "Same", "Different")) %>% 
            group_by(same_survID_eff_fish,same_survID_eff_catch,same_survID_catch_fish) %>% 
            count()
          
            mn_dat[is.na(lake_name), lake_name := ifelse(is.na(lake_name_effort), lake_name_indivfish, lake_name_effort)]
              mn_dat[ , c("lake_name_effort","lake_name_indivfish") := NULL]
              
#date
              sort(names(mn_dat))
              
                  mn_dat %>% 
      select(contains("date")) %>% 
      colnames()
              
              mn_dat[ , .N , is.na(date_survey)]
              
              mn_dat[ , length(unique(date.2_effort)) , .(lake_id,date_survey) ]
              
              mn_dat[ , .N , .(is.na(date_survey), is.na(date.1_indivfish))]
              
              #dump all extra dates:
              #"date.1_effort" and "date.2_indivfish" are equivalent to date_survey, date.2_effort (sample date) came with the fish as date_sample, previously (date.1_indivfish), , 
              
              mn_dat[ , c("date.1_effort", "date.2_indivfish", "date.2_effort", "date.1_indivfish", "date", "end_date") := NULL , ]
              
              
              
# add an effort_ident column "lake_id", "date_clean" , "sampling_method_abbrev", "survey_type.1"effort_ident := .GRP
    mn_dat[ , .N ,.(lake_id, date_survey, survey_type, sampling_method  #survey 
                                                        ) ]  
        #verify extra fields align with this level - they dont
        mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
                         total_effort_1, total_effort_unit, nothing_caught 
                                                        ) ] 
        
        #which don't?
                mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
                         total_effort_1, total_effort_unit, nothing_caught 
                                                        ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1]
        #Grab out full sets from summaries: this shows that the key issue here is that there are two total effort vals for some of these
                mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, and grab out the ones that our key fails on 
                  mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
                                   total_effort_1, total_effort_unit, nothing_caught #addt'l key metrics
                  ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1,
                                                                                      .(lake_id, date_survey, sampling_method)],
                  on = .(lake_id, date_survey, sampling_method), ]
                
        # snag a sample for review: this suggests that there's soem kind of mis-aggregation or miscalculation fo total efforts in the CPUE dataset that is effing these up. I'm going to flag them and press on        
          # a <-       mn_dat[mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, 
          #         mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
          #                          total_effort_1, total_effort_unit, nothing_caught 
          #         ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1, .(lake_id, date_survey, sampling_method)], on = .(lake_id, date_survey, sampling_method), ][1:2] ,
          #         on = .(lake_id, date_survey, sampling_method) , ]       
          #       
        
        #flag records where two total efforts are happenign in one survey for a gear: 20988 records 
                mn_dat[mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, 
                  mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method,
                                   total_effort_1, total_effort_unit, nothing_caught
                  ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1, .(lake_id, date_survey, sampling_method)], on = .(lake_id, date_survey, sampling_method), ] ,
                  on = .(lake_id, date_survey, sampling_method) , .N ,  ]
                
                mn_dat[mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, 
                  mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey
                                   total_effort_1, total_effort_unit, nothing_caught
                  ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1, .(lake_id, date_survey, sampling_method)], on = .(lake_id, date_survey, sampling_method), ] ,
                  on = .(lake_id, date_survey, sampling_method) , flag := ifelse(is.na(flag),
                                               "Questionable total effort, 2 values in one survey" ,
                                               paste(flag, "Questionable total effort, 2 values in one survey", sep = ";" ))  ]
                
                

        #adds in our effort_ident field
        mn_dat[ , total_effort_ident := .GRP ,.( lake_id, date_survey, survey_type, #survey 
                        sampling_method, total_effort_1, total_effort_unit, nothing_caught 
                                                        ) ] 
        
        #add a total effort nothing caught
mn_dat[ , all(nothing_caught)  , .(total_effort_ident)][ ,summary(V1)] #1876 seems reasonable, ~3% of TEIs found nothing.
mn_dat[ , all(is.na(species))  , .(total_effort_ident)][ ,summary(V1)]#check against another method of calculation

mn_dat[ , total_effort_nothing_caught :=  all(is.na(species))  , .(total_effort_ident)]

mn_dat[ , .N , total_effort_nothing_caught == nothing_caught]
mn_dat[ ,nothing_caught := NULL]

```


#notes cols tidying
```{r}
    

#CPUE Y/N column
  
  mn_dat[ , .N , notes.1 ]
  mn_dat[ notes.1 == "CPUE:N" , flag := ifelse(is.na(flag),
                                               "Do not use in CPUE calcs" ,
                                               paste(flag, "Do not use in CPUE calcs", sep = ";" ))]
  mn_dat[  , notes.1 := NULL , ]
  
#representative sampling column
  
  mn_dat[ , .N , notes ]
  mn_dat[ notes %in% c("REPRESENTATIVE_SAMPLING:N", "REPRESENTATIVE_SAMPLING:N, Y", "REPRESENTATIVE_SAMPLING:Y, N") , flag := ifelse(is.na(flag),
                                                                          "Not representative sampling" ,
                                                                          paste(flag, "Not representative sampling", sep = ";" ))]
  mn_dat[  , notes := NULL , ]  
  
#other notes  
  mn_dat[ , .N , notes.2] #send to workshop
  # workship results
  note_revisions <- fread("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_comments_notes_reviewed.csv")
  
  note_revisions[ , .N , flag]
  
  note_revisions <- note_revisions[flag != "", .(flag,notes.2) ]
  
  setnames(note_revisions, "flag", "note_flag")
  
  mn_dat[note_revisions, on = .(notes.2), note_flag := note_flag]
    mn_dat[  , .N , note_flag]
  
  mn_dat[ !is.na(note_flag) , flag := ifelse(is.na(flag),
                                                                          note_flag ,
                                                                          paste(flag, note_flag, sep = ";" ))]
  mn_dat[ , .(n = .N, n_surv = length(unique(total_effort_ident))) , flag]
  
  mn_dat[  , notes.2 := NULL , ]  
  
  mn_dat[ ,note_flag := NULL , ]
    
    
  
  mn_dat[ , .N , garbage_bin_notes.1_indivfish] #junk
  mn_dat[ , .N , garbage_bin_notes.1_effort]#junk
  mn_dat[ , .N , garbage_bin_notes.2_effort] #junk
  mn_dat[ , .N , garbage_bin_notes.2_indivfish] #junk
  mn_dat[ , .N , garbage_bin_notes.3] # junk

  mn_dat[ , `:=` (garbage_bin_notes.1_indivfish = NULL, garbage_bin_notes.1_effort = NULL, garbage_bin_notes.2_effort = NULL, garbage_bin_notes.3 = NULL)]

  
      
#gear notes  
  mn_dat[ , gear_data_notes := paste(gear_data_notes,
                                     gear_data_notes.1,
                                     gear_data_notes.2,
                                     gear_data_notes.3,
                                     gear_data_notes.4,
                                     gear_data_notes.5, sep = "_") ,]
  
  mn_dat[ , .N, gear_data_notes]
  mn_dat[ , .N, gear_data_notes.1]
  mn_dat[ , .N, gear_data_notes.2]
  mn_dat[ , .N, gear_data_notes.3]
  mn_dat[ , .N, gear_data_notes.4]
  mn_dat[ , .N, gear_data_notes.5] #junk
   # mn_dat[!is.na(gear_data_notes.5) , .N , .(gear_data_notes.5, total_effort, total_effort_unit) ]
  
    mn_dat[ , gear_data_notes.1 := NULL , ]  
    mn_dat[ , gear_data_notes.2 := NULL , ]
    mn_dat[ , gear_data_notes.3 := NULL , ]
    mn_dat[ , gear_data_notes.4 := NULL , ]
    mn_dat[ , gear_data_notes.5 := NULL , ]
  
#location notes  
  mn_dat[!(location_notes %in% c("LOC_DESC:","LOC_DESC:null")|is.na(location_notes)) , .N , location_notes ]  #keep
  
#lake type
  mn_dat[ , .N , lake_type ] #minimal info, but keep for consistency to other states
  
#aging str
  mn_dat[ , .N , .(aging_structure.1, aging_structure.2) ]
  mn_dat[ aging_structure.1 == "" , aging_structure.1 := NA , ]
  mn_dat[aging_structure.2 == "" , aging_structure.2 := NA , ]
  mn_dat[!is.na(aging_structure.2), aging_structure.1 := aging_structure.2]
  
  mn_dat[ , aging_structure.2 := NULL , ]

## Now grab exportable cols for the workshop
  names(mn_dat)
  
  
# #what cols go out to the team? (We'll carry these forward & round up at the end)
#   # gear stuff
#         dat <- mn_dat[ , .("records" = .N, "surveys" = length(unique(total_effort_ident))) , by= .(gear_data_notes, gear_data_notes.1, gear_data_notes.2, gear_data_notes.3, gear_data_notes.4,sampling_method, sampling_method_desc)]
#         
#         # Function to extract the first word from a string
#         extract_first_word <- function(x) {
#           non_na_values <- na.omit(x)
#           if (length(non_na_values) > 0) {
#             first_word <- strsplit(as.character(non_na_values[1]), " ")[[1]][1]
#             return(first_word)
#           }
#           return(NULL)
#         }
#         
#         # Loop through each column and update column name
#         for (col in names(dat)[1:3]) {
#           first_word <- extract_first_word(dat[[col]])
#           if (!is.null(first_word)) {
#             setnames(dat, col, first_word)
#           }
#         }
# 
#       # Function to strip the first word from a string
#       strip_first_word <- function(x) {
#         if (!is.na(x)) {
#           stripped_string <- sub("^[^:]+:", "", as.character(x))
#           return(stripped_string)
#         }
#         return(x)
#       }
#       
#       # Loop through each column and update values
#       for (col in names(dat)[1:3]) {
#         dat[, (col) := lapply(.SD[[col]], strip_first_word)]
#       }
#       
#       #fix column names
#       names(dat) <- word(names(dat), 1, sep = fixed(":"))
#       
#       # Print the updated data.table#write to file
#       
#       # fwrite(dat, file = "MN_gear_notes.csv")
#   
# # other generic notes
#         dat <- mn_dat[ , .("records" = .N, "surveys" = length(unique(total_effort_ident))) , by= .(garbage_bin_notes.2_indivfish, location_notes, notes, notes.2)]
#         
#         # Function to extract the first word from a string
#         extract_first_word <- function(x) {
#           non_na_values <- na.omit(x)
#           if (length(non_na_values) > 0) {
#             first_word <- strsplit(as.character(non_na_values[1]), " ")[[1]][1]
#             return(first_word)
#           }
#           return(NULL)
#         }
#         
#         # Loop through each column and update column name
#         for (col in names(dat)[1:3]) {
#           first_word <- extract_first_word(dat[[col]])
#           if (!is.null(first_word)) {
#             setnames(dat, col, first_word)
#           }
#         }
# 
#       # Function to strip the first word from a string
#       strip_first_word <- function(x) {
#         if (!is.na(x)) {
#           stripped_string <- sub("^[^:]+:", "", as.character(x))
#           return(stripped_string)
#         }
#         return(x)
#       }
#       
#       # Loop through each column and update values
#       for (col in names(dat)[1:3]) {
#         dat[, (col) := lapply(.SD[[col]], strip_first_word)]
#       }
#       
#       #fix column names
#       names(dat) <- word(names(dat), 1, sep = fixed(":"))
#       
#       # Print the updated data.table#write to file
#       
#       # fwrite(dat, file = "MN_comments_notes.csv")
#   
#   rm(col, first_word, dat)
#   
#   
#   
#   
  
# This section moved to the 
# # adds columns needed for state filtering
#   #adding lake area to the data
# area <- read_csv("Data_and_Scripts/Data/input/mn_lake_list.csv") %>%
#   select(DOW, LAKE_AREA_DOW_ACRES) %>%
#   mutate(DOW = as.integer(DOW))
# #joins fish data with area data
# mn_dat <- left_join(mn_dat, area, by = c("lake_id" = "DOW")) %>%
#   #hard codes lakes that do not have area in the area file (areas taken from lake finder)
#   mutate(LAKE_AREA_DOW_ACRES = as.numeric(case_when(
#         lake_id == "56038800" ~ "1288",
#         lake_id == "55006400" ~ "20",
#         lake_id == "76014601" ~ "467",
#         lake_id == "76014602" ~ "218",
#         TRUE ~ as.character(LAKE_AREA_DOW_ACRES)))) %>%
#   #creates binned area grouping used for minimum effort calculation
#   mutate(area.group = case_when(LAKE_AREA_DOW_ACRES < 100 ~ "<100",
#                                 LAKE_AREA_DOW_ACRES >= 100 & LAKE_AREA_DOW_ACRES <= 300 ~ "100-300",
#                                 LAKE_AREA_DOW_ACRES >= 301 & LAKE_AREA_DOW_ACRES <= 600 ~ "301-600",
#                                 LAKE_AREA_DOW_ACRES >= 601 & LAKE_AREA_DOW_ACRES <= 1500 ~ "601-1500",
#                                 LAKE_AREA_DOW_ACRES >= 1501 ~ "1501+",
#                                 TRUE ~ "NA")) %>%
#   #orders groupings so they plot well
#   mutate(area.group = factor(area.group, levels = c("<100", "100-300", "301-600", "601-1500", "1501+")))
# 
# 
#   setnames(mn_dat, c("LAKE_AREA_DOW_ACRES", "area.group"), c("lake_area_dow_acres", "area_group"))
  

  
#consolidate original filenames
  
  mn_dat[ , original_file_names := str_remove_all(paste(original_file_name_effort,original_file_name_indivfish,original_file_name, sep = ", "),
                                               ", NA"),]
  mn_dat[ , `:=` ("original_file_name_effort" = NULL ,"original_file_name_indivfish" = NULL , "original_file_name" = NULL) ,]
  
  mn_dat[ , .N ,  original_file_names]
  
  mn_dat[ , original_file_names := str_remove_all(original_file_names, "NA, ") , ]
  

#yoy as age class
  mn_dat[ , .N  ,  young_of_year]
  mn_dat[ young_of_year == "Y", age_class := "young_of_year"]
  mn_dat[ young_of_year == "N", age_class := "not_young_of_year"]
  mn_dat[ ,young_of_year := NULL]
    
  
  
```


#sampling_methods
```{r}
mn_dat <- clean_names(mn_dat) #clean up names a touch

# here's what were after: 
#sampling_method_simple, sampling_method, sampling_method_2, gear_data_notes
mn_dat[ , .N , .(sampling_method_abbrev, sampling_method, sampling_method_desc)]

#and we need to add some data that we'll peel from here:

mn_dat[ , .N , gear_data_notes ]
  mn_dat[ , mesh := str_extract(gear_data_notes, "(?<=MESH:)[^_]+")]
 # the biggest mesh GN I could find was 8" at Duluth Nets, I think it's too ambiguous for me to interpret literally as a numeric inches. 
    mn_dat[ , .N , .(mesh) ][order(mesh)]
    mn_dat[mesh == "NA", mesh := NA]
  mn_dat[ , n_netters := as.numeric(str_extract(gear_data_notes, "(?<=NUMBER_OF_NETTERS:)[^_]+"))]
    mn_dat[ ,summary(n_netters) ]
  mn_dat[ , daylight_sampling := str_extract(gear_data_notes, "(?<=DAYLIGHT_SAMPLING:)[^_]+")]
    mn_dat[ , .N, .(daylight_sampling) ]
    mn_dat[ daylight_sampling %in% c("N, Y", "Y, N"), daylight_sampling := NA , ]
  
    
    
    
#gear xwalk
gear_xwalk <- fread("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/crosswalks/gears_by_state.csv", na.strings = "")%>% 
  filter(state == "Minnesota")

setnames(mn_dat, old = c("sampling_method", "sampling_method_desc"), new = c("gear_1", "gear_2"))

mn_dat[gear_xwalk, on = .(gear_1, gear_2), `:=` (sampling_method_simple = i.sampling_method_simple,
                                                 sampling_method_1 = i.sampling_method_1,
                                                 sampling_method_2 = i.sampling_method_2
                                                 )
       ]
mn_dat[ , .N , sampling_method_simple ]
mn_dat[ , .N , sampling_method_1 ]



#now bring over the gear notes
mn_dat[daylight_sampling == "Y", ':=' (sampling_method_1 = paste(sampling_method_1,"day",sep = "_"),
                                       sampling_method_2 = paste(sampling_method_2,"day",sep = "_")), ]

mn_dat[daylight_sampling == "N", ':=' (sampling_method_1 = paste(sampling_method_1,"night",sep = "_"),
                                       sampling_method_2 = paste(sampling_method_2,"night",sep = "_")) , ]
mn_dat[!is.na(n_netters), .N , sampling_method_simple ]
mn_dat[!is.na(n_netters),  ':=' (sampling_method_2 = paste(sampling_method_2, paste(n_netters, "netters",sep = "_"), sep = "; ")), ]

mn_dat[!is.na(mesh), .N , sampling_method_simple ]
mn_dat[!is.na(mesh) ,mesh := paste( "mesh", mesh, sep = ":" ) , ]
#moved to notes_1 iaw strategy from Large Lakes datasets
setnames( mn_dat ,  old = "mesh" , new = "notes_1" )
mn_dat[ , .N , .(notes_1 , sampling_method_simple) ]

#remove old cols and clean up workspace

mn_dat[ , ':='(gear_1 = NULL, gear_2 = NULL, n_netters = NULL, daylight_sampling = NULL)  ,]

rm(gear_xwalk, gear_table_mn, species_codes_mn, note_revisions, a, cde)

```


#NHDids
```{r}

  #use mwlaxeref to fetch NHDIDs
  # mwlaxeref::mn_to_nhdhr(mn_dat, from_colname = "lake_id")[,nhdhr.id]
  
  mn_dat[ , nhdhr_id := mwlaxeref::mn_to_nhdhr(mn_dat, from_colname = "lake_id")[,nhdhr.id] , ]
    

  #remaining unmatched lakes:
  mn_dat[is.na(nhdhr_id), .N , lake_id ]
  


msu_xwalk <- fread("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/crosswalks/MGLP_FISH_LAKES_12Aug24.csv") %>% 
  filter(STATE == "MN") 
setDT(msu_xwalk)

msu_xwalk[, lake_id := as.integer(STATE_ID) ,]

#add locs and NHDs from MSU
mn_dat[msu_xwalk, on = .(lake_id = lake_id) , ':=' (nhdhr_id_MSU = gsub("nhdhr_", "", NHDHR_ID), latitude_lake_centroid = LAT, longitude_lake_centroid = LON) ]

  #remaining unmatched lakes:
  mn_dat[, .N , .(lake_id, nhdhr_id, nhdhr_id_MSU) ][ , .N , .(is.na(nhdhr_id), is.na(nhdhr_id_MSU)) ]
  #agreement? 71 not aligned...
  mn_dat[, .N , .(lake_id, nhdhr_id, nhdhr_id_MSU) ][!is.na(nhdhr_id) & !is.na(nhdhr_id_MSU), .N , .(nhdhr_id == nhdhr_id_MSU) ]

#great. still missing ~134 lakes
  mn_dat[is.na(nhdhr_id), nhdhr_id := nhdhr_id_MSU] #fill where missing
  mn_dat[ !is.na(nhdhr_id) & !is.na(nhdhr_id_MSU) & nhdhr_id_MSU != nhdhr_id, .N , .(lake_id, nhdhr_id, nhdhr_id_MSU) ]  #some lakes have mismatches:
  #for these we oughtta saver both?
  mn_dat[!is.na(nhdhr_id) & !is.na(nhdhr_id_MSU) & nhdhr_id_MSU != nhdhr_id , notes_1 :=  ifelse(is.na(notes_1),
                                               paste("Secondary_nhdhrid", nhdhr_id_MSU , sep = ":" ) ,
                                               paste(notes_1, paste("Secondary_nhdhrid", nhdhr_id_MSU , sep = ":" ), sep = "; " )), ]
  
  mn_dat[ , nhdhr_id_MSU:= NULL , ]

# more lat/lons?
usgs <- fread("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/crosswalks/usgs_lake_metadata.csv") %>% 
  filter(state == "MN")

usgs[ , nhdhr_id := gsub("nhdhr_", "", site_id ) , ]  

 #add locs and NHDs from MSU
mn_dat[usgs, on = .(nhdhr_id = nhdhr_id) ,  ':=' (latitude_lake_centroid_USGS = centroid_lat, longitude_lake_centroid_USGS = centroid_lon) ]
 
#remaining unknown location lakes: 126.
  mn_dat[, .N , .(lake_id, nhdhr_id, latitude_lake_centroid, longitude_lake_centroid, latitude_lake_centroid_USGS, longitude_lake_centroid_USGS) ][ , .N , .(is.na(latitude_lake_centroid), is.na(latitude_lake_centroid_USGS)) ]
  
mn_dat[ is.na(latitude_lake_centroid), ':=' (latitude_lake_centroid = latitude_lake_centroid_USGS, longitude_lake_centroid = longitude_lake_centroid_USGS)]
  
mn_dat[ , ':=' (latitude_lake_centroid_USGS = NULL, longitude_lake_centroid_USGS = NULL )  ,  ]  
  
  
  #remaining unmatched lakes:
  mn_dat[is.na(nhdhr_id) | is.na(latitude_lake_centroid), .N , .(lake_id, lake_name, county_effort) ]


```

#dates
```{r}

mn_dat[ , .SD , .SDcols = colnames(mn_dat)[str_detect(colnames(mn_dat), "date")]]
#how many survey dates in a TEI? ONE ONLY ONE! HOORAY
mn_dat[ , length(unique(date_survey)) , total_effort_ident ][order(-V1)]
mn_dat[ , length(unique(date_sample)) , total_effort_ident ][order(-V1)]
mn_dat[ , date_total_effort_ident := date_survey ]







```

#Column Names
```{r}
rm(usgs, msu_xwalk)

#Denver's clean up of columns and names
mn_dat <- mn_dat %>% 
  mutate(
    state = "Minnesota",
    county = county_effort,
    lake_name = lake_name,
    lake_id = as.character(lake_id),
    nhdhr_id = nhdhr_id,
    
    latitude_lake_centroid = latitude_lake_centroid,
    longitude_lake_centroid = longitude_lake_centroid,
    
    date_survey = date_survey,
    date_total_effort_ident = date_total_effort_ident,
    date_sub_effort_ident = as.IDate(NA),
    date_sample = date_sample,
    year = year(date_survey),
    month = month(date_survey),
    
    survey_id = as.character(NA),
    survey_type = survey_type,
    survey_type_2 = as.character(NA),
    survey_type_3 = as.character(NA),
    survey_type_4 = as.character(NA),
    
    sampling_method_abbrev = NULL,
    sampling_method_simple = sampling_method_simple,
    sampling_method = sampling_method_1,
    sampling_method_2 = sampling_method_2,
    gear_data_notes = gear_data_notes,
    target_species = target_species,
    
    total_effort_ident = total_effort_ident,
    total_effort_1 = total_effort_1,
    total_effort_2 = as.numeric(NA),
    total_effort_1_units = total_effort_unit,
    total_effort_2_units = as.character(NA),
    total_effort_nothing_caught = total_effort_nothing_caught,
    
    water_temp = water_temp,
    water_temp_units = case_when(!is.na(water_temp) ~ "fahrenheit", TRUE ~ NA),
    water_clarity = as.numeric(NA),
    water_clarity_units = as.character(NA),
    
    lat_start = as.numeric(NA),
    lon_start = as.numeric(NA),
    lat_end = as.numeric(NA),
    lon_end = as.numeric(NA),
    site_id = site_id,
    
    sub_effort_ident = as.character(NA),
    sub_effort_1 = as.numeric(NA),
    sub_effort_1_units = as.character(NA),
    sub_effort_2 = as.numeric(NA),
    sub_effort_2_units = as.character(NA),
    sub_effort_nothing_caught = NA,
    
    species_1 = species,
    length_1 = as.numeric(length),
    length_unit_1 = length_unit, 
    length_bin = as.character(NA),
    length_bin_unit = as.character(NA),
    
    age = age,
    aging_structure_1 = aging_structure_1,
    
    weight_1 = weight,
    weight_unit_1 = weight_unit,
    batch_weight = as.character(NA),
    batch_weight_unit = as.character(NA),
    
    sex = sex,
    age_class = age_class,
    
    flag = flag,
    original_file_names = original_file_names,
    ind_fish_ident = as.character(sample_id),
    # lakesize = as.numeric(NA),
    # lakesize_units = as.numeric(NA),
    lat_unspec = as.numeric(NA),
    lon_unspec = as.numeric(NA),
    waterbody_type = lake_type,
    location_notes_1 = location_notes,
    obs_id = as.character(row_number())
  ) %>% 
  unite("notes_1", c(notes_1,
                     garbage_bin_notes_2_indivfish,
                     reproductive_condition_notes),
                     remove = T,
                     na.rm = T) %>% 
  setDT()

# intersect(names(mw_data), names(mn_dat))
# 
# setdiff(names(mw_data), names(mn_dat)
#            )

schema_names <- c(
  "state", "county", "lake_name", "lake_id", "nhdhr_id",

  "latitude_lake_centroid", "longitude_lake_centroid",

  "date_survey", "date_total_effort_ident", "date_sub_effort_ident", "date_sample", "year", "month",

  "survey_id", "survey_type", "survey_type_2", "survey_type_3", "survey_type_4",

  "sampling_method_simple", "sampling_method", "sampling_method_2", "gear_data_notes","target_species",

  "total_effort_ident", "total_effort_1", "total_effort_2", "total_effort_1_units", "total_effort_2_units", "total_effort_nothing_caught",

  "water_temp", "water_temp_units", "water_clarity", "water_clarity_units",

  "lat_start", "lon_start", "lat_end", "lon_end",

  "site_id", "sub_effort_ident", "sub_effort_1", "sub_effort_1_units", "sub_effort_2","sub_effort_2_units", "sub_effort_nothing_caught",

  "species_1", "length_1", "length_unit_1", "length_bin", "length_bin_unit",

  "age", "aging_structure_1",

  "weight_1", "weight_unit_1", "batch_weight", "batch_weight_unit",

  "sex", "age_class",

  "flag", "original_file_names",

  "ind_fish_ident",
  "lat_unspec", "lon_unspec",
  "waterbody_type",
  "location_notes_1", "notes_1", "obs_id"
)

setcolorder(mn_dat, neworder = schema_names)


# match column classes
glimpse(mn_dat)

dropcols <- colnames(mn_dat)[!colnames(mn_dat)%in%schema_names]

mn_dat[ , (dropcols):= NULL , ]


# column classes

chr_cols <- c("state", "county", "lake_name", "lake_id", "nhdhr_id", "survey_id", "survey_type",
  "survey_type_2", "survey_type_3", "survey_type_4", "sampling_method_simple",
  "sampling_method", "sampling_method_2", "gear_data_notes", "target_species",
  "total_effort_ident","total_effort_1_units",
  "total_effort_2_units", "water_temp_units", "water_clarity_units","site_id", "sub_effort_ident",
  "sub_effort_1_units","sub_effort_2_units","species_1","length_unit_1", "length_bin", "length_bin_unit","aging_structure_1",
   "weight_unit_1", "batch_weight", "batch_weight_unit", "sex", "age_class", "flag",
  "original_file_names", "ind_fish_ident", "waterbody_type", "location_notes_1", "notes_1", "obs_id")

mn_dat[, (chr_cols) := lapply(.SD, as.character), .SDcols = chr_cols]

num_cols <- c("latitude_lake_centroid", "longitude_lake_centroid" , "year", "month", "total_effort_1", "total_effort_2",
              "water_temp", "water_clarity", "lat_start", "lon_start", "lat_end", "lon_end", "sub_effort_1", "sub_effort_2",
              "length_1", "age", "weight_1",  "lat_unspec", "lon_unspec")

mn_dat[, (num_cols) := lapply(.SD, as.numeric), .SDcols = num_cols]

mn_dat[ , .N , .(sampling_method_simple, sampling_method_2, flag) ]



```

#checks, flags, schema match
```{r}
#effort
mn_dat[ , .N , .(total_effort_1_units)]
mn_dat[is.na(total_effort_1) , .N , total_effort_1_units  ]
mn_dat[is.na(total_effort_1) , total_effort_1_units := NA  ]



#creating an auto flagging system
flag_xwalk <- mn_dat %>% 
  distinct(total_effort_ident, .keep_all = TRUE) %>% 
  group_by(sampling_method_simple) %>% 
  mutate(mean_effort = mean(total_effort_1, na.rm = TRUE),
         sd_effort = sd(total_effort_1, na.rm = TRUE),
         z_score = (total_effort_1 - mean_effort) / sd_effort,
         flag = NA, #creates a flag column so I can case_when the flag generation 
         flag = case_when(total_effort_1 < 0 ~ paste("invalid effort (negative)", coalesce(flag, "")),
                          total_effort_1 == 0 ~ paste("effort reported as 0", ";", coalesce(flag, "")),
                          z_score > 10 ~ paste("high effort", ";", coalesce(flag, "")),
                          is.na(z_score) ~ coalesce(flag, ""),
                          TRUE ~ coalesce(flag, "")
         ),
         flag = na_if(flag, "")) %>% 
  select(total_effort_ident,
         sampling_method_simple,
         flag,
         total_effort_1,
         mean_effort,
         sd_effort,
         z_score) %>% 
  filter(!is.na(flag))

flag_xwalk_2 <- mn_dat %>% 
  distinct(total_effort_ident, .keep_all = TRUE) %>%
  filter(!(total_effort_ident %in% flag_xwalk$total_effort_ident)) %>% 
  group_by(sampling_method_simple) %>% 
  mutate(mean_effort = mean(total_effort_1, na.rm = TRUE),
         sd_effort = sd(total_effort_1, na.rm = TRUE),
         z_score = (total_effort_1 - mean_effort) / sd_effort,
         flag = NA,
         flag = case_when(total_effort_1 < 0 ~ paste("invalid effort (negative)", coalesce(flag, "")),
                          total_effort_1 == 0 ~ paste("effort reported as 0", ";", coalesce(flag, "")),
                          z_score > 10 ~ paste("high effort", ";", coalesce(flag, "")),
                          is.na(z_score) ~ coalesce(flag, ""),
                          TRUE ~ coalesce(flag, "")
         ),
         flag = na_if(flag, "")) %>% 
  select(total_effort_ident,
         sampling_method_simple,
         flag,
         total_effort_1,
         mean_effort,
         sd_effort,
         z_score) %>% 
  filter(!is.na(flag))

flag_xwalk_effort <- rbind(flag_xwalk, flag_xwalk_2) %>% 
  ungroup() %>% 
  select(total_effort_ident, 
         flag)
rm(flag_xwalk, flag_xwalk_2)

setnames(flag_xwalk_effort,old="flag" ,new="tei_flag" )

mn_dat[flag_xwalk_effort , on = .(total_effort_ident = total_effort_ident), tei_flag:= tei_flag ]
mn_dat[ , .N , .(tei_flag) ]
mn_dat[str_detect(tei_flag, " ; ") , .N ,tei_flag]
mn_dat[ , tei_flag := gsub(" ; ", "", tei_flag) , ]

mn_dat[, flag := do.call(paste, c(replace(.SD, is.na(.SD), ""), sep = "; ")), .SDcols= c("flag", "tei_flag")]

# Remove "; " from the beginning or end
mn_dat[!is.na(flag), flag := gsub("^\\s*;+|;+\\s*$", "", gsub(";\\s*;+", "; ", flag))]

mn_dat[ , .N , flag]

mn_dat %>% 
  filter(total_effort_ident %in% flag_xwalk_effort$total_effort_ident)
#flag gets applied to 869077 fish records
rm(flag_xwalk_effort)

#flag tables for fish
#summarize length based stats (and trim them to only species where we can do this work)
length_stats <- mn_dat[, .(stat = c(names(summary(length_1)), "sd"),
            value = c(summary(length_1), sd(length_1, na.rm = T)),
            units = "length_inches"), .(species_1) ]
length_stats[is.na(value) , .N , species_1 ]
length_stats <- length_stats[!species_1 %in% length_stats[is.na(value) , unique(species_1) ]]

#zero lengths
mn_dat[length_1 == 0 , len_flag := "length reported as zero"]

#when your length is >10sd from mean for species, get flagged as unlikely length
length_stats <- dcast(length_stats, formula = species_1 ~ ...)
length_stats[ , high_lim := Mean_length_inches + 10*sd_length_inches ,]
setcolorder(length_stats, neworder= c("species_1", "Mean_length_inches", "high_lim"))


mn_dat[length_stats , on = .(species_1, length_1 > high_lim) , len_flag := "length > 10sd from species mean in state" ]
mn_dat[ , .N , len_flag]

rm(length_stats)

#summarize weight based stats
weight_stats <- mn_dat[, .(stat = c(names(summary(weight_1)), "sd"),
            value = c(summary(weight_1), sd(weight_1, na.rm = T)),
            units = "weight_pounds"), .(species_1) ]
weight_stats[is.na(value) , .N , species_1 ]
weight_stats <- weight_stats[!species_1 %in% weight_stats[is.na(value) , unique(species_1) ]]

#zero lengths
mn_dat[weight_1 <= 0, .N]
mn_dat[weight_1 == 0 , weight_flag := "weight reported as zero"]

#when your wt is >10sd from mean for species, get flagged as unlikely weight
weight_stats <- dcast(weight_stats, formula = species_1 ~ ...)
weight_stats[ , high_lim := Mean_weight_pounds + 10*sd_weight_pounds ,]
setcolorder(weight_stats, neworder= c("species_1", "Mean_weight_pounds", "high_lim"))


mn_dat[weight_stats , on = .(species_1, weight_1 > high_lim) , weight_flag := "length > 10sd from species mean in state" ]
mn_dat[ , .N , weight_flag]

rm(weight_stats)

#summarize age based stats
age_stats <- mn_dat[, .(stat = c(names(summary(age)), "sd"),
            value = c(summary(age), sd(age, na.rm = T)),
            units = "age_years"), .(species_1) ]
age_stats[is.na(value) , .N , species_1 ]
age_stats <- age_stats[!species_1 %in% age_stats[is.na(value) , unique(species_1) ]]

#when your age is >10sd from mean for species, get flagged as unlikely age
age_stats <- dcast(age_stats, formula = species_1 ~ ...)
age_stats[ , high_lim := Mean_age_years + 10*sd_age_years ,]
setcolorder(age_stats, neworder= c("species_1", "Mean_age_years", "high_lim"))


mn_dat[age_stats , on = .(species_1, age > high_lim) , age_flag := "age > 10sd from species mean in state" ]
mn_dat[ , .N , age_flag]


mn_dat[, flag := do.call(paste, c(replace(.SD, is.na(.SD), ""), sep = "; ")), .SDcols= c("flag", "len_flag", "weight_flag", "age_flag")]
mn_dat[, .N , flag ]

# Remove "; " from the beginning or end & compress multiples
mn_dat[!is.na(flag), flag := gsub("^\\s*;+|;+\\s*$", "", gsub(";\\s*;+", "; ", flag))]


mn_dat[ , c("len_flag", "weight_flag", "age_flag") := NULL , ]



#check speceis richness seems reasonable
mn_dat[!is.na(total_effort_ident) , length(unique(species_1)) , total_effort_ident ][, hist(V1, main = "species richness per survey") , ]

#check fish counts:
mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , total_effort_ident ][, hist(V1, main = "N indiv fish sampled in survey") ,]

mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>100000 , ,]



# Remove "; " from the beginning or end and compress multiples
mn_dat[!is.na(flag), flag := trimws(gsub("^\\s*;+|;+\\s*$", "", gsub(";\\s*;+", "; ", flag)))]
mn_dat[flag == "", flag := NA]

mn_dat[!is.na(species_1) & !is.na(total_effort_ident) &is.na(flag) , length(unique(obs_id)) , total_effort_ident ][, hist(V1, main = "N indiv fish sampled in survey") ,]

mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>20000 , ,]

#review other high catch records
mn_dat[ total_effort_ident %in% 
                                        mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>20000 , total_effort_ident, ],
.N, .(total_effort_ident, lake_name, species_1, length_1)]

#flag them
mn_dat[ total_effort_ident %in% 
                                        mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>20000 , total_effort_ident, ],
                                      flag := ifelse(is.na(flag), "very high catch is suspected error" , paste(flag, "very high catch is suspected error", sep = ";" ))]




rm(age_stats, a, mw_data, chr_cols, col_name, del_col, i, new, new_colnames, num_cols, old, schema_names)


#units matched to hive and NAs where vcalues missing?
mn_dat[ , .N , .(length_unit_1, weight_unit_1) ]
#length
mn_dat[is.na(length_1) , .N , length_unit_1]
  mn_dat[is.na(length_1) ,length_unit_1 := NA]

mn_dat[ length_unit_1 == "mm" , length_unit_1 := "millimeters" ,]
#weight
mn_dat[is.na(weight_1) , .N , weight_unit_1]
  mn_dat[is.na(weight_1) ,weight_unit_1 := NA]

mn_dat[ weight_unit_1 == "g" , weight_unit_1 := "grams" ,]

#effort
mn_dat[is.na(total_effort_1) , .N , .(total_effort_1_units, total_effort_2_units) ]


#flags

mn_dat[ , .N, flag ]



```




# Save Parquet
```{r}

fwrite(mn_dat, "Data_and_Scripts/Data/output/mn_data.csv" , na="NA" )
rm(mn_dat)
mn_dat <- fread("Data_and_Scripts/Data/output/mn_data.csv", na.strings = "NA")

mn_dat[ , date_sub_effort_ident := as.IDate(NA) , ]
mn_dat %>% 
  mutate(date_sub_effort_ident = as.IDate(NA)) %>% 
  {mn_dat <<- .}

#needs date_sub_effort_ident converted back to date


# write_dataset(dataset = mn_dat, path = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "hive_update", "state=Minnesota"), format = "parquet")
# 
# mn_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "hive_update", "state=Minnesota"))
# 
# glimpse(mn_data)

```

#sleuth out the MN duplicated/missing erroneously data issues:
```{r}

#we have a number of cases where we ended up with >1 record for a single survey. Some are clearly effed up. Historical surveys that sometimes had the data sometimes not, and match to a "population assessment" that sometimes has the data, sometimes not. The following code works tyowards a flagging system, but does not accomplish what we want. 


# mn_dat %>% 
#   group_by(lake_name, lake_id, date_total_effort_ident, sampling_method, sampling_method_2) %>% 
#   collect() %>% 
#   summarise(n = n_distinct(survey_type)) %>%
#   filter(n>1) %>% 
#   collect() %>% {. ->> dups}
# 
# mn_dat %>% 
#   # mutate(year = year(date)) %>%
#   inner_join(dups) %>%
#   group_by(lake_name, lake_id, date_total_effort_ident, sampling_method,sampling_method_2, survey_type, total_effort_ident) %>% 
#   summarise(n_catch_all_spp = sum(!is.na(species_1))) %>%
#   collect() %>%
#   filter(!lake_name == "Winnibigoshish") %>% 
#   {. ->> dups_with_surveytype_and_catch}
# 
# setorder(dups_with_surveytype_and_catch, lake_id, date_total_effort_ident ,-n_catch_all_spp)
# 
# #After excluding the indfish/effort mismatch
# dups_with_surveytype_and_catch <- dups_with_surveytype_and_catch[!(survey_type %in% c("External Management Survey", "External Mgmt Survey"))]
# 
# 
# #we've got some bad data coming in from DNR, wherever we see a lake - year -gear
# mn_dat[lake_name == "Lake of the Woods", .N , .(lake_name, lake_id, date_total_effort_ident, sampling_method, survey_type) ][order(date_total_effort_ident, sampling_method )]
# 
# #within a year - lake - gear, if a catch value is 1/10th of other TEIs, flag that TEI
# dup_gears <- mn_dat[ , .("n_TEIs" = length(unique(total_effort_ident)), "nfish" = sum(!is.na(species_1))) , .(lake_id,lake_name, year, sampling_method, sampling_method_2) ][n_TEIs>1]
# 
# two_surveys <- mn_dat[dup_gears, on = .(lake_id, year, sampling_method)][ , .("nfish" = sum(!is.na(species_1))) , .(lake_id,lake_name, year, date_total_effort_ident,  sampling_method, sampling_method_2, survey_type , total_effort_ident, total_effort_1, total_effort_1_units) ]
# 
# two_surveys[ , max_catch_inyear := max(nfish) , .(lake_id, year, sampling_method_2) ]
# 
# two_surveys[ , flag_test := nfish/max_catch_inyear < 1/100 , ]
# 
# 
# ##### another way to find them easily?
# 
# nocatch <- mn_dat[total_effort_nothing_caught == T , .N   ,.(lake_id,lake_name, year, date_total_effort_ident,  sampling_method, sampling_method_2, survey_type , total_effort_ident, total_effort_1, total_effort_1_units) ]
# 
# 
# nocatch_lakes <- mn_dat[nocatch, on = .(lake_id,lake_name,sampling_method),
#        ][  ,.N   ,.(lake_id,lake_name, year, date_total_effort_ident,  sampling_method, sampling_method_2, survey_type , total_effort_ident, total_effort_1, total_effort_1_units, total_effort_nothing_caught, original_file_names)][order(lake_id,  sampling_method_2,date_total_effort_ident, total_effort_nothing_caught)]
# 
















```

