---
title: "MN_Flat_File_Aggregation"
author: "Holly Kundel & Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Preamble
updated 11/29
To Do:

DONE 4. Line 229 might offer a key to the large lakes missing data issue. Had a "total.count" val for those-- doubtful, but recheck this. 

DONE 1. where gear notes marked with dippers, bring back dipper # into end of gear desc.
DONE 2. Consider grabbing the mesh var?
DONE 3. Daylight EFing (gear data notes 2 - keep it around for filtering)
DONE 9. Move DL's hiving code up into the actual data agg scripts

DONE 11.We have numerous TEIs with multiple total efforts. That's not how total effort works. Fix.
DONE 10.Chris Rounds is finding Fall EF dates that are actually reporting the lakefinder date for this survey rather than the actual observation date.
        How are we currently handling date, are we using the “survey_date” and dumping the “observation_date” and losing info?
        See hive date setups
DONE 13. MOve lake area data over to the agg script (trims lots of NAs in other states)
DONE 15. Clean up sex column Male, Female, Unknown, NA (not reported, not sexed, data lost, all those things)
DONE 16. QC on age column
DONE 17. negative vals on sample ids (whats that mean and should they be removed?)
DONE 18. Add sampling method simple column to hive
DONE 24. Add flags for 0 efforts and high efforts
DONE 19. Fix units to match hive structure, make sure anything with na values are also na for the unit (e.g., units like net-nights -> net_nights)
DONE 25. Auto Flagging: (build a flagging template from the MW Hive then apply it state-by-state)
          0 and negative efforts
          10SDs above mean effort? (by gear/state?)
          Lengths (prob group by age/survey/something) - big and small
          Weights (prob group by age/survey/something)
          Counts (by TEI)
DONE 20. Where values are NA, units should also be NA (efforts, lengths, weights, others?)
DONE 22. Check N species in a survey (is 44 spp a reasonable finding?)
DONE 23. Check that our flagging is spelling out implications
DONE 7. Lake 28003200|1996|Trap has 2 idents one TEInothing caught=T and TEInothingcaught=F
        
        
NOT COMPLETELY RESOLVED:        
12.We have many TEIs tagged with nothing caught)
          Large Lakes data R/R with Kristen [somethings]'s data that she agg'd 
          Built a script in the project that lays out a few other cases of weirdness (white bear lake 2022 trap nets, Willard Lake, Cougar Lake) 
          Corey sees the same un-entered data we do in their data entry dashboard
ACTION    Send all Nothing caught = True total efforts off to Jon Hansen
CHECK:    Historical WAE survey duplication
          - check on outcome of fixes (section name: sleuth out the MN duplicated data issues)

5. In 13 Aug cpue export, State ==NA, total effort ident = NA, Note that in the filtered cpe data file these records are NA-laden
    Pulled example: feather lake, 06/10/1992, gill net, total effort ident = 44170. Only one observation, species is “no particular species”, survey type is historical

14. Grace found whole surveys-lake-year that were excluded from chris custer cpe output but were present in lake finder 
    Examples sent to denver to explore what might be happening
      LOW 2013, 1997, and 1996 not showing up in CPE
        2013 has a flag
        96 and 97 have NA dates - all historical surveys?
      Artichoke 2021 gillnet not showing up in CPE
        Total effort of 3 with an area of 1501+ (effort is confirmed on lake finder)





libraries
```{r}
library(arrow)
library(readr)
library(tidyverse)
library(stringr)
library(data.table)
library(janitor)
library(tidyr)
library(lubridate)
library(bit64)
library(ggplot2)
library(mwlaxeref)

options(scipen = 999)
```




# Load Data

* note depending on your local settings, file paths need correction to match local GDrive letter

```{r}
##build functions  
remove_unnecessary_suffixes <- function(dt) {
  # Identify columns with numeric suffixes
  numeric_suffix_cols <- grep("\\.\\d+$", names(dt), value = TRUE)
  
  # Create a list to store prefixes
  prefixes <- unique(sub("\\.\\d+$", "", numeric_suffix_cols))
  
  # Remove unnecessary numeric suffixes
  for (prefix in prefixes) {
    cols <- grep(paste0("^", prefix, "\\.\\d+$"), names(dt), value = TRUE)
    if (length(cols) == 1) {
      setnames(dt, cols, sub("\\.\\d+$", "", cols))
    }
  }
  
  return(dt)
}

remove_unnecessary_suffixes_2 <- function(dt) {
  # Identify columns with numeric suffixes
  numeric_suffix_cols <- grep("\\_\\d+$", names(dt), value = TRUE)
  
  # Create a list to store prefixes
  prefixes <- unique(sub("\\_\\d+$", "", numeric_suffix_cols))
  
  # Remove unnecessary numeric suffixes
  for (prefix in prefixes) {
    cols <- grep(paste0("^", prefix, "\\_\\d+$"), names(dt), value = TRUE)
    if (length(cols) == 1) {
      setnames(dt, cols, sub("\\_\\d+$", "", cols))
    }
  }
  
  return(dt)
}

#generate a file list to import
files_list <- list.files(path = "G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files
files_list

#object for use in loop (simple length of file list)
n <- length(files_list)

for(i in 1:n) {
  #i = 3
  outer_break <- FALSE# clean up so that break tool will refresh
  
  filei <- word(gsub(".csv","", files_list[i]), start = -1, sep = fixed("/"))
  #this does those two steps in one package
  assign(filei ,
          fread(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/",
                                          files_list[i])))
  
  # if the file is a crosswalk, do not rename anything, just loop to the confirm import line
  if(str_detect(filei, "crosswalk")) {  #confirm import of files:  
    print(paste(filei ,"added to workspace" ))  
    #confirm import of files:  
    print(paste(i ,"files added to workspace" )) ; next}
  
  
  # note we want to review a sorted list of column names to check misspelling etc.
  # we still need to use the columns with names like col_name_length_in, or known_units
  
  
  cde %>% # call data explainer file
    filter(`new_file_name`== filei)%>% #keep only the row relevant to this file
    select_if(~ !any(is.na(.))) %>% 
    data.table::transpose(keep.names = "newname") %>% 
    rename("oldname" = V1) %>% 
    assign("names", ., envir = .GlobalEnv)
  
  #see if any column names will not have a match! 
  # IF any pop FALSE, force stop and revist of data explainer ()
  # - e.g., named something "total catch" when actual column name was "total_catch"
  print(
    cbind(colnames(get(filei)),
          colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]
    )
  )
  
  
  # break the loop if the current file has column names not in the data explainer
  # if (all(cbind(colnames(get(filei)),  colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ])[,2]) == FALSE ) break
  if (all(colnames(get(filei)) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]) == FALSE ) {
    warning(paste("Warning: Import renaming failed at file", filei))
    outer_break <-  TRUE
    break}
  
  
  # append old col names into new "notes" columns:
  get(filei)[ , (names[ str_detect(newname, "notes") , oldname   ,  ]) := Map(paste, colnames(.SD), .SD, sep = ':') , .SDcols =  names[ str_detect(newname, "notes") , oldname   ,  ] ]
  
  #now rename that file's colnames
  setnames(get(filei), colnames(get(filei)), names[!str_detect(newname,"unique_row_key")] [match(names(get(filei)),names[!str_detect(newname,"unique_row_key"),oldname]), newname] )
  
  #append all other data from data explainer
  unusedbits <- 
    data.table(
      matrix(
        rep(names[ !newname %in% colnames(get(filei)) , oldname , ],
            each = nrow(get(filei))
        ),
        nrow = nrow(get(filei)),
        dimnames = list(rep(NA,nrow(get(filei))),
                        names[ !newname %in% colnames(get(filei)) , newname , ])
        )
      )
  
  #add all not yet used columns from data explainer:
  get(filei)[ , (names[ !newname %in% colnames(get(filei)) , newname , ]) := unusedbits[] ]

  rm(names, unusedbits)
  
  
  #remove the unneeded suffixes from data
  assign(filei, remove_unnecessary_suffixes_2(remove_unnecessary_suffixes(get(filei))))
  
  #confirm import of files:  
  print(paste(filei ,"added to workspace" ))  
  #confirm import of files:  
  print(paste(i ,"files added to workspace" )) 

  
} 
if (outer_break) {print("Data import issue");rm(outer_break)
  } else {
  #confirm import of files:  
  print(paste(i ,"files added to workspace" ))
  #confirm import of files:  
  print(paste(n-i ,"remaining to be added" )) 
}
  
  
```


# Effort

```{r}

# effort dataset has a unique key in the combo of (lake_id, date.2(survey date), gear, survey type)
mn_effort_18aug2023[          , .N , .(lake_id, date.1, sampling_method, survey_type) ]

#note that there is a component count column in here that seems like it's related to effort, but not exactly clear to me:
mn_effort_18aug2023[ , summary(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2))) ,]
mn_effort_18aug2023[ , hist(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), breaks = 100) ,]
#actually, these appear to be like total.count
mn_effort_18aug2023[ ,plot(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), total_count.1) , ]#CPUE included count
mn_effort_18aug2023[ ,plot(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), total_count.2) , ]#cpue excluded count
mn_effort_18aug2023[ ,plot(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)), total_count.1+total_count.2) , ] #boom! there is is, component count = the sum of catch between fish marked to exclude form CPUE and those marked to include
mn_effort_18aug2023[ !(as.numeric(gsub("COMPONENT_COUNT:","",gear_data_notes.2)) == total_count.1+total_count.2) ,, ]


# we have two total effort cols, one is EF seconds, which I think is redundant to the main effort column:
mn_effort_18aug2023[ , total_effort_2 := gsub("EF_SECONDS:", "" , total_effort_2), ]
mn_effort_18aug2023[ total_effort_2 == "NA" , total_effort_2 := NA  ,]
mn_effort_18aug2023[  , total_effort_2 := as.numeric(total_effort_2) ,] #one weird one we'll ignore (retain total.count, drop component count implicity by process)


mn_effort_18aug2023[ , .N , .("te2NA" = is.na(total_effort_2), "te1NA" = is.na(total_effort_1))] # any cases where all dat in total.effort.2
#NOPE

#here we can see that ef time is covered by the total effort 1 column
plot(mn_effort_18aug2023[ !is.na(total_effort_2) , total_effort_2,]/3600~ mn_effort_18aug2023[ !is.na(total_effort_2) , total_effort_1 ,])
#what about the one's where they don't jive?
mn_effort_18aug2023[ round(total_effort_2/3600, 2) != total_effort_1, plot(total_effort_2/3600~total_effort_1), ]
abline(0,1)

#other weirdness?
mn_effort_18aug2023[ , summary(total_effort_1) , ]
mn_effort_18aug2023[total_effort_1 == 0, ]#review of these records indicates they oughtta stay in, effort moved over to tot eff 1
mn_effort_18aug2023[total_effort_1 == 0, total_effort_1 := total_effort_2]

mn_effort_18aug2023[is.na(total_effort_1)]


#we don't need that extra total eff column
mn_effort_18aug2023[ , total_effort_2 := NULL , ]

#the DNR sent along two values for counts, one where they count only cases where CPUE was marked yes in the indiv fish data and one count including only where cpue was marked no in the indiv fish data
#in the effort data thius is the number of units of gear that were used (and thus number of units marked us/don't use for cpues)

mn_effort_18aug2023[total_count.2>0 , .(total_count.1,total_count.2) , ]
mn_effort_18aug2023[ , plot(total_count.1~total_count.2) , ]

#are there any effort data with NO associated count (i.e., nothing caught)? 
mn_effort_18aug2023[   , .N , .("tc1_0" = total_count.1 == 0, "tc2_0" = total_count.2 == 0) ]
#NOPE

# Crazy that there are "no zeros". However, because these data are gear level, this is agg'd across species so relatively useless.

#drop these columns
mn_effort_18aug2023[ , `:=` ("total_count.1" = NULL,"total_count.2" = NULL) , ]


#fix up dates
mn_effort_18aug2023[ , date_survey :=  as.IDate(word(date.1, 1, sep = fixed(" ")), format = "%m/%d/%Y"),]
mn_effort_18aug2023[ is.na(date_survey), ]

mn_effort_18aug2023[ , hist(yday(date_survey)) ,]
mn_effort_18aug2023[ , hist(year(date_survey)) ,]

#check other fields
names(mn_effort_18aug2023)


#tidy sampling methods
mn_effort_18aug2023[ , .N , sampling_method ]
mn_effort_18aug2023[ , .N , .(sampling_method, sampling_method_abbrev)]

#keep the abbreviation/code column because that's the only thing in the CPUE file
mn_effort_18aug2023[ , .N , sampling_method_abbrev , ]
mn_cpue_21aug2023[ , .N , sampling_method_abbrev]

any(is.na(match(mn_effort_18aug2023[ , unique(sampling_method_abbrev),], mn_cpue_21aug2023[ , unique(sampling_method_abbrev) , ])))
#all abbreviations are covered (we could come back to that to add full gears names field to the cpues)

sort(names(mn_effort_18aug2023))

#year
mn_effort_18aug2023[ , unique(year) , ]
mn_effort_18aug2023[ is.na(year),]
 sum(mn_effort_18aug2023[ ,year != year(date_survey)])
#non-issue, delete year column
mn_effort_18aug2023[ , year := NULL] 

#effort units
mn_effort_18aug2023[ , .N , sampling_method ]


mn_effort_18aug2023[ , summary(total_effort_1) ,]

mn_effort_18aug2023[ total_effort_1 <0 , , ]

ggplot( mn_effort_18aug2023[!total_effort_1 < 0 ], aes(total_effort_1, group = sampling_method))+
  geom_density()+
  facet_wrap(~sampling_method, scales = "free")

# we can get units for these effort data from here (p45ish):
# https://files.dnr.state.mn.us/publications/fisheries/special_reports/180.pdf




```



# CPUE 2(3) ways
We have CPUE directly form the DNR, we can generate our own cpue using the individual fish data (both including and excluding the fish marked for CPUE inclusion)
Update 15 Sep 2023: We recieved a new data rip from MN DNR in three pieces: effort, catch, and cpue. Why catch and CPUE you ask? Because the DNR tells me that there are historical surveys (pre ~1985) that have CPUEs but no data for the indiv fish caught in those surveys. 

```{r}

  #cpue data
  mn_cpue_21aug2023[  , .N , .(lake_id, date, sampling_method_abbrev, survey_type, species) ]
  sum(duplicated(mn_cpue_21aug2023))
  #not really sure why there would be duplicated cpues in there... 
  mn_cpue_21aug2023[duplicated(mn_cpue_21aug2023) , .N, .(lake_name, lake_id, date) ]
  #if these cpues are calc'd per net or instance, that could cause it. how many of each gear in each survey?
  mn_cpue_21aug2023[ , .N, .(lake_name, lake_id, date, species, sampling_method_abbrev) ][, summary(N)]
  mn_cpue_21aug2023[ ,mean(total_effort), .(sampling_method_abbrev) ]
  mn_cpue_21aug2023[ , .N, .(lake_name, lake_id, date, species, sampling_method_abbrev) ][N>1, ]

  #not likely a problem. I think sometimes multiple rows are reported for a given gear on a survey, when the catch is == between two of them you get an aparrent "duplicate". This would be mitigated by a sub_effort_ident coming from the db. 
  
  #we'll be circling back to this dataset as a way to get the historical un-lengthed fish data into our fish-obs-as-rows format
  
  
  #fix up dates
  mn_cpue_21aug2023[ , date_survey :=  as.IDate(date),]
  mn_cpue_21aug2023[ is.na(date_survey), ]

  mn_cpue_21aug2023[ , hist(yday(date_survey)) ,]
  mn_cpue_21aug2023[ , .N ,.(lake_name,lake_id,year(date_survey))][ , hist(year) ,]
  
  #compared to the effort file, it's obvious there is overlap in these and that the CPUE stuff has more old data
  mn_effort_18aug2023[ ,.N ,.(lake_name,lake_id,year(date_survey))][ , hist(year) ,]
  
  
  #check overlap betweek effort and CPUE:
  mn_cpue_21aug2023[            , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]
  mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]

   mn_cpue_21aug2023[            , .("cpueN" = .N) , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ][
     mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , effortN := "N" 
   ][ , .N , .("missingCPUE" = is.na(cpueN), "missing_effort" = is.na(effortN))] #
   

   
  mn_cpue_21aug2023[            , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]
  mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ]

   mn_cpue_21aug2023[            , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ][
     mn_effort_18aug2023[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , 
   ]
  
  # Awesome-- the effort data are entirely contained in the cpue file. So here's a plan: 
   # 1. we use the fish obs and effort to set up our obs level file
   #  2a. Identify CPUE rows unmatched in the obs level surveys (for which we have nothing in effort or indiv fish). 
   #  2b. expand those rows into obs level format (but no wieghts or lengths)
   #  2c. bring them back to that main obs level file
  
```


# Prep & Scope Indiv Fish Data

```{r}

#inspect data
colnames(mn_indiv_fish_23nov2023)
sort(colnames(mn_indiv_fish_23nov2023))

setcolorder(mn_indiv_fish_23nov2023, c("lake_id", "date.2", "date.1", "sampling_method_abbrev", "survey_type")) 

mn_indiv_fish_23nov2023[ , .N , .(sampling_method_abbrev, species)  ]
mn_indiv_fish_23nov2023[ species == "WAE", .N , .(sampling_method_abbrev)  ] 
mn_indiv_fish_23nov2023[ species == "WAE", .N , .(sampling_method_abbrev)  ][ , sum(N) , ]

#to-do:
# make these clean: "lake_id", "date_clean", "sampling_method", "survey_type.1"

#dates (from data explainer we know that date.2 is survey date, we prob want to join on that one to connect to the effort data)
mn_indiv_fish_23nov2023[ , .(date.2, date.1), ]
mn_indiv_fish_23nov2023[ , .N , .(date1=is.na(date.1), date2=is.na(date.2)) ]


#change dates to Idate:
mn_indiv_fish_23nov2023[ , date_survey := as.IDate(date.2, format = "%m/%d/%Y") ,]
mn_indiv_fish_23nov2023[ , date_sample := as.IDate(date.1, format = "%m/%d/%Y") ,]


mn_indiv_fish_23nov2023[  , summary(date_sample - date_survey)   ,  ]
mn_indiv_fish_23nov2023[ (abs(date_sample - date_survey) < 400) == TRUE , hist(date_sample - date_survey)   ,  ]

mn_indiv_fish_23nov2023[ , summary(date_survey) , ]
mn_indiv_fish_23nov2023[is.na(date_survey) , .N ]
#no missingness!

mn_indiv_fish_23nov2023[ , summary(date_sample) , ]
mn_indiv_fish_23nov2023[is.na(date_sample) , .N ]
#no missingness!


#lake_id
mn_indiv_fish_23nov2023[ , summary(lake_id) ,]
mn_indiv_fish_23nov2023[ , str(lake_id), ]
mn_effort_18aug2023[ , str(lake_id) ,]
#no issues here


#tidy sampling methods
mn_indiv_fish_23nov2023[ , .N , .(sampling_method_abbrev)]
mn_effort_18aug2023[ , .N , .(sampling_method, sampling_method_abbrev)]


any(is.na(match(mn_effort_18aug2023[ , unique(sampling_method_abbrev),], mn_indiv_fish_23nov2023[ , unique(sampling_method_abbrev) , ])))
#all abbreviations are covered (we could come back to that to add full gears to the indiv fish, but that'll happen in the merge, so skipped here. )


#survey.type
mn_indiv_fish_23nov2023[ , .N , survey_type ]
mn_effort_18aug2023[ , .N , survey_type]

mn_indiv_fish_23nov2023[ , .N , survey_type ][ ,survey_type, ]


# generate a type table (these codes can be found in MN_Data folder)
# Assuming you have the data in a vector or a data frame column named 'codes'
codes <- c("IS", "RS", "PA", "SA", "CR", "WK", "DO", "GM", "CP", "LL", "SS", "NR", "OT", "RE", "XM", "XR", "SD", "TS")
full_names <- c("Initial Survey", "Re-Survey", "Population Assessment", "Special Assessment", "Creel Survey", "Winter Kill Assessment", 
                "Dissolved Oxygen Check", "Game Survey", "Collectors Permit", "Large Lake Survey", "Subsurvey", "Natural Reproduction Check", 
                "Investigational Survey", "Research Survey", "External Management Survey", "External Research Survey", "Standard Survey", "Targeted Survey")

# Create a data.table
survey_type_key <- data.table(code = codes, fullname = full_names)

# Print the resulting data.table
print(survey_type_key)


#execute
any(is.na(match(mn_indiv_fish_23nov2023[ , survey_type], survey_type_key[,code]))) #add codes are in key
# survey_type_key[match(mn_indiv_fish_23nov2023[ , survey_type], survey_type_key[,code]), fullname]

mn_indiv_fish_23nov2023[ , survey_type := survey_type_key[match(mn_indiv_fish_23nov2023[ , survey_type], survey_type_key[,code]), fullname]   ,]
rm(survey_type_key)

#uncount indiv_fish
mn_indiv_fish_23nov2023[, summary(total_count) , ]
mn_indiv_fish_23nov2023[total_count == 0 , .N , ]

#expected result size:
mn_indiv_fish_23nov2023[ , sum(total_count) , ] + mn_indiv_fish_23nov2023[total_count == 0 , .N , ]


#execute dataset expansion/uncount
mn_indiv_fish_23nov2023 <- 
rbind(
  mn_indiv_fish_23nov2023[total_count == 0 , , ], 
  uncount(mn_indiv_fish_23nov2023, weights = total_count, .remove = F)
)



#summarize ages
mn_indiv_fish_23nov2023[ , .N , age ]
mn_indiv_fish_23nov2023[ , .N , .(age=!is.na(age))]


```



# Merge effort and indiv fish
```{r}
#merge the effort and indiv fish data


#scope the merge:
#99% of the wb in the effort data are represneted in indiv fish data
sum(mn_indiv_fish_23nov2023[ , unique(lake_id) , ] %in% mn_effort_18aug2023[ , unique(lake_id)])/mn_effort_18aug2023[, length(unique(lake_id)) ,] 

#100% of the wb in the indiv fish data are represneted in  effort data
sum(mn_indiv_fish_23nov2023[ , unique(lake_id) , ] %in% mn_effort_18aug2023[ , unique(lake_id)])/mn_indiv_fish_23nov2023[, length(unique(lake_id)) ,] 

#merge the effort and indiv fish data

intersect(names(mn_effort_18aug2023), names(mn_indiv_fish_23nov2023))

setdiff(names(mn_effort_18aug2023), names(mn_indiv_fish_23nov2023))

# setnames(mn_effort_18aug2023, old = "date_clean", new = "date_survey")

#MEMORY HOG PINCH POINT
#we know that dates are a problem, can we schmooze that join? 
#we leave many column names out of this merge key to avoid disconnecting things where one dataset has empty info
flatfish <- merge(mn_effort_18aug2023,mn_indiv_fish_23nov2023, by = c("lake_id", "date_survey" , "sampling_method_abbrev", "survey_type"), suffixes = c("_effort", "_indivfish"), all = T)

rm(mn_indiv_fish_23nov2023,mn_effort_18aug2023)

flatfish[ , .N , sampling_method]
flatfish[ , .N, species]


#check on the annual #s of aged, lengthed, both, and neither numbers
a <- setnames(
  dcast(flatfish[ (species == "WAE" & sampling_method %in% c("Standard gill net sets")), .N, .(year = year(date_survey), age = !is.na(age), length = !is.na(length))][order(year, -N)],
        formula = year ~ age + length, fill = 0 ),
  old = c("FALSE_FALSE", "FALSE_TRUE", "TRUE_TRUE"), new = c("neither", "length_only", "both" ))


names(flatfish)

#how many unique surveys exist for each category of has effort and has indiv fish data?
flatfish[ , length(unique(paste(lake_id, date_survey , sampling_method, survey_type))) , .(effort = !is.na(state_effort), indiv_fish_dat = !is.na(state_indivfish)) ]
  
#sum of all
  flatfish[ , length(unique(paste(lake_id, date_survey , sampling_method, survey_type))) , .(effort = !is.na(state_effort), indiv_fish_dat = !is.na(state_indivfish)) ][ , sum(V1) , ]

  
  
#now roll the catch CPUE data back to this bigger file:
  
##Context for this next operation:
#   3. Historical WAE survey duplications show same gear/lake/date cathing many (from CPUE) and zero (from indivfish)
# - Consider dropping the effort table info where we have a HISTORICAL cpue record (or at least scope this strategy)


     
  #We've got a bad survey type thingy going on here where cpue is historical:
   mn_cpue_21aug2023[ , .N , survey_type ]
   flatfish[, .N , survey_type]
   
   # consider dropping records from the effort table where we have a cpue in a historical?
   #send this off to Corey Geving to review 
   
   # setnames(mn_cpue_21aug2023, old = "date_clean", new = "date_survey")
   
  flatfish[mn_cpue_21aug2023[survey_type == "HISTORICAL", , ], on = .(lake_id, date_survey, sampling_method_abbrev) , nomatch = NULL ][ , .(neffcatch = .N, n_cpue = sum(i.total_count), nlen = sum(!is.na(length))) , .(lake_id, date_survey, sampling_method_abbrev, survey_type, i.survey_type) ][order(lake_id,date_survey)] 
  
  
  #for these records we want to push the flatfish record out completely. 
  badflatfish <- flatfish[mn_cpue_21aug2023[survey_type == "HISTORICAL", , ], on = .(lake_id, date_survey, sampling_method_abbrev) , nomatch = NULL ][ , .(neffcatch = .N, n_cpue = sum(i.total_count), nlen = sum(!is.na(length))) , .(lake_id, date_survey, sampling_method_abbrev, survey_type, i.survey_type) ][order(lake_id,date_survey)][nlen %in% c(0), ]
  
  
  flatfish <- flatfish[!badflatfish, on = .(lake_id, date_survey, sampling_method_abbrev,survey_type) ]
  
  rm(badflatfish)

  #subset of MN_cpue where lake/date/method are not matched:
   glimpse(mn_cpue_21aug2023[ 
     !flatfish[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , 
   ])
  
  #trim to only the not already covered data:
  mn_cpue_21aug2023 <- mn_cpue_21aug2023[ 
     !flatfish[          , .N , .(lake_id, date_survey, sampling_method_abbrev, survey_type) ], 
     on = .(lake_id, date_survey, sampling_method_abbrev, survey_type) , 
   ]
  
  mn_cpue_21aug2023[ , summary(total_count) , ]
  mn_cpue_21aug2023[total_count == 9999 ]
  
  
  
  #execute dataset expansion/uncount
mn_cpue_21aug2023 <- 
rbind(
  mn_cpue_21aug2023[total_count == 0 , , ], 
  uncount(mn_cpue_21aug2023, weights = total_count, .remove = F)
)

names(flatfish)


  #MEMORY HOG STEP
  #now bring that together with the flatfish data:
  flatfish <- rbindlist(list(flatfish, mn_cpue_21aug2023 ),
                               fill = TRUE,
                               use.names = TRUE)
  
  rm(mn_cpue_21aug2023)
  
  #for each of those with both effort and fish data, summarize catch by species (bring effort along too!)   
flatfish[ !is.na(state_effort)&!is.na(state_indivfish) , .N , .(lake_id, date_survey , sampling_method, survey_type, species, total_effort_1) ]


  
```


# Save an rds to work from 
```{r}





# saveRDS(flatfish, file = "Data_and_Scripts\\Data\\output\\mn_flat_effort_indivfish_merge.rds")


#this flatfish file is big AF and will slow down Mike's Hansen Lab Computer (32gb RAM)
# flatfish <- readRDS("Data_and_Scripts/Data/output/mn_flat_effort_indivfish_merge.rds")






```


#Gear codes, species codes, state, county, length age units, target spp
```{r}
#we need to
# 1. revises the data in a few columns: sampling method, nothing caught, species names, state, units, target species, effort,  
# 2. export notes for the workshop. (delete, flag, clarify, keep for workshop) then dump those columns
# 3. match to schema (https://docs.google.com/spreadsheets/d/1zfevASMxRMxMYNWm3Hq1yR2Qk0zcJV_JnO0LBeODaPQ/edit#gid=0)

mn_dat <- flatfish
rm(flatfish)
names(mn_dat)

# dump some known trash

mn_dat[, c("targeted_or_standard_effort", "targeted_or_standard_indivfish", "targeted_or_standard",
           "new_file_name_effort", "new_file_name_indivfish", "new_file_name",
           "source_contact_effort", "source_agency", "source_agency_effort", "source_contact_indivfish", "source_agency_indivfish", "source_contact",
           "file_status_effort", "file_type_effort", "file_status_indivfish", "file_type_indivfish", "file_status", "file_type",
           "jan2023_inventoried_effort", "jan2023_inventoried_indivfish", "jan2023_inventoried",
           "data_type_effort", "data_type_indivfish", "data_type",
           "date_recieved_effort", "date_recieved_indivfish", "date_recieved", 
           "unique_row_key_effort", "unique_row_key_indivfish", "unique_row_key",
           "total_count") := NULL, ]


#gear full names need filling out
    mn_dat[ , .N , .(sampling_method_abbrev, sampling_method)]
    
    #borrow codes as key from other rows
    gear_table_mn <- mn_dat[!is.na(sampling_method), .N , .(sampling_method_abbrev,sampling_method) ]
    
    
    #apply these to all rows
    #drop old col
    mn_dat[ , sampling_method := NULL ,]
    #rejoin to complete gear table
    mn_dat[gear_table_mn,
      on = .(sampling_method_abbrev), sampling_method := sampling_method
          ]
    #checkwork
    mn_dat[ , .N , .(sampling_method_abbrev, sampling_method)]
    
 
#add nothing_caught
    mn_dat[is.na(species), nothing_caught := T , ] #The only species == NA are places where lake/date/gear/surveytype found no species
     
    
#species names are mixed codes and names
    #what cols are have species in name?
    mn_dat %>%
      select(contains("species"))
    #show problem
    mn_dat[ , .N , species ]
    
    # Review "codes" as defined by 3 letter species names
    print(mn_dat[nchar(species) == 3, .N , species ],
          nrow(mn_dat[nchar(species) == 3, .N , species ]))
    
    # choose only species.1 length of 3, assign these to a species_code
    mn_dat[nchar(species) == 3, `:=` (species_abbrev = species, species = NA) , ]
    
    a <- mn_dat[ ,.N , species_abbrev ]
    
    #generate a species code table (will also use later to fix target species cols)    
      species_codes_mn <-  mn_dat[!is.na(species_abbrev), .(species_abbrev = unique(species_abbrev)) ,  ] 
    
    #adds names to that table (uses key provided by Jon Hansen, MNDNR on 10 Oct 2023: https://docs.google.com/spreadsheets/d/1ZakE4S2zfSdYNEWHfikFqILNnD1gYpNf?rtpof=true&usp=drive_fs )    
        #This chunk cleans codes and names within the species column, references? for codes?
      species_codes_mn[ , species := 
                          case_when(# first fish codes are converted,
                            species_abbrev == "LMB" ~ "largemouth_bass",
                            species_abbrev == "SMB" ~ "smallmouth_bass",
                            species_abbrev == "WAE" ~ "walleye", 
                            species_abbrev == "WTS" ~ "white_sucker",
                            species_abbrev == "BLC" ~ "black_crappie",
                            species_abbrev == "YEP" ~ "yellow_perch",
                            species_abbrev == "BLG" ~ "bluegill",
                            species_abbrev == "NOP" ~ "northern_pike",
                            species_abbrev == "PMK" ~ "pumpkinseed", 
                            species_abbrev == "RKB" ~ "rock_bass",
                            species_abbrev == "BKF" ~ "banded_killifish",
                            species_abbrev == "IOD" ~ "iowa_darter",
                            species_abbrev == "BCS" ~ "blackchin_shiner",
                            species_abbrev == "BLB" ~ "black_bullhead",
                            species_abbrev == "JND" ~ "johnny_darter",
                            species_abbrev == "GOS" ~ "golden_shiner",
                            species_abbrev == "YEB" ~ "yellow_bullhead",
                            species_abbrev == "BRB" ~ "brown_bullhead",
                            species_abbrev == "CNM" ~ "central_mudminnow",
                            species_abbrev == "NRD" ~ "northern_redbelly_dace",
                            species_abbrev == "CSH" ~ "chinook_salmon",
                            species_abbrev == "MTS" ~ "mottled_sculpin",
                            species_abbrev == "TPM" ~ "tadpole_madtom",
                            species_abbrev == "SHR" ~ "shorthead_redhorse",
                            species_abbrev == "HSF" ~ "hybrid_sunfish",
                            species_abbrev == "BOF" ~ "bowfin",
                            species_abbrev == "WDS" ~ "weed_shiner",
                            species_abbrev == "BNM" ~ "bluntnose_minnow",
                            species_abbrev == "BND" ~ "blacknose_dace",
                            species_abbrev == "OTM" ~ "minnows",
                            species_abbrev == "FHM" ~ "fathead_minnow",
                            species_abbrev == "FND" ~ "finescale_dace",
                            species_abbrev == "SDM" ~ "slender_madtom",
                            species_abbrev == "SPO" ~ "spottail_shiner",
                            species_abbrev == "BNS" ~ "blacknose_shiner",
                            species_abbrev == "BKT" ~ "brook_trout",
                            species_abbrev == "SPT" ~ "splake",
                            species_abbrev == "RBT" ~ "rainbow_trout",
                            species_abbrev == "BNT" ~ "brown_trout",
                            species_abbrev == "BUB" ~ "burbot",
                            species_abbrev == "BIB" ~ "bigmouth_buffalo",
                            species_abbrev == "BLH" ~ "bullheads",
                            species_abbrev == "TLC" ~ "cisco",
                            species_abbrev == "CCF" ~ "channel_catfish",
                            species_abbrev == "BSD" ~ "blackside_darter",
                            species_abbrev == "BKS" ~ "brook_silverside",
                            species_abbrev == "LED" ~ "least_darter",
                            species_abbrev == "LGP" ~ "logperch",
                            species_abbrev == "SLR" ~ "silver_redhorse",
                            species_abbrev == "HHC" ~ "hornyhead_chub",
                            species_abbrev == "MUE" ~ "muskellunge",
                            species_abbrev == "TRP" ~ "trout_perch",
                            species_abbrev == "RHS" ~ "redhorse",
                            species_abbrev == "CAP" ~ "common_carp",
                            species_abbrev == "MMS" ~ "mimic_shiner",
                            species_abbrev == "EMS" ~ "emerald_shiner",
                            species_abbrev == "SUN" ~ "sunfish",
                            species_abbrev == "CRC" ~ "creek_chub",
                            species_abbrev == "GSF" ~ "green_sunfish",
                            species_abbrev == "SFS" ~ "spotfin_shiner",
                            species_abbrev == "SHI" ~ "shiners",
                            species_abbrev == "BST" ~ "brook_stickleback",
                            species_abbrev == "CSR" ~ "central_stoneroller",
                            species_abbrev == "SDS" ~ "sand_shiner",
                            species_abbrev == "GRR" ~ "greater_redhorse",
                            species_abbrev == "PGS" ~ "pugnose_shiner",
                            species_abbrev == "SMS" ~ "slimy_sculpin",
                            species_abbrev == "LND" ~ "longnose_dace",
                            species_abbrev == "FRD" ~ "freshwater_drum",
                            species_abbrev == "WHC" ~ "white_crappie",
                            species_abbrev == "DAR" ~ "darters",
                            species_abbrev == "SAR" ~ "sauger",
                            species_abbrev == "FCF" ~ "flathead_catfish",
                            species_abbrev == "BMS" ~ "bigmouth_shiner",
                            species_abbrev == "LKS" ~ "lake_sturgeon",
                            species_abbrev == "LKW" ~ "lake_whitefish",
                            species_abbrev == "BRM" ~ "brassy_minnow",
                            species_abbrev == "SLM" ~ "mississippi_silvery_minnow",
                            species_abbrev == "GLR" ~ "golden_redhorse",
                            species_abbrev == "OTS" ~ "suckers",
                            species_abbrev == "NST" ~ "ninespine_stickleback",
                            species_abbrev == "GOE" ~ "goldeye",
                            species_abbrev == "QBS" ~ "quillback",
                            species_abbrev == "RVS" ~ "river_shiner",
                            species_abbrev == "LES" ~ "longear_sunfish",
                            species_abbrev == "CHL" ~ "chestnut_lamprey",
                            species_abbrev == "RRH" ~ "river_redhorse",
                            species_abbrev == "WHB" ~ "white_bass",
                            species_abbrev == "OSS" ~ "orangespotted_sunfish",
                            species_abbrev == "STC" ~ "stonecat",
                            species_abbrev == "SNG" ~ "shortnose_gar",
                            species_abbrev == "RFS" ~ "rosyface_shiner",
                            species_abbrev == "CMS" ~ "carmine_shiner",
                            species_abbrev == "SAB" ~ "smallmouth_buffalo",
                            species_abbrev == "GIS" ~ "gizzard_shad",
                            species_abbrev == "LNG" ~ "longnose_gar",
                            species_abbrev == "RCS" ~ "river_carpsucker",
                            species_abbrev == "SLS" ~ "shovelnose_sturgeon",
                            species_abbrev == "LAT" ~ "lake_trout",
                            species_abbrev == "TME" ~ "tiger_muskellunge",
                            species_abbrev == "PRD" ~ "pearl_dace",
                            species_abbrev == "PLS" ~ "pallid_shiner",
                            species_abbrev == "RBS" ~ "rainbow_smelt",
                            species_abbrev == "RBD" ~ "rainbow_darter",
                            species_abbrev == "LNS" ~ "longnose_sucker",
                            species_abbrev == "SJC" ~ "shortjaw_cisco",
                            species_abbrev == "DWS" ~ "deepwater_sculpin",
                            species_abbrev == "SHS" ~ "spoonhead_sculpin",
                            species_abbrev == "BDD" ~ "banded_darter",
                            species_abbrev == "MOE" ~ "mooneye",
                            species_abbrev == "BLS" ~ "blue_sucker",
                            species_abbrev == "SLC" ~ "silver_chub",
                            species_abbrev == "HFS" ~ "highfin_carpsucker",
                            species_abbrev == "MCP" ~ "mirror_carp",
                            species_abbrev == "WAS" ~ "walleye/sauger",
                            species_abbrev == "AME" ~ "american_eel",
                            species_abbrev == "GOF" ~ "goldfish",
                            species_abbrev == "SPS" ~ "spotted_sucker",
                            species_abbrev == "PAH" ~ "paddlefish",
                            species_abbrev == "RVD" ~ "river_darter",
                            species_abbrev == "NHS" ~ "northern_hog_sucker",
                            species_abbrev == "BHM" ~ "bullhead_minnow",
                            species_abbrev == "WSD" ~ "western_sand_darter",
                            species_abbrev == "MDD" ~ "mud_darter",
                            species_abbrev == "SIL" ~ "silver_lamprey",
                            species_abbrev == "SKJ" ~ "skipjack_herring",
                            species_abbrev == "PGM" ~ "pugnose_minnow",
                            species_abbrev == "BKB" ~ "black_buffalo",
                            species_abbrev == "FTD" ~ "fantail_darter",
                            species_abbrev == "YLB" ~ "yellow_bass",
                            species_abbrev == "PRP" ~ "pirate_perch",
                            species_abbrev == "LKC" ~ "lake_chub",
                            species_abbrev == "RIR" ~ "river_ruffe",
                            species_abbrev == "WHP" ~ "white_perch",
                            species_abbrev == "ALW" ~ "alewife",
                            species_abbrev == "SEL" ~ "sea_lamprey",
                            species_abbrev == "BHC" ~ "bighead_carp",
                            species_abbrev == "GLD" ~ "gilt_darter",
                            species_abbrev == "CRD" ~ "crystal_darter",
                            species_abbrev == "SMT" ~ "speckled_madtom",
                            species_abbrev == "RDS" ~ "red_shiner",
                            species_abbrev == "WAM" ~ "warmouth",
                            species_abbrev == "CIS" ~ "cisco",
                            species_abbrev == "CPS" ~ "carpsucker",
                            species_abbrev == "CRP" ~ "crappie",
                            species_abbrev == "HCR" ~ "hybrid_crappie",
                            species_abbrev == "SCU" ~ "sculpin",
                            species_abbrev == "SHD" ~ "slenderhead_darter",
                            species_abbrev == "SIP" ~ "northern_pike_silver_phase",
                            species_abbrev == "UK1" ~ "unknown_fish",
                            species_abbrev == "UK2" ~ "unknown_fish",
                            # below unknown codes specified as such
                            species_abbrev == "LXB" ~ "see_ambiguous_species_code",
                            TRUE ~ species_abbrev
                          )
      ]
      
      #some dulicated common names (two codes for one thing)
      species_codes_mn[duplicated(species_codes_mn$species), , ]
      
      setnames(species_codes_mn, "species" , "common.name" )
      
        #fill species names
      mn_dat[species_codes_mn,
             on = .(species_abbrev=species_abbrev),
             species.fix := common.name 
      ]
        #fill into species.1 col, drop fix column
        mn_dat[is.na(species), species := species.fix , ]
        mn_dat[ , species.fix := NULL ,]
       
        
        #clean up common names
        mn_dat[ , species := str_replace_all( str_replace_all( str_replace_all(tolower(species),
                                                                                 " ", "_") ,
                                                                 "\\)" , "") ,
                                                "\\(" , "")]
        
        mn_dat[ , species := 
                  case_when(
                    species == "tullibee_cisco" ~ "cisco",
                    species == "trout-perch" ~ "trout_perch",
                    species == "bowfin dogfish" ~ "bowfin",
                    # unknown species
                    species == "no_particular_species" ~ "no_particular_species",
                    species == "unknown_fish_one" ~ "unknown_fish",
                    species == "walleye/sauger" ~ "walleye_x_sauger",
                    TRUE ~ species
                  )
        ]
        
        #check work
        print(mn_dat[ , .N , species][order(species)], nrow(mn_dat[ , .N , species])) 
        
        #still 1876 NAs here (nothing caught surveyXgears)
        mn_dat[is.na(species) , .N , .(nothing_caught)  ]
        
        
        #drop species codes
        species_codes_mn
        mn_dat[, .N , .(species_abbrev,species) ]
        mn_dat[ , `:=` (species_abbrev = NULL)]

        
# fix state
       #what cols are have species in name?
    mn_dat[ , .N ,  .(state, state_effort, state_indivfish) ]
    
    #collapse all
    mn_dat[ , `:=` (state = "Minnesota", state_effort = NULL, state_indivfish = NULL) , ]
    
#fix county
    mn_dat[ , .N ,  .(county_effort, county_indivfish) ]
    mn_dat[is.na(county_effort), county_effort := county_indivfish ]
              mn_dat[ , c("county_indivfish") := NULL]
    

# fix unit names for length and age
    #formats the length unit for clarity
    mn_dat[ , .N , length_unit ]
      mn_dat[ !is.na(length) ,.N  , length_unit ]
    
    mn_dat[ , .N , weight_unit]
      mn_dat[ !is.na(weight) ,.N  , weight_unit ]
    
    mn_dat[!is.na(length_unit), length_unit := "mm" ,]
    mn_dat[!is.na(weight_unit), weight_unit := "g" ,]

# target species
    #targeted species - needs cleaning, has species codes (clean up with the species code table generated above)
    mn_dat[ , .N , target_species ]
    mn_dat[ , c("T1","T2", "T3", "T4", "T5")  :=   tstrsplit(target_species, split = ", "), ]
    
    mn_dat[species_codes_mn, on = .(T1=species_abbrev) , target_species_1 := common.name ]
    mn_dat[species_codes_mn, on = .(T2=species_abbrev) , target_species_2 := common.name ]
    mn_dat[species_codes_mn, on = .(T3=species_abbrev) , target_species_3 := common.name ]
    mn_dat[species_codes_mn, on = .(T4=species_abbrev) , target_species_4 := common.name ]
    mn_dat[species_codes_mn, on = .(T5=species_abbrev) , target_species_5 := common.name ]
    
    mn_dat[ !is.na(target_species), unique(paste(target_species_1,target_species_2, target_species_3, target_species_4,target_species_5))]
    
    mn_dat[ , target_species := str_remove_all(paste(target_species_1,target_species_2, target_species_3, target_species_4,target_species_5, sep = ","),
                                               ",NA"),]
    
    mn_dat[ , c("T1","T2", "T3", "T4", "T5",
                "target_species_1","target_species_2", "target_species_3", "target_species_4", "target_species_5") := NULL]
    
#fix sexes
mn_dat[,.N, sex]

mn_dat[ sex == "U", sex := "Unknown"]
mn_dat[ sex == "M", sex := "Male"]
mn_dat[ sex == "F", sex := "Female"]
mn_dat[ sex == "N" | sex == "" | is.na(sex), sex := NA]
    
    
```

# Efforts
```{r}
# effort needs attention. Currently we have a negative effort and no units (nab code to add units from Holly based on gear type)
    #effort
    mn_dat %>% 
      select(contains("effort")) %>% 
      colnames()
    #total_effort_1.1
    
    gear_table_mn[order(sampling_method_abbrev)]
    
    #we got one total eff from the cpues, one from the fish/effort
    mn_dat[ ,.N  , .(is.na(total_effort_1), is.na( total_effort))]
    
    
    mn_dat[ !is.na(total_effort), .(mean = mean(total_effort), min =min(total_effort), max = max(total_effort)) , ]
    mn_dat[ !is.na(total_effort_1), .(mean = mean(total_effort_1), min =min(total_effort_1), max = max(total_effort_1)) , ]
    
    mn_dat[ !is.na(total_effort), .(mean = mean(total_effort), min =min(total_effort), max = max(total_effort)) , .(sampling_method_abbrev)]
    mn_dat[ !is.na(total_effort_1), .(mean = mean(total_effort_1), min =min(total_effort_1), max = max(total_effort_1)) , .(sampling_method_abbrev) ]
    
    #consolidate cols
    mn_dat[is.na(total_effort_1), total_effort_1 := total_effort]
    mn_dat[ ,total_effort := NULL]
    
    #no effort units, we have a negative effort? 
    
    # we can get units for these effort data from here (p45ish):
    # https://files.dnr.state.mn.us/publications/fisheries/special_reports/180.pdf
    #manually added the units for AOW, TMF, and GFM -- assuming same as similar gear categories
    # effort_table <- data.table( CODE = c(c("AIF", "TOU", "ATL", "STL", "EXT", "XXX", "MT", "SA", "ZO", "AOW"),#NONE
    #                              c("TR" , "STR", "SEF", "EFB", "EW", "EF"), #HOUR 
    #                              c("SSE", "SE", "S18", "S58"), #HAUL
    #                              c("GN", "GDE", "GSH", "GSM", "GST", "GSU", "GNP", "FCI", "TML", "SGN", "GNA", "VGN", "TN", "LTN", "DTN", "TQU", "T38", "THA", "T34", "T1", "STN", "GFM", "TMF")),#SET  
    #                     effort_unit = c(rep("no_unit", 10), rep("hours", 6), rep("haul", 4), rep("set", 23)))
    
    effort_table <- data.table(
  Code = c("GN", "GDE", "GSH", "GSM", "GST", "GSU", "GNP", "FCI", "TML", "SGN", "GNA", "VGN", "TN", "LTN", "DTN", "TQU", "T38", "THA", "T34", "T1", "STN", "EF", "EW", "EFB", "SEF", "S58", "S18", "SE", "SSE", "AIF", "AOW", "TOU", "ATL", "STL", "TR", "STR", "EXT", "XXX", "MT", "SA", "ZO"),
  Method = c("Gill net Standard graduated-mesh gill net", "Gill net Deep GN net, in waters colder than 12 o C", "Gill net Shallow GN net, in waters warmer than 12 o C", "Gill net Small-mesh (0.375 and 0.5-in-bar) net", "Gill net Short-term sets, monofilament net", "Gill net Suspended set, any gill net type", "Gill net Large-mesh (1.5, 2.0, 2.5-in bar) net", "Gill net Fish Community Index nets (Ontario)", "Gill net Trammel net", "Gill net Other (special) gill net", "Gill net Standard North American gill net", "Gill net Vertical gill net", "Trap net Standard 0.75-in-mesh double-frame net", "Trap net Large-frame (5x6 ft) net", "Trap net Pair of TN nets, set in double-pot configuration", "Trap net 0.25-in-mesh net", "Trap net 0.375-in-mesh net", "Trap net 0.5-in-mesh net", "Trap net 0.75-in-mesh single-frame net", "Trap net 1.0-in-mesh net", "Trap net Other (special) trap net", "Electrofishing Standard Boat electrofishing", "Electrofishing Fall electrofishing targeting Walleye", "Electrofishing Backpack electrofishing", "Electrofishing Other (special) electrofishing", "Seine 50-ft bag seine, 0.125-in mesh", "Seine 15-ft seine, 0.125-in mesh", "Seine 50-ft bag seine, 0.25-in mesh", "Seine Other (special) seining", "Angling Ice fishing", "Angling Open-water angling", "Angling Tournament angling", "Trot line Normal trot-line configuration", "Trot line Other (special) trot line", "Trawl Large lake trawl", "Trawl Other (special) trawl", "Miscellaneous Fish collected from or by an external source", "Miscellaneous Fish taken by miscellaneous means", "Miscellaneous 0.25-in-mesh minnow trap", "Miscellaneous Fish spawning trap", "Zooplankton Zooplankton sampling gear"),
  effort_unit = c("net_night", "net_night", "net_night", "net_night", "sets", "net_night", "net_night", "net_night", "net_night", "sets", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night", "net_night","sets", "hours", "hours", "hours", "hours", "hauls", "hauls", "hauls", "hauls", "none", "none", "none", "none", "none", "hours", "hours", "none", "none", "none", "none", "none")
)
    
    
    #add effort units to gear tbale       
    gear_table_mn[effort_table, on = .(sampling_method_abbrev=Code), `:=` (total_effort_unit = effort_unit, sampling_method_desc = Method)]
    
    mn_dat[ , .N , .(is.na(sampling_method_abbrev), is.na(sampling_method)) ]
    
    #and push those over to the main df
    mn_dat[gear_table_mn, on = .(sampling_method_abbrev), `:=` (total_effort_unit = total_effort_unit, sampling_method_desc = sampling_method_desc)]
    
    #cleanup and checkwork
    rm(effort_table)
    mn_dat[ , .N , .(sampling_method,sampling_method_desc,  total_effort_unit)]
    
    mn_dat[ total_effort_1 <= 0 ][order(site_id)] #weird, but not diagnosable/fixable IMO.
    mn_dat[ total_effort_1 <= 0, flag := "effort value error" ]
    
    
#survey_id 
    mn_dat %>%
      select(contains("survey_")) %>% 
      colnames()
    
    mn_dat[ ,  .N, .(is.na(survey_id), is.na(survey_id_effort))]
    
    #something looks wonky in these survey Ids but the MNDNR says they're not useful anyways... see Minnnesota_README (https://docs.google.com/document/d/1sXag9RDvbKN0pUB6uRlADz36-kFNqRPwHVGlkJgyWTM/edit)
    
    mn_dat[ , `:=` (survey_id = NULL, survey_id_effort = NULL, survey_id_indivfish = NULL) , ]
    
    
#lake_name
              mn_dat[, .N, .(catchisNA = is.na(lake_name),
                                                                effortisNA = is.na(lake_name_effort),
                                                                fishisNA = is.na(lake_name_indivfish))
                 ]
            mn_dat %>% 
            mutate(same_survID_eff_fish = ifelse(lake_name_effort == lake_name_indivfish, "Same", "Different")) %>% 
            mutate(same_survID_eff_catch = ifelse(lake_name_effort == lake_name, "Same", "Different")) %>% 
            mutate(same_survID_catch_fish = ifelse(lake_name == lake_name_indivfish, "Same", "Different")) %>% 
            group_by(same_survID_eff_fish,same_survID_eff_catch,same_survID_catch_fish) %>% 
            count()
          
            mn_dat[is.na(lake_name), lake_name := ifelse(is.na(lake_name_effort), lake_name_indivfish, lake_name_effort)]
              mn_dat[ , c("lake_name_effort","lake_name_indivfish") := NULL]
              
#date
              sort(names(mn_dat))
              
                  mn_dat %>% 
      select(contains("date")) %>% 
      colnames()
              
              mn_dat[ , .N , is.na(date_survey)]
              
              mn_dat[ , length(unique(date.2_effort)) , .(lake_id,date_survey) ]
              
              mn_dat[ , .N , .(is.na(date_survey), is.na(date.1_indivfish))]
              
              #dump all extra dates:
              #"date.1_effort" and "date.2_indivfish" are equivalent to date_survey, date.2_effort (sample date) came with the fish as date_sample, previously (date.1_indivfish), , 
              
              mn_dat[ , c("date.1_effort", "date.2_indivfish", "date.2_effort", "date.1_indivfish", "date", "end_date") := NULL , ]
              
              
              
# add an effort_ident column "lake_id", "date_clean" , "sampling_method_abbrev", "survey_type.1"effort_ident := .GRP
    mn_dat[ , .N ,.(lake_id, date_survey, survey_type, sampling_method  #survey 
                                                        ) ]  
        #verify extra fields align with this level - they dont
        mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
                         total_effort_1, total_effort_unit, nothing_caught 
                                                        ) ] 
        
        #which don't?
                mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
                         total_effort_1, total_effort_unit, nothing_caught 
                                                        ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1]
        #Grab out full sets from summaries: this shows that the key issue here is that there are two total effort vals for some of these
                mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, and grab out the ones that our key fails on 
                  mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
                                   total_effort_1, total_effort_unit, nothing_caught #addt'l key metrics
                  ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1,
                                                                                      .(lake_id, date_survey, sampling_method)],
                  on = .(lake_id, date_survey, sampling_method), ]
                
        # snag a sample for review: this suggests that there's soem kind of mis-aggregation or miscalculation fo total efforts in the CPUE dataset that is effing these up. I'm going to flag them and press on        
          # a <-       mn_dat[mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, 
          #         mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey 
          #                          total_effort_1, total_effort_unit, nothing_caught 
          #         ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1, .(lake_id, date_survey, sampling_method)], on = .(lake_id, date_survey, sampling_method), ][1:2] ,
          #         on = .(lake_id, date_survey, sampling_method) , ]       
          #       
        
        #flag records where two total efforts are happenign in one survey for a gear: 20988 records 
                mn_dat[mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, 
                  mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method,
                                   total_effort_1, total_effort_unit, nothing_caught
                  ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1, .(lake_id, date_survey, sampling_method)], on = .(lake_id, date_survey, sampling_method), ] ,
                  on = .(lake_id, date_survey, sampling_method) , .N ,  ]
                
                mn_dat[mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, total_effort_1, total_effort_unit, nothing_caught ) ][ #take the summary table, 
                  mn_dat[ , .N ,.( lake_id, date_survey, survey_type, sampling_method, #survey
                                   total_effort_1, total_effort_unit, nothing_caught
                  ) ] [ , .N , .(lake_id, date_survey, survey_type, sampling_method)][N>1, .(lake_id, date_survey, sampling_method)], on = .(lake_id, date_survey, sampling_method), ] ,
                  on = .(lake_id, date_survey, sampling_method) , flag := ifelse(is.na(flag),
                                               "Questionable total effort, 2 values in one survey" ,
                                               paste(flag, "Questionable total effort, 2 values in one survey", sep = ";" ))  ]
                
                

        #adds in our effort_ident field
        mn_dat[ , total_effort_ident := .GRP ,.( lake_id, date_survey, survey_type, #survey 
                        sampling_method, total_effort_1, total_effort_unit, nothing_caught 
                                                        ) ] 
        
        #add a total effort nothing caught
mn_dat[ , all(nothing_caught)  , .(total_effort_ident)][ ,summary(V1)] #1876 seems reasonable, ~3% of TEIs found nothing.
mn_dat[ , all(is.na(species))  , .(total_effort_ident)][ ,summary(V1)]#check against another method of calculation

mn_dat[ , total_effort_nothing_caught :=  all(is.na(species))  , .(total_effort_ident)]

mn_dat[ , .N , total_effort_nothing_caught == nothing_caught]
mn_dat[ ,nothing_caught := NULL]

```


#notes cols tidying
```{r}
    

#CPUE Y/N column
  
  mn_dat[ , .N , notes.1 ]
  mn_dat[ notes.1 == "CPUE:N" , flag := ifelse(is.na(flag),
                                               "Do not use in CPUE calcs" ,
                                               paste(flag, "Do not use in CPUE calcs", sep = ";" ))]
  mn_dat[  , notes.1 := NULL , ]
  
#representative sampling column
  
  mn_dat[ , .N , notes ]
  mn_dat[ notes %in% c("REPRESENTATIVE_SAMPLING:N", "REPRESENTATIVE_SAMPLING:N, Y", "REPRESENTATIVE_SAMPLING:Y, N") , flag := ifelse(is.na(flag),
                                                                          "Not representative sampling" ,
                                                                          paste(flag, "Not representative sampling", sep = ";" ))]
  mn_dat[  , notes := NULL , ]  
  
#other notes  
  mn_dat[ , .N , notes.2] #send to workshop
  # workship results
  note_revisions <- fread("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_comments_notes_reviewed.csv")
  
  note_revisions[ , .N , flag]
  
  note_revisions <- note_revisions[flag != "", .(flag,notes.2) ]
  
  setnames(note_revisions, "flag", "note_flag")
  
  mn_dat[note_revisions, on = .(notes.2), note_flag := note_flag]
    mn_dat[  , .N , note_flag]
  
  mn_dat[ !is.na(note_flag) , flag := ifelse(is.na(flag),
                                                                          note_flag ,
                                                                          paste(flag, note_flag, sep = ";" ))]
  mn_dat[ , .(n = .N, n_surv = length(unique(total_effort_ident))) , flag]
  
  mn_dat[  , notes.2 := NULL , ]  
  
  mn_dat[ ,note_flag := NULL , ]
    
    
  
  mn_dat[ , .N , garbage_bin_notes.1_indivfish] #junk
  mn_dat[ , .N , garbage_bin_notes.1_effort]#junk
  mn_dat[ , .N , garbage_bin_notes.2_effort] #junk
  mn_dat[ , .N , garbage_bin_notes.2_indivfish] #junk
  mn_dat[ , .N , garbage_bin_notes.3] # junk

  mn_dat[ , `:=` (garbage_bin_notes.1_indivfish = NULL, garbage_bin_notes.1_effort = NULL, garbage_bin_notes.2_effort = NULL, garbage_bin_notes.3 = NULL)]

  
      
#gear notes  
  mn_dat[ , gear_data_notes := paste(gear_data_notes,
                                     gear_data_notes.1,
                                     gear_data_notes.2,
                                     gear_data_notes.3,
                                     gear_data_notes.4,
                                     gear_data_notes.5, sep = "_") ,]
  
  mn_dat[ , .N, gear_data_notes]
  mn_dat[ , .N, gear_data_notes.1]
  mn_dat[ , .N, gear_data_notes.2]
  mn_dat[ , .N, gear_data_notes.3]
  mn_dat[ , .N, gear_data_notes.4]
  mn_dat[ , .N, gear_data_notes.5] #junk
   # mn_dat[!is.na(gear_data_notes.5) , .N , .(gear_data_notes.5, total_effort, total_effort_unit) ]
  
    mn_dat[ , gear_data_notes.1 := NULL , ]  
    mn_dat[ , gear_data_notes.2 := NULL , ]
    mn_dat[ , gear_data_notes.3 := NULL , ]
    mn_dat[ , gear_data_notes.4 := NULL , ]
    mn_dat[ , gear_data_notes.5 := NULL , ]
  
#location notes  
  mn_dat[!(location_notes %in% c("LOC_DESC:","LOC_DESC:null")|is.na(location_notes)) , .N , location_notes ]  #keep
  
#lake type
  mn_dat[ , .N , lake_type ] #minimal info, but keep for consistency to other states
  
#aging str
  mn_dat[ , .N , .(aging_structure.1, aging_structure.2) ]
  mn_dat[ aging_structure.1 == "" , aging_structure.1 := NA , ]
  mn_dat[aging_structure.2 == "" , aging_structure.2 := NA , ]
  mn_dat[!is.na(aging_structure.2), aging_structure.1 := aging_structure.2]
  
  mn_dat[ , aging_structure.2 := NULL , ]

## Now grab exportable cols for the workshop
  names(mn_dat)
  
  
# #what cols go out to the team? (We'll carry these forward & round up at the end)
#   # gear stuff
#         dat <- mn_dat[ , .("records" = .N, "surveys" = length(unique(total_effort_ident))) , by= .(gear_data_notes, gear_data_notes.1, gear_data_notes.2, gear_data_notes.3, gear_data_notes.4,sampling_method, sampling_method_desc)]
#         
#         # Function to extract the first word from a string
#         extract_first_word <- function(x) {
#           non_na_values <- na.omit(x)
#           if (length(non_na_values) > 0) {
#             first_word <- strsplit(as.character(non_na_values[1]), " ")[[1]][1]
#             return(first_word)
#           }
#           return(NULL)
#         }
#         
#         # Loop through each column and update column name
#         for (col in names(dat)[1:3]) {
#           first_word <- extract_first_word(dat[[col]])
#           if (!is.null(first_word)) {
#             setnames(dat, col, first_word)
#           }
#         }
# 
#       # Function to strip the first word from a string
#       strip_first_word <- function(x) {
#         if (!is.na(x)) {
#           stripped_string <- sub("^[^:]+:", "", as.character(x))
#           return(stripped_string)
#         }
#         return(x)
#       }
#       
#       # Loop through each column and update values
#       for (col in names(dat)[1:3]) {
#         dat[, (col) := lapply(.SD[[col]], strip_first_word)]
#       }
#       
#       #fix column names
#       names(dat) <- word(names(dat), 1, sep = fixed(":"))
#       
#       # Print the updated data.table#write to file
#       
#       # fwrite(dat, file = "MN_gear_notes.csv")
#   
# # other generic notes
#         dat <- mn_dat[ , .("records" = .N, "surveys" = length(unique(total_effort_ident))) , by= .(garbage_bin_notes.2_indivfish, location_notes, notes, notes.2)]
#         
#         # Function to extract the first word from a string
#         extract_first_word <- function(x) {
#           non_na_values <- na.omit(x)
#           if (length(non_na_values) > 0) {
#             first_word <- strsplit(as.character(non_na_values[1]), " ")[[1]][1]
#             return(first_word)
#           }
#           return(NULL)
#         }
#         
#         # Loop through each column and update column name
#         for (col in names(dat)[1:3]) {
#           first_word <- extract_first_word(dat[[col]])
#           if (!is.null(first_word)) {
#             setnames(dat, col, first_word)
#           }
#         }
# 
#       # Function to strip the first word from a string
#       strip_first_word <- function(x) {
#         if (!is.na(x)) {
#           stripped_string <- sub("^[^:]+:", "", as.character(x))
#           return(stripped_string)
#         }
#         return(x)
#       }
#       
#       # Loop through each column and update values
#       for (col in names(dat)[1:3]) {
#         dat[, (col) := lapply(.SD[[col]], strip_first_word)]
#       }
#       
#       #fix column names
#       names(dat) <- word(names(dat), 1, sep = fixed(":"))
#       
#       # Print the updated data.table#write to file
#       
#       # fwrite(dat, file = "MN_comments_notes.csv")
#   
#   rm(col, first_word, dat)
#   
#   
#   
#   
  
# This section moved to the 
# # adds columns needed for state filtering
#   #adding lake area to the data
# area <- read_csv("Data_and_Scripts/Data/input/mn_lake_list.csv") %>%
#   select(DOW, LAKE_AREA_DOW_ACRES) %>%
#   mutate(DOW = as.integer(DOW))
# #joins fish data with area data
# mn_dat <- left_join(mn_dat, area, by = c("lake_id" = "DOW")) %>%
#   #hard codes lakes that do not have area in the area file (areas taken from lake finder)
#   mutate(LAKE_AREA_DOW_ACRES = as.numeric(case_when(
#         lake_id == "56038800" ~ "1288",
#         lake_id == "55006400" ~ "20",
#         lake_id == "76014601" ~ "467",
#         lake_id == "76014602" ~ "218",
#         TRUE ~ as.character(LAKE_AREA_DOW_ACRES)))) %>%
#   #creates binned area grouping used for minimum effort calculation
#   mutate(area.group = case_when(LAKE_AREA_DOW_ACRES < 100 ~ "<100",
#                                 LAKE_AREA_DOW_ACRES >= 100 & LAKE_AREA_DOW_ACRES <= 300 ~ "100-300",
#                                 LAKE_AREA_DOW_ACRES >= 301 & LAKE_AREA_DOW_ACRES <= 600 ~ "301-600",
#                                 LAKE_AREA_DOW_ACRES >= 601 & LAKE_AREA_DOW_ACRES <= 1500 ~ "601-1500",
#                                 LAKE_AREA_DOW_ACRES >= 1501 ~ "1501+",
#                                 TRUE ~ "NA")) %>%
#   #orders groupings so they plot well
#   mutate(area.group = factor(area.group, levels = c("<100", "100-300", "301-600", "601-1500", "1501+")))
# 
# 
#   setnames(mn_dat, c("LAKE_AREA_DOW_ACRES", "area.group"), c("lake_area_dow_acres", "area_group"))
  

  
#consolidate original filenames
  
  mn_dat[ , original_file_names := str_remove_all(paste(original_file_name_effort,original_file_name_indivfish,original_file_name, sep = ", "),
                                               ", NA"),]
  mn_dat[ , `:=` ("original_file_name_effort" = NULL ,"original_file_name_indivfish" = NULL , "original_file_name" = NULL) ,]
  
  mn_dat[ , .N ,  original_file_names]
  
  mn_dat[ , original_file_names := str_remove_all(original_file_names, "NA, ") , ]
  

#yoy as age class
  mn_dat[ , .N  ,  young_of_year]
  mn_dat[ young_of_year == "Y", age_class := "young_of_year"]
  mn_dat[ young_of_year == "N", age_class := "not_young_of_year"]
  mn_dat[ ,young_of_year := NULL]
    
  
  
```


#sampling_methods
```{r}
mn_dat <- clean_names(mn_dat) #clean up names a touch

# here's what were after: 
#sampling_method_simple, sampling_method, sampling_method_2, gear_data_notes
mn_dat[ , .N , .(sampling_method_abbrev, sampling_method, sampling_method_desc)]

#and we need to add some data that we'll peel from here:

mn_dat[ , .N , gear_data_notes ]
  mn_dat[ , mesh := str_extract(gear_data_notes, "(?<=MESH:)[^_]+")]
 # the biggest mesh GN I could find was 8" at Duluth Nets, I think it's too ambiguous for me to interpret literally as a numeric inches. 
    mn_dat[ , .N , .(mesh) ][order(mesh)]
    mn_dat[mesh == "NA", mesh := NA]
  mn_dat[ , n_netters := as.numeric(str_extract(gear_data_notes, "(?<=NUMBER_OF_NETTERS:)[^_]+"))]
    mn_dat[ ,summary(n_netters) ]
  mn_dat[ , daylight_sampling := str_extract(gear_data_notes, "(?<=DAYLIGHT_SAMPLING:)[^_]+")]
    mn_dat[ , .N, .(daylight_sampling) ]
    mn_dat[ daylight_sampling %in% c("N, Y", "Y, N"), daylight_sampling := NA , ]
  
    
    
    
#gear xwalk
gear_xwalk <- fread("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/crosswalks/gears_by_state.csv", na.strings = "")%>% 
  filter(state == "Minnesota")

setnames(mn_dat, old = c("sampling_method", "sampling_method_desc"), new = c("gear_1", "gear_2"))

mn_dat[gear_xwalk, on = .(gear_1, gear_2), `:=` (sampling_method_simple = i.sampling_method_simple,
                                                 sampling_method_1 = i.sampling_method_1,
                                                 sampling_method_2 = i.sampling_method_2
                                                 )
       ]
mn_dat[ , .N , sampling_method_simple ]
mn_dat[ , .N , sampling_method_1 ]



#now bring over the gear notes
mn_dat[daylight_sampling == "Y", ':=' (sampling_method_1 = paste(sampling_method_1,"day",sep = "_"),
                                       sampling_method_2 = paste(sampling_method_2,"day",sep = "_")), ]

mn_dat[daylight_sampling == "N", ':=' (sampling_method_1 = paste(sampling_method_1,"night",sep = "_"),
                                       sampling_method_2 = paste(sampling_method_2,"night",sep = "_")) , ]
mn_dat[!is.na(n_netters), .N , sampling_method_simple ]
mn_dat[!is.na(n_netters),  ':=' (sampling_method_2 = paste(sampling_method_2, paste(n_netters, "netters",sep = "_"), sep = "; ")), ]

mn_dat[!is.na(mesh), .N , sampling_method_simple ]
mn_dat[!is.na(mesh) ,mesh := paste( "mesh", mesh, sep = ":" ) , ]
#moved to notes_1 iaw strategy from Large Lakes datasets
setnames( mn_dat ,  old = "mesh" , new = "notes_1" )
mn_dat[ , .N , .(notes_1 , sampling_method_simple) ]

#remove old cols and clean up workspace

mn_dat[ , ':='(gear_1 = NULL, gear_2 = NULL, n_netters = NULL, daylight_sampling = NULL)  ,]

rm(gear_xwalk, gear_table_mn, species_codes_mn, note_revisions, a, cde)

```


#NHDids
```{r}

  #use mwlaxeref to fetch NHDIDs
  # mwlaxeref::mn_to_nhdhr(mn_dat, from_colname = "lake_id")[,nhdhr.id]
  
  mn_dat[ , nhdhr_id := mwlaxeref::mn_to_nhdhr(mn_dat, from_colname = "lake_id")[,nhdhr.id] , ]
    

  #remaining unmatched lakes:
  mn_dat[is.na(nhdhr_id), .N , lake_id ]
  


msu_xwalk <- fread("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/crosswalks/MGLP_FISH_LAKES_12Aug24.csv") %>% 
  filter(STATE == "MN") 
setDT(msu_xwalk)

msu_xwalk[, lake_id := as.integer(STATE_ID) ,]

#add locs and NHDs from MSU
mn_dat[msu_xwalk, on = .(lake_id = lake_id) , ':=' (nhdhr_id_MSU = gsub("nhdhr_", "", NHDHR_ID), latitude_lake_centroid = LAT, longitude_lake_centroid = LON) ]

  #remaining unmatched lakes:
  mn_dat[, .N , .(lake_id, nhdhr_id, nhdhr_id_MSU) ][ , .N , .(is.na(nhdhr_id), is.na(nhdhr_id_MSU)) ]
  #agreement? 71 not aligned...
  mn_dat[, .N , .(lake_id, nhdhr_id, nhdhr_id_MSU) ][!is.na(nhdhr_id) & !is.na(nhdhr_id_MSU), .N , .(nhdhr_id == nhdhr_id_MSU) ]

#great. still missing ~134 lakes
  mn_dat[is.na(nhdhr_id), nhdhr_id := nhdhr_id_MSU] #fill where missing
  mn_dat[ !is.na(nhdhr_id) & !is.na(nhdhr_id_MSU) & nhdhr_id_MSU != nhdhr_id, .N , .(lake_id, nhdhr_id, nhdhr_id_MSU) ]  #some lakes have mismatches:
  #for these we oughtta saver both?
  mn_dat[!is.na(nhdhr_id) & !is.na(nhdhr_id_MSU) & nhdhr_id_MSU != nhdhr_id , notes_1 :=  ifelse(is.na(notes_1),
                                               paste("Secondary_nhdhrid", nhdhr_id_MSU , sep = ":" ) ,
                                               paste(notes_1, paste("Secondary_nhdhrid", nhdhr_id_MSU , sep = ":" ), sep = "; " )), ]
  
  mn_dat[ , nhdhr_id_MSU:= NULL , ]

# more lat/lons?
usgs <- fread("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/crosswalks/usgs_lake_metadata.csv") %>% 
  filter(state == "MN")

usgs[ , nhdhr_id := gsub("nhdhr_", "", site_id ) , ]  

 #add locs and NHDs from MSU
mn_dat[usgs, on = .(nhdhr_id = nhdhr_id) ,  ':=' (latitude_lake_centroid_USGS = centroid_lat, longitude_lake_centroid_USGS = centroid_lon) ]
 
#remaining unknown location lakes: 126.
  mn_dat[, .N , .(lake_id, nhdhr_id, latitude_lake_centroid, longitude_lake_centroid, latitude_lake_centroid_USGS, longitude_lake_centroid_USGS) ][ , .N , .(is.na(latitude_lake_centroid), is.na(latitude_lake_centroid_USGS)) ]
  
mn_dat[ is.na(latitude_lake_centroid), ':=' (latitude_lake_centroid = latitude_lake_centroid_USGS, longitude_lake_centroid = longitude_lake_centroid_USGS)]
  
mn_dat[ , ':=' (latitude_lake_centroid_USGS = NULL, longitude_lake_centroid_USGS = NULL )  ,  ]  
  
  
  #remaining unmatched lakes:
  mn_dat[is.na(nhdhr_id) | is.na(latitude_lake_centroid), .N , .(lake_id, lake_name, county_effort) ]


```

#dates
```{r}

mn_dat[ , .SD , .SDcols = colnames(mn_dat)[str_detect(colnames(mn_dat), "date")]]
#how many survey dates in a TEI? ONE ONLY ONE! HOORAY
mn_dat[ , length(unique(date_survey)) , total_effort_ident ][order(-V1)]
mn_dat[ , length(unique(date_sample)) , total_effort_ident ][order(-V1)]
mn_dat[ , date_total_effort_ident := date_survey ]







```

#Column Names
```{r}
rm(usgs, msu_xwalk)

#Denver's clean up of columns and names
mn_dat <- mn_dat %>% 
  mutate(
    state = "Minnesota",
    county = county_effort,
    lake_name = lake_name,
    lake_id = as.character(lake_id),
    nhdhr_id = nhdhr_id,
    
    latitude_lake_centroid = latitude_lake_centroid,
    longitude_lake_centroid = longitude_lake_centroid,
    
    date_survey = date_survey,
    date_total_effort_ident = date_total_effort_ident,
    date_sub_effort_ident = as.IDate(NA),
    date_sample = date_sample,
    year = year(date_survey),
    month = month(date_survey),
    
    survey_id = as.character(NA),
    survey_type = survey_type,
    survey_type_2 = as.character(NA),
    survey_type_3 = as.character(NA),
    survey_type_4 = as.character(NA),
    
    sampling_method_abbrev = NULL,
    sampling_method_simple = sampling_method_simple,
    sampling_method = sampling_method_1,
    sampling_method_2 = sampling_method_2,
    gear_data_notes = gear_data_notes,
    target_species = target_species,
    
    total_effort_ident = total_effort_ident,
    total_effort_1 = total_effort_1,
    total_effort_2 = as.numeric(NA),
    total_effort_1_units = total_effort_unit,
    total_effort_2_units = as.character(NA),
    total_effort_nothing_caught = total_effort_nothing_caught,
    
    water_temp = water_temp,
    water_temp_units = case_when(!is.na(water_temp) ~ "fahrenheit", TRUE ~ NA),
    water_clarity = as.numeric(NA),
    water_clarity_units = as.character(NA),
    
    lat_start = as.numeric(NA),
    lon_start = as.numeric(NA),
    lat_end = as.numeric(NA),
    lon_end = as.numeric(NA),
    site_id = site_id,
    
    sub_effort_ident = as.character(NA),
    sub_effort_1 = as.numeric(NA),
    sub_effort_1_units = as.character(NA),
    sub_effort_2 = as.numeric(NA),
    sub_effort_2_units = as.character(NA),
    sub_effort_nothing_caught = NA,
    
    species_1 = species,
    length_1 = as.numeric(length),
    length_unit_1 = length_unit, 
    length_bin = as.character(NA),
    length_bin_unit = as.character(NA),
    
    age = age,
    aging_structure_1 = aging_structure_1,
    
    weight_1 = weight,
    weight_unit_1 = weight_unit,
    batch_weight = as.character(NA),
    batch_weight_unit = as.character(NA),
    
    sex = sex,
    age_class = age_class,
    
    flag = flag,
    original_file_names = original_file_names,
    ind_fish_ident = as.character(sample_id),
    # lakesize = as.numeric(NA),
    # lakesize_units = as.numeric(NA),
    lat_unspec = as.numeric(NA),
    lon_unspec = as.numeric(NA),
    waterbody_type = lake_type,
    location_notes_1 = location_notes,
    obs_id = as.character(row_number())
  ) %>% 
  unite("notes_1", c(notes_1,
                     garbage_bin_notes_2_indivfish,
                     reproductive_condition_notes),
                     remove = T,
                     na.rm = T) %>% 
  setDT()

# intersect(names(mw_data), names(mn_dat))
# 
# setdiff(names(mw_data), names(mn_dat)
#            )

schema_names <- c(
  "state", "county", "lake_name", "lake_id", "nhdhr_id",

  "latitude_lake_centroid", "longitude_lake_centroid",

  "date_survey", "date_total_effort_ident", "date_sub_effort_ident", "date_sample", "year", "month",

  "survey_id", "survey_type", "survey_type_2", "survey_type_3", "survey_type_4",

  "sampling_method_simple", "sampling_method", "sampling_method_2", "gear_data_notes","target_species",

  "total_effort_ident", "total_effort_1", "total_effort_2", "total_effort_1_units", "total_effort_2_units", "total_effort_nothing_caught",

  "water_temp", "water_temp_units", "water_clarity", "water_clarity_units",

  "lat_start", "lon_start", "lat_end", "lon_end",

  "site_id", "sub_effort_ident", "sub_effort_1", "sub_effort_1_units", "sub_effort_2","sub_effort_2_units", "sub_effort_nothing_caught",

  "species_1", "length_1", "length_unit_1", "length_bin", "length_bin_unit",

  "age", "aging_structure_1",

  "weight_1", "weight_unit_1", "batch_weight", "batch_weight_unit",

  "sex", "age_class",

  "flag", "original_file_names",

  "ind_fish_ident",
  "lat_unspec", "lon_unspec",
  "waterbody_type",
  "location_notes_1", "notes_1", "obs_id"
)

setcolorder(mn_dat, neworder = schema_names)


# match column classes
glimpse(mn_dat)

dropcols <- colnames(mn_dat)[!colnames(mn_dat)%in%schema_names]

mn_dat[ , (dropcols):= NULL , ]


# column classes

chr_cols <- c("state", "county", "lake_name", "lake_id", "nhdhr_id", "survey_id", "survey_type",
  "survey_type_2", "survey_type_3", "survey_type_4", "sampling_method_simple",
  "sampling_method", "sampling_method_2", "gear_data_notes", "target_species",
  "total_effort_ident","total_effort_1_units",
  "total_effort_2_units", "water_temp_units", "water_clarity_units","site_id", "sub_effort_ident",
  "sub_effort_1_units","sub_effort_2_units","species_1","length_unit_1", "length_bin", "length_bin_unit","aging_structure_1",
   "weight_unit_1", "batch_weight", "batch_weight_unit", "sex", "age_class", "flag",
  "original_file_names", "ind_fish_ident", "waterbody_type", "location_notes_1", "notes_1", "obs_id")

mn_dat[, (chr_cols) := lapply(.SD, as.character), .SDcols = chr_cols]

num_cols <- c("latitude_lake_centroid", "longitude_lake_centroid" , "year", "month", "total_effort_1", "total_effort_2",
              "water_temp", "water_clarity", "lat_start", "lon_start", "lat_end", "lon_end", "sub_effort_1", "sub_effort_2",
              "length_1", "age", "weight_1",  "lat_unspec", "lon_unspec")

mn_dat[, (num_cols) := lapply(.SD, as.numeric), .SDcols = num_cols]

mn_dat[ , .N , .(sampling_method_simple, sampling_method_2, flag) ]



```

#checks, flags, schema match
```{r}
#effort
mn_dat[ , .N , .(total_effort_1_units)]
mn_dat[is.na(total_effort_1) , .N , total_effort_1_units  ]
mn_dat[is.na(total_effort_1) , total_effort_1_units := NA  ]



#creating an auto flagging system
flag_xwalk <- mn_dat %>% 
  distinct(total_effort_ident, .keep_all = TRUE) %>% 
  group_by(sampling_method_simple) %>% 
  mutate(mean_effort = mean(total_effort_1, na.rm = TRUE),
         sd_effort = sd(total_effort_1, na.rm = TRUE),
         z_score = (total_effort_1 - mean_effort) / sd_effort,
         flag = NA, #creates a flag column so I can case_when the flag generation 
         flag = case_when(total_effort_1 < 0 ~ paste("invalid effort (negative)", coalesce(flag, "")),
                          total_effort_1 == 0 ~ paste("effort reported as 0", ";", coalesce(flag, "")),
                          z_score > 10 ~ paste("high effort", ";", coalesce(flag, "")),
                          is.na(z_score) ~ coalesce(flag, ""),
                          TRUE ~ coalesce(flag, "")
         ),
         flag = na_if(flag, "")) %>% 
  select(total_effort_ident,
         sampling_method_simple,
         flag,
         total_effort_1,
         mean_effort,
         sd_effort,
         z_score) %>% 
  filter(!is.na(flag))

flag_xwalk_2 <- mn_dat %>% 
  distinct(total_effort_ident, .keep_all = TRUE) %>%
  filter(!(total_effort_ident %in% flag_xwalk$total_effort_ident)) %>% 
  group_by(sampling_method_simple) %>% 
  mutate(mean_effort = mean(total_effort_1, na.rm = TRUE),
         sd_effort = sd(total_effort_1, na.rm = TRUE),
         z_score = (total_effort_1 - mean_effort) / sd_effort,
         flag = NA,
         flag = case_when(total_effort_1 < 0 ~ paste("invalid effort (negative)", coalesce(flag, "")),
                          total_effort_1 == 0 ~ paste("effort reported as 0", ";", coalesce(flag, "")),
                          z_score > 10 ~ paste("high effort", ";", coalesce(flag, "")),
                          is.na(z_score) ~ coalesce(flag, ""),
                          TRUE ~ coalesce(flag, "")
         ),
         flag = na_if(flag, "")) %>% 
  select(total_effort_ident,
         sampling_method_simple,
         flag,
         total_effort_1,
         mean_effort,
         sd_effort,
         z_score) %>% 
  filter(!is.na(flag))

flag_xwalk_effort <- rbind(flag_xwalk, flag_xwalk_2) %>% 
  ungroup() %>% 
  select(total_effort_ident, 
         flag)
rm(flag_xwalk, flag_xwalk_2)

setnames(flag_xwalk_effort,old="flag" ,new="tei_flag" )

mn_dat[flag_xwalk_effort , on = .(total_effort_ident = total_effort_ident), tei_flag:= tei_flag ]
mn_dat[ , .N , .(tei_flag) ]
mn_dat[str_detect(tei_flag, " ; ") , .N ,tei_flag]
mn_dat[ , tei_flag := gsub(" ; ", "", tei_flag) , ]

mn_dat[, flag := do.call(paste, c(replace(.SD, is.na(.SD), ""), sep = "; ")), .SDcols= c("flag", "tei_flag")]

# Remove "; " from the beginning or end
mn_dat[!is.na(flag), flag := gsub("^\\s*;+|;+\\s*$", "", gsub(";\\s*;+", "; ", flag))]

mn_dat[ , .N , flag]

mn_dat %>% 
  filter(total_effort_ident %in% flag_xwalk_effort$total_effort_ident)
#flag gets applied to 869077 fish records
rm(flag_xwalk_effort)

#flag tables for fish
#summarize length based stats (and trim them to only species where we can do this work)
length_stats <- mn_dat[, .(stat = c(names(summary(length_1)), "sd"),
            value = c(summary(length_1), sd(length_1, na.rm = T)),
            units = "length_inches"), .(species_1) ]
length_stats[is.na(value) , .N , species_1 ]
length_stats <- length_stats[!species_1 %in% length_stats[is.na(value) , unique(species_1) ]]

#zero lengths
mn_dat[length_1 == 0 , len_flag := "length reported as zero"]

#when your length is >10sd from mean for species, get flagged as unlikely length
length_stats <- dcast(length_stats, formula = species_1 ~ ...)
length_stats[ , high_lim := Mean_length_inches + 10*sd_length_inches ,]
setcolorder(length_stats, neworder= c("species_1", "Mean_length_inches", "high_lim"))


mn_dat[length_stats , on = .(species_1, length_1 > high_lim) , len_flag := "length > 10sd from species mean in state" ]
mn_dat[ , .N , len_flag]

rm(length_stats)

#summarize weight based stats
weight_stats <- mn_dat[, .(stat = c(names(summary(weight_1)), "sd"),
            value = c(summary(weight_1), sd(weight_1, na.rm = T)),
            units = "weight_pounds"), .(species_1) ]
weight_stats[is.na(value) , .N , species_1 ]
weight_stats <- weight_stats[!species_1 %in% weight_stats[is.na(value) , unique(species_1) ]]

#zero lengths
mn_dat[weight_1 <= 0, .N]
mn_dat[weight_1 == 0 , weight_flag := "weight reported as zero"]

#when your wt is >10sd from mean for species, get flagged as unlikely weight
weight_stats <- dcast(weight_stats, formula = species_1 ~ ...)
weight_stats[ , high_lim := Mean_weight_pounds + 10*sd_weight_pounds ,]
setcolorder(weight_stats, neworder= c("species_1", "Mean_weight_pounds", "high_lim"))


mn_dat[weight_stats , on = .(species_1, weight_1 > high_lim) , weight_flag := "length > 10sd from species mean in state" ]
mn_dat[ , .N , weight_flag]

rm(weight_stats)

#summarize age based stats
age_stats <- mn_dat[, .(stat = c(names(summary(age)), "sd"),
            value = c(summary(age), sd(age, na.rm = T)),
            units = "age_years"), .(species_1) ]
age_stats[is.na(value) , .N , species_1 ]
age_stats <- age_stats[!species_1 %in% age_stats[is.na(value) , unique(species_1) ]]

#when your age is >10sd from mean for species, get flagged as unlikely age
age_stats <- dcast(age_stats, formula = species_1 ~ ...)
age_stats[ , high_lim := Mean_age_years + 10*sd_age_years ,]
setcolorder(age_stats, neworder= c("species_1", "Mean_age_years", "high_lim"))


mn_dat[age_stats , on = .(species_1, age > high_lim) , age_flag := "age > 10sd from species mean in state" ]
mn_dat[ , .N , age_flag]


mn_dat[, flag := do.call(paste, c(replace(.SD, is.na(.SD), ""), sep = "; ")), .SDcols= c("flag", "len_flag", "weight_flag", "age_flag")]
mn_dat[, .N , flag ]

# Remove "; " from the beginning or end & compress multiples
mn_dat[!is.na(flag), flag := gsub("^\\s*;+|;+\\s*$", "", gsub(";\\s*;+", "; ", flag))]


mn_dat[ , c("len_flag", "weight_flag", "age_flag") := NULL , ]



#check speceis richness seems reasonable
mn_dat[!is.na(total_effort_ident) , length(unique(species_1)) , total_effort_ident ][, hist(V1, main = "species richness per survey") , ]

#check fish counts:
mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , total_effort_ident ][, hist(V1, main = "N indiv fish sampled in survey") ,]

mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>100000 , ,]



# Remove "; " from the beginning or end and compress multiples
mn_dat[!is.na(flag), flag := trimws(gsub("^\\s*;+|;+\\s*$", "", gsub(";\\s*;+", "; ", flag)))]
mn_dat[flag == "", flag := NA]

mn_dat[!is.na(species_1) & !is.na(total_effort_ident) &is.na(flag) , length(unique(obs_id)) , total_effort_ident ][, hist(V1, main = "N indiv fish sampled in survey") ,]

mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>20000 , ,]

#review other high catch records
mn_dat[ total_effort_ident %in% 
                                        mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>20000 , total_effort_ident, ],
.N, .(total_effort_ident, lake_name, species_1, length_1)]

#flag them
mn_dat[ total_effort_ident %in% 
                                        mn_dat[!is.na(species_1) & !is.na(total_effort_ident) , length(unique(obs_id)) , .(total_effort_ident,lake_name,date_survey) ][V1>20000 , total_effort_ident, ],
                                      flag := ifelse(is.na(flag), "very high catch is suspected error" , paste(flag, "very high catch is suspected error", sep = ";" ))]




rm(age_stats, a, mw_data, chr_cols, col_name, del_col, i, new, new_colnames, num_cols, old, schema_names)


#units matched to hive and NAs where vcalues missing?
mn_dat[ , .N , .(length_unit_1, weight_unit_1) ]
#length
mn_dat[is.na(length_1) , .N , length_unit_1]
  mn_dat[is.na(length_1) ,length_unit_1 := NA]

mn_dat[ length_unit_1 == "mm" , length_unit_1 := "millimeters" ,]
#weight
mn_dat[is.na(weight_1) , .N , weight_unit_1]
  mn_dat[is.na(weight_1) ,weight_unit_1 := NA]

mn_dat[ weight_unit_1 == "g" , weight_unit_1 := "grams" ,]

#effort
mn_dat[is.na(total_effort_1) , .N , .(total_effort_1_units, total_effort_2_units) ]


#flags

mn_dat[ , .N, flag ]



```

#Large Lakes addition
Data gathered from the large lakes was not as abundant as we hoped. In communication with MN DNR staff, we collected a more inclusive large lake data file. Data gathered with gill net surveys from the 10 MN lakes. Each lake ranges in data availability. The process goes:
1. scope the effort data reported
2. scope the fish data reported
3. combine fish and effort data
4. aggregate historical pepin data (1965-2005) that were not included in LL agency update 
5. combine ll and historical pepin data
6. remove previous surveys from the statwide data base where they get an update here
7. add in new ll data 

Items to do:
1. ensure that leaving columns off won't screw the hive
2. ensure columns are the correct type 


#Large Lake overall agregation
```{r}
#effort data 
ll_effort <- fread("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/large_lakes_data_2025/02tbl_GN_Effort.csv") %>% 
  rename(SURVEY_ID = ID) %>% 
  mutate(original_file = "02tbl_GN_Effort.csv")

glimpse(ll_effort)

ll_effort %>% 
  summarise(n_distinct_ID = n_distinct(SURVEY_ID))
#survey ID appears to be a unique identifier for surveys in MN,
#however, those surveys appear to be at a net level - our sub effort

ll_effort %>% 
  group_by(NAME, Year) %>% 
  summarise(nets = n_distinct(STATION_CODE),
            n_ids = n_distinct(SURVEY_ID)) %>% 
  filter(nets != n_ids)
#there are only 12 times within a lake-year that there are not the same number of ids as nets 
#there are always more ids than nets for those cases


ll_effort %>% 
  group_by(SAMP_STA_TYPE_NAME, GEAR_TP) %>% 
  count()
#exploring the naming of sampling methods 

ll_effort %>% 
  group_by(Year, NAME, STATION_CODE) %>% 
  count() %>% 
  filter(n > 1) %>% 
  print(n = nrow(.))
#looks like year-name-station code is unique to each net, but there are sometimes multiple records for the same net within a year

ll_effort %>% 
  group_by(Year, NAME, STATION_CODE) %>% 
  summarise(n_dates = n_distinct(SURVEY_DATE),
            n_start_date = n_distinct(START_DATE),
            n_end_date = n_distinct(END_DATE),
            n_sample_station_number = n_distinct(SAMP_STATION_NBR),
            n = n()) %>% 
  filter(n >1) %>% 
  print(n = nrow(.))

#look at an example 
check <- ll_effort %>% 
  filter(NAME == "Pepin" & Year == "1998" & STATION_CODE == "GN1")
#this shows repetitive combos differ by survey date 

#lets check another example
check <- ll_effort %>% 
  filter(NAME == "Mille Lacs" & Year == "1995" & STATION_CODE == "GN11")
#the only thing that really varies between the two records is the bottom depth and old id - do we just trust the survey id code? Will have to explore the data more
rm(check)

##############################################################################
#reading in total effort reported by agency 
total_effort <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/large_lakes_data_2025/total_nets.csv") %>%
  mutate(Lake = case_when(Lake == "Red" ~ "Red (Upper Red)",
                          TRUE ~ Lake)) %>% 
  mutate(WB_NM = Lake) %>% 
  rename(ID_total_effort = ID)

#############################################################################
#fish data
ll_fish <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/large_lakes_data_2025/MN_LL_FISH.csv") %>% 
  rename(STATION_CODE = STATION) %>% 
  mutate(original_file = "MN_LL_FISH.csv")

glimpse(ll_fish)
#columns that maybe be useful in a survey Xwalk: STATION_CODE, GEARTP, WB_NM, SURVEY_ID, ID

ll_fish %>% 
  distinct(SURVEY_ID)

ll_fish %>% 
  distinct(SURVEY_ID, STATION_CODE)

#lets look at an examples
ll_effort %>% 
  filter(SURVEY_ID %in% c("1", "50", "200", "500", "5000")) %>% 
  select(SURVEY_ID, NAME, Year, STATION_CODE)

ll_fish %>% 
  filter(SURVEY_ID %in% c("1", "50", "200", "500", "5000")) %>% 
  distinct(SURVEY_ID, WB_NM, STATION_CODE)
#bomb, looks like we should be able to join on survey id

#############################################################################
#combined data
ll_fish_effort <- ll_fish %>% 
  left_join(ll_effort, by = c("SURVEY_ID"), suffix = c("_fish", "_effort"))

ll_fish_effort <- ll_fish_effort %>% 
  left_join(total_effort, by = c("WB_NM", "Year"))

ll_fish_effort %>% 
  group_by(Lake, Year, is.na(Number_Nets)) %>% 
  count() %>% 
  filter(`is.na(Number_Nets)` == "TRUE")


glimpse(ll_fish_effort)

ll_fish_effort %>% 
  filter(STATION_CODE_fish != STATION_CODE_effort)

ll_fish_effort %>% 
  distinct(SURVEY_ID, STATION_CODE_fish, STATION_CODE_effort)

ll_fish_effort %>% 
  distinct(SURVEY_ID)

#holy cow - it looks like the join actually worked - there are quite a few nothing caught that will be worth exploring

ll_fish_effort %>%
  group_by(NAME, Year, STATION_CODE_fish, SURVEY_ID) %>% 
  count() %>% 
  group_by(NAME, Year, STATION_CODE_fish) %>% 
  count() %>% 
  filter(n >1)
#this shows that NAME, Year, STATION_CODE applies to more than one station number some times

ll_fish_effort %>% 
  group_by(NAME, Year, STATION_CODE_effort) %>% 
  summarise(n_survey_ids = n_distinct(SURVEY_ID)) %>% 
  filter(n_survey_ids >1)

#now we need to find out what is happening to those left over surveys
ll_effort %>% 
  anti_join(ll_fish, by = c("SURVEY_ID", "STATION_CODE")) %>% 
  group_by(original_file) %>% 
  count()

#do the number of unique nets match the the total effort reported?
ll_fish_effort %>% 
  group_by(WB_NM, Year, Number_Nets) %>% 
  summarise(n_stations = n_distinct(STATION_CODE_fish)) %>% 
  filter(Number_Nets != n_stations) %>% 
  print(n = nrow(.))
#not excatly 
  
#from above we have 11443 unique survey ids from the effort file and 10756 from the fish file (diff of 687)
#that is the exact amount from the anti join - meaning all fish found an effort home

# nofish <- ll_effort %>% 
#   anti_join(ll_fish, by = c("SURVEY_ID", "STATION_CODE"))
#the code above uses the effort file to to fine nets that did not catch fish - the latest version uses the total nets instead 

# nofish %>% 
#   group_by(NAME, Year) %>% 
#   count() %>% 
#   arrange(NAME, desc(Year)) %>% 
#   print(n = nrow(.))

##creating flag for no fish surveys
# nofish <- nofish %>% 
#   mutate(flag = "no catch reported by agency, 0s could be misleading") %>% 
#   rename(original_file_effort = original_file,
#          STATION_CODE_effort = STATION_CODE)

#now row bind the nothing caught surveys

ll_fish_effort %>% 
  group_by(is.na(SP)) %>% 
  count()

ll_fish_effort %>% 
  filter(is.na(SP)) %>% 
  glimpse()
#two na fish species - that's not going to fly
#its hard to make any quick sense of what they could be - let's just remove them

ll_fish_effort <- ll_fish_effort %>% 
  filter(!is.na(SP))

#now add in no fish efforts and make nothing caught columns - after discussion of using overall total effort this is commented out
# ll_fish_effort <- ll_fish_effort %>% 
#   bind_rows(nofish)

check <- ll_fish %>% 
  mutate(FISH_COUNT = case_when(is.na(FISH_COUNT) ~ 1,
                                TRUE ~ FISH_COUNT)) %>% 
  group_by(FISH_COUNT) %>% 
  count() %>%
  mutate(total = FISH_COUNT * n,
         left_over = total - n) %>% 
  print(n = nrow(.))
sum(check$left_over)
rm(check)
#we should uncount 11492 fish 

ll_fish %>% 
  mutate(FISH_COUNT = case_when(is.na(FISH_COUNT) ~ 1,
                                TRUE ~ FISH_COUNT)) %>% 
  filter(FISH_COUNT != 1) %>% 
  filter(!is.na(WT_G)) %>% 
  group_by(SP, WT_G) %>% 
  count() %>% 
  print(n = nrow(.))
#sweet, doesn't look like we have any batched weight fish - all fish batch counted do not have a weight

########################## Let's Hive #####################################
ll_fish_effort <- ll_fish_effort %>% 
  mutate(FISH_COUNT = case_when(is.na(FISH_COUNT) ~ 1,
                                TRUE ~ FISH_COUNT)) %>% 
  uncount(FISH_COUNT)

#lets create a total effort ident - we currently have sub effort idents
ll_fish_effort <- ll_fish_effort%>% 
  # group_by(NAME, Year, STATION_CODE_effort) %>% 
  # mutate(sub_effort_ident = cur_group_id(),
  #        sub_effort_ident = paste0(sub_effort_ident, SURVEY_ID),
  #        sub_effort_1 = 1) %>% 
  group_by(NAME, Year) %>% 
  mutate(total_effort_ident = cur_group_id(),
         #total_effort_1 = sum(n_distinct(sub_effort_ident))
         total_effort_1 = Number_Nets
         ) %>% 
  ungroup()

#above I took the reported total effort from the DNR - lets compare this to the unique net ids within a total effort ident to see how they compare
ll_fish_effort %>% 
  group_by(total_effort_ident, WB_NM, Year, total_effort_1) %>% 
  summarise(n_nets = n_distinct(STATION_CODE_effort)) %>% 
  ungroup() %>% 
  summarise(n_agree = sum(total_effort_1 == n_nets),
            n_diff = sum(total_effort_1 != n_nets),
            total_surveys = n_agree + n_diff,
            prop_diff = n_diff/total_surveys)
#roughly 10% of total effort idents different between the number of sub effort nets and the total effort reported by the DNR - lets look at those cases 

ll_fish_effort %>% 
  group_by(total_effort_ident, WB_NM, Year, total_effort_1) %>% 
  summarise(n_nets = n_distinct(STATION_CODE_effort)) %>% 
  filter(total_effort_1 != n_nets) %>% 
  print(n = nrow(.))

#lets create a flag for these surveys
effort_flag <- ll_fish_effort %>% 
  group_by(total_effort_ident, WB_NM, Year, total_effort_1) %>% 
  summarise(n_nets = n_distinct(STATION_CODE_effort)) %>% 
  filter(total_effort_1 != n_nets) %>%
  mutate(flag = "unique # of nets does not match reported total effort, use caution in generating cpue") %>% 
  select(-n_nets)

ll_fish_effort <- ll_fish_effort %>% 
  left_join(effort_flag) 
rm(effort_flag)

ll_fish_effort %>% 
  group_by(total_effort_ident, flag) %>% 
  count() %>% 
  filter(!is.na(flag))
#we see here that 39 total effort idents pick up a flag - perfect


#fish species - this is a long one
ll_fish_effort <- ll_fish_effort %>% 
  mutate(species_1 = case_when(# first fish codes are converted,
                            SP == "LMB" ~ "largemouth_bass",
                            SP == "SMB" ~ "smallmouth_bass",
                            SP == "WAE" ~ "walleye", 
                            SP == "WTS" ~ "white_sucker",
                            SP == "BLC" ~ "black_crappie",
                            SP == "YEP" ~ "yellow_perch",
                            SP == "BLG" ~ "bluegill",
                            SP == "NOP" ~ "northern_pike",
                            SP == "PMK" ~ "pumpkinseed", 
                            SP == "RKB" ~ "rock_bass",
                            SP == "BKF" ~ "banded_killifish",
                            SP == "IOD" ~ "iowa_darter",
                            SP == "BCS" ~ "blackchin_shiner",
                            SP == "BLB" ~ "black_bullhead",
                            SP == "JND" ~ "johnny_darter",
                            SP == "GOS" ~ "golden_shiner",
                            SP == "YEB" ~ "yellow_bullhead",
                            SP == "BRB" ~ "brown_bullhead",
                            SP == "CNM" ~ "central_mudminnow",
                            SP == "NRD" ~ "northern_redbelly_dace",
                            SP == "CSH" ~ "chinook_salmon",
                            SP == "MTS" ~ "mottled_sculpin",
                            SP == "TPM" ~ "tadpole_madtom",
                            SP == "SHR" ~ "shorthead_redhorse",
                            SP == "HSF" ~ "hybrid_sunfish",
                            SP == "BOF" ~ "bowfin",
                            SP == "WDS" ~ "weed_shiner",
                            SP == "BNM" ~ "bluntnose_minnow",
                            SP == "BND" ~ "blacknose_dace",
                            SP == "OTM" ~ "minnows",
                            SP == "FHM" ~ "fathead_minnow",
                            SP == "FND" ~ "finescale_dace",
                            SP == "SDM" ~ "slender_madtom",
                            SP == "SPO" ~ "spottail_shiner",
                            SP == "BNS" ~ "blacknose_shiner",
                            SP == "BKT" ~ "brook_trout",
                            SP == "SPT" ~ "splake",
                            SP == "RBT" ~ "rainbow_trout",
                            SP == "BNT" ~ "brown_trout",
                            SP == "BUB" ~ "burbot",
                            SP == "BIB" ~ "bigmouth_buffalo",
                            SP == "BLH" ~ "bullheads",
                            SP == "TLC" ~ "cisco",
                            SP == "CCF" ~ "channel_catfish",
                            SP == "BSD" ~ "blackside_darter",
                            SP == "BKS" ~ "brook_silverside",
                            SP == "LED" ~ "least_darter",
                            SP == "LGP" ~ "logperch",
                            SP == "SLR" ~ "silver_redhorse",
                            SP == "HHC" ~ "hornyhead_chub",
                            SP == "MUE" ~ "muskellunge",
                            SP == "TRP" ~ "trout_perch",
                            SP == "RHS" ~ "redhorse",
                            SP == "CAP" ~ "common_carp",
                            SP == "MMS" ~ "mimic_shiner",
                            SP == "EMS" ~ "emerald_shiner",
                            SP == "SUN" ~ "sunfish",
                            SP == "CRC" ~ "creek_chub",
                            SP == "GSF" ~ "green_sunfish",
                            SP == "SFS" ~ "spotfin_shiner",
                            SP == "SHI" ~ "shiners",
                            SP == "BST" ~ "brook_stickleback",
                            SP == "CSR" ~ "central_stoneroller",
                            SP == "SDS" ~ "sand_shiner",
                            SP == "GRR" ~ "greater_redhorse",
                            SP == "PGS" ~ "pugnose_shiner",
                            SP == "SMS" ~ "slimy_sculpin",
                            SP == "LND" ~ "longnose_dace",
                            SP == "FRD" ~ "freshwater_drum",
                            SP == "WHC" ~ "white_crappie",
                            SP == "DAR" ~ "darters",
                            SP == "SAR" ~ "sauger",
                            SP == "FCF" ~ "flathead_catfish",
                            SP == "BMS" ~ "bigmouth_shiner",
                            SP == "LKS" ~ "lake_sturgeon",
                            SP == "LKW" ~ "lake_whitefish",
                            SP == "BRM" ~ "brassy_minnow",
                            SP == "SLM" ~ "mississippi_silvery_minnow",
                            SP == "GLR" ~ "golden_redhorse",
                            SP == "OTS" ~ "suckers",
                            SP == "NST" ~ "ninespine_stickleback",
                            SP == "GOE" ~ "goldeye",
                            SP == "QBS" ~ "quillback",
                            SP == "RVS" ~ "river_shiner",
                            SP == "LES" ~ "longear_sunfish",
                            SP == "CHL" ~ "chestnut_lamprey",
                            SP == "RRH" ~ "river_redhorse",
                            SP == "WHB" ~ "white_bass",
                            SP == "OSS" ~ "orangespotted_sunfish",
                            SP == "STC" ~ "stonecat",
                            SP == "SNG" ~ "shortnose_gar",
                            SP == "RFS" ~ "rosyface_shiner",
                            SP == "CMS" ~ "carmine_shiner",
                            SP == "SAB" ~ "smallmouth_buffalo",
                            SP == "GIS" ~ "gizzard_shad",
                            SP == "LNG" ~ "longnose_gar",
                            SP == "RCS" ~ "river_carpsucker",
                            SP == "SLS" ~ "shovelnose_sturgeon",
                            SP == "LAT" ~ "lake_trout",
                            SP == "TME" ~ "tiger_muskellunge",
                            SP == "PRD" ~ "pearl_dace",
                            SP == "PLS" ~ "pallid_shiner",
                            SP == "RBS" ~ "rainbow_smelt",
                            SP == "RBD" ~ "rainbow_darter",
                            SP == "LNS" ~ "longnose_sucker",
                            SP == "SJC" ~ "shortjaw_cisco",
                            SP == "DWS" ~ "deepwater_sculpin",
                            SP == "SHS" ~ "spoonhead_sculpin",
                            SP == "BDD" ~ "banded_darter",
                            SP == "MOE" ~ "mooneye",
                            SP == "BLS" ~ "blue_sucker",
                            SP == "SLC" ~ "silver_chub",
                            SP == "HFS" ~ "highfin_carpsucker",
                            SP == "MCP" ~ "mirror_carp",
                            SP == "WAS" ~ "walleye/sauger",
                            SP == "AME" ~ "american_eel",
                            SP == "GOF" ~ "goldfish",
                            SP == "SPS" ~ "spotted_sucker",
                            SP == "PAH" ~ "paddlefish",
                            SP == "RVD" ~ "river_darter",
                            SP == "NHS" ~ "northern_hog_sucker",
                            SP == "BHM" ~ "bullhead_minnow",
                            SP == "WSD" ~ "western_sand_darter",
                            SP == "MDD" ~ "mud_darter",
                            SP == "SIL" ~ "silver_lamprey",
                            SP == "SKJ" ~ "skipjack_herring",
                            SP == "PGM" ~ "pugnose_minnow",
                            SP == "BKB" ~ "black_buffalo",
                            SP == "FTD" ~ "fantail_darter",
                            SP == "YLB" ~ "yellow_bass",
                            SP == "PRP" ~ "pirate_perch",
                            SP == "LKC" ~ "lake_chub",
                            SP == "RIR" ~ "river_ruffe",
                            SP == "WHP" ~ "white_perch",
                            SP == "ALW" ~ "alewife",
                            SP == "SEL" ~ "sea_lamprey",
                            SP == "BHC" ~ "bighead_carp",
                            SP == "GLD" ~ "gilt_darter",
                            SP == "CRD" ~ "crystal_darter",
                            SP == "SMT" ~ "speckled_madtom",
                            SP == "RDS" ~ "red_shiner",
                            SP == "WAM" ~ "warmouth",
                            SP == "CIS" ~ "cisco",
                            SP == "CPS" ~ "carpsucker",
                            SP == "CRP" ~ "crappie",
                            SP == "HCR" ~ "hybrid_crappie",
                            SP == "SCU" ~ "sculpin",
                            SP == "SHD" ~ "slenderhead_darter",
                            SP == "SIP" ~ "northern_pike_silver_phase",
                            SP == "UK1" ~ "unknown_fish",
                            SP == "UK2" ~ "unknown_fish",
                            # below unknown codes specified as such
                            SP == "LXB" ~ "see_ambiguous_species_code",
                            TRUE ~ SP
                          ))

ll_fish_effort %>% 
  group_by(STD_SAMPLING_FLAG, CPUE) %>% 
  count()

#note columns
#gear data notes - BOTTOM_DEPTH_1ST_END_FEET, BOTTOM_DEPTH_2nd_END_FEET, MESH_SIZE_1st_end_inches, Start_Time, End_Time
#location notes - there could be some location notes in the crew notes/comments.. is it worth moving
#notes_1 - MESH, CREW_NOTES, COMMENTS
ll_fish_effort %>% 
  mutate(gear_data_notes = paste0("start_time: ", Start_Time, "; ",
                                  "end_time: ", End_Time, "; ",
                                  "depth_end_1_ft: ", BOTTOM_DEPTH_1ST_END_FEET, "; ",
                                   "depth_end_2_ft: ", BOTTOM_DEPTH_2ND_END_FEET, "; ",
                                  "mesh_end_1_in: ", MESH_SIZE_1ST_END_INCHES, "; "),
         notes_1 = paste0("mesh_fish_capture: ", MESH, "; ",
                          "crew_notes: ", CREW_NOTES, "; ",
                          "comments: ", COMMENTS, "; ")) %>% 
  distinct(gear_data_notes, notes_1) %>% 
  glimpse()

ll_fish_effort <- ll_fish_effort %>% 
  mutate(
    gear_data_notes = paste0(
      ifelse(!is.na(Start_Time), paste0("start_time: ", Start_Time, "; "), ""),
      ifelse(!is.na(End_Time), paste0("end_time: ", End_Time, "; "), ""),
      ifelse(!is.na(BOTTOM_DEPTH_1ST_END_FEET), paste0("depth_end_1_ft: ", BOTTOM_DEPTH_1ST_END_FEET, "; "), ""),
      ifelse(!is.na(BOTTOM_DEPTH_2ND_END_FEET), paste0("depth_end_2_ft: ", BOTTOM_DEPTH_2ND_END_FEET, "; "), ""),
      ifelse(!is.na(MESH_SIZE_1ST_END_INCHES), paste0("mesh_end_1_in: ", MESH_SIZE_1ST_END_INCHES, "; "), "")
    ),
    notes_1 = paste0(
      ifelse(!is.na(MESH), paste0("mesh_fish_capture: ", MESH, "; "), ""),
      ifelse(!is.na(CREW_NOTES), paste0("crew_notes: ", CREW_NOTES, "; "), ""),
      ifelse(!is.na(COMMENTS), paste0("comments: ", COMMENTS, "; "), "")
    )
  ) %>% 
  mutate(
    gear_data_notes = str_remove(gear_data_notes, "; $"),  # Remove trailing semicolon
    notes_1 = str_remove(notes_1, "; $"),  # Remove trailing semicolon
    gear_data_notes = na_if(gear_data_notes, ""),  # Replace empty string with NA
    notes_1 = na_if(notes_1, "")  # Replace empty string with NA
  )

ll_fish_effort %>% 
  group_by(species_1) %>% 
  count() %>% 
  print(n = nrow(.))

ll_fish_effort %>% 
  group_by(flag) %>% 
  count()

#auto flags for length and weight
length_weight_flag_1 <- ll_fish_effort %>% 
  #filter out surveys that have no catch
  filter(!is.na(species_1)) %>% 
  #group by species to create z score for weight and length
  #flag created for fish more or less than 10 sd 
  #add this flag to the flag column with coalesce
  group_by(species_1) %>% 
  mutate(mean_length = mean(LEN_MM, na.rm = TRUE),
         mean_weight = mean(WT_G, na.rm = TRUE),
         sd_length = sd(LEN_MM, na.rm = TRUE),
         sd_weight = sd(WT_G, na.rm = TRUE),
         z_score_length = (LEN_MM - mean_length) / sd_length,
         z_score_weight = (WT_G - mean_weight) / sd_weight,
         flag = case_when(z_score_length >= 10 | z_score_length <= -10 ~ paste0("unlikely length", "; ", coalesce(flag, "")),
                          z_score_weight >= 10 | z_score_weight <= -10 ~ paste0("unlikely weight", "; ", coalesce(flag, "")),
                          TRUE ~ coalesce(flag, "")),
         flag = case_when(LEN_MM == 0 | WT_G == 0 ~ paste("length or weight reported as 0", "; ", coalesce(flag, "")),
                          TRUE ~ coalesce(flag, "")),
         flag = na_if(flag, "")) %>% 
  ungroup()

length_weight_flag_2 <- length_weight_flag_1 %>% 
  filter((is.na(z_score_length) | (z_score_length > -10 & z_score_length < 10)) & 
           (is.na(z_score_weight) | (z_score_weight > -10 & z_score_weight < 10))) %>% 
  group_by(species_1) %>% 
  mutate(mean_length = mean(LEN_MM, na.rm = TRUE),
         mean_weight = mean(WT_G, na.rm = TRUE),
         sd_length = sd(LEN_MM, na.rm = TRUE),
         sd_weight = sd(WT_G, na.rm = TRUE),
         z_score_length = (LEN_MM - mean_length) / sd_length,
         z_score_weight = (WT_G - mean_weight) / sd_weight,
         flag = case_when(z_score_length >= 10 | z_score_length <= -10 ~ paste0("unlikely length", "; ", coalesce(flag, "")),
                          z_score_weight >= 10 | z_score_weight <= -10 ~ paste0("unlikely weight", "; ", coalesce(flag, "")),
                          TRUE ~ coalesce(flag, "")),
         flag = na_if(flag, "")) %>% 
  ungroup()

# no_effort_flags <- ll_fish_effort %>% 
#   filter(is.na(species_1))

combined_flags <- length_weight_flag_1 %>% 
   filter(!((is.na(z_score_length) | (z_score_length > -10 & z_score_length < 10)) & 
           (is.na(z_score_weight) | (z_score_weight > -10 & z_score_weight < 10)))) %>% 
  bind_rows(length_weight_flag_2) %>% 
  #bind_rows(no_effort_flags) %>% 
  select(total_effort_ident,
         #sub_effort_ident,
         species_1,
         LEN_MM,
         WT_G,
         flag) %>% 
  distinct()

combined_flags %>% 
  group_by(flag) %>% 
  count()

length_weight_flag_1 %>% 
  group_by(flag) %>% 
  count()

length_weight_flag_2 %>% 
  group_by(flag) %>% 
  count()

length_weight_flag_1 %>% 
  filter(str_detect(flag, "unlikely")) %>% 
  select(flag)

length_weight_flag_2 %>% 
  filter(str_detect(flag, "unlikely")) %>% 
  select(flag)

combined_flags %>% 
  filter(str_detect(flag, "unlikely")) %>% 
  select(flag)

combined_flags %>% 
  filter(str_detect(flag, "#")) %>% 
  group_by(total_effort_ident) %>% 
  count()

rm(length_weight_flag_1, length_weight_flag_2)

#we can left join these flags to the data 
ll_fish_effort <- ll_fish_effort %>% 
  ungroup() %>% 
  select(-flag) %>% 
  left_join(combined_flags, relationship = "many-to-one") 

ll_fish_effort %>% 
  group_by(flag) %>% 
  count()

rm(combined_flags)

ll_fish_effort %>% 
  filter(str_detect(flag, "unlikely") | str_detect(flag, "reported as")) %>% 
  select(flag)

ll_fish_effort %>% 
  filter(str_detect(flag, "#")) %>% 
  select(flag)

ll_fish_effort %>% 
  group_by(is.na(flag)) %>% 
  count()

#missing spacial data
ll_fish_effort <- ll_fish_effort %>% 
  group_by(NAME) %>% 
  mutate(lake_id = DOW_KITTLE) %>% 
  fill(lake_id, .direction = "updown") %>% 
  ungroup()

#nothing caught columns
ll_fish_effort <- ll_fish_effort %>% 
  group_by(total_effort_ident) %>% 
  mutate(total_effort_nothing_caught = is.na(species_1)) %>% 
  ungroup()

# ll_fish_effort %>% 
#   group_by(total_effort_ident) %>% 
#   summarise(
#     num_sub_efforts = n_distinct(sub_effort_ident),
#     num_nothing_caught = sum(sub_effort_nothing_caught == T),
#     total_flag = first(total_effort_nothing_caught)
#   ) %>% 
#   filter(num_nothing_caught != 0) %>% 
#   print(n = nrow(.))
#we can see that when all sub efforts are na within a total effort than total effort nothing caught is T, otherwise it is false 

#as a quick check, our 783323 total fish records make a difference of 11490 from the original file - this is perfect given our uncount and removal of two records with NA species. 

# age and aging structure
aging_check <- ll_fish_effort %>% 
  distinct(NAME, Year, STATION_CODE_effort, LEN_MM, WT_G, species_1, OFF_AGE, AGE_SRC, CALCIFIED_STRUC_TYPE_NAME, SERIAL) %>% 
  arrange(STATION_CODE_effort, LEN_MM, WT_G, OFF_AGE, SERIAL)
#when you arrange by net, length, weight, species, serial - it is pretty clear the serial number is for identifying individual fish
#it looks like we have a ton of obs where there is more than one observation of a fish within a sub effort ident (which includes length, weight, and age). This leads me to think there are multiple fish with different aging types 

check <- aging_check %>% 
  filter(STATION_CODE_effort == "GN1" & NAME == "Vermilion" & Year == 2018 & species_1 == "walleye")
rm(check)
#in this example of a net, SERIAL looks like a unique fish identifier for when a fish is aged - but is that always the case?

# what about for cases when the serial is na, null, or lengths vary within a serial?

# filters for where there are multiple length values within a net, serial 
check <- aging_check %>% 
  filter(SERIAL != "null" & !is.na(SERIAL)) %>% 
  group_by(NAME, Year, STATION_CODE_effort, species_1, SERIAL) %>% 
  filter(n_distinct(LEN_MM) > 1) %>% 
  ungroup() %>% 
  arrange(NAME, Year, STATION_CODE_effort, species_1, SERIAL, LEN_MM)
# most of these samples come from cass where it looks like the serial numbers are off by one fish - for example serial 2015 has an excat matching weight of a fish for serial 2016
#the other cases either look like they could be within measurement error (the fish was measured twice with some variance) or who knows...

#what about cases where the serial is absent? can we just assume 1.) those fish were not aged or 2.) they were aged but not duplicates 
check <- ll_fish_effort %>% 
  mutate(check= case_when(is.na(SERIAL) | SERIAL == "null" ~ "absent",
                          TRUE ~ "present")) %>% 
  filter(check == "absent" & !is.na(OFF_AGE)) %>% 
  select(NAME, Year, STATION_CODE_effort, SERIAL, species_1, LEN_MM, WT_G, AGE_SRC, OFF_AGE) %>% 
  group_by(NAME, Year, STATION_CODE_effort, species_1, LEN_MM, WT_G) %>% 
  count() %>% 
  filter(n > 1)

check<- ll_fish_effort %>% 
  mutate(check= case_when(is.na(SERIAL) | SERIAL == "null" ~ "absent",
                          TRUE ~ "present")) %>% 
  filter(check == "absent" & !is.na(OFF_AGE)) %>% 
  filter(!is.na(AGE_SRC)) %>% 
  select(STATION_CODE_effort, NAME, Year, SERIAL, species_1, LEN_MM, WT_G, AGE_SRC, OFF_AGE) %>% 
  arrange(STATION_CODE_effort, species_1, LEN_MM, WT_G) 
rm(check, aging_check)
#this shows a ton of records where a serial is absent and there appears to be a fish aged in different ways
#long story short, I'm going to have to make a decision here

# I am going to move forward assuming the serial column identifies unique fish, any fish without an id is unique. Any duplicate weight/leng/spec combo for a net is a complete separate fish. If a fish were aged without a serial number we are going to assume that is also a separate fish 

#what do our structures look like?
ll_fish_effort %>% 
  mutate(aging_structure_1 = case_when(CALCIFIED_STRUC_TYPE_NAME == "N/A" ~ NA,
                                       CALCIFIED_STRUC_TYPE_NAME == "null" ~ NA,
                                       TRUE ~ CALCIFIED_STRUC_TYPE_NAME),
         aging_structure_1 = tolower(aging_structure_1),
         aging_structure_1 = str_replace(aging_structure_1, " ", "_")) %>% 
  group_by(CALCIFIED_STRUC_TYPE_NAME, aging_structure_1) %>% 
  count()

aging_rank <- c("otolith" = 1, "cleithrum" = 2, "opercle" = 3, 
                "pectoral_spine" = 4, "dorsal_spine" = 5, "fin_ray" = 6, "scale" = 7)



check <- ll_fish_effort %>% 
  mutate(fish_id = case_when(is.na(SERIAL) | SERIAL == "null" ~ NA,
                             TRUE ~ SERIAL),
         fish_id = case_when(is.na(fish_id) ~ as.character(row_number()),
                             TRUE ~ fish_id)) %>% 
  group_by(NAME, Year, STATION_CODE_effort, fish_id) %>% 
  count() %>% 
  group_by(n) %>% 
  count() %>%  
  mutate(total_obs = n * nn,
         remaining_fish = nn) 
#it is valid that a fish might be aged 3+ times... or 17 times?
sum(check$remaining_fish)
#after the slice we should have 690277 fish left

#here we see an example of a serial number that repeats 17 times 
#was it really aged 17 times, mostly all by a scale? we will just have to roll with it
ll_fish_effort %>% 
  mutate(fish_id = case_when(is.na(SERIAL) | SERIAL == "null" ~ NA,
                             TRUE ~ SERIAL),
         fish_id = case_when(is.na(fish_id) ~ as.character(row_number()),
                             TRUE ~ fish_id)) %>% 
  group_by(NAME, Year, STATION_CODE_effort, fish_id, LEN_MM, WT_G) %>% 
  count() %>% 
  filter(n == 17)
#every fish aged 17 times comes from Mille Lacs in 2019, lets look at just one example

ll_fish_effort %>% 
  mutate(fish_id = case_when(is.na(SERIAL) | SERIAL == "null" ~ NA,
                             TRUE ~ SERIAL),
         fish_id = case_when(is.na(fish_id) ~ as.character(row_number()),
                             TRUE ~ fish_id)) %>% 
  filter(fish_id == "10030" & STATION_CODE_effort == "GN18" & Year == "2019") %>% 
  select(NAME, Year, STATION_CODE_effort, fish_id, LEN_MM, WT_G, SX, OFF_AGE, AGE_SRC, CALCIFIED_STRUC_TYPE_NAME) %>% 
  print(n = nrow(.))

  

#clean up serial, structure, and slice for only one unique fish obs
ll_fish_effort <- ll_fish_effort %>% 
  #make fish id column NA where it is null, make sure there is a unique id for those that are NA
  mutate(fish_id = case_when(is.na(SERIAL) | SERIAL == "null" ~ NA,
                             TRUE ~ SERIAL),
         fish_id = case_when(is.na(fish_id) ~ as.character(row_number()),
                             TRUE ~ fish_id)) %>% 
  #clean up aging structure 
  mutate(aging_structure_1 = case_when(CALCIFIED_STRUC_TYPE_NAME == "N/A" ~ NA,
                                       CALCIFIED_STRUC_TYPE_NAME == "null" ~ NA,
                                       TRUE ~ CALCIFIED_STRUC_TYPE_NAME),
         aging_structure_1 = tolower(aging_structure_1),
         aging_structure_1 = str_replace(aging_structure_1, " ", "_")) %>% 
  #bring in aging rank
  mutate(aging_rank = aging_rank[aging_structure_1]) %>% 
  #group by sub effort - the net the fish was caught in - and the serial number
  #order by the aging structure rank
  #this will take the best aging structure for serial numbers that repeat in a net
  group_by(NAME, Year, STATION_CODE_effort, fish_id) %>% 
  arrange(aging_rank, .by_group = TRUE) %>% 
  slice(1) %>% 
  ungroup()
#the obs in here match the above calc for remaining fish, lets see how this handled a fish aged 17 times

ll_fish_effort %>% 
  mutate(fish_id = case_when(is.na(SERIAL) | SERIAL == "null" ~ NA,
                             TRUE ~ SERIAL),
         fish_id = case_when(is.na(fish_id) ~ as.character(row_number()),
                             TRUE ~ fish_id)) %>% 
  filter(fish_id == "10030" & STATION_CODE_effort == "GN18" & Year == "2019") %>% 
  select(NAME, Year, STATION_CODE_effort, fish_id, LEN_MM, WT_G, SX, OFF_AGE, AGE_SRC, CALCIFIED_STRUC_TYPE_NAME) %>% 
  print(n = nrow(.))
#perfect we can see it just grabbed one record and it was the highest ranking structure 

ll_fish_effort %>% 
  group_by(aging_structure_1) %>% 
  count() %>% 
  arrange(desc(n))
#we can see that most fish are not aged, but the ranking looks correct here in cases where they are

#Kristin tipped us off to a few data anomalies that we should explore
#1. 1986-1989 in Mille Lacs only has WAE, NOP, MUE
ll_fish_effort %>% 
  filter(NAME == "Mille Lacs" & Year %in% c("1986", "1987", "1988", "1989")) %>% 
  group_by(Year, species_1) %>% 
  count()
#that is correct - lets make a flag for this 

#2. bulk weighted fish 1983-2001 for yep and tlc
ll_fish_effort %>% 
  filter(species_1 %in% c("yellow_perch", "cisco") & !is.na(LEN_MM) & !is.na(WT_G)) %>% 
  ggplot() +
  geom_point(aes(LEN_MM, WT_G, color = species_1)) +
  geom_smooth(aes(LEN_MM, WT_G, color = species_1)) +
  facet_wrap(~NAME, scales = "free")
#most of these look pretty good - there are certainly a few outliers - large weights for a length

ll_fish_effort %>%
  filter(species_1 %in% c("yellow_perch", "cisco") & !is.na(LEN_MM) & !is.na(WT_G) & !(LEN_MM == 0 | WT_G == 0)) %>% 
  mutate(lw_ratio = WT_G/LEN_MM) %>% 
  ggplot() +
  geom_histogram(aes(lw_ratio, fill = species_1)) +
  facet_wrap(~NAME, scale = "free")

check <- ll_fish_effort %>% 
 filter(species_1 %in% c("yellow_perch", "cisco") & !is.na(LEN_MM) & !is.na(WT_G) & !(LEN_MM == 0 | WT_G == 0)) %>% 
  mutate(lw_ratio = WT_G/LEN_MM) %>%
  filter(lw_ratio > 3) %>%
  arrange(NAME, Year, STATION_CODE_effort ,species_1, LEN_MM, WT_G) %>% 
  select(NAME, Year, STATION_CODE_effort ,species_1, LEN_MM, WT_G, lw_ratio, flag)
#there are some clear outliers, the most egregious seem to already pick up a flag - lets just move on
rm(check)

additional_flags <- ll_fish_effort %>% 
  mutate(flag = case_when(WB_NM == "Mille Lacs" & 
                     Year %in% c("1986", "1987", "1988", "1989") ~ paste0("only walleye, pike and muskie collected - all other species should not be considered", "; ", coalesce(flag, "")),
                     TRUE ~ coalesce(flag, "")),
         flag = na_if(flag, "")) %>% 
  filter(!is.na(flag))

additional_flags %>% 
  group_by(flag) %>% 
  count()

ll_fish_effort <- ll_fish_effort %>% 
  ungroup() %>% 
  select(-flag) %>% 
  left_join(additional_flags, relationship = "many-to-one") 

ll_fish_effort %>% 
  group_by(flag) %>% 
  count()

rm(additional_flags)


#making sure we have all the hived columns
ll_fish_effort <- ll_fish_effort %>% 
  mutate(state = "Minnesota",
         lake_name = NAME) %>% 
  #the date headache 
  #survey date - when reported just use that, other wise take the first of start date or end date
  #date total effort ident - first record of date in start date and then the first record of end date if start date NA
  #date sub effort idnet - start date if present or then end date
  #date sample - end date when present or NA
  mutate(SURVEY_DATE = mdy(SURVEY_DATE),
         START_DATE = mdy(START_DATE),
         END_DATE = mdy(END_DATE)) %>% 
  mutate(lake_id = as.character(lake_id)) %>% 
  group_by(total_effort_ident) %>% 
  mutate(date_survey = case_when(!is.na(SURVEY_DATE) ~ SURVEY_DATE,
                                 is.na(SURVEY_DATE) & !is.na(START_DATE) ~ first(START_DATE),
                                 is.na(SURVEY_DATE) & is.na(START_DATE) & !is.na(END_DATE) ~ first(END_DATE),
                                 TRUE ~ NA),
         date_total_effort_ident = case_when(!is.na(START_DATE) ~ first(START_DATE),
                                             is.na(START_DATE) & !is.na(END_DATE) ~ (END_DATE),
                                             TRUE ~ NA),
         # date_sub_effort_ident = case_when(!is.na(START_DATE) ~ START_DATE,
         #                                   is.na(START_DATE) & !is.na(END_DATE) ~ END_DATE,
         #                                   TRUE ~ NA),
         # date_sample = case_when(!is.na(END_DATE) ~ END_DATE,
         #                         TRUE ~ NA),
         year = case_when(!is.na(date_survey) ~ year(date_survey),
                          TRUE ~ Year),
         month = case_when(!is.na(date_total_effort_ident) ~ month(date_total_effort_ident),
                           TRUE ~ NA)) %>% 
  ungroup() %>% 
  #reporting agency given survey id - this is really the sub effort ident those so many its not needed
  mutate(survey_id = SURVEY_ID,
         survey_type = SRV_TYPE) %>% 
  #first make a clean sampling method - these are all gill nets
  mutate(sampling_method_simple = "gill_net",
         sampling_method = "gill_net_standard",
         sampling_method_2 = "gill_net_standard_graduated_mesh") %>% 
  #total effort 1 units
  mutate(total_effort_1_units = "net_nights") %>% 
  #water temp
  mutate(water_temp = Temperature,
         water_temp_units = "celcius") %>% 
  #site id - making it the net number 
  mutate(site_id = STATION_CODE_fish) %>% 
  #sub effort units
  #mutate(sub_effort_1_units = "net_nets") %>% 
  #length
  mutate(length_1 = LEN_MM,
         length_unit_1 = "mm") %>% 
  #weight
  mutate(weight_1 = WT_G,
         weight_unit_1 = "g") %>% 
  #sex
  mutate(sex = case_when(SX == "N/A" ~ NA,
                         SX == "Not Examined" ~ NA,
                         SX == "Unknown" ~ NA,
                         SX == "male" ~ "Male",
                         TRUE ~ SX)) %>%
  #indv fish idents - reported by agency
  mutate(ind_fish_ident = SERIAL,
         ind_fish_ident = case_when(ind_fish_ident == "null" ~ NA,
                                    TRUE ~ ind_fish_ident)) %>% 
  #original file
  mutate(original_file_names = paste0(original_file_effort, "; ", original_file_fish)) %>% 
  rename(age = OFF_AGE) %>% 
  #making sure everything is in the correct type
  mutate(survey_id = as.character(survey_id),
         total_effort_ident = as.character(total_effort_ident)) %>% 
  select(state, 
         #county, 
         lake_name,
         lake_id,
         #nhdhr_id,
         #latitude_lake_centroid,
         #longitude_lake_centroid,
         date_survey,
         date_total_effort_ident,
         #date_sub_effort_ident,
         #date_sample,
         year,
         month,
         survey_id,
         survey_type,
         sampling_method_simple,
         sampling_method,
         sampling_method_2,
         gear_data_notes,
         #target_species,
         total_effort_ident,
         total_effort_1, 
         #total_effort_2,
         total_effort_1_units,
         #total_effort_2_units,
         total_effort_nothing_caught,
         water_temp,
         water_temp_units,
         #water_clarity,
         #water_clarity_units,
         #lat_start,
         #lon_start,
         #lat_end,
         #lon_end,
         site_id,
         #sub_effort_ident,
         #sub_effort_1,
         #sub_effort_1_units,
         #sub_effort_2,
         #sub_effort_2_units,
         #sub_effort_nothing_caught,
         species_1,
         length_1,
         length_unit_1,
         #length_bin,
         #length_bin_unit,
         age,
         aging_structure_1,
         weight_1,
         weight_unit_1,
         #batch_weight,
         #batch_weight_unit,
         sex,
         #age_class,
         flag,
         original_file_names,
         ind_fish_ident,
         #lakesize,
         #lakesize_units,
         #lat_unspec,
         #lon_unspec,
         #waterbody_type,
         #location_notes_1,
         notes_1
         )

glimpse(ll_fish_effort)
```

# Historical Pepin aggregation 
```{r}
#read in Pepin effort and fish data
pep_fish <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/large_lakes_data_2025/pepin_fish_65_05.csv") %>% 
  mutate(original_file_name = "pepin_fish_65_05.csv")

total_effort <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/large_lakes_data_2025/total_nets.csv") %>% 
  mutate(Lake = case_when(Lake == "pepin" ~ "Pepin",
                          TRUE ~ Lake)) %>% 
  filter(Lake == "Pepin" & 
           (Year >=1965 & Year <= 2005)) %>% 
  mutate(original_file_name = "total_nets.csv")


################ effort data ###############################
glimpse(total_effort)

total_effort <- total_effort %>% 
  rename(year = Year) %>% 
  select(-ID)

####################fish data###############################
glimpse(pep_fish)

pep_fish <- pep_fish %>% 
  #clean species
  mutate(species_1 = tolower(Species),
         species_1 = str_replace(species_1, " ", "_")) %>% 
  #clean length and weight 
  mutate(length_1 = `LEN (mm)`,
         length_unit_1 = "mm",
         weight_1 = `WEIGHT (g)`,
         weight_unit_1 = "g") %>% 
  #clean date
  mutate(date = mdy(Date)) %>% 
  #rename some columns 
  rename(sex = Sex,
         station = STATION,
         fish_count = `FISH COUNT`) %>% 
  #dates
  mutate(year = year(date),
         month = month(date)) %>% 
  select(`Mesh Size (in)`,
         species_1,
         length_1,
         length_unit_1,
         weight_1,
         weight_unit_1,
         sex,
         Maturity,
         date,
         year,
         month,
         station,
         original_file_name,
         fish_count)


pep_fish %>% 
  filter(year== 1999 & species_1 ==  "freshwater_drum" & weight_1 >2000) %>% 
  select(date, species_1, length_1, weight_1, fish_count)

pep_fish %>% 
  filter(fish_count > 1 & !is.na(weight_1)) %>% 
  select(date, species_1, length_1, weight_1, fish_count) %>% 
  print(n = nrow(.))
#it looks like fish that were counted have a batch weight 

pep_fish %>% 
  filter(!is.na(length_1) & !is.na(weight_1) & fish_count > 1) %>% 
  glimpse()
#I'm moving forward assuming that all fish with a weight but not length and count grater an 1 are batch weighted
#fish that a length and weight but have multiple fish I will note as individual fish 


#this code below creates a grouping id for fish that were batched weighted together
#I then add those weights to a batch weight column, making the weight column na
#all other fish will be treated as individual fish measures 
pep_fish <- pep_fish %>% 
  group_by(year, species_1, fish_count) %>% 
  mutate(batch_weight_id = cur_group_id(),
         batch_weight_id = case_when(is.na(length_1) & !is.na(weight_1) & fish_count > 1 ~ batch_weight_id,
                                     TRUE ~ NA)) %>%
  mutate(batch_weight = case_when(is.na(length_1) & !is.na(weight_1) & fish_count > 1 ~ weight_1,
                                  TRUE ~ NA),
         batch_weight_unit = "g",
         weight_1 = case_when(!is.na(batch_weight_id) ~ NA,
                              TRUE ~ weight_1)) %>% 
  ungroup()

#check it
pep_fish %>% 
  filter((!is.na(length_1) & is.na(batch_weight) & fish_count >1) |
           (is.na(length_1) & !is.na(batch_weight) & fish_count > 1)) %>% 
  select(batch_weight_id, date, species_1, length_1, weight_1, fish_count, batch_weight) %>% 
  print(n = nrow(.))
#this shows that in cases where a weight was reported without length and the fish count was greater than 1, the weight was moved to a batch weight 
#call other cases retain individual fish level 


################ combining year with effort ############################
pep_fish_effort <- pep_fish %>% 
  left_join(total_effort, by = "year", suffix = c("_fish", "_effort"))

pep_fish_effort %>% 
  filter(is.na(Number_Nets)) %>% 
  group_by(year) %>% 
  count()
#all fish records have effort - sweet

#now lets see if the number of unique stations in a year is the same as the total effort
pep_fish_effort %>% 
  group_by(year, Number_Nets) %>% 
  summarise(n_stations = n_distinct(station)) %>% 
  distinct(year, .keep_all = T) %>% 
  filter(Number_Nets != n_stations)
#we have one case where there are more nets than total effort reported - 1991
#we have two cases where the total effort is more than the number of stations - a potential nothing caught - 1993 and 1996
#given we did not get a sub-effort level effort, i am just going to roll with the total effort given by the number of nets 

glimpse(pep_fish_effort)

pep_fish_effort <- pep_fish_effort %>% 
  uncount(fish_count)

pep_fish_effort %>% 
  group_by(is.na(weight_1), is.na(batch_weight), is.na(batch_weight_id)) %>% 
  count()
#this shows that all batch weight fish have been uncounted and do not have a weight_1. All fish that have a weight but are individual fish do not have a batch weight 

#total_effort_ident
pep_fish_effort <- pep_fish_effort %>% 
  group_by(year) %>% 
  mutate(total_effort_ident = paste0(cur_group_id(), "_pep"))

pep_fish_effort %>% 
  group_by(total_effort_ident) %>% 
  summarise(n_distinct(date)) %>% 
  filter(`n_distinct(date)` >1)

#auto flags for length and weight
length_weight_flag_1 <- pep_fish_effort %>% 
  #filter out surveys that have no catch
  filter(!is.na(species_1)) %>% 
  #group by species to create z score for weight and length
  #flag created for fish more or less than 10 sd 
  #add this flag to the flag column with coalesce
  group_by(species_1) %>% 
  mutate(mean_length = mean(length_1, na.rm = TRUE),
         mean_weight = mean(weight_1, na.rm = TRUE),
         sd_length = sd(length_1, na.rm = TRUE),
         sd_weight = sd(weight_1, na.rm = TRUE),
         z_score_length = (length_1 - mean_length) / sd_length,
         z_score_weight = (weight_1 - mean_weight) / sd_weight,
         #we have not created a flag for this the pepin data yet, so here we cannot refer a previously made column like we do in other ag scripts - removed the coalesce fuction in the case when for length and w
         flag = case_when(z_score_length >= 10 | z_score_length <= -10 ~ paste0("unlikely length", ";"),
                          TRUE ~ NA),
         flag = case_when(z_score_weight >= 10 | z_score_weight <= -10 ~ paste0("unlikely weight", "; ", coalesce(flag, "")),
                          TRUE ~ coalesce(flag, "")),
         flag = case_when(length_1 == 0 | weight_1 == 0 ~ paste("length or weight reported as 0", "; ", coalesce(flag, "")),
                          TRUE ~ coalesce(flag, "")),
         flag = na_if(flag, "")) %>% 
  ungroup()

length_weight_flag_1 %>% 
  group_by(flag) %>% 
  count()

length_weight_flag_2 <- length_weight_flag_1 %>% 
  filter((is.na(z_score_length) | (z_score_length > -10 & z_score_length < 10)) & 
           (is.na(z_score_weight) | (z_score_weight > -10 & z_score_weight < 10))) %>% 
  group_by(species_1) %>% 
  mutate(mean_length = mean(length_1, na.rm = TRUE),
         mean_weight = mean(weight_1, na.rm = TRUE),
         sd_length = sd(length_1, na.rm = TRUE),
         sd_weight = sd(weight_1, na.rm = TRUE),
         z_score_length = (length_1 - mean_length) / sd_length,
         z_score_weight = (weight_1 - mean_weight) / sd_weight,
         flag = case_when(z_score_length >= 10 | z_score_length <= -10 ~ paste0("unlikely length", "; ", coalesce(flag, "")),
                          z_score_weight >= 10 | z_score_weight <= -10 ~ paste0("unlikely weight", "; ", coalesce(flag, "")),
                          TRUE ~ coalesce(flag, "")),
         flag = na_if(flag, "")) %>% 
  ungroup()

length_weight_flag_2 %>% 
  group_by(flag) %>% 
  count()

no_effort_flags <- pep_fish_effort %>% 
  filter(is.na(species_1))
#no no effort surveys
rm(no_effort_flags)

combined_flags <- length_weight_flag_1 %>% 
   filter(!((is.na(z_score_length) | (z_score_length > -10 & z_score_length < 10)) & 
           (is.na(z_score_weight) | (z_score_weight > -10 & z_score_weight < 10)))) %>% 
  bind_rows(length_weight_flag_2) %>% 
  select(total_effort_ident,
         species_1,
         length_1,
         weight_1,
         flag) %>% 
  distinct()

pep_fish_effort <- pep_fish_effort %>% 
  left_join(combined_flags) 

pep_fish_effort %>% 
  group_by(flag) %>% 
  count()

pep_fish_effort.trial <- pep_fish_effort %>%
  ungroup() %>% 
  mutate(state = "Minnesota",
         lake_name = "Pepin",
         lake_id = "25000100") %>% 
  #dates
  mutate(date_survey = first(date),
         date_total_effort_ident = first(date),
         # date_sub_effort_ident = date,
         # date_sample = date,
         month = month(date_total_effort_ident)) %>% 
  #sampling method 
  mutate(sampling_method_simple = "gill_net",
         sampling_method = "gill_net_standard",
         sampling_method_2 = "gill_net_standard_graduated_mesh") %>% 
  # total effort 
  mutate(total_effort_1 = Number_Nets,
         total_effort_1_units = "net_nights") %>% 
  #total effort nothing caught
  group_by(total_effort_ident) %>% 
  mutate(total_effort_nothing_caught = case_when(is.na(species_1) ~ TRUE,
                                                 TRUE ~ FALSE)) %>% 
  ungroup() %>% 
  # site id 
  mutate(site_id = station)  %>% 
  #original file name
  mutate(original_file_names = paste0(original_file_name_fish, "; ", original_file_name_effort)) %>%
  #notes_1
  mutate(notes_1 = paste0(ifelse(!is.na(`Mesh Size (in)`), paste0("mesh_fish_capture: ", `Mesh Size (in)`, "; "), ""),
                          ifelse(!is.na(Maturity), paste0("fish_maturity: ", Maturity, "; "), ""),
                          ifelse(!is.na(batch_weight_id), paste0("batch_weight_id: ", batch_weight_id, "; "), "")),
         notes_1 = str_remove(notes_1, "; $"),
         notes_1 = na_if(notes_1, "")) %>% 
    select(state, 
         #county, 
         lake_name,
         lake_id,
         #nhdhr_id,
         #latitude_lake_centroid,
         #longitude_lake_centroid,
         date_survey,
         date_total_effort_ident,
         #date_sub_effort_ident,
         #date_sample,
         year,
         month,
         #survey_id,
         #survey_type,
         sampling_method_simple,
         sampling_method,
         sampling_method_2,
         #gear_data_notes,
         #target_species,
         total_effort_ident,
         total_effort_1, 
         #total_effort_2,
         total_effort_1_units,
         #total_effort_2_units,
         total_effort_nothing_caught,
         #water_temp,
         #water_temp_units,
         #water_clarity,
         #water_clarity_units,
         #lat_start,
         #lon_start,
         #lat_end,
         #lon_end,
         site_id,
         #sub_effort_ident,
         #sub_effort_1,
         #sub_effort_1_units,
         #sub_effort_2,
         #sub_effort_2_units,
         #sub_effort_nothing_caught,
         species_1,
         length_1,
         length_unit_1,
         #length_bin,
         #length_bin_unit,
         #age,
         #aging_structure_1,
         weight_1,
         weight_unit_1,
         batch_weight,
         batch_weight_unit,
         sex,
         #age_class,
         flag,
         original_file_names,
         #ind_fish_ident,
         #lakesize,
         #lakesize_units,
         #lat_unspec,
         #lon_unspec,
         #waterbody_type,
         #location_notes_1,
         notes_1
         )

glimpse(pep_fish_effort.trial)
```

# combine large lakes and historical pepin
```{r}
rm(combined_flags, 
   length_weight_flag_1, 
   length_weight_flag_2,
   ll_effort, 
   ll_fish, pep_fish,
   pep_fish_effort, 
   total_effort,
   aging_rank)

large_lakes <- bind_rows(ll_fish_effort, pep_fish_effort.trial) %>% 
  mutate(length_unit_1 = case_when(is.na(length_1) ~ NA,
                                   TRUE ~ length_unit_1),
         weight_unit_1 = case_when(is.na(weight_1) ~ NA,
                                   TRUE ~ weight_unit_1),
         total_effort_1_units = case_when(is.na(total_effort_1) ~ NA,
                                   TRUE ~ total_effort_1_units),
         water_temp_units = case_when(is.na(water_temp) ~ NA,
                                   TRUE ~ water_temp_units),
         batch_weight_unit = case_when(is.na(batch_weight) ~ NA,
                                   TRUE ~ batch_weight_unit))
glimpse(large_lakes)
```

# Lets get this into the mn hive
1. First we need to check the lake years of gn data that already exist and how to compares to the update
2. we need to update some spatial data in ll update from the existing db
3. then we need to make sure the columns align (date structure, survey date, make total effort ident unique within ll so there is none repeated, new obs id)
```{r}
glimpse(mn_dat)

#what do we have for data?
dows <- large_lakes %>% 
  distinct(lake_id, lake_name, survey_id)

#do we have the dows?
mn_dat %>% 
  filter(lake_id %in% dows$lake_id) %>% 
  distinct(lake_id, lake_name)
#okay everything looks good here

#compare how many years of data exist in the mn state pull vs the large lakes pull
mn_dat %>% 
  filter(lake_id %in% dows$lake_id) %>%
  filter(sampling_method_simple == "gill_net") %>% 
  group_by(lake_name, lake_id, survey_type) %>% 
  summarise(year_min = min(year),
            year_max = max(year),
            n_years = n_distinct(year))%>% 
  print(n = nrow(.))

large_lakes %>% 
  group_by(lake_name, lake_id) %>% 
  summarise(year_min = min(year),
            year_max = max(year),
            n_years = n_distinct(year)) 
#looks like kab already has some extra surveys that are labeled as historical and not just Large Lakes?

no_match <- mn_dat %>% 
  filter(lake_id %in% dows$lake_id) %>% 
  filter(sampling_method_simple == "gill_net") %>% 
  anti_join(large_lakes, by = c("lake_name", "year", "sampling_method_simple"))

no_match %>% 
  group_by(lake_name, year, survey_type) %>% 
  count()
#there are 17 lake years that the older file has that the new one doesn't
#looks like they are are historical outside of the one erroneous lake of the woods "nothing caught"

no_match %>% 
  group_by(is.na(age)) %>% 
  count()
#i bet it is because they didn't have age data so Kristin left those data out... we can check places where they do share data to see if they have age 

glimpse(no_match)

large_lakes %>% 
  anti_join(mn_dat, by = c("lake_name", "year", "sampling_method_simple")) %>% 
  group_by(lake_name, year, is.na(age)) %>% 
  count() %>% 
  print(n = nrow(.))
 #for the data in large lakes but not in the original database, there are always ages, we knew historical pepin did not have age data. This gives me more confidence that Kristin gave us only surveys with age data and may not have included historical surveys without age
 #this leads me to thinking we will still want those residual surveys from the original data pull 
 #1980 kab, 1980 rainy, 1981 red, all but 81 and 83 vermilion do not show in lake finder
 
org_efforts <- mn_dat %>% 
  filter(lake_id %in% dows$lake_id) %>%
  filter(sampling_method_simple == "gill_net") %>% 
  distinct(lake_name, year, total_effort_1, sampling_method_2, date_total_effort_ident)

ll_effort <- large_lakes %>% 
  distinct(lake_name, year, total_effort_1) %>% 
  rename(effort_ll = total_effort_1)

check <- org_efforts %>% 
  left_join(ll_effort)

ll_fish_effort %>% 
  filter(lake_name == "Winnibigoshish" & year == "2014") %>%
  group_by(date_total_effort_ident) %>% 
  summarise(n_nets = n_distinct(site_id))
#lake finder shows two targeted surveys from Winni in some years - thus Mike has split this effort out into the respective groupings correctly. It appears that Kristin has not done this and I need to go back and see if I could retain that information in some way

winni <- ll_effort %>% 
  filter(NAME == "Winnibigoshish" & Year == "2014")
#okay this is from the original file - we can see that there are multipule survey types in winni in 2014 so the effort should be split up between them respectively... but that makes a headache for using the total effort tab because it is not split up like that 
```


# Save Parquet
```{r}

fwrite(mn_dat, "Data_and_Scripts/Data/output/mn_data.csv" , na="NA" )
rm(mn_dat)
mn_dat <- fread("Data_and_Scripts/Data/output/mn_data.csv", na.strings = "NA")

mn_dat[ , date_sub_effort_ident := as.IDate(NA) , ]
mn_dat %>% 
  mutate(date_sub_effort_ident = as.IDate(NA)) %>% 
  {mn_dat <<- .}

#needs date_sub_effort_ident converted back to date


# write_dataset(dataset = mn_dat, path = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "hive_update", "state=Minnesota"), format = "parquet")
# 
# mn_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "hive_update", "state=Minnesota"))
# 
# glimpse(mn_data)

```

#sleuth out the MN duplicated/missing erroneously data issues:
```{r}

#we have a number of cases where we ended up with >1 record for a single survey. Some are clearly effed up. Historical surveys that sometimes had the data sometimes not, and match to a "population assessment" that sometimes has the data, sometimes not. The following code works tyowards a flagging system, but does not accomplish what we want. 


# mn_dat %>% 
#   group_by(lake_name, lake_id, date_total_effort_ident, sampling_method, sampling_method_2) %>% 
#   collect() %>% 
#   summarise(n = n_distinct(survey_type)) %>%
#   filter(n>1) %>% 
#   collect() %>% {. ->> dups}
# 
# mn_dat %>% 
#   # mutate(year = year(date)) %>%
#   inner_join(dups) %>%
#   group_by(lake_name, lake_id, date_total_effort_ident, sampling_method,sampling_method_2, survey_type, total_effort_ident) %>% 
#   summarise(n_catch_all_spp = sum(!is.na(species_1))) %>%
#   collect() %>%
#   filter(!lake_name == "Winnibigoshish") %>% 
#   {. ->> dups_with_surveytype_and_catch}
# 
# setorder(dups_with_surveytype_and_catch, lake_id, date_total_effort_ident ,-n_catch_all_spp)
# 
# #After excluding the indfish/effort mismatch
# dups_with_surveytype_and_catch <- dups_with_surveytype_and_catch[!(survey_type %in% c("External Management Survey", "External Mgmt Survey"))]
# 
# 
# #we've got some bad data coming in from DNR, wherever we see a lake - year -gear
# mn_dat[lake_name == "Lake of the Woods", .N , .(lake_name, lake_id, date_total_effort_ident, sampling_method, survey_type) ][order(date_total_effort_ident, sampling_method )]
# 
# #within a year - lake - gear, if a catch value is 1/10th of other TEIs, flag that TEI
# dup_gears <- mn_dat[ , .("n_TEIs" = length(unique(total_effort_ident)), "nfish" = sum(!is.na(species_1))) , .(lake_id,lake_name, year, sampling_method, sampling_method_2) ][n_TEIs>1]
# 
# two_surveys <- mn_dat[dup_gears, on = .(lake_id, year, sampling_method)][ , .("nfish" = sum(!is.na(species_1))) , .(lake_id,lake_name, year, date_total_effort_ident,  sampling_method, sampling_method_2, survey_type , total_effort_ident, total_effort_1, total_effort_1_units) ]
# 
# two_surveys[ , max_catch_inyear := max(nfish) , .(lake_id, year, sampling_method_2) ]
# 
# two_surveys[ , flag_test := nfish/max_catch_inyear < 1/100 , ]
# 
# 
# ##### another way to find them easily?
# 
# nocatch <- mn_dat[total_effort_nothing_caught == T , .N   ,.(lake_id,lake_name, year, date_total_effort_ident,  sampling_method, sampling_method_2, survey_type , total_effort_ident, total_effort_1, total_effort_1_units) ]
# 
# 
# nocatch_lakes <- mn_dat[nocatch, on = .(lake_id,lake_name,sampling_method),
#        ][  ,.N   ,.(lake_id,lake_name, year, date_total_effort_ident,  sampling_method, sampling_method_2, survey_type , total_effort_ident, total_effort_1, total_effort_1_units, total_effort_nothing_caught, original_file_names)][order(lake_id,  sampling_method_2,date_total_effort_ident, total_effort_nothing_caught)]
# 

#we had a question about the validity of using site IDs as one net-night of effort (assume within a survey each site was only sampled for one net-night). While that pattern does dominate the data, it is not consistent enough to use based on this exploration: 

# mn_dat %>%
#   filter(sampling_method_2 %in% c("gill_net_standard_graduated_mesh", "trap_net; 0.75-in-mesh net; double-frame net")) %>% 
#   group_by(total_effort_ident, lake_name, lake_id, year, total_effort_1, sampling_method_simple) %>% 
#   summarise( n_sites = length(unique(site_id))) %>%
#   collect() %>% 
#   mutate(effort_diff = total_effort_1-n_sites ) %>% 
#   {effort_assumptions <<- .}
#   
# 
# ggplot(effort_assumptions, aes( effort_diff, group = sampling_method_simple))+
#   geom_histogram(breaks = c(-30:100))+
#   scale_y_log10()












```

