---
title: "c_Iowa_Filtering_CPUE"
author: "Denver Link"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
ia_data <- open_dataset(sources = file.path("G:",
                                            "Shared drives",
                                            "Hansen Lab",
                                            "RESEARCH PROJECTS",
                                            "Fish Survey Data",
                                            "Parquet files",
                                            "hive_update"), partitioning = c("state")) %>% 
  filter(state == "Iowa")
glimpse(ia_data)
```

#target species
```{r}
ia_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect()

ia_data %>% 
  mutate(species_included = case_when(str_detect(target_species, "walleye") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  group_by(target_species, species_included) %>% 
  count() %>% 
  arrange(species_included) %>% 
  collect()
```
We will make to make sure that we filter for surveys were there is no target species or the species of interest is the target. See above how we make a y or n for filtering when the species of interest should be included in the good surveys. This will be undergone for each species within their section of code. 

#effort flag check
```{r}
ia_data %>% 
  group_by(flag) %>% 
  count() %>% 
  collect() %>%
  arrange(flag) %>% 
  print(n = nrow(.))

ia_data %>% 
  group_by(cpue_invalid, flag) %>%
  count() %>% 
  arrange(cpue_invalid) %>% 
  collect() %>% 
  print(n = nrow(.))
```
There are several flags in Iowa that we will want to weed out for relative abundance. Any fish level flag is fine because we are just counting. Here are two flags that are concerning - high effort and agency flagged surveys. We see that our invalid cpue flags do a good job picking up all survey level flags.

#effort units filtering
```{r}
ia_data %>% 
  group_by(sampling_method, sampling_method_2, total_effort_1_units) %>% 
  count() %>% 
  collect() %>% 
  arrange(sampling_method) %>% 
  print(n = nrow(.))
```
Hours is the standard Iowan effort unit - we need to make sure that we do not use distance. There appear to be some NA unit values, all of those have an adequacy flag tagged on them and it is not needed to worry about filtering on that.

#filtering walleye
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "walleye")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "walleye") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y", 
                                        TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_walleye <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  #remove sub sample surveys that caught nothing but total effort did - generates NAs in cpe calc
  #the effort is still accounted for in the total effort
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE")) %>% 
  ungroup()

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included, cpue_invalid) %>% 
  count() 

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_walleye %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, cpue_invalid) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, cpue_invalid) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_walleye)
```

#filtering black crappie
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "black_crappie")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "black crappie") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y",
                                        TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_black_crappie <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE"))

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included, cpue_invalid) %>% 
  count() 

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_black_crappie %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, flag) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, flag) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_black_crappie)
```

#filtering bluegill
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "bluegill")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "bluegill") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y", 
                              TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_bluegill <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE"))

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included, cpue_invalid) %>% 
  count() 

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_bluegill %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, flag) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, flag) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_bluegill_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_bluegill)
```

#filtering largemouth bass
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "largemouth_bass")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "largemouth") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y", 
                              TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_largemouth_bass <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE"))

#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included) %>% 
  count() 

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_largemouth_bass %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, flag) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, flag) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth_bass %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_largemouth_bass_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_largemouth_bass)
```

#filtering smallmouth bass
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "smallmouth_bass")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "smallmouth") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y",
                                        TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_smallmouth_bass <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE"))

#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included, cpue_invalid) %>% 
  count() 

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_smallmouth_bass %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, flag) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, flag) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth_bass %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_smallmouth_bass_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_smallmouth_bass)
```

#filtering northern pike
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "northern_pike")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "pike") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y", 
                              TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_northern_pike <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE"))

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included, cpue_invalid) %>% 
  count() 

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_northern_pike %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, flag) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, flag) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_northern_pike)
```

#filtering yellow perch
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "yellow_perch")  %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- ia_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(str_detect(target_species, "perch") | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  #filter for good survey type
  mutate(survey_type_filter = case_when(survey_type %in% c("General", 
                                                           "Comprehensive", 
                                                           "Trend") ~ "y", 
                              TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method_2")) %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           total_effort_1_units,
           species_included,
           survey_type_filter,
           cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
        month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        species_included == "y",
        survey_type_filter == "y",
        cpue_invalid == "n") %>% 
  collect() %>% 
  filter(total_effort_1_units == "hours")

#collecting fish from good surveys
adult_yellow_perch <- ia_data %>% 
  right_join(good_surveys) %>% 
  collect() %>% 
  #removing one survey that had na and lmb targeted species values at the species level (not survey level)
  group_by(total_effort_ident) %>% 
  filter(n_distinct(target_species) == 1) %>% 
  filter(!(total_effort_nothing_caught == "FALSE" & sub_effort_nothing_caught == "TRUE"))

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

good_surveys %>% 
  group_by(survey_type_filter, species_included, cpue_invalid) %>% 
  count() 

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)), 
            min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            n = n())

adult_yellow_perch %>% 
  distinct(total_effort_ident,target_species, species_included,survey_type, survey_type_filter, flag) %>%
  group_by(target_species, species_included,survey_type, survey_type_filter, flag) %>% 
   collect() %>% 
   count() 

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(total_effort_ident,
           sampling_method_2,
           date_total_effort_ident,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           species_included,
           survey_type_filter,
           target_species) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue)) 

rm(adult_yellow_perch)
```

#combining species
```{r}
ia_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(ia_filtered_species, "IA_all_cpue_filtered.csv")
```
