---
title: "c_Iowa_Filtering_CPUE"
author: "Denver Link"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

To do:
-figure out joining on total effort columns
    -nas appears removed from the output (everything works when I join on effort ident)
-figure out the date format
-several surveys that have na or 0 total effort? leads to na cpue
       -it appears many of these efforts are recorded in distance rather than time
-cpue for electofishing still na after removing na total effort
-ia_data seems to be missing some fish (ex. only 35 lmb)

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
filter_table <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(state == "Illinois")

ia_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "ia_file_arrow"))
glimpse(ia_data)
```

#filtering walleye
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "walleye") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "walleye") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
#  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_walleye <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_walleye %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_walleye)
```

#filtering black crappie
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "black_crappie") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "black_crappie") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
#  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_black_crappie <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_black_crappie %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_black_crappie_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_black_crappie)
```

#filtering bluegill
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "bluegill") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "bluegill") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
# mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_bluegill <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_bluegill %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_bluegill_cpue %>% 
  filter(!is.na(total_effort_1)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()
```

#filtering largemouth bass
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "largemouth_bass") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "largemouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
# mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_largemouth <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_largemouth %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_largemouth_cpue <- adult_largemouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_largemouth_cpue %>% 
  filter(!is.na(total_effort_1)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_largemouth_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_largemouth_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_largemouth)
```

#filtering smallmouth bass
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "smallmouth_bass") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "largemouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
# mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_smallmouth <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_smallmouth %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_smallmouth_cpue <- adult_smallmouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_smallmouth_cpue %>% 
  filter(!is.na(total_effort_1)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_smallmouth_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_smallmouth_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_smallmouth)
```

#filtering northern pike
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "northern_pike") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "northern_pike") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
#  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_northern_pike <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_northern_pike %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_northern_pike_cpue %>% 
  filter(!is.na(total_effort_1)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_northern_pike_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_northern_pike)
```

#filtering yellow perch
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "yellow_perch") %>% 
  select(-metric,
         -survey_type,
         -species_1,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -effort_min,
         -survey_type_effort,
         -target_species_1,
         -target_species_2)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Iowa" & species_1 == "yellow_perch") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- ia_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
#           total_effort_1,
#           total_effort_1_units,
#          total_effort_2,
#          total_effort_2_units
#           effort_min,
#           month_min,
#           month_max
           ) %>% 
#  filter(total_effort_1.1 >= effort_min,
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect() 
#  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S"))

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month((date.1))))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

adult_yellow_perch <- ia_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_yellow_perch %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  mutate(date.1 = as.POSIXct(date.1, format = "%m/%d/%Y %H:%M:%S")) %>% 
  summarise(min.month = min(month(date.1)), max.month = max(month(date.1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           date.1,
           sampling_method,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           total_effort_2,
           total_effort_2_units,
           count,
           cpue) %>% 
  collect()

#range of cpues?
adult_yellow_perch_cpue %>% 
  filter(!is.na(total_effort_1)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))
#an efishing value has an effort of 0?

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue)) %>% 
  print(n = nrow(.))

#surveys that have nothing caught?
adult_yellow_perch_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_yellow_perch)

```

#combining species
```{r}
ia_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(ia_filtered_species, "IA_all_cpue_filtered.csv")
```