---
title: "c_IL_Filtering_for_CPUE"
author: "Denver Link"
date: "2023-10-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
il_data <- open_dataset(sources = file.path("G:",
                                            "Shared drives",
                                            "Hansen Lab",
                                            "RESEARCH PROJECTS",
                                            "Fish Survey Data",
                                            "Parquet files",
                                            "hive_update"), partitioning = c("state")) %>% 
  filter(state == "Illinois")
glimpse(il_data)
```

#targeted species check
```{r}
il_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect()
#here we can see that Illinois has unique ways of marking a survey as targeted:
#1.) the survey targeted that fish and only that species should be used in cpe generation
#2.) the survey actively didn't target that species and that species should NOT be used in cpe generation while the others can be
#see the explanation below for more info

il_data %>% 
  mutate(bluegill_relevant =case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "bluegillY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "bluegillN") ~ "y",
                                      T ~ "n")) %>% 
  group_by(target_species, bluegill_relevant) %>% 
  count() %>% 
  arrange(bluegill_relevant) %>% 
  collect()
```
We can see here that Illinois structures efishing surveys differently than other states, they do "sub sampling" where in the first amount of time they specifically DO NOT target a particular species and then in the 2nd part of the survey they ONLY target that species. We have aggregated this data so that each "sub sampling" event is its own total effort ident. Each total effort ident then has a target species where a species is either targeted "[insert species]Y" or it is not targeted "[insert species]N". We have to do a little clever coding so that we only include total efforts where a species is targeted OR not excluded 

#effort flag check
```{r}
il_data %>% 
  group_by(flag) %>% 
  count() %>% 
  collect()

il_data %>% 
  group_by(flag, cpue_invalid) %>% 
  count() %>% 
  collect()
```
We can see here that all of our flags concerning effort have been spotted by the cpue_invalid column

#walleye filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "walleye") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(walleye_relevant = case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "walleyeY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "walleyeN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           walleye_relevant,
           cpue_invalid) %>% 
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         walleye_relevant == "y",
         cpue_invalid == "n"
         ) %>% 
    collect() 

#grabbing the fish data
adult_walleye <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, walleye_relevant) %>% 
  count()

#all walleye targeted surveys have month or effort weeding them out
il_data %>% 
  filter(target_species == "walleyeY") %>% 
  distinct(total_effort_ident, date_total_effort_ident, total_effort_1, flag, sampling_method) %>% 
  collect() 

il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_walleye %>% 
  group_by(species_1) %>% 
  count()

adult_walleye %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_walleye %>% 
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           walleye_relevant,
           cpue_invalid) %>% 
  glimpse()

adult_walleye %>% 
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           #sampling_method_2
           ) %>% 
  glimpse()
#here we see that removing the sampling method 2 gets us to the same number of obs as the cpe file, this is because there are some surveys that have _Wisconsin applied to the end of the value and some records do within a total effort ident. These stay around in the good surveys and adult species column in order to retain the fish records for calculation, but the cpe is made at the total effort ident level and then only the first record of the sampling method appears when they are collapsed

#sub surveys happen across multiple days so there are many sub surveys within a total effort ident
adult_walleye %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

adult_walleye_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_walleye_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_walleye)
```

#black crappie filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "black_crappie") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(crappie_relevant = case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "black_crappieY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "black_crappieN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           crappie_relevant,
           cpue_invalid) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         crappie_relevant == "y",
         cpue_invalid == "n") %>% 
    collect() 

#grabbing the fish data
adult_black_crappie <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, crappie_relevant) %>% 
  count()


il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  group_by(sampling_method_2, month, total_effort_1) %>% 
  count() %>% 
  collect()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_black_crappie %>% 
  group_by(species_1) %>% 
  count()

adult_black_crappie %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_black_crappie %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()
#the additional records in good_surveys relative to the cpe file is due to some records within a total effort ident containing _wisconsin while some records don't - this is fine becuase we retain all records from the total effort ident and then we generate cpe at the total effort ident level across all gear types

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_black_crappie %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

adult_black_crappie_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_black_crappie_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_black_crappie)
```

#bluegill filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "bluegill") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(bluegill_relevant =case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "bluegillY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "bluegillN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species,
           bluegill_relevant,
           cpue_invalid) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         bluegill_relevant == "y",
         cpue_invalid == "n") %>% 
    collect() 

#grabbing the fish data
adult_bluegill <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, bluegill_relevant) %>% 
  count()

il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  group_by(sampling_method, month, total_effort_1) %>% 
  count() %>% 
  collect()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_bluegill %>% 
  group_by(species_1) %>% 
  count()

adult_bluegill %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_bluegill %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_bluegill %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()
#see notes in walleye and black crappie section - the difference in obs when including sampling method 2 vs just sampling method is inconsequential to the cpe calculation 

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_bluegill %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_bluegill_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

adult_bluegill_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method_2, scales = "free") 

adult_bluegill_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_bluegill)
```

#largemouth bass filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "largemouth_bass") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(largemouth_relevant =case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "largemouth_bassY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "largemouth_bassN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           largemouth_relevant,
           cpue_invalid) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         largemouth_relevant == "y",
         cpue_invalid == "n"
         ) %>% 
    collect() 

#grabbing the fish data
adult_largemouth_bass <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, largemouth_relevant) %>% 
  count()

il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  group_by(sampling_method_2) %>% 
  count() %>% 
  collect()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_largemouth_bass %>% 
  group_by(species_1) %>% 
  count()

adult_largemouth_bass %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_largemouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method, sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_largemouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_largemouth_bass %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()
#see notes in above species about the difference between good surves and cpe obs

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_largemouth_bass %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_largemouth_bass_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue))

adult_largemouth_bass_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_largemouth_bass_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_largemouth_bass)
```

#smallmouth bass filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "smallmouth_bass") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(smallmouth_relevant =case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "smallmouth_bassY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "smallmouth_bassN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           smallmouth_relevant,
           cpue_invalid) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         smallmouth_relevant == "y",
         cpue_invalid == "n") %>% 
    collect() 

#grabbing the fish data
adult_smallmouth_bass <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, smallmouth_relevant) %>% 
  count()

il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  group_by(sampling_method, month, total_effort_1) %>% 
  count() %>% 
  collect()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_smallmouth_bass %>% 
  group_by(species_1) %>% 
  count()

adult_smallmouth_bass %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method, sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_smallmouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_smallmouth_bass %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_smallmouth_bass %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_smallmouth_bass_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue))

adult_smallmouth_bass_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_smallmouth_bass_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_smallmouth_bass)
```

#northern pike filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "northern_pike") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(pike_relevant =case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "northern_pikeY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "northern_pikeN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           pike_relevant,
           cpue_invalid) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         pike_relevant == "y",
         cpue_invalid == "n") %>% 
    collect() 

#grabbing the fish data
adult_northern_pike <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, pike_relevant) %>% 
  count()

il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  group_by(sampling_method, month, total_effort_1) %>% 
  count() %>% 
  collect()

#checking final filter
adult_northern_pike %>% 
  group_by(species_1) %>% 
  count()

adult_northern_pike %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method, sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_northern_pike %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multiple days so there are many sub surveys within a total effort ident
adult_northern_pike %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

adult_northern_pike_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_northern_pike_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_northern_pike_cpue %>% 
  filter((total_effort_1 < 100) & sampling_method == "Boat electrofishing DC (Day)") %>% 
  filter(cpue != 0) %>% 
  select(lake_id, year, cpue, count, total_effort_1) %>% 
  arrange(total_effort_1) %>% 
  print(n = nrow(.))

rm(adult_northern_pike)
```

#yellow perch filtering
```{r}
###filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "yellow_perch") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(perch_relevant =case_when(is.na(target_species) ~ "y",
                                      str_detect(target_species, "yellow_perchY") ~ "y",
                                      str_detect(target_species, "N") & !str_detect(target_species, "yellow_percheN") ~ "y",
                                      T ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type,
           date_total_effort_ident, 
           total_effort_ident,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           perch_relevant,
           cpue_invalid) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         perch_relevant == "y",
         cpue_invalid == "n") %>% 
    collect() 

#grabbing the fish data
adult_yellow_perch <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method_2, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(target_species, perch_relevant) %>% 
  count()

il_data %>% 
  filter(target_species == "yellow_perchY") %>% 
  select(total_effort_1, date_total_effort_ident, total_effort_ident) %>% 
  glimpse()
#these surveys are filtered out by date

il_data %>% 
  filter(total_effort_nothing_caught == TRUE) %>% 
  group_by(sampling_method, month, total_effort_1) %>% 
  count() %>% 
  collect()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_yellow_perch %>% 
  group_by(species_1) %>% 
  count()

adult_yellow_perch %>% 
  group_by(survey_type) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method, sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#how does the minimum effort look?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_yellow_perch %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(lake_id,
           lake_name,
           survey_type,
           survey_id,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multiple days so there are many sub surveys within a total effort ident
adult_yellow_perch %>% 
  distinct(total_effort_ident, date_total_effort_ident) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

adult_yellow_perch_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_yellow_perch_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_yellow_perch)
```

#combining species
```{r}
il_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(il_filtered_species, "IL_all_cpue_filtered.csv")
```
