---
title: "c_IL_Filtering_for_CPUE"
author: "Denver Link"
date: "2023-10-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

update 1/31:
-walleye filtering has been updated to work with new filters and script 
-now the same filtering done on walleye needs to be applied to all other fish
-one thing I might have to consider is at what point did I make sampling_method names in a survey
   -if it is before i make the total effort ident there is a chance that there are two total effort idents with in the same survey

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
il_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "il_file_arrow"))
glimpse(il_data)
```

#walleye filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "walleye") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Walleye", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Walleye", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_walleye <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

#all walleye targeted surveys have month or effort weeding them out
il_data %>% 
  filter(target_species == "Walleye") %>% 
  distinct(total_effort_ident, date, total_effort_1) %>% 
  compute() %>% 
  glimpse()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_walleye %>% 
  group_by(species_1) %>% 
  count()

adult_walleye %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_walleye %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_walleye %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

adult_walleye_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_walleye_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_walleye)
```

#black crappie filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "black_crappie") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Black Crappie", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Black Crappie", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_black_crappie <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Black Crappie", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Black Crappie", no_target) ~ "y",
                                      TRUE ~ "n")) %>%
  group_by(target_species, no_target, species_included) %>% 
  count() %>% 
  collect()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_black_crappie %>% 
  group_by(species_1) %>% 
  count()

adult_black_crappie %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_black_crappie %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_black_crappie %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

adult_black_crappie_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_black_crappie_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_black_crappie)
```

#bluegill filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "bluegill") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Bluegill", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Bluegill", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_bluegill <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Bluegill", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Bluegill", no_target) ~ "y",
                                      TRUE ~ "n")) %>%
  group_by(species_included, no_target, target_species) %>% 
  count() %>% 
  arrange(species_included) %>% 
  collect()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_bluegill %>% 
  group_by(species_1) %>% 
  count()

adult_bluegill %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_bluegill %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_bluegill %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_bluegill_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

adult_bluegill_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_bluegill_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_bluegill)
```

#largemouth bass filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "largemouth_bass") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Largemouth Bass", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Largemouth Bass", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_largemouth_bass <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Largemouth Bass", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Largemouth Bass", no_target) ~ "y",
                                      TRUE ~ "n")) %>%
  group_by(species_included, no_target, target_species) %>% 
  count() %>% 
  arrange(species_included) %>% 
  collect()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_largemouth_bass %>% 
  group_by(species_1) %>% 
  count()

adult_largemouth_bass %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_largemouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_largemouth_bass %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_largemouth_bass %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_largemouth_bass_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue))

adult_largemouth_bass_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_largemouth_bass_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_largemouth_bass)
```

#smallmouth bass filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Smallmouth Bass", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Smallmouth Bass", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_smallmouth_bass <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Smallmouth Bass", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Smallmouth Bass", no_target) ~ "y",
                                      TRUE ~ "n")) %>%
  group_by(species_included, no_target, target_species) %>% 
  count() %>% 
  arrange(species_included) %>% 
  collect()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_smallmouth_bass %>% 
  group_by(species_1) %>% 
  count()

adult_smallmouth_bass %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_smallmouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_smallmouth_bass %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multipule days so there are many sub surveys within a total effort ident
adult_smallmouth_bass %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_smallmouth_bass_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue))

adult_smallmouth_bass_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_smallmouth_bass_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_smallmouth_bass)
```

#northern pike filtering
```{r}
##filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "northern_pike") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Northern Pike", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Northern Pike", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_northern_pike <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Northern Pike", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Northern Pike", no_target) ~ "y",
                                      TRUE ~ "n")) %>%
  group_by(species_included, no_target, target_species) %>% 
  count() %>% 
  arrange(species_included) %>% 
  collect()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_northern_pike %>% 
  group_by(species_1) %>% 
  count()

adult_northern_pike %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_northern_pike %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multiple days so there are many sub surveys within a total effort ident
adult_northern_pike %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

adult_northern_pike_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_northern_pike_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_northern_pike_cpue %>% 
  filter((total_effort_1 < 100) & sampling_method == "Boat electrofishing DC (Day)") %>% 
  filter(cpue != 0) %>% 
  select(lake_id, year, cpue, count, total_effort_1) %>% 
  arrange(total_effort_1) %>% 
  print(n = nrow(.))

rm(adult_northern_pike)
```

#yellow perch filtering
```{r}
###filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "yellow_perch") %>% 
  rename(survey_type_1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

#selecting good surveys
good_surveys <- il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Yellow Perch", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Yellow Perch", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  #joining the filter table and data - this will select survey types and methods that are appropriate 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method")) %>% 
  #collapse data to survey level while keeping important information for filtering below (.keep_all = T not supported in arrow)
  #date must be retained here so that all fish are brought in later, this will show surveys with sub surveys
  distinct(lake_id,
           survey_type_1,
           date, 
           total_effort_ident,
           sampling_method,
           total_effort_1,
           effort_min,
           effort_max,
           month_min,
           month_max,
           target_species, 
           no_target,
           species_included) %>%
  #filtering on certain columns 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date) >= month_min,
         month(date) <= month_max,
         species_included == "y") %>% 
    collect() 

#grabbing the fish data
adult_yellow_perch <- il_data %>% 
  collect() %>% #join functions ignore the na_match call in parquet form - must collect before joining
  inner_join(good_surveys) 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking good surveys filtering
good_surveys %>% 
  ggplot() +
  geom_histogram(aes(total_effort_1)) +
  geom_vline(xintercept = c(30, 1440, 2880), color = "red") +
  facet_wrap(~sampling_method, scales = "free") 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

good_surveys %>% 
  group_by(no_target, target_species, species_included) %>% 
  count()

il_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("Yellow Perch", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Yellow Perch", no_target) ~ "y",
                                      TRUE ~ "n")) %>%
  group_by(species_included, no_target, target_species) %>% 
  count() %>% 
  arrange(species_included) %>% 
  collect()

il_data %>% 
  filter(total_survey_nothing_caught == TRUE) %>% 
  glimpse()
#nothing caught surveys are filtered out by month or min effort

#checking final filter
adult_yellow_perch %>% 
  group_by(species_1) %>% 
  count()

adult_yellow_perch %>% 
  group_by(survey_type_1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date)), max.month = max(month(date)))

#how does the minimum effort look?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_yellow_perch %>% 
  distinct(lake_id,
           lake_name,
           survey_type_1,
           survey_id,
           date,
           sampling_method,
           total_effort_ident,
           total_effort_1) %>% 
  glimpse()

#sub surveys happen across multiple days so there are many sub surveys within a total effort ident
adult_yellow_perch %>% 
  distinct(total_effort_ident, date) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

adult_yellow_perch_cpue %>% 
  ggplot() +
  geom_histogram(aes(cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_yellow_perch_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_yellow_perch)
```

#combining species
```{r}
il_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(il_filtered_species, "IL_all_cpue_filtered.csv")
```


#Data exploration
```{r}
il_data %>% 
  collect() %>% 
  mutate(species_included = case_when(grepl("Walleye", target_species) & is.na(no_target) ~ "y",
                                      is.na(target_species) & is.na(no_target) ~ "y", 
                                      is.na(target_species) & !grepl("Walleye", no_target) ~ "y",
                                      TRUE ~ "n")) %>% 
  group_by(target_species, no_target, species_included) %>% 
  count()

il_data %>% 
  group_by(no_target) %>% 
  count() %>% 
  collect()

il_data %>% 
  group_by(species_1) %>% 
  count() %>% 
  collect()

il_data %>% 
  filter(is.na(species_1)) %>% 
  collect()

#targeted surveys?
il_data %>% 
  group_by(survey_type_1) %>% 
  count() %>% 
  collect()

il_data %>% 
  filter(survey_type_1 == "Targeted Species") %>% 
  group_by(total_effort_ident, species_1) %>%
  count() %>% 
  collect() %>% 
  print(n = nrow(.)) %>% 
  group_by(total_effort_ident) %>% 
  count() %>% 
  filter(n >1)
#some surveys report more than one species in the targeted surveys but it doesn't look universal

il_data %>% 
  filter(survey_type_1 == "Targeted Species") %>% 
  group_by(total_effort_ident, species_1) %>%
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(survey_type_1 ,sub_sample_effort) %>% 
  filter(sub_sample_effort != "Subsample Time:0") %>% 
  count() %>% 
  collect() %>% 
  print(n= nrow(.))
#all sub sampling appears to come from community samples

#minimum effort 
il_data %>% 
  group_by(total_effort_1_units) %>% 
  count() %>% 
  collect()

il_data %>% 
  group_by(sampling_method_1, total_effort_1_units) %>% 
  count() %>% 
  collect()

il_data %>% 
  group_by(sampling_method_1) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1)) %>% 
  collect()

high_total <- il_data %>% 
  filter(total_effort_1 == 115200) %>% 
  collect() %>% 
  glimpse()
rm(high_total)

#making net nights for nets
il_data %>% 
  group_by(sampling_method_1, total_effort_1_units) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1)) %>% 
  collect()

il_filtered_species %>%
  ggplot() +
  geom_point(aes(total_effort_1, cpue)) + 
  facet_grid(species_1~sampling_method, scales = "free")

il_filtered_species %>%
  ggplot() +
  geom_histogram(aes(cpue)) + 
  facet_grid(species_1~sampling_method, scales = "free")
ggsave()

il_filtered_species %>%
  filter(str_detect("Trap net",sampling_method)) %>% 
  ggplot() +
  geom_histogram(aes(cpue)) + 
  facet_grid(species_1~sampling_method, scales = "free")


il_filtered_species %>% 
  group_by(sampling_method, species_1) %>% 
  arrange(species_1) %>% 
  count() %>% 
  print(n = nrow(.))

il_filtered_species %>% 
  filter(species_1 == "yellow_perch") %>% 
  ggplot() +
  geom_histogram(aes(cpue)) + 
  facet_grid(~sampling_method, scales = "free")

il_data %>% 
  group_by(species_1) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.)) 

il_data %>% 
  filter(sampling_method == "Boat electrofishing AC (Day)" | sampling_method == "Boat electrofishing DC (Day)") %>% 
  collect() %>% 
  ggplot() + 
  geom_histogram(aes(length_1, fill = sampling_method), alpha = .5) + 
  facet_wrap(~species_1, scales = "free")
```

