---
title: "c_IL_Filtering_for_CPUE"
author: "Denver Link"
date: "2023-10-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

To do:
-targeted species surveys filtering?
-sub sample in electrofishing
   -non subsampled species should use the total effort associated with the total effort ident
   -subsampled species should use the sub sample times

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
filter_table <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(state == "Illinois")

il_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "il_file_arrow"))
glimpse(il_data)

il_data %>% 
  group_by(species_1) %>% 
  count() %>% 
  collect()

il_data %>% 
  filter(is.na(species_1)) %>% 
  collect()

#targeted surveys?
il_data %>% 
  group_by(survey_type_1) %>% 
  count() %>% 
  collect()

il_data %>% 
  filter(survey_type_1 == "Targeted Species") %>% 
  group_by(total_effort_ident, species_1) %>%
  count() %>% 
  collect() %>% 
  print(n = nrow(.)) %>% 
  group_by(total_effort_ident) %>% 
  count() %>% 
  filter(n >1)
#some surveys report more than one species in the targeted surveys but it doesn't look universal

il_data %>% 
  filter(survey_type_1 == "Targeted Species") %>% 
  group_by(total_effort_ident, species_1) %>%
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

il_data %>% 
  group_by(survey_type_1 ,sub_sample_effort) %>% 
  filter(sub_sample_effort != "Subsample Time:0") %>% 
  count() %>% 
  collect() %>% 
  print(n= nrow(.))
#all sub sampling appears to come from community samples

#minimum effort 
il_data %>% 
  group_by(total_effort_1_units) %>% 
  count() %>% 
  collect()

il_data %>% 
  group_by(sampling_method_1, total_effort_1_units) %>% 
  count() %>% 
  collect()

il_data %>% 
  group_by(sampling_method_1) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1)) %>% 
  collect()

high_total <- il_data %>% 
  filter(total_effort_1 == 115200) %>% 
  collect() %>% 
  glimpse()
rm(high_total)

#making net nights for nets
il_data %>% 
  group_by(sampling_method_1, total_effort_1_units) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1)) %>% 
  collect()
```

#walleye filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "walleye") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "walleye") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 


good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_walleye <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_walleye %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_walleye %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_walleye %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

rm(adult_walleye)
```

#black crappie filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "black_crappie") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "black_crappie") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_black_crappie <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_black_crappie %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_black_crappie %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_black_crappie %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

rm(adult_black_crappie)
```

#bluegill filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "bluegill") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "bluegill") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_bluegill <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_bluegill %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_bluegill %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_bluegill %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_bluegill_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

rm(adult_bluegill)
```

#largemouth bass filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "largemouth_bass") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "largemouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_largemouth_bass <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_largemouth_bass %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_largemouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_largemouth_bass %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_largemouth_bass_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue))

rm(adult_largemouth_bass)
```

#smallmouth bass filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_smallmouth_bass <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_smallmouth_bass %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_smallmouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_smallmouth_bass %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_smallmouth_bass_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue))

rm(adult_smallmouth_bass)
```

#northern pike filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "northern_pike") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "northern_pike") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_northern_pike <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_northern_pike %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_northern_pike %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_northern_pike %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

rm(adult_northern_pike)
```

#yellow perch filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "yellow_perch") %>% 
  rename(survey_type_1 = survey_type,
         sampling_method_2 = sampling_method) %>% 
  select(-metric,
         -species_1,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Illinois" & species_1 == "yellow_perch") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- il_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type_1",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units,
           effort_min,
           month_min,
           month_max) %>%
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
    collect() 

#checking filtering
#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type_1) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))


adult_yellow_perch <- il_data %>% 
  #filter data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_yellow_perch %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#how does the minimum effort look?
adult_yellow_perch %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1), max.effort = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys
adult_yellow_perch %>% 
  distinct(lake_id,
           lake_name_1,
           survey_type_1,
           survey_id,
         date_1,
           sampling_method_1,
           sampling_method_2,
           total_effort_ident,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse() 


#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>%
  collect() %>% 
  select(state,
         county,
         lake_name_1,
         lake_id,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

rm(adult_yellow_perch)
```

#combining species
```{r}
il_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(il_filtered_species, "IL_all_cpue_filtered.csv")
```
