---
title: "c_IN_Filtering_CPUE"
author: "Denver Link"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
in_data <- open_dataset(sources = file.path("G:",
                                            "Shared drives",
                                            "Hansen Lab",
                                            "RESEARCH PROJECTS",
                                            "Fish Survey Data",
                                            "Parquet files",
                                            "hive_update"), partitioning = c("state")) %>% 
  filter(state == "Indiana")
glimpse(in_data)
```

#checking for effort flags
```{r}
in_data %>% 
  group_by(flag) %>% 
  count() %>% 
  collect()

in_data %>% 
  summarise(n_effort_flags = sum(str_detect(flag, "high") |
                                   str_detect(flag, "no sampling") |
                                   str_detect(flag, "missing") |
                                   str_detect(flag, "no match"), na.rm = T),
            n_cpue_invalid = sum(cpue_invalid == "y", na.rm = T),
            n_cpue_invalid_flags = sum(cpue_invalid == "y" & (str_detect(flag, "high") |
                                   str_detect(flag, "no sampling") |
                                   str_detect(flag, "missing")|
                                     str_detect(flag, "no match")), na.rm = T)) %>% 
  collect()

in_data %>% 
  filter(cpue_invalid == "y") %>% 
  distinct(flag) %>% 
  collect()
```
Looks like we have some effort flags, some noting missing effort and some noting high effort. We will not have to worry about the missing effort as it will get weeded out naturally, but we will have to filter out the high effort case. We can filter out records we don't want using the invalid cpue column, selecting "n" cases. 

#filtering walleye
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "walleye") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("walleye", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_walleye <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_walleye %>% 
  group_by(target_species, waterbody_type) %>% 
  count()


adult_walleye %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_walleye_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_walleye_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

adult_walleye %>% 
  group_by(waterbody_type, survey_type) %>% 
  count()

adult_walleye %>% 
  group_by(sampling_method, survey_type) %>% 
  count()

rm(adult_walleye)
```

#filtering black crappie
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "black_crappie") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)


good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("black_crappie", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_black_crappie <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_black_crappie %>% 
  group_by(target_species) %>% 
  count()

adult_black_crappie %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_black_crappie_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_black_crappie_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_black_crappie_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

in_data %>% 
  filter(str_detect(target_species, "black_crappie")) %>% 
  group_by(month, total_effort_1, sampling_method) %>% 
  count() %>% 
  collect()
#confirms we should not have any targeted black crappie 

rm(adult_black_crappie)
```

#filtering bluegill
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "bluegill") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)


good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("bluegill", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>% 
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_bluegill <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_bluegill %>% 
  group_by(target_species) %>% 
  count()

adult_bluegill %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_bluegill_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_bluegill_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_bluegill_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 


in_data %>% 
  filter(str_detect(target_species, "bluegill")) %>% 
  group_by(month, total_effort_1, sampling_method) %>% 
  count() %>% 
  collect()

in_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect()
#confirms we should not have any targeted 

rm(adult_bluegill)
```

#filtering largemouth bass
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "largemouth_bass") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("argemouth", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect() 

adult_largemouth_bass <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_largemouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_largemouth_bass %>% 
  group_by(target_species) %>% 
  count()

adult_largemouth_bass %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_largemouth_bass_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_largemouth_bass_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_largemouth_bass_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 


rm(adult_largemouth_bass)
```

#filtering smallmouth bass
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "smallmouth_bass") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)


good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("mallmouth", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_smallmouth_bass <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_smallmouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_smallmouth_bass %>% 
  group_by(target_species) %>% 
  count()

adult_smallmouth_bass %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_smallmouth_bass_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_smallmouth_bass_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_smallmouth_bass_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

rm(adult_smallmouth_bass)
```

#filtering northern pike
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "northern_pike") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)

good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("pike", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_northern_pike <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_northern_pike %>% 
  group_by(target_species) %>% 
  count()

adult_northern_pike %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_northern_pike_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_northern_pike_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_northern_pike_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

in_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect()
#no targeted pike surveys 
  
rm(adult_northern_pike)
```

#filtering yellow perch
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "yellow_perch") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)


good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("yellow_perch", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_yellow_perch <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_yellow_perch %>% 
  group_by(target_species) %>% 
  count()

adult_yellow_perch %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_yellow_perch_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_yellow_perch_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_yellow_perch_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

in_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect()
#no targeted surveys so that checks 

rm(adult_yellow_perch)
```

#filtering cisco
```{r}
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Indiana" & species_1 == "cisco") %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -area_group)


good_surveys <- in_data %>% 
  #creating a y or n for if we could even expect the species to be present in the survey
  mutate(species_included = case_when(grepl("cisco", target_species) | is.na(target_species) ~ "y",
                                      TRUE ~ "n")) %>%
  right_join(filter_table_effort, by = c("state",
                                        "sampling_method",
                                        "survey_type")) %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           month,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min,
           month_min,
           month_max,
           species_included,
           cpue_invalid)%>% 
  filter(total_effort_1 >= effort_min,
         month >= month_min,
         month <= month_max,
         species_included == "y",
         cpue_invalid == "n") %>% 
  collect()

adult_cisco <- in_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'cisco')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking surveys
good_surveys %>% 
  group_by(survey_type, sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), n = n())

adult_cisco %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), 
            max = max(total_effort_1), 
            mean = mean(total_effort_1), 
            n = n()) 

adult_cisco %>% 
  group_by(target_species) %>% 
  count()

adult_cisco %>% 
  distinct(lake_id,
           sampling_method,
           survey_type,
           total_effort_1,
           total_effort_1_units,
           total_effort_ident,
           effort_min) %>% 
  glimpse()

adult_cisco %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_cisco_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_cisco_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_cisco_cpue %>% 
  ggplot() +
  geom_smooth(aes(total_effort_1, cpue)) +
  geom_point(aes(total_effort_1, cpue)) +
  facet_wrap(~sampling_method, scales = "free") 

in_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect()

rm(adult_cisco)
```

#combining cpes
```{r}
in_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_cisco_cpue %>% 
                                   mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(in_filtered_species, "IN_all_cpue_filtered.csv")
```

