---
title: "c_MI_Filtering_for_CPUE"
author: "Denver Link"
date: "2023-10-29"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
mi_data <- open_dataset(sources = file.path("G:",
                                            "Shared drives",
                                            "Hansen Lab",
                                            "RESEARCH PROJECTS",
                                            "Fish Survey Data",
                                            "Parquet files", 
                                            "hive_update"), partitioning = c("state")) %>% 
  filter(state == "Michigan")
glimpse(mi_data)
```

#check for effort flags
```{r}
mi_data %>% 
  group_by(flag) %>% 
  count() %>% 
  collect()

mi_data %>% 
  filter(flag == "no effort data") %>% 
  glimpse()

mi_data %>% 
  summarise(n_flag_effort = sum(str_detect(flag, "no effort"), na.rm = T),
            n_invalid = sum(str_detect(flag, "no effort") & cpue_invalid == "y", na.rm = T)) %>% 
  collect()
```
There are no effort flags outside of there being no effort data in a survey. We will not have to worry about filtering those out because they do not have total effort idents and will not show up in our good surveys generation. We could filter out by the column, cpue_invalid but it is an unnessary step here. 

#walleye filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "walleye") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           total_effort_ident,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
 filter(total_effort_1 >= effort_min,
        total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_walleye <- mi_data %>% 
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))


#checking filters after collecting fish
adult_walleye %>% 
  group_by(flag) %>% 
  count()
#no effort flags in the fish doc - we are good

adult_walleye %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method,
           total_effort_1) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_walleye %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count()

#remove larger fish object to save memory
rm(adult_walleye)
```

#black crappie filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "black_crappie") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max,
#           month_min,
#           month_max
           ) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#collecting fish from good surveys
adult_black_crappie <- mi_data %>% 
  #filter the data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1))


#checking filtering after collecting fish
adult_black_crappie %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method,
           total_effort_1) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_black_crappie_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_black_crappie %>% 
  group_by(total_effort_nothing_caught, species_1) %>% 
  count()

#removing larger fish object to save memory
rm(adult_black_crappie)
```

#bluegill filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "bluegill") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           total_effort_ident,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
 filter(total_effort_1 >= effort_min,
        total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#collecting fish from good surveys
adult_bluegill <- mi_data %>% 
  #filter the data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1))


#checking final filter
adult_bluegill %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1) %>% 
  group_by(sampling_method_2) %>% 
  glimpse()

#range of cpues?
adult_bluegill_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_bluegill_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(sampling_method_2, cpue) %>% 
  count()

adult_bluegill %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count()

#removing large object to save memory
rm(adult_bluegill)
```

#cisco filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "cisco") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
  filter(total_effort_1>= effort_min,
         total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

adult_cisco <- mi_data %>% 
  #filter the data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'cisco')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1))




#checking final filter
adult_cisco %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_cisco %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_cisco %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_cisco %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1) %>% 
  group_by(sampling_method_2) %>% 
  glimpse()

#range of cpues?
adult_cisco_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_cisco_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(sampling_method_2, cpue) %>% 
  count()

adult_cisco %>% 
  group_by(total_effort_nothing_caught, species_1) %>% 
  count()

#removing large fish object to save memory
rm(adult_cisco)
```

#largemouth filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "largemouth_bass") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           total_effort_ident,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
 filter(total_effort_1 >= effort_min,
        total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_largemouth_bass <- mi_data %>% 
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))


#checking filters after collecting fish
adult_largemouth_bass %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth_bass %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method,
           total_effort_1) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_largemouth_bass_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_largemouth_bass_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_largemouth_bass %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count()

#remove larger fish object to save memory
rm(adult_largemouth_bass)
```

#smallmouth filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "smallmouth_bass") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           total_effort_ident,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
 filter(total_effort_1 >= effort_min,
        total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_smallmouth_bass <- mi_data %>% 
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))


#checking filters after collecting fish
adult_smallmouth_bass %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth_bass %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method,
           total_effort_1) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_smallmouth_bass_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_smallmouth_bass_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_smallmouth_bass %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count()

#remove larger fish object to save memory
rm(adult_smallmouth_bass)
```

#northern pike filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "northern_pike") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           total_effort_ident,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
 filter(total_effort_1 >= effort_min,
        total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_northern_pike <- mi_data %>% 
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))


#checking filters after collecting fish
adult_northern_pike %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method,
           total_effort_1) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_northern_pike_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_northern_pike %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count()

#remove larger fish object to save memory
rm(adult_northern_pike)
```

#yellow perch filtering 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Michigan" & species_1 == "yellow_perch") %>% 
  select(-metric,
         -species_1,
         -area_group,
         -month_min,
         -month_max,
         -water_temp_min,
         -water_temp_max,
         -area_group) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- mi_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "survey_type",
                                         "sampling_method_2")) %>% 
  distinct(lake_id,
           lake_name,
           total_effort_ident,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method_2,
           total_effort_1,
           effort_min,
           effort_max
#           month_min,
#           month_max
           ) %>% 
 filter(total_effort_1 >= effort_min,
        total_effort_1 <= effort_max) %>% 
#         month(date_clean) >= month_min,
#         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_yellow_perch <- mi_data %>% 
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) 

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))


#checking filters after collecting fish
adult_yellow_perch %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(lake_id,
           lake_name,
           date_total_effort_ident,
           survey_id,
           year,
           survey_type,
           sampling_method,
           total_effort_1) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method_2) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_yellow_perch_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_yellow_perch %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count()

#remove larger fish object to save memory
rm(adult_yellow_perch)
```

#combining species
```{r}
mi_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_cisco_cpue %>% 
                                   mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(mi_filtered_species, "MI_all_cpue_filtered.csv")
```
