---
title: "c_MN_Filtering_for_CPUE"
author: "Denver Link"
date: "2023-10-28"
output: html_document
editor_options: 
  chunk_output_type: console
---

12/29 filter cleaning:
1. selecting columns to retain for filtering... is that necessary? 
2. order of operation in creation of good surveys
      -could call the distinct function first
      -then I could join by filter table by specific columns and values (even numeric)
      -collect
      -could minimize the number of things I call to make a survey distinct (only include total effort ident)
3. I have realized filtering by fish level parameters is not the best way to go
      -instead, if a fish level parameter is of intrest - include that variable in counting for cpue
4. only use total effort ident in the grouping for cpue
5. change the call of na species to nothing caught
6. collect first before calling the distinct function
  -that way we can retain all columns with .keep_all = T



#Library
```{r}
library(tidyverse)
library(arrow)
```

#Data
```{r}
mn_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "mn_file_arrow"))

glimpse(mn_data)
```

General structure for each filtering chunks:
1. gather the filter tables (survey level and fish level)
2. filter for "good surveys"
3. gather fish data from the good surveys
4. calculate cpue - ensuring you are counting fish of the correct parameters
   -if fish level counting within certain parameters is desired (bin lengths) apply that here
5. check survey level filtering
6. check filtering post gathering fish
7. check cpue generation
8. remove large individual fish object to save memory 

#walleye filtering
```{r}
#filters for metric and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "walleye") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

#filters for surveys that are good measures of abundance for given fish species
#first, merge the filter table with the data - selects for non-numeric parameters important for abundance 
#second, collapse the data into survey level
#third, filter the surveys based on numeric values 
good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

#gather fish from the surveys deemed good for abundance of the given species
adult_walleye <- mn_data %>% 
  #filter the data for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#create cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by survey level parameters to get survey level counts
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()


#checking filtering

#survey level checks
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))


#checks after collecting all fish data

#checking final filter
adult_walleye %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue

#range of cpues?
adult_walleye_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

adult_walleye %>% 
  group_by(species.1, nothing_caught) %>% 
  count() %>% 
  print(n = nrow(.))

#removes individual fish object to save memory
rm(adult_walleye)
```

#black crappie filtering
```{r}
#filters for metric and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "black_crappie") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_black_crappie <- mn_data %>% 
  #gathering fish surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()


#checking filtering

#survey level checks
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))
#there is only 1 lake in the <100 area grouping that has an effort of 4
#this explains why the min effort for this group is not the same as the min we set
good_surveys %>% 
  filter(sampling_method == "Standard gill nets, set shallow in stratified assessment" & area_group == "<100")



#checks after gathering all fish
adult_black_crappie %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_black_crappie_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_black_crappie_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#removing large fish level object to save memory
rm(adult_black_crappie)
```

#bluegill filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "bluegill") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

#gathering fish from good surveys
adult_bluegill <- mn_data %>% 
  #gathering fish from surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()

#checking survey level filterings
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))
#there is only 1 lake in the <100 area grouping that has an effort of 4
#this explains why the min effort for this group is not the same as the min we set
good_surveys %>% 
  filter(sampling_method == "Standard gill nets, set shallow in stratified assessment" & area_group == "<100")


#checking filtering after collecting fish
adult_bluegill %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_bluegill_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_bluegill_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#removing large fish data object to save memory
rm(adult_bluegill)
```

#largemouth filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "largemouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)


good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

adult_largemouth <- mn_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect()

#creating cpue for species of interest
adult_largemouth_cpue <- adult_largemouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()

#checking survey level filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))
#there is only 1 lake in the <100 area grouping that has an effort of 4
#this explains why the min effort for this group is not the same as the min we set
good_surveys %>% 
  filter(sampling_method == "Standard gill nets, set shallow in stratified assessment" & area_group == "<100")

 
#checks after gathering fish
adult_largemouth %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_largemouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_largemouth_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#remove fish object to save memory
rm(adult_largemouth)
```

#smallmouth filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

adult_smallmouth <- mn_data %>% 
  #gather fish from surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_smallmouth_cpue <- adult_smallmouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()

#survey level checks
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))
#there is only 1 lake in the <100 area grouping that has an effort of 4
#this explains why the min effort for this group is not the same as the min we set
good_surveys %>% 
  filter(sampling_method == "Standard gill nets, set shallow in stratified assessment" & area_group == "<100")


#checks after gathering fish
adult_smallmouth %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_smallmouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_smallmouth_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#remove large fish object to save memory
rm(adult_smallmouth)
```

#cisco filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "cisco") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

adult_cisco <- mn_data %>% 
  #gathering fish from surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'cisco')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()

#checking survey level filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))



#checks after gathering fish data
adult_cisco %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_cisco %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_cisco %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_cisco %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_cisco_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_cisco_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#remove fish object to save memory
rm(adult_cisco)
```

#northern pike filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "northern_pike") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

adult_northern_pike <- mn_data %>% 
  #gather fish from surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()

#survey level checks
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))
#this explains why the min effort for this group is not the same as the min we set
good_surveys %>% 
  filter(sampling_method == "Standard gill nets, set shallow in stratified assessment" & area_group == "<100")


#checks after gathering fish data
adult_northern_pike %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_northern_pike_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#remove large fish object to save memory
rm(adult_northern_pike)
```

#yellow perch filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Minnesota" & species_1 == "yellow_perch") %>% 
  rename(survey_type.1 = survey_type) %>% 
  select(-metric,
         -species_1,
         -water_temp_min,
         -water_temp_max,
         -target_species_1,
         -target_species_2,
         -survey_type_effort,
         -length_min,
         -length_max)

good_surveys <- mn_data %>% 
  right_join(filter_table_effort, by = c("state",
                                        "survey_type.1",
                                        "sampling_method",
                                        "area_group")) %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min,
           month_min,
           month_max) %>% 
  filter(total_effort_1.1 >= effort_min,
         month(date_clean) >= month_min,
         month(date_clean) <= month_max) %>% 
  collect()

adult_yellow_perch <- mn_data %>% 
  #gather fish from surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species.1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species.1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1.1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(lake_id,
           lake_name.1,
           nhdhr.id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           nothing_caught,
           area_group,
           effort_min,
           count,
           cpue) %>% 
  collect()

#survey level checks
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#survey purpose
good_surveys %>% 
  group_by(survey_type.1) %>% 
  count()

#do I have all area groupings?
good_surveys %>% 
  group_by(sampling_method, area_group) %>% 
  count()

#min effort by grouping
good_surveys %>% 
  group_by(area_group, sampling_method) %>% 
  summarise(min_effort = min(total_effort_1.1))
#this explains why the min effort for this group is not the same as the min we set
good_surveys %>% 
  filter(sampling_method == "Standard gill nets, set shallow in stratified assessment" & area_group == "<100")



#checks after gathering fish data
adult_yellow_perch %>% 
  group_by(species.1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_clean)), max.month = max(month(date_clean)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(lake_id,
           date_clean,
           survey_type.1,
           sampling_method,
           total_effort_1.1,
           effort_units.1,
           area_group,
           effort_min) %>% 
  group_by(sampling_method) %>% 
  glimpse()


#checks after calculating cpue
#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

#surveys that have nothing caught?
adult_yellow_perch_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#removing large fish object to retain memory
rm(adult_yellow_perch)
```

#combining species
```{r}
mn_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_cisco_cpue %>% 
                                   mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(mn_filtered_species, "MN_all_cpue_filtered.csv")
```

