---
title: "c_SD_Filtering_CPUE"
author: "Denver Link"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

to do (12/4):
1. figure out how to count nothing caught surveys
   -our nothing caught is at the sub-survey level and causes issue counting for cpe when a survey has sub levels where some caught something and some caught nothing
2. figure out if we have cisco data - if not remove cisco filter

#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
filter_table <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(state == "South Dakota")

sd_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "sd_file_arrow"))
glimpse(sd_data)

sd_data %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  collect() %>% 
  arrange(desc(n)) %>% 
  print(n = nrow(.))

sd_data %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  collect() %>% 
  arrange(desc(n)) %>% 
  print(n = nrow(.))

sd_data %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1)) %>% 
  collect() %>% 
  print(n = nrow(.))
#looks like nets are in net night and efishing is in minutes

sd_data %>% 
  filter(nothing_caught == "TRUE") %>% 
  collect() %>% 
  distinct(total_effort_ident, .keep_all = T) %>% 
  group_by(sampling_method) %>% 
  summarise(effort.min = min(total_effort_ident),
            date.min = min(month(date_1)),
            date.max = max(month(date_1)),
            count = n()) %>% 
  collect() %>% 
  print(n=nrow(.))

data <- sd_data %>% 
  filter(total_effort_ident == 1 & species_1 == "yellow_perch") %>% 
  collect()

sd_data %>% 
  group_by(species_1) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))
```

#filtering walleye
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "walleye") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "walleye") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
          effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

check <- sd_data %>% 
  filter(total_effort_ident == 2 & species_1 == "walleye") %>% 
  collect()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1))

adult_walleye <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

adult_walleye %>% 
  group_by(month(date_1)) %>% 
  count()

#checking final filter
adult_walleye %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(total_effort_1), max = max(total_effort_1), mean = mean(total_effort_1)) 


#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()
#this includes subsurvey level

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 
  
#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_walleye_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_walleye)
```

#filtering black crappie
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "black_crappie") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "black_crappie") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
          effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_black_crappie <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_black_crappie %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_black_crappie_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_black_crappie_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_black_crappie)
```

#filtering bluegill
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "bluegill") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "bluegill") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
           effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_bluegill <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_bluegill %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_bluegill_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_bluegill_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_bluegill)
```

#filtering largemouth bass
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "largemouth_bass") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "largemouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
          effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_largemouth_bass <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_largemouth_bass %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth_bass %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_largemouth_bass_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_largemouth_bass_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_largemouth_bass)
```

#filtering smallmouth bass
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "smallmouth_bass") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
           effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_smallmouth_bass <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_smallmouth_bass %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_smallmouth_bass %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth_bass %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_smallmouth_bass_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_smallmouth_bass_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_smallmouth_bass)
```

#filtering northern pike
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "northern_pike") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "northern_pike") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
           effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_northern_pike <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_northern_pike %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_northern_pike_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_northern_pike_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_northern_pike)
```

#filtering yellow perch
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "yellow_perch") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "yellow_perch") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
          effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_yellow_perch <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_yellow_perch %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_yellow_perch_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_yellow_perch_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_yellow_perch)
```

#filtering cisco
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "cisco") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -survey_type_effort,
         -target_species_1,
         -target_species_2,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -area_group,
         -length_min,
         -length_max)

#selects fish level filters to be applied in the second filtering step
filter_table_fish <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "cisco") %>% 
  rename(survey_type.1 = survey_type) %>% 
  #fish level parameters of interest 
  select(state) %>% 
  distinct() 

#date is not a good grouping variable for survey because some surveys span multiple dates 
good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("state",
                                         "sampling_method")) %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units,
           effort_min,
          month_min,
          month_max,
          nothing_caught) %>% 
  filter(total_effort_1 >= effort_min,
         month(date_1) >= month_min,
        month(date_1) <= month_max) %>% 
  collect() %>% 
  group_by(total_effort_ident) %>% 
  #removing sub-survey level (nets/runs) that did not catch fish but there were fish caught in the total effort ident
  mutate(mixed.bag = n_distinct(nothing_caught) > 1) %>% 
  filter(!(mixed.bag == TRUE & nothing_caught == TRUE)) %>% 
  #only keep one record of total effort idents that caught nothing in all nets
  slice(if (any(nothing_caught)) 1 else 1:n()) %>% 
  ungroup()

#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()


adult_cisco <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  #filter for fish level items (length, etc)
  right_join(filter_table_fish) %>% 
  collect() 

#checking final filter
adult_cisco %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_cisco %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_cisco %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_cisco %>% 
  distinct(lake_name_1,
           total_effort_ident,
           date_1,
           sampling_method,
           total_effort_1,
           total_effort_1_units) %>% 
  glimpse()

adult_cisco %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
  
#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'cisco')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T) %>% 
  collect() %>% 
  select(state, 
         county, 
         lake_name_1,
         lake_id,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         nothing_caught,
         count,
         cpue)

#range of cpues?
adult_cisco_cpue %>%
  filter(!is.na(cpue)) %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_cisco_cpue %>% 
  filter(nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_cisco)
```

#combining species
```{r}
sd_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 #adult_cisco_cpue %>% 
                                   #mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(sd_filtered_species, "SD_all_cpue_filtered.csv")
```
