---
title: "c_SD_Filtering_CPUE"
author: "Denver Link"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---


#Library
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
sd_data <- open_dataset(sources = file.path("G:",
                                            "Shared drives",
                                            "Hansen Lab",
                                            "RESEARCH PROJECTS",
                                            "Fish Survey Data",
                                            "Parquet files",
                                            "hive_update"), partitioning = c("state")) %>% 
  filter(state == "South_Dakota")
glimpse(sd_data)
```

#effort flag check
```{r}
sd_data %>% 
  group_by(flag) %>% 
  count() %>% 
  collect()

sd_data %>% 
  group_by(flag, cpue_invalid) %>% 
  count() %>% 
  collect()
```
Looks like we have some high effort surveys that we need to be sure to remove, this is acccomplished by the cpue invalid column

#filtering walleye
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "walleye") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_walleye <- sd_data %>% 
  right_join(good_surveys) %>% 
  collect() %>%
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variables so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#the difference between good surveys and adult walleye cpe is that the surveys where total effort ident spans into November or december are filtered out
  
#range of cpues?
adult_walleye_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

adult_walleye %>% 
  filter(total_effort_ident == "1550") %>% 
  group_by(total_effort_nothing_caught, sub_effort_nothing_caught) %>% 
  count()
#this gives a record of a total effort of catching something but a sub effort did not

adult_walleye_cpue %>% 
  filter(total_effort_ident == "1550") %>% 
  glimpse()
#this shows we were able to generate a good cpe even tho the effort ident had a sub that caught nothing

#surveys that have nothing caught?
adult_walleye_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

#exploring cases of sub efforts spanning the minimum or maximum ranges 
sd_data %>% 
  filter(lake_name == "Francis Case" & year == 2021) %>% 
  group_by(month(date_total_effort_ident),
           month(date_sub_effort_ident),
           total_effort_ident) %>% 
  count() %>% 
  collect()

adult_walleye %>% 
  filter(total_effort_ident == "2")
#We see here that in total effort ident 2, the total effort spans months 3, 8, and 8. Month is created by taking the first month from the total effort ident. Thus, all sub efforts are filtered out here.

sd_data %>% 
  filter(lake_name == "Poinsett" & year == 2016) %>% 
  group_by(total_effort_ident, 
           month(date_total_effort_ident), 
           month(date_sub_effort_ident)) %>% 
  count() %>% 
  collect()

adult_walleye_cpue %>% 
  filter(total_effort_ident == "3831") %>% 
  group_by(total_effort_ident, 
           month(date_total_effort_ident), 
           month(date_sub_effort_ident)) %>% 
  count()
#filtered out here 

adult_walleye %>% 
  mutate(sub_month = month(date_sub_effort_ident)) %>% 
  distinct(total_effort_ident, sub_month) %>%
  group_by(total_effort_ident) %>% 
  filter(n_distinct(sub_month) > 1, any(sub_month > 10)) %>% 
  ungroup()
#no cases of overlap here 

sd_data %>% 
  filter(total_effort_ident == "2724") %>% 
  group_by(total_effort_ident, 
           month(date_total_effort_ident),
           month(date_sub_effort_ident),
           species_1) %>% 
  count() %>% 
  collect()

adult_walleye %>% 
  filter(total_effort_ident == "2724")

rm(adult_walleye)
```

#filtering black crappie
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "black_crappie") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_black_crappie <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()
#this includes subsurvey level

adult_black_crappie %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_black_crappie %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_black_crappie_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_black_crappie_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_black_crappie)
```

#filtering bluegill
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "bluegill") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_bluegill <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_bluegill %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_bluegill_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()


#surveys that have nothing caught?
adult_bluegill_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_bluegill)
```

#filtering largemouth bass
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "largemouth_bass") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_largemouth_bass <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_largemouth_bass_cpue <- adult_largemouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_largemouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth_bass %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_largemouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_largemouth_bass %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_largemouth_bass_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_bass_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()


#surveys that have nothing caught?
adult_largemouth_bass_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_largemouth_bass)
```

#filtering smallmouth bass
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "smallmouth_bass") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_smallmouth_bass <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_smallmouth_bass_cpue <- adult_smallmouth_bass %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_smallmouth_bass %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth_bass %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_smallmouth_bass %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_smallmouth_bass %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_smallmouth_bass_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_bass_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()


#surveys that have nothing caught?
adult_smallmouth_bass_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_smallmouth_bass)
```

#filtering northern pike
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "northern_pike") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_northern_pike <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()
#this includes subsurvey level

adult_northern_pike %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_northern_pike %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_northern_pike_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()


#surveys that have nothing caught?
adult_northern_pike_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_northern_pike)
```

#filtering yellow perch
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "yellow_perch") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_yellow_perch <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()
#this includes subsurvey level

adult_yellow_perch %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_yellow_perch %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_yellow_perch_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_yellow_perch_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_yellow_perch)
```

#filtering cisco
*returns no cisco catch because cisco are not an included speices* 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "South Dakota" & species_1 == "cisco") %>% 
  select(-metric,
         -species_1,
         -survey_type,
         -area_group,
         -water_temp_min,
         -water_temp_max,
         -state) %>% 
  rename(sampling_method_2 = sampling_method)

good_surveys <- sd_data %>% 
  right_join(filter_table_effort, by = c("sampling_method_2")) %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method_2,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max,
          cpue_invalid) %>% 
  filter(total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         month(date_total_effort_ident) >= month_min,
        month(date_total_effort_ident) <= month_max,
        cpue_invalid == "n") %>% 
  collect() 

adult_cisco <- sd_data %>% 
  #filter for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  collect() %>% 
  #remove teis that have a good starting month but a sub extends outside of the range
  mutate(sub_month = month(date_sub_effort_ident)) %>%
  group_by(total_effort_ident) %>%
  mutate(
    multi_month = n_distinct(sub_month) > 1,
    has_late_month = any(sub_month > 10),
    bad_group = multi_month & has_late_month
  ) %>%
  ungroup() %>%
  filter(!bad_group)

#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #creating non na species 1 variabesl so that no na cpe values are returned for sub effort nothing caught true
  mutate(species_1 = case_when(is.na(species_1) ~ "na_taxon",
                               TRUE ~ species_1)) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'cisco')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue (also works as a select function)
  distinct(total_effort_ident,
           .keep_all = T)


#checking filtering
#month range
good_surveys %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#is each sampling method correct for the filter?
adult_cisco %>% 
  group_by(sampling_method_2) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min = min(total_effort_1),
            max = max(total_effort_1),
            n = n())

#do I retain all of the good surveys? - rows here should match good surveys 
adult_cisco %>% 
  distinct(total_effort_ident,
           date_total_effort_ident,
           sampling_method,
           total_effort_1,
          effort_min,
          effort_max,
          month_min,
          month_max) %>% 
  glimpse()
#this includes subsurvey level

adult_cisco %>% 
  distinct(total_effort_ident,
           date_total_effort_ident) %>% 
  glimpse()

adult_cisco %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#survey level records 

adult_cisco %>% 
  group_by(flag) %>% 
  count()
  
#range of cpues?
adult_cisco_cpue %>%
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue)) %>% 
  glimpse()

#surveys that have nothing caught?
adult_cisco_cpue %>% 
  filter(total_effort_nothing_caught == "TRUE") %>% 
  group_by(cpue) %>% 
  count()

rm(adult_cisco)
```

#combining species
```{r}
sd_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_bass_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_bass_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 #adult_cisco_cpue %>% 
                                   #mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(sd_filtered_species, "SD_all_cpue_filtered.csv")
```

