---
title: "Filtering for CPUEs"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Libraries 
```{r}
library(arrow)
library(tidyverse)
```

# Data
```{r}
wi_data <- open_dataset(sources = file.path("G:",
                                            "Shared drives",
                                            "Hansen Lab",
                                            "RESEARCH PROJECTS",
                                            "Fish Survey Data",
                                            "Parquet files",
                                            "hive_update"), partitioning = c("state")) %>% 
  filter(state == "Wisconsin")
glimpse(wi_data)
```

#General steps for filtering
all steps below happen with specific parameters for each species 
1. load in filtering table 
2. filter for good surveys based on species specific conditions
     -in WI its important to ensure that the species are targeted within the survey
     -a binary (Y/N) columns is created to keep track if a species is targeted based
     -a filter table join makes sure the correct sampling methods are used and imports numeric filtering values
     -data is collapsed into survey-level information
     -values are used to filter for month, water_temp, minimum effort, and if the species are targeted in the survey
3. gather all fish records from the good surveys
4. calculate cpe for surveys
    -use survey grouping variable and sum all of the species of interest in a given survey 
       (additional fish level constraints can be used here (ex. length bins))
    -ensure surveys that caught nothing are counted as 0 cpe
    -collapse data into survey level and ensure you retain columns of interest
5. Run checks on all previous steps

#targeted species
```{r}
wi_data %>% 
  group_by(target_species) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

wi_data %>% 
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "yellow_perch") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "escocid_percid") ~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        TRUE ~ "n")) %>% 
  group_by(target_species, targeted_in_survey) %>% 
  count() %>% 
  arrange(targeted_in_survey, target_species) %>% 
  collect() %>% 
  print(n = nrow(.))
```
Wisconsin is heavy on the targeted sampling, so we need to be savvy in picking the surveys that would be valid for abundance measures. See the example above that will be applied to in various forms to all species filtering below. This code creates a yes or no if we should keep that survey for a cpe calculation based on if the species was targeted. We have to be a little creative with the string detection because we want to collect "gamefish" and "gamefish_species" but we DONT want to collect "non-gamefish" if searching for yellow perch. Note that we are more picky here, we do not include if the the survey is NA for targeted species - a Wisconsin type of thing were we cannot just assume that a survey would be good to measure walleye cpe if it were not stated 

#effort flags
```{r}
wi_data %>% 
  group_by(flag) %>% 
  count() %>% 
  collect() %>% 
  arrange(flag) %>% 
  print(n = nrow(.))

wi_data %>% 
  distinct(cpue_invalid, flag) %>% 
  arrange(cpue_invalid) %>% 
  collect() %>% 
  print(n = nrow(.))
```
We see here that the cpue invalid filtering will keep only good abundance measures 


#effort units
```{r}
wi_data %>% 
  group_by(sampling_method_simple, total_effort_1_units) %>% 
  count() %>% 
  collect() %>% 
  arrange(sampling_method_simple) %>% 
  print(n = nrow(.))
#it looks like for any given gear there could be multiple types of effort reported
#we will need to make sure this is consistent

wi_data %>% 
  distinct(total_effort_ident, sampling_method, total_effort_1_units) %>% 
  filter(sampling_method %in% c("fyke_net") | str_detect(sampling_method, "boat_electrofishing")) %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count() %>%
  arrange(sampling_method) %>% 
  collect()

wi_data %>% 
  filter(sampling_method %in% c("fyke_net") | str_detect(sampling_method, "boat_electrofishing")) %>% 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  group_by(sampling_method, total_effort_1_units, units_flag) %>% 
  count() %>%
  arrange(sampling_method) %>% 
  collect()
```
For fyke nets it is really obvious that it is standard to report in net_nights and we will proceed only that for that gear. For electrofishing the most used unit is minutes, we will need to make sure that only minutes show up for electrofishing and only hours for fykes.

#Walleye filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "walleye") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "walleye") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "(^|,)(gamefish|gamefish_species)(,|$)") ~ "y",
                                        str_detect(target_species, "escocid_percid") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>% 
  #removing na water temps
  filter(!is.na(water_temp)) %>% 
  #remove surveys without total effort ident
  filter(!is.na(total_effort_ident)) %>% 
  #first join the filter table data to the Wisconsin data
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>% 
    
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag)  %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect() 

#checking why we have dup total effort idents
more <- good_surveys %>% 
  group_by(total_effort_ident) %>% 
  count()  %>% 
  filter(n >1)

good_surveys %>% 
  filter(total_effort_ident %in% more$total_effort_ident)  %>% 
  group_by(total_effort_ident) %>% 
  summarise(n_temps = n_distinct(water_temp)) %>% 
  ungroup() %>% 
  summarise(n_more = sum(n_temps>=2),
            n_one = sum(n_temps < 2))
rm(more)
#all duplicate total effort idents in good surveys comes from water temps

good_surveys %>% 
  distinct(total_effort_ident)  %>% 
  count()

good_surveys %>% 
  distinct(total_effort_ident, water_temp) %>% 
  count()

#second step of filtering 
##gathering good records associated with "good surveys" 
adult_walleye <- wi_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>%
  collect()

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T) 

#survey level filtering checks
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)))


#checks after collecting fish back
#what is the survey good for?
adult_walleye %>% 
  distinct(total_effort_ident, survey_type) %>% 
  group_by(survey_type) %>% 
  count() %>% 
  print(n = nrow(.))

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min.temp = min(water_temp, na.rm = T), 
            max.temp = max(water_temp, na.rm = T))

adult_walleye %>% 
  group_by(target_species, targeted_in_survey) %>% 
  count() %>% 
  print(n = nrow(.))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey)

adult_walleye %>% 
  distinct(date_total_effort_ident,
    total_effort_ident) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()
#the difference between good surveys and cpe obs is there being multiple water temps within a total effort ident 


#checks after calculating cpe
adult_walleye_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

adult_walleye_cpue %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count()

#removing large fish file to save memory
rm(adult_walleye)
```

#black crappie filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "black_crappie") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  #note that the targeted species here does not included a "non-gamefish,panfish" - I moved on as it was a limited number of obs 
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "centrarchid") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "black_crappie")~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        str_detect(target_species, "panfish") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  #removing na water temps
  filter(!is.na(water_temp)) %>% 
  #remove surveys without total effort ident
  filter(!is.na(total_effort_ident)) %>% 
  #first join the filter table dta to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect() 

good_surveys %>% 
  distinct(total_effort_ident) %>% 
  count()
#1627

good_surveys %>% 
  distinct(total_effort_ident, water_temp) %>% 
  count()
#all extra obs are attributed to multiple water temps within a total effort ident

#second step of filtering 
##gathering good records associated with "good surveys" 
adult_black_crappie <- wi_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>%
  collect()

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
   group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T)

#checking survey level filtering
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)))


#checks after collecting fish back
#what is the survey good for?
adult_black_crappie %>% 
  group_by(survey_type) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min.temp = min(water_temp, na.rm = T), 
            max.temp = max(water_temp, na.rm = T))

adult_black_crappie %>% 
  group_by(target_species, targeted_in_survey) %>% 
  count() %>% 
  print(n = nrow(.))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident,water_temp) %>% 
  glimpse()

#checks after calculating cpe
adult_black_crappie_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

adult_black_crappie_cpue %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count()

#removing large fish file to save memory
rm(adult_black_crappie)
```

#bluegill filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "bluegill") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type) 

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "centrarchid") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "bluegill")~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        str_detect(target_species, "panfish") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>% 
  #removing na water temps
  filter(!is.na(water_temp)) %>% 
  #remove surveys without total effort ident
  filter(!is.na(total_effort_ident)) %>% 
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         !is.na(water_temp) & water_temp >= water_temp_min,
         !is.na(water_temp) & water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

#duplicates of total effort idents?
good_surveys %>% 
  group_by(total_effort_ident) %>% 
  count() 

good_surveys %>% 
  group_by(total_effort_ident, water_temp) %>% 
  count()

#gathering fish from good surveys
adult_bluegill <- wi_data %>% 
  right_join(good_surveys) %>%
  collect() 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T)


#filtering checks

good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)))


#checks after collecting fish back
#what is the survey good for?
adult_bluegill %>% 
  group_by(survey_type) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1),
            min.month = min(month(date_total_effort_ident)), 
            max.month = max(month(date_total_effort_ident)),
            min.temp = min(water_temp, na.rm = T), 
            max.temp = max(water_temp, na.rm = T))

adult_bluegill %>% 
  distinct(total_effort_ident, target_species, targeted_in_survey) %>% 
  group_by(target_species, targeted_in_survey) %>% 
  count() %>% 
  print(n = nrow(.))

adult_bluegill %>% 
  group_by(total_effort_ident) %>% 
  summarise(n_gears = n_distinct(sampling_method)) %>% 
  filter(n_gears >1)
#this shows no multi gears

#do I retain all of the good surveys? - rows here should match good surveys 
good_surveys %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(
    total_effort_ident,
    water_temp) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident,
           sampling_method,
           water_temp) %>% 
  count()


#checks after calculating cpe
adult_bluegill_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

adult_bluegill_cpue %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count()

rm(adult_bluegill, good_surveys, filter_table_effort)
```

#largemouth filtering
Per correspondence with agency folks, bass in Wisconsin are targeted in spring and fall sampling. The first filtering represents standard sampling where bass would be counted. The second is represented of more targeted surveys for bass. We filter separately because we have different thresholds of temp and effort for the surveys in the fall for the same sampling methods 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort.standard <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "largemouth_bass") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type) %>% 
  filter(water_temp_min != 0)

filter_table_effort.extra <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "largemouth_bass") %>%
  select(-area_group,
         -metric,
         -species_1,
         -survey_type) %>% 
  filter(water_temp_min == 0)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys_standard <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "centrarchid") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "largemouth_bass")~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        str_detect(target_species, "basses") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  filter(!is.na(water_temp)) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.standard, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

good_surveys_extra <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "centrarchid") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "largemouth_bass")~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        str_detect(target_species, "basses") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  filter(!is.na(water_temp)) %>% 
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.extra, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

#bind surveys
good_surveys <- bind_rows(good_surveys_extra, good_surveys_standard)

#duplicate total effort idents?
good_surveys %>% 
  distinct(total_effort_ident) %>% 
  count()

good_surveys %>% 
  distinct(total_effort_ident, water_temp) %>% 
  count()
#all attributed to water temp

 #gathering fish data from good surveys
adult_largemouth <- wi_data %>% 
   right_join(good_surveys) %>%
  collect()

#creating cpue for species of interest
adult_largemouth_cpue <- adult_largemouth %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T)


#checking survey level filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

good_surveys %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(sampling_method, year(date_total_effort_ident)) %>% 
  count() %>% 
  print(n = nrow(.))

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()

adult_largemouth %>% 
  distinct(total_effort_ident, target_species) %>% 
  group_by(target_species) %>% 
  count() %>% 
  print(n = nrow(.))

adult_largemouth %>%
  filter(sampling_method_simple == "boat_electrofishing") %>% 
  distinct(total_effort_ident, survey_type) %>% 
  group_by(survey_type) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  print(n = nrow(.))

#water temp range by sampling method
adult_largemouth %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T), min.effort = min(total_effort_1), max = max(total_effort_1)) %>% 
  print(n = nrow(.))

#do I retain all of the good surveys? - rows here should match good surveys 
good_surveys %>% 
  distinct(total_effort_ident) %>% 
  count()

adult_largemouth %>% 
  distinct(total_effort_ident,
    sampling_method,
    water_temp)  %>% 
  count()

adult_largemouth %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#checks after calculating cpe
adult_largemouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_cpue %>% 
  filter(is.na(cpue))

adult_largemouth %>% 
  group_by(species_1, total_effort_nothing_caught) %>%
  count() %>% 
  filter(total_effort_nothing_caught == "TRUE")

adult_largemouth_cpue %>% 
  group_by(sampling_method_simple, total_effort_1_units) %>% 
  count()

#removing large object to save memory
rm(adult_largemouth)
```

#smallmouth filtering
Per correspondence with agency folks, bass in Wisconsin are targeted in spring and fall sampling. The first filtering represents standard sampling where bass would be counted. The second is represented of more targeted surveys for bass. We filter separately because we have different thresholds of temp and effort for the surveys in the fall for the same sampling methods 
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort.standard <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "smallmouth_bass") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type) %>% 
  filter(water_temp_min != 0)

filter_table_effort.extra <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "smallmouth_bass") %>%
  select(-area_group,
         -metric,
         -species_1,
         -survey_type) %>% 
  filter(water_temp_min == 0)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys_standard <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "centrarchid") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "smallmouth_bass")~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        str_detect(target_species, "basses") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  filter(!is.na(water_temp)) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.standard, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

good_surveys_extra <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "centrarchid") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "smallmouth_bass")~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        str_detect(target_species, "basses") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  filter(!is.na(water_temp)) %>% 
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.extra, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

good_surveys <- bind_rows(good_surveys_extra, good_surveys_standard)

good_surveys %>% 
  distinct(total_effort_ident) %>% 
  count()

good_surveys %>% 
  distinct(total_effort_ident,
           water_temp) %>% 
  count()
#duplicate total efforts accounted for by water temp and sampling method

#gathering fish data from good surveys
adult_smallmouth <- wi_data %>% 
   right_join(good_surveys) %>%
  collect()

#creating cpue for species of interest
adult_smallmouth_cpue <- adult_smallmouth %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T)


#checking survey level filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

good_surveys %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()


adult_smallmouth %>% 
  group_by(target_species) %>% 
  count() %>% 
  print(n = nrow(.))

#water temp range by sampling method
adult_smallmouth %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T), min.effort = min(total_effort_1), max = max(total_effort_1)) %>% 
  print(n = nrow(.))

#do I retain all of the good surveys? - rows here should match good surveys 
good_surveys %>% 
  distinct(total_effort_ident)

adult_smallmouth %>% 
  distinct(sampling_method,
    total_effort_ident,
    water_temp) %>% 
  count()


adult_smallmouth %>% 
  distinct(total_effort_ident) %>% 
  count()

#checks after calculating cpe
adult_smallmouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_cpue %>% 
  filter(is.na(cpue))

adult_smallmouth %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count() %>% 
  filter(total_effort_nothing_caught == "TRUE")

adult_smallmouth_cpue %>% 
  group_by(sampling_method_simple, total_effort_1_units) %>% 
  count()

#removing large object to save memory
rm(adult_smallmouth, good_surveys_extra, good_surveys_standard, filter_table_effort.extra, filter_table_effort.standard)
```

#cisco filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "cisco") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "cisco") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        is.na(target_species) ~ "y",
                                        TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  mutate(units_flag = case_when(sampling_method == "gill_net_vertical" & total_effort_1_units == "net_nights" ~ "y",
                                TRUE ~ "n")) %>%
  #filter(!is.na(water_temp)) %>% 
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    #water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         #water_temp >= water_temp_min,
         #water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()
#ignoring water temp for these surveys due to data structure and limited samples(note limited samples are not due to temperature restriction, it just makes the filtering easier)

#gathering fish from good surveys
adult_cisco <- wi_data %>% 
  right_join(good_surveys) %>%
  collect() 

#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(str_detect(species_1, "cisco"))) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T) 


#checks at the survey level filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#is each sampling method correct for the filter?
adult_cisco %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)),
            max.month = max(month(date_total_effort_ident)),
            na_temp = sum(is.na(water_temp)),
            water_temp_min = min(water_temp, na.rm = T),
            water_temp_max = max(water_temp, na.rm = T))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_cisco %>% 
  distinct(date_total_effort_ident,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max) %>% 
  group_by(sampling_method) %>% 
  glimpse()

adult_cisco %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#checks after calculating cpe
adult_cisco_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue))

adult_cisco_cpue %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count()

#removing large fish object to save memory
rm(adult_cisco)
```

#northern pike filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "northern_pike") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "northern_pike") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "escocid_percid") ~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  filter(!is.na(water_temp)) %>% 
  #first join the filter table data to the Wisconsin data
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey, 
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

#duplicate total effort idents?
good_surveys %>% 
  distinct(total_effort_ident) %>% 
  count()

good_surveys %>% 
  distinct(total_effort_ident,
           water_temp) %>% 
  count()
#all accounted for by water temp

#collecting fish from good surveys
adult_northern_pike <- wi_data %>% 
  right_join(good_surveys) %>%
  collect()

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/total_effort_1) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

good_surveys %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>% 
  count()

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()

adult_northern_pike %>% 
  group_by(target_species) %>% 
  count() %>% 
  print(n = nrow(.))

#water temp range by sampling method
adult_northern_pike %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T), min.effort = min(total_effort_1), max = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys 
good_surveys %>% 
  distinct(total_effort_ident)

adult_northern_pike %>% 
  distinct(total_effort_ident,
    water_temp) %>% 
  count()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#checks after calculating cpe
adult_northern_pike_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

adult_northern_pike %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count() %>% 
  filter(total_effort_nothing_caught == "TRUE")

adult_northern_pike_cpue %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count()

#removing large object to save memory
rm(adult_northern_pike)
```

#yellow perch filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv(file.path("G:",
                                          "Shared drives",
                                          "Hansen Lab",
                                          "RESEARCH PROJECTS",
                                          "Fish Survey Data",
                                          "manuscript_products",
                                          "abundance",
                                          "data",
                                          "input",
                                          "cpue_inclusion_table.csv")) %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "yellow_perch") %>% 
  select(-area_group,
         -metric,
         -species_1,
         -survey_type)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(str_detect(target_species, "non_gamefish") ~ "n",
                                        str_detect(target_species, "yellow_perch") ~ "y",
                                        str_detect(target_species, "all_species") ~ "y",
                                        str_detect(target_species, "escocid_percid") ~ "y",
                                        str_detect(target_species, "gamefish") ~ "y",
                                        TRUE ~ "n")) %>%
  #units check 
  mutate(units_flag = case_when(sampling_method == "fyke_net" & total_effort_1_units == "net_nights" ~ "y",
                                str_detect(sampling_method, "boat_electrofishing") & total_effort_1_units == "minutes" ~ "y",
                                TRUE ~ "n")) %>%
  filter(!is.na(total_effort_ident)) %>% 
  filter(!is.na(water_temp)) %>% 
  #first join the filter table data to the Wisconsin data 
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(date_total_effort_ident,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    effort_max,
    targeted_in_survey,
    cpue_invalid,
    units_flag) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_total_effort_ident) >= month_min,
         month(date_total_effort_ident) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         total_effort_1 <= effort_max,
         targeted_in_survey == "y",
         cpue_invalid == "n",
         units_flag == "y") %>% 
  #select(-water_temp) %>% 
  collect()

#duplicate total effort idents?
good_surveys %>% 
  distinct(total_effort_ident) %>% 
  count()

good_surveys %>% 
  distinct(total_effort_ident,
           sampling_method, 
           water_temp) %>% 
  count()
#duplicates are accounted for by water temp and sampling method 

#collecting fish from good surveys
adult_yellow_perch <- wi_data %>% 
  right_join(good_surveys) %>%
  collect() 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(total_effort_ident) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(total_effort_nothing_caught == "TRUE" ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(total_effort_ident, .keep_all = T) 

#checking filtering
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1),
            max.effort = max(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_total_effort_ident)), max.month = max(month(date_total_effort_ident)))

good_surveys %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>% 
  count()

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()

adult_yellow_perch %>% 
  group_by(target_species) %>% 
  count() %>% 
  print(n = nrow(.))

#water temp range by sampling method
adult_yellow_perch %>% 
  group_by(month = month(date_total_effort_ident), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T), min.effort = min(total_effort_1), max = max(total_effort_1))

#do I retain all of the good surveys? - rows here should match good surveys 
good_surveys %>% 
  distinct(total_effort_ident)

adult_yellow_perch %>% 
  distinct(sampling_method,
    total_effort_ident,
    water_temp) %>% 
  count()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  count()

#checks after calculating cpe
adult_yellow_perch_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

adult_yellow_perch %>% 
  group_by(species_1, total_effort_nothing_caught) %>% 
  count() %>% 
  filter(total_effort_nothing_caught == "TRUE")

adult_yellow_perch_cpue %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count()

#removing large fish object
rm(adult_yellow_perch)
```

#combining species
```{r}
wi_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_cisco_cpue %>% 
                                   mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))
write_csv(wi_filtered_species, "WI_all_cpue_filtered.csv")
```

