---
title: "Filtering for CPUEs"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

1/5 updates:
-everything works great besides the order in which I have to collect the SMB and LMB good surveys

# Libraries 
```{r}
library(arrow)
library(tidyverse)
```

# Data
```{r}
wi_data <- open_dataset(sources = file.path("D:", "Shared drives", "Hansen Lab", "RESEARCH PROJECTS", "Fish Survey Data", "Parquet files", "wi_file_arrow"))
glimpse(wi_data)
```

#General steps for filtering
-all steps below happen with specific parameters for each species 
1. load in filtering table 
2. filter for good surveys based on species specific conditions
     -in WI its important to ensure that the species are targeted within the survey
     -a binary (Y/N) columns is created to keep track if a species is targeted based
     -a filter table join makes sure the correct sampling methods are used and imports numeric filtering values
     -data is collapsed into survey-level information
     -values are used to filter for month, water_temp, minimum effort, and if the species are targeted in the survey
3. gather all fish records from the good surveys
4. calculate cpe for surveys
    -use survey grouping variable and sum all of the species of interest in a given survey 
       (additional fish level constraints can be used here (ex. length bins))
    -ensure surveys that caught nothing are counted as 0 cpe
    -collapse data into survey level and ensure you retain columns of interest
5. Run checks on all previous steps
     

#Walleye filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "walleye") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("walleye", 
                                                                     "all_species", 
                                                                     "gamefish_species", 
                                                                     "escocid_percid") ~ "Y",
                                        secondary_target_species %in% c("walleye", 
                                                                        "all_species", 
                                                                        "gamefish_species", 
                                                                        "escocid_percid") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    nothing_caught,
    targeted_in_survey) %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp) %>% 
  collect() 

#second step of filtering 
##gathering good records associated with "good surveys" 
adult_walleye <- wi_data %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>%
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect()

#creating cpue for species of interest
adult_walleye_cpue <- adult_walleye %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'walleye')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()

#survey level filtering checks
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(nothing_caught) %>% 
  count()

#checks after collecting fish back
#what is the survey good for?
adult_walleye %>% 
  group_by(survey_type_1_indivfish) %>% 
  count()

#is the state label the same for all fish
adult_walleye %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_walleye %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#water temp range by sampling method
adult_walleye %>% 
  group_by(sampling_method) %>%
  summarise(min.temp = min(water_temp), max.temp = max(water_temp))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_walleye %>% 
  distinct(lake_name_1,
    survey_type_1_indivfish,
    survey_type_1_effort,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught) %>% 
  group_by(sampling_method) %>% 
  glimpse()

adult_walleye %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#checks after calculating cpe
adult_walleye_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_walleye_cpue %>% 
  filter(is.na(cpue))

#removing large fish file to save memory
rm(adult_walleye)
```

#black crappie filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "black_crappie") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("panfish", 
                                                                     "all_species", 
                                                                     "gamefish_species", 
                                                                     "panfish_gamefish", 
                                                                     "black_crappie",
                                                                     "centrarchids",
                                                                     "centrarchid_panfish") ~ "Y",
                                        secondary_target_species %in% c("panfish", 
                                                                        "all_species",
                                                                        "gamefish_species", 
                                                                        "panfish_gamefish", 
                                                                        "black_crappie", 
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish") ~ "Y",
                                        TRUE ~ "N")) %>%  
  #first join the filter table dta to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    nothing_caught,
    targeted_in_survey) %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  #do not include surveys without effort
  select(-water_temp) %>% 
  collect() 

#gather fish records from good surveys
adult_black_crappie <- wi_data %>%
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>%
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect() 

#creating cpue for species of interest
adult_black_crappie_cpue <- adult_black_crappie %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'black_crappie')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()


#checking survey level filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#is the state label the same for all fish
adult_black_crappie %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_black_crappie %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#water temp range by sampling method
adult_black_crappie %>% 
  group_by(sampling_method) %>%
  summarise(min.temp = min(water_temp), max.temp = max(water_temp))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_black_crappie %>% 
  distinct(lake_name_1,
    survey_type_1_effort,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    effort_min,
    nothing_caught,
    targeted_in_survey) %>% 
  glimpse()

adult_black_crappie %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#checks for after calculating cpe
adult_black_crappie_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_black_crappie_cpue %>% 
  filter(is.na(cpue))

#removing large fish object to save memory
rm(adult_black_crappie)
```

#bluegill filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "bluegill") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max) 

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("panfish", 
                                                                     "all_species", 
                                                                     "gamefish_species", 
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish") ~ "Y",
                                        secondary_target_species %in% c("panfish", 
                                                                        "all_species", 
                                                                        "gamefish_species", 
                                                                        "panfish_gamefish", 
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    effort_min,
    targeted_in_survey) %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp) %>% 
  collect() 

#gathering fish from good surveys
adult_bluegill <- wi_data %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>%
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect() 

#creating cpue for species of interest
adult_bluegill_cpue <- adult_bluegill %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'bluegill')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()


#checks at the survey level
good_surveys %>%
  mutate_all(~ ifelse(is.na(.), "NA", as.character(.))) %>%
  gather(key = "variable", value = "value") %>%
  filter(value == "NA")

#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#checks at after collecting fish data
adult_bluegill %>% 
  group_by(species_1, nothing_caught) %>% 
  count()

#is the state label the same for all fish
adult_bluegill %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_bluegill %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#water temp range by sampling method
adult_bluegill %>% 
  group_by(sampling_method) %>%
  summarise(min.temp = min(water_temp), max.temp = max(water_temp))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_bluegill %>% 
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught) %>% 
  glimpse()

adult_bluegill %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#checks after calculating cpe
adult_bluegill_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_bluegill_cpue %>% 
  filter(is.na(cpue))

#remove fish object to save memory
rm(adult_bluegill)
```

#largemouth filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort.standard <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "largemouth_bass") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max) %>% 
  filter(!is.na(water_temp_min))

filter_table_effort.extra <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "largemouth_bass") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  filter(is.na(water_temp_min)) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max,
         -water_temp_min,
         -water_temp_max) 

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys_standard <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("all_species", 
                                                                     "basses",
                                                                     "gamefish_species",
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish",
                                                                     "largemouth_bass") ~ "Y",
                                        secondary_target_species %in% c("all_species",
                                                                        "basses", 
                                                                        "gamefish_species",
                                                                        "panfish_gamefish",
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish", 
                                                                        "largemouth_bass") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.standard, by = c("state", 
                                        "sampling_method")) %>% 
  #collapse data into unique surveys 
    distinct(lake_name_1,
             date_1,
             lake_id,
             survey_id,
             sampling_method,
             total_effort_ident,
             total_effort_1,
             total_effort_1_units,
             nothing_caught,
             targeted_in_survey,
             month_min,
             month_max,
             water_temp_min,
             water_temp_max,
             effort_min,
             water_temp
             ) %>% 
mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         (is.na(water_temp_min) | water_temp >= water_temp_min),
         (is.na(water_temp_max) | water_temp <= water_temp_max),
         total_effort_1 >= effort_min,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp,
         -water_temp_min,
         -water_temp_max) %>% 
  collect()

good_surveys_extra <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("all_species", 
                                                                     "basses",
                                                                     "gamefish_species",
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish",
                                                                     "largemouth_bass") ~ "Y",
                                        secondary_target_species %in% c("all_species",
                                                                        "basses", 
                                                                        "gamefish_species",
                                                                        "panfish_gamefish",
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish", 
                                                                        "largemouth_bass") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.extra, by = c("state", 
                                        "sampling_method")) %>% 
  #collapse data into unique surveys 
    distinct(lake_name_1,
             date_1,
             lake_id,
             survey_id,
             sampling_method,
             total_effort_ident,
             total_effort_1,
             total_effort_1_units,
             nothing_caught,
             targeted_in_survey,
             month_min,
             month_max,
             effort_min
             ) %>% 
mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         total_effort_1 >= effort_min,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  collect()

good_surveys <- bind_rows(good_surveys_extra, good_surveys_standard)

 #gathering fish data from good surveys
adult_largemouth <- wi_data %>% 
   mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect()

#creating cpue for species of interest
adult_largemouth_cpue <- adult_largemouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()


#checking survey level filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

good_surveys %>% 
  group_by(month(date_1), sampling_method) %>% 
  count()

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()

 
#checking filtering after collecting fish
adult_largemouth %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth %>% 
  group_by(state) %>% 
  count()

adult_largemouth %>% 
  group_by(target_species_effort, secondary_target_species) %>% 
  count() %>% 
  print(n = nrow(.))

#water temp range by sampling method
adult_largemouth %>% 
  group_by(month(date_1), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth %>% 
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught) %>% 
  group_by(sampling_method) %>% 
  glimpse()


adult_largemouth %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#checks after calculating cpe
adult_largemouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_cpue %>% 
  filter(is.na(cpue))

adult_largemouth %>% 
  group_by(species_1, nothing_caught) %>% 
  count()

#removing large object to save memory
rm(adult_largemouth)
```

#smallmouth filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort.standard <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max) %>% 
  filter(!is.na(water_temp_min))

filter_table_effort.extra <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "smallmouth_bass") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  filter(is.na(water_temp_min)) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max,
         -water_temp_min,
         -water_temp_max) 

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys_standard <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("all_species", 
                                                                     "basses",
                                                                     "gamefish_species",
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish",
                                                                     "largemouth_bass") ~ "Y",
                                        secondary_target_species %in% c("all_species",
                                                                        "basses", 
                                                                        "gamefish_species",
                                                                        "panfish_gamefish",
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish", 
                                                                        "largemouth_bass") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.standard, by = c("state", 
                                        "sampling_method")) %>% 
  #collapse data into unique surveys 
    distinct(lake_name_1,
             date_1,
             lake_id,
             survey_id,
             sampling_method,
             total_effort_ident,
             total_effort_1,
             total_effort_1_units,
             nothing_caught,
             targeted_in_survey,
             month_min,
             month_max,
             water_temp_min,
             water_temp_max,
             effort_min,
             water_temp
             ) %>% 
mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         (is.na(water_temp_min) | water_temp >= water_temp_min),
         (is.na(water_temp_max) | water_temp <= water_temp_max),
         total_effort_1 >= effort_min,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp,
         -water_temp_min,
         -water_temp_max) %>% 
  collect()

good_surveys_extra <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("all_species", 
                                                                     "basses",
                                                                     "gamefish_species",
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish",
                                                                     "largemouth_bass") ~ "Y",
                                        secondary_target_species %in% c("all_species",
                                                                        "basses", 
                                                                        "gamefish_species",
                                                                        "panfish_gamefish",
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish", 
                                                                        "largemouth_bass") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.extra, by = c("state", 
                                        "sampling_method")) %>% 
  #collapse data into unique surveys 
    distinct(lake_name_1,
             date_1,
             lake_id,
             survey_id,
             sampling_method,
             total_effort_ident,
             total_effort_1,
             total_effort_1_units,
             nothing_caught,
             targeted_in_survey,
             month_min,
             month_max,
             effort_min
             ) %>% 
mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         total_effort_1 >= effort_min,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  collect()

good_surveys <- bind_rows(good_surveys_extra, good_surveys_standard)

#second step of filtering 
##first - filter entire Wisconsin for records of fish that were included in a "good" survey for adult abundance 
##second - apply fish level filters (species, fish lengths, etc.)
adult_smallmouth <- wi_data %>% 
   mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect() 

#creating cpue for species of interest
adult_smallmouth_cpue <- adult_smallmouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'smallmouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()

#checks for survey level items
#min effort
good_surveys %>% 
  group_by(month(date_1), sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))

#month range
#sampling method
good_surveys %>% 
  group_by(month(date_1), sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()

adult_smallmouth %>% 
  group_by(target_species_effort, secondary_target_species) %>% 
  count() %>% 
  print(n = nrow(.))

#is the state label the same for all fish
adult_smallmouth %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_smallmouth %>% 
  group_by(month(date_1), sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#water temp range by sampling method
adult_smallmouth %>% 
  group_by(month(date_1), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_smallmouth %>% 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught,
    effort_min,
    targeted_in_survey) %>% 
  group_by(sampling_method) %>% 
  glimpse()

adult_smallmouth %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#checks after calcuating cpe
adult_smallmouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_smallmouth_cpue %>% 
  filter(is.na(cpue))

adult_smallmouth %>% 
  group_by(species_1, nothing_caught) %>% 
  count()

#removing large fish object
rm(adult_smallmouth)
```

#cisco filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "cisco") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max,
         -effort_min,
         -water_temp_min,
         -water_temp_max)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  #first join the filter table dta to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max) %>% 
  select(-water_temp) %>% 
  collect() 

#gathering fish from good surveys
adult_cisco <- wi_data %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect() 

#creating cpue for species of interest
adult_cisco_cpue <- adult_cisco %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'cisco')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()


#checks at the survey level filtering
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#is the state label the same for all fish
adult_cisco %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_cisco %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_cisco %>% 
  distinct(lake_name_1,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught) %>% 
  group_by(sampling_method) %>% 
  glimpse()

adult_cisco %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#checks after calculating cpe
adult_cisco_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_cisco_cpue %>% 
  filter(is.na(cpue))

#removing large fish object to save memory
rm(adult_cisco)
```

#northern pike filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "northern_pike") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("northern_pike", 
                                                                     "all_species", 
                                                                     "gamefish_species", 
                                                                     "northern_pike_x_muskellunge", 
                                                                     "escocid_percid") ~ "Y",
                                        secondary_target_species %in% c("northern_pike", 
                                                                        "all_species", 
                                                                        "gamefish_species", 
                                                                        "northern_pike_x_muskellunge", 
                                                                        "escocid_percid") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table dta to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    effort_min,
    targeted_in_survey) %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp) %>% 
  collect() 

#collecting fish from good surveys
adult_northern_pike <- wi_data %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>%
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect() 

#creating cpue for species of interest
adult_northern_pike_cpue <- adult_northern_pike %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'northern_pike')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()


#checking filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#checking final filter
adult_northern_pike %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_northern_pike %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_northern_pike %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#water temp range by sampling method
adult_northern_pike %>% 
  group_by(sampling_method) %>%
  summarise(min.temp = min(water_temp), max.temp = max(water_temp))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_northern_pike %>% 
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught) %>% 
  group_by(sampling_method) %>% 
  glimpse()

adult_northern_pike %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#range of cpues?
adult_northern_pike_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_northern_pike_cpue %>% 
  filter(is.na(cpue))

adult_northern_pike %>% 
  group_by(species_1, nothing_caught) %>% 
  count()

#removing large fish object to save memory
rm(adult_northern_pike)
```

#yellow perch filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "yellow_perch") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max)

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("yellow_perch", 
                                                                     "all_species",
                                                                     "gamefish_species",
                                                                     "escocid_percid") ~ "Y",
                                        secondary_target_species %in% c("yellow_perch", 
                                                                        "all_species", 
                                                                        "gamefish_species", 
                                                                        "escocid_percid") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table dta to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort, by = c("state", 
                                        "sampling_method")) %>%
  #collapse data into unique surveys 
  distinct(lake_name_1,
    lake_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    effort_min,
    targeted_in_survey) %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         water_temp >= water_temp_min,
         water_temp <= water_temp_max,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp) %>% 
  collect() 

#collecting fish from good surveys
adult_yellow_perch <- wi_data %>% 
  mutate(total_effort_1 = as.numeric(total_effort_1)) %>%
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect() 

#creating cpue for species of interest
adult_yellow_perch_cpue <- adult_yellow_perch %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'yellow_perch')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()



#checking filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))
#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

#checking after gathering fish
adult_yellow_perch %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_yellow_perch %>% 
  group_by(state) %>% 
  count()

#is each sampling method correct for the filter?
adult_yellow_perch %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

#water temp range by sampling method
adult_yellow_perch %>% 
  group_by(sampling_method) %>%
  summarise(min.temp = min(water_temp), max.temp = max(water_temp))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_yellow_perch %>% 
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    water_temp_min,
    water_temp_max,
    nothing_caught) %>% 
  glimpse()

adult_yellow_perch %>% 
  distinct(total_effort_ident) %>% 
  glimpse()

#range of cpues?
adult_yellow_perch_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_yellow_perch_cpue %>% 
  filter(is.na(cpue))

#removing large fish object
rm(adult_yellow_perch)
```

#combining species
```{r}
wi_filtered_species <- bind_rows(adult_walleye_cpue %>% 
                                   mutate(species_1 = "walleye"), 
                                 adult_bluegill_cpue %>% 
                                   mutate(species_1 = "bluegill"), 
                                 adult_black_crappie_cpue %>% 
                                   mutate(species_1 = "black_crappie"),
                                 adult_largemouth_cpue %>% 
                                   mutate(species_1 = "largemouth_bass"), 
                                 adult_smallmouth_cpue %>% 
                                   mutate(species_1 = "smallmouth_bass"), 
                                 adult_cisco_cpue %>% 
                                   mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
                                   mutate(species_1 = "cisco"), 
                                 adult_northern_pike_cpue %>% 
                                   mutate(species_1 = "northern_pike"), 
                                 adult_yellow_perch_cpue %>% 
                                   mutate(species_1 = "yellow_perch"))

write_csv(wi_filtered_species, "WI_all_cpue_filtered.csv")
```


#general data exploration
```{r}
filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  group_by(state) %>% 
  distinct(state, sampling_method, species_1, area_group, water_temp_min, .keep_all = T)

filter_table_effort <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  group_by(state, survey_type) %>% 
  count() %>% 
  select(-n)

#fall efishing and spring netting for bass 
wi_data %>% 
  group_by(target_species_effort, secondary_target_species) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))
#never conflict but effort gives more info

wi_data %>% 
  group_by(target_species_effort) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))


wi_data %>% 
  group_by(secondary_target_species) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

wi_data %>% 
  group_by(survey_type_1_effort) %>% 
  count() %>% 
  collect()

#what species are in some targeted surveys/survey purposes?
target_species <- wi_data %>% 
  group_by(target_species_effort, secondary_target_species, species_1) %>% 
  count() %>% 
  collect() 

survey_purpose <- wi_data %>% 
  group_by(survey_type_1_indivfish, species_1) %>% 
  count() %>% 
  collect()

#effort in wi?
wi_data %>% 
  group_by(sampling_method, total_effort_1_units) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

wi_data %>% 
  group_by(sampling_method, total_effort_2_units) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

wi_data %>% 
  filter(sampling_method == "fyke_net") %>% 
  group_by(total_effort_1_units, total_effort_2_units) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

#########attempted fix for net nights#####
wi_data %>% 
  mutate(total_effort_1 = case_when(sampling_method == "fyke_net" ~ as.numberic(total_effort_1/24) ~
                                    TRUE ~ total_effort_1)) %>% 
  mutate(total_effort_1_units= case_when(sampling_method == "fyke_net" ~ "nights" ~
                                    TRUE ~ total_effort_1_units)) %>%
  mutate(total_effort = case_when(sampling_method == "fyke_net" ~ total_effort_1*total_effort_2,
                                  TRUE ~ total_effort_1)) %>% 
  mutate(total_effort_units = case_when(sampling_method == "fyke_net" ~ "net_nights",
                                        TRUE ~ total_effort_1_units)) %>% 
  group_by(total_effort, total_effort_units) %>% 
  summarise(min = min(total_effort_2, na.rm = T), max = max(total_effort_2, na.rm = T), mean = mean(total_effort_2, na.rm = T)) %>% 
  collect() %>% 
  print(n = nrow(.))

odd_survey <- wi_data %>% 
  filter(nhdhr_id == "nhdhr_77358700" & date_1 == "2000-04-04") %>% 
  collect()
  distinct(total_effort_ident, total_effort_1)
```

#largemouth filtering
```{r}
#filters for metric (adult abundance) and selects columns needed for survey filtering
filter_table_effort.standard <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "largemouth_bass") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max) %>% 
  filter(!is.na(water_temp_min))

filter_table_effort.extra <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSkGZinLPrSuv1DorBrAkV2JNUZsOlpG7jk_WsnGyZX56z9JMVcGHdCPlsxInjOdjUH0tXb4nBwt9Js/pub?output=csv") %>% 
  filter(metric == "adult_abundance" & state == "Wisconsin" & species_1 == "largemouth_bass") %>% 
  rename(survey_type_1_indivfish = survey_type,
         survey_type_1_effort = survey_type_effort) %>% 
  filter(is.na(water_temp_min)) %>% 
  select(-area_group,
         -metric,
         -species_1,
         -target_species_1,
         -target_species_2,
         -survey_type_1_effort,
         -survey_type_1_indivfish,
         -length_min,
         -length_max,
         -water_temp_min,
         -water_temp_max) 

#filters entire Wisconsin data for surveys representative of adult abundance 
good_surveys_standard <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("all_species", 
                                                                     "basses",
                                                                     "gamefish_species",
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish",
                                                                     "largemouth_bass") ~ "Y",
                                        secondary_target_species %in% c("all_species",
                                                                        "basses", 
                                                                        "gamefish_species",
                                                                        "panfish_gamefish",
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish", 
                                                                        "largemouth_bass") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.standard, by = c("state", 
                                        "sampling_method")) %>% 
  #collapse data into unique surveys 
    distinct(lake_name_1,
             date_1,
             lake_id,
             survey_id,
             sampling_method,
             total_effort_ident,
             total_effort_1,
             total_effort_1_units,
             nothing_caught,
             targeted_in_survey,
             month_min,
             month_max,
             water_temp_min,
             water_temp_max,
             effort_min,
             water_temp
             ) %>% 
mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         (is.na(water_temp_min) | water_temp >= water_temp_min),
         (is.na(water_temp_max) | water_temp <= water_temp_max),
         total_effort_1 >= effort_min,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  select(-water_temp,
         -water_temp_min,
         -water_temp_max) %>% 
  collect()

good_surveys_extra <- wi_data %>%
  mutate(targeted_in_survey = case_when(target_species_effort %in% c("all_species", 
                                                                     "basses",
                                                                     "gamefish_species",
                                                                     "panfish_gamefish", 
                                                                     "centrarchids", 
                                                                     "centrarchid_panfish",
                                                                     "largemouth_bass") ~ "Y",
                                        secondary_target_species %in% c("all_species",
                                                                        "basses", 
                                                                        "gamefish_species",
                                                                        "panfish_gamefish",
                                                                        "centrarchids", 
                                                                        "centrarchid_panfish", 
                                                                        "largemouth_bass") ~ "Y",
                                        TRUE ~ "N")) %>%
  #first join the filter table data to the Wisconsin data (while also filtering for survey purpose)
  right_join(filter_table_effort.extra, by = c("state", 
                                        "sampling_method")) %>% 
  #collapse data into unique surveys 
    distinct(lake_name_1,
             date_1,
             lake_id,
             survey_id,
             sampling_method,
             total_effort_ident,
             total_effort_1,
             total_effort_1_units,
             nothing_caught,
             targeted_in_survey,
             month_min,
             month_max,
             effort_min
             ) %>% 
mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter surveys that meet criteria for a species 
  filter(month(date_1) >= month_min,
         month(date_1) <= month_max,
         total_effort_1 >= effort_min,
         total_effort_1 >= effort_min,
         targeted_in_survey == "Y") %>% 
  collect()

good_surveys <- bind_rows(good_surveys_extra, good_surveys_standard)

 #gathering fish data from good surveys
adult_largemouth <- wi_data %>% 
   mutate(total_effort_1 = as.numeric(total_effort_1)) %>% 
  #filter the Wisconsin data set for surveys that were good for species abundance
  right_join(good_surveys) %>% 
  mutate(state = "Wisconsin") %>%
  collect()

#creating cpue for species of interest
adult_largemouth_cpue <- adult_largemouth %>% 
  #grouping by variables to get survey fish count 
  group_by(lake_name_1,
    lake_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1) %>% 
  #counts fish of a particular species in a survey 
  mutate(count = sum(species_1 == 'largemouth_bass')) %>% 
  #surveys that did not catch fish (species column is na) return a catch of 0
  mutate(count = case_when(is.na(species_1) ~ 0,
                           TRUE ~ count)) %>% 
  #generates cpue from counts and efforts associated with that survey count
  mutate(cpue = count/as.numeric(total_effort_1)) %>%
  #collapse data into survey level cpue
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    date_1,
    nhdhr_id,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught,
    count,
    cpue) %>% 
  collect()


#checking survey level filtering
#min effort
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.effort = min(total_effort_1))

#month range
good_surveys %>% 
  group_by(sampling_method) %>% 
  summarise(min.month = min(month(date_1)), max.month = max(month(date_1)))

good_surveys %>% 
  group_by(month(date_1), sampling_method) %>% 
  count()

#sampling method
good_surveys %>% 
  group_by(sampling_method) %>% 
  count()

good_surveys %>% 
  group_by(targeted_in_survey) %>% 
  count()

 
#checking filtering after collecting fish
adult_largemouth %>% 
  group_by(species_1) %>% 
  count()

#is the state label the same for all fish
adult_largemouth %>% 
  group_by(state) %>% 
  count()

adult_largemouth %>% 
  group_by(target_species_effort, secondary_target_species) %>% 
  count() %>% 
  print(n = nrow(.))

#water temp range by sampling method
adult_largemouth %>% 
  group_by(month(date_1), sampling_method) %>%
  summarise(min.temp = min(water_temp, na.rm = T), max.temp = max(water_temp, na.rm = T))

#do I retain all of the good surveys? - rows here should match good surveys 
adult_largemouth %>% 
  distinct(lake_name_1,
    survey_type_1_indivfish,
    lake_id,
    nhdhr_id,
    date_1,
    survey_id,
    sampling_method,
    total_effort_ident,
    total_effort_1,
    total_effort_1_units,
    water_temp,
    month_min, 
    month_max,
    nothing_caught) %>% 
  group_by(sampling_method) %>% 
  glimpse()


adult_largemouth %>% 
  distinct(total_effort_ident) %>% 
  glimpse()


#checks after calculating cpe
adult_largemouth_cpue %>% 
  group_by(sampling_method) %>% 
  summarise(min = min(cpue), max = max(cpue))

#any na cpue?
adult_largemouth_cpue %>% 
  filter(is.na(cpue))

adult_largemouth %>% 
  group_by(species_1, nothing_caught) %>% 
  count()

#removing large object to save memory
rm(adult_largemouth)
```