---
title: "CASC_Data_Explainer_Work"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
---

2/2/23
First attempt to use the CASC Data Explainer File to rename columns and create units columns where needed. Uses WI data for test case

```{r}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(arrow)
library(data.table)
  # update_dev_pkg()
library(googledrive)
library(janitor)
library(tidyr)
 # install.packages("tidyr")

```

Read in CASC Data Explainer
```{r}

cde <- data.table(read_excel("E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/CASC_Data_Explainer_for_R_Data_Manipulation.xlsx")) #set as DT on the fly

```

Prep the data explainer
```{r}
#tidy up those column names
# colnames(cde)
  cde <- clean_names(cde) #all lwr case, only _ and letters & numbers


# even we can't get our data straight! r/r commas with semicolons 
  cols = colnames(cde)[colnames(cde) != "source_agency" ] # can choose which cols to do action on here skipping source agency b/c "South Dakota Game, Fish and Parks"
  
  #r/r commas to semicolons
  cde[, (cols) := #rebuild each column ()
        lapply(.SD, function(x){str_replace_all(x , pattern = ",", replacement = ";")}), #by looking through each cell and r/r commas with semicolons
      .SDcols = cols ]  # do the rebuild on cols selected in the cols object

#expand each col to the maxN of col names in any row that column
  #get a vector of how many columns named in each cde column:
    maxn <- unlist(cde[, 
        lapply(.SD,function(x){max(lengths(str_split(x, pattern = ";")))}), # how many units are in here? each unit is sep by a semicolon, count col names, then use the max
        .SD ][1], use.names = F)
    
  #split each multi-content column into many:
    for (i in 1:length(colnames(cde)[maxn > 1])) {
      # i =3
      cde[, paste(colnames(cde)[maxn > 1][i],1:maxn[maxn > 1][i], sep = ".") := tstrsplit(get(colnames(cde)[maxn > 1][i]), ";", fixed=TRUE)]
    }

colnames(cde)





```





Read in state Data (starting with 2 files)
From GDrive, not doing this for now, having issues with direct read from GDrive
```{r}  

WI_files_list <- list.files(path = "E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/WI_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files
WI_files_list

n <- length(WI_files_list)

for(i in 1:n) {
  assign(gsub(".csv","", WI_files_list[i]),
         data.table(read_csv_arrow(paste0("E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/WI_raw_disaggregated_data/",
                          WI_files_list[i]))))
}  


```
  





Get new column names from data explainer for all files
```{r}

CDE_cols <-    #all possible column names, including garbage column

  
#oldnames
 CASC_Data_explainer %>% # call data explainer file
  filter(`New File Name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
  
  
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  make_clean_names()
  select(14:25) #keep only columns of relevance (remove our file tracking and urls)

# note for files that have items in the "garbage bin" column, we NEED a different process where we remove garbage columns after 'filter' but before dropping na columns

# note we want to review a sorted list of column names to check misspelling etc. 

  
#newnames
  CASC_Data_explainer %>% # call data explainer file, I abbreviate it as CDE
    filter(`New File Name`== sub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
    select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
    clean_names() %>%  # clean up new names from data explainer
    colnames() #print only CDE column names
  

    
    

colnames(all_gde_gsh) <- (colnames(gde_gsh_col_id))
```


```

# Rename using a named vector and `all_of()`
lookup <- c(pl = "Petal.Length", sl = "Sepal.Length")
rename(iris, all_of(lookup))


# Using data.table
library(data.table)

# rename all columns for old to new
# Rename columns from list
setnames(my_dataframe, old = c('c1','c2','c3','c4','c5'), 
         new = c('id','pages','name','chapters','price'))

