---
title: "CASC_Data_Explainer_Work"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---

2/2/23
First attempt to use the CASC Data Explainer File to rename columns and create units columns where needed

```{r}
library(readr)
library(dplyr)
library(stringr)
library(arrow)
library(data.table)
```

Read in CASC Data Explainer
```{r}
CASC_Data_explainer_02Feb23 <- read_csv_arrow("Data_and_Scripts/Data/Data_Explainer/CASC_Data_Explainer_for_R_Data_Manipulation_02Feb23.csv")
  
  #"G:\.shortcut-targets-by-id\1eHXdJvNMq46tBKURFTeViKWGdSBhWsL_/Data/CASC_Data_Explainer_for_R_manipulation.gsheet"
```

Read in MN Data (starting with 2 files)
```{r}
all_gn_tn <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/5.3.22_update/ALL-GNTN-SURVEYS.csv")

all_gde_gsh <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/5.3.22_update/ALL-GDEGSH-SURVEYS.csv")
```

Get new column names from data explainer
```{r}
CDE_cols <- colnames(CASC_Data_explainer_02Feb23)

gde_gsh_cols <- colnames(all_gde_gsh)

all_gde_gsh_col <- setnames(all_gde_gsh, old = gde )
```

# Using data.table
library(data.table)

# rename all columns for old to new
# Rename columns from list
setnames(my_dataframe, old = c('c1','c2','c3','c4','c5'), 
         new = c('id','pages','name','chapters','price'))

