---
title: "d_filtered_state_combining"
author: "Denver Link"
date: "2023-10-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

#library 
```{r}
library(tidyverse)
library(arrow)
```

#2/5 cpe data dump
```{r}
#reading in individual state cpe 
wi <- read_csv("Filtered_CPEs/WI_all_cpue_filtered.csv")
mn <- read_csv("Filtered_CPEs/MN_all_cpue_filtered.csv")
mi <- read_csv("Filtered_CPEs/MI_all_cpue_filtered.csv")
il <- read_csv("Filtered_CPEs/IL_all_cpue_filtered.csv") 
sd <- read_csv("Filtered_CPEs/SD_all_cpue_filtered.csv")
ia <- read_csv("Filtered_CPEs/IA_all_cpue_filtered.csv")
indy <- read_csv("Filtered_CPEs/IN_all_cpue_filtered.csv")

#make sure i get ride of columns that are not survey level or helpful to Chris
cross_state <- row_bins(wi, mn, mi, il, sd, ia, indy)

#write final cpe
write_csv(cross_state, "all_state_cpue_8Dec23.csv")
```






#data
```{r}
wi <- read_csv("WI_all_cpue_filtered.csv") %>%
  select(lake_id, nhdhr_id, sampling_method, date_1, total_effort_1, species_1, count, cpue, total_effort_1_units) %>% 
  mutate(lake_id = as.character(lake_id),
         state = "Wisconsin") %>% 
  rename(nhdhr.id = nhdhr_id,
         effort_units.1 = total_effort_1_units)
mn <- read_csv("MN_all_cpue_filtered.csv") %>% 
  select(lake_id, nhdhr.id, sampling_method, date_clean, total_effort_1.1, count, cpue, species_1, effort_units.1)  %>% 
  mutate(lake_id = as.character(lake_id),
         state = "Minnesota") %>% 
  rename(date_1 = date_clean, 
         total_effort_1 = total_effort_1.1)
mi <- read_csv("MI_all_cpue_filtered.csv") %>% 
  select(lake_id, nhdhr.id, sampling_method, date.1, total_effort_1.1_effort, count, cpue, species_1, effort_units.1) %>% 
  rename(date_1 = date.1,
         total_effort_1 = total_effort_1.1_effort) %>% 
  mutate(state = "Michigan")

#combined
cross_state <- bind_rows(wi, mn, mi)
write_csv(cross_state, "mn_wi_mi_filtered_cpue_11NOV23.csv")
glimpse(cross_state)

cross_state <- read_csv("all_state_cpue_8Dec23.csv")
cross_state %>% 
  filter(!is.na(nhdhr_id)) %>% 
  count()
cross_state %>% 
  group_by(state) %>% 
  count()
```

#smashing cpes across states together
```{r}
wi <- read_csv("Filtered_CPEs/WI_all_cpue_filtered.csv") %>% 
  select(lake_id,
         lake_name_1,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         count,
         cpue,
         species_1) %>% 
  mutate(state = "Wisconsin",
         lake_id = as.character(lake_id))
mn <- read_csv("Filtered_CPEs/MN_all_cpue_filtered.csv") %>% 
  select(lake_id,
         lake_name.1,
         nhdhr.id,
         date_clean,
         sampling_method,
         total_effort_1.1,
         effort_units.1,
         count,
         cpue,
         species_1) %>% 
  rename(lake_name_1 = lake_name.1,
         nhdhr_id = nhdhr.id,
         date_1 = date_clean,
         total_effort_1 = total_effort_1.1,
         total_effort_1_units = effort_units.1) %>% 
  mutate(state = "Minnesota",
         lake_id = as.character(lake_id))
mi <- read_csv("Filtered_CPEs/MI_all_cpue_filtered.csv") %>% 
  select(lake_id,
         lake_name.1,
         nhdhr.id,
         date.1,
         sampling_method,
         total_effort_1.1_effort,
         effort_units.1,
         count,
         cpue,
         species_1) %>% 
  rename(lake_name_1 = lake_name.1,
         nhdhr_id = nhdhr.id,
         date_1 = date.1,
         total_effort_1 = total_effort_1.1_effort,
         total_effort_1_units = effort_units.1) %>% 
  mutate(state = "Michigan")
il <- read_csv("Filtered_CPEs/IL_all_cpue_filtered.csv") %>% 
  select(state,
         lake_id,
         lake_name_1,
         date_1,
         sampling_method_2,
         total_effort_1,
         total_effort_1_units,
         count,
         cpue,
         species_1) %>% 
  rename(sampling_method = sampling_method_2) %>% 
  mutate(lake_id = as.character(lake_id),
         nhdhr_id = NA)
sd <- read_csv("Filtered_CPEs/SD_all_cpue_filtered.csv") %>% 
  select(state,
         lake_id,
         lake_name_1,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         count,
         cpue,
         species_1)
ia <- read_csv("Filtered_CPEs/IA_all_cpue_filtered.csv") %>% 
  mutate(state = "Iowa",
         nhdhr_id = NA,
         date_1 = as.Date(date.1)) %>% 
  select(state,
         lake_id,
         lake_name.1,
         nhdhr_id,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         count,
         cpue,
         species_1) %>% 
  rename(lake_name_1 = lake_name.1) 
indy <- read_csv("Filtered_CPEs/IN_all_cpue_filtered.csv") %>% 
  select(state,
         lake_id,
         lake_name_1,
         date_1,
         sampling_method,
         total_effort_1,
         total_effort_1_units,
         count,
         cpue,
         species_1)

#combining 
cross_state <- bind_rows(wi, mn, mi, il, sd, ia, indy)

write_csv(cross_state, "all_state_cpue_8Dec23.csv")
wi <- read_csv("all_state_cpue_8Dec23.csv") %>% 
  filter(state == "Wisconsin") %>% 
  group_by(species_1) %>% 
  count()
```

#basic review of combination
```{r}
cross_state %>% 
  group_by(state, species_1) %>% 
  count() %>% 
  print(n = nrow(.))

cross_state %>% 
  group_by(state, sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))

cross_state %>% 
  group_by(state, species_1, sampling_method) %>% 
  summarise(min.cpe = min(cpue), mean.cpe = mean(cpue), max.cpe = max(cpue), n = n()) %>% 
  print(n = nrow(.))

cross_state %>% 
  ggplot() +
  geom_histogram(aes(x = cpue), bins = 50) +
  facet_grid(state ~ species_1, scales = "free")

cross_state %>% 
  ggplot() +
  geom_violin(aes(x = species_1, y = cpue, color = species_1)) +
  facet_wrap(~state, scales = "free")

cross_state %>% 
  filter(state == "Minnesota" & species_1 %in%  c("walleye", "northern_pike", "black_crappie", "largemouth_bass")) %>% 
  filter(sampling_method == "Standard gill net sets") %>% 
  filter(total_effort_1 < 20) %>% 
  ggplot() +
  geom_smooth(aes(x = total_effort_1, y = cpue)) +
  facet_wrap(species_1~sampling_method, scales = "free")
```

#mergeing cross-state to covariates
-data taken from temp projections summaries
 -clarity form the metadata
 -temp data from the temp metrics feather file
```{r}
#cross state cpue data
cross_state <- read_csv("all_state_cpue_8Dec23.csv") %>% 
  mutate(year = year(date_1))

#covariate data
temp <- read_feather("lake_temperature_metrics_GLM_NLDAS.feather", col_select = NULL, as_data_frame = TRUE, mmap = TRUE) %>% 
  select(site_id, year, stratification_onset_yday, stratification_duration, mean_surf_JulAugSep, ice_on_date, ice_off_date) %>% 
  rename(nhdhr_id = site_id)

clarity <- read_csv("glm_lake_metadata.csv") %>% 
  select(site_id, max_depth, area, clarity, elevation) %>% 
  rename(nhdhr_id = site_id)

cross_state_covary <- cross_state %>% 
  left_join(clarity) %>% 
  left_join(temp, by = c("nhdhr_id",
                         "year"))

cross_state_covary %>% 
  filter(is.na(nhdhr_id)) %>% 
  glimpse()

write_csv(cross_state_covary, "all_state_cpue_covariates.csv")
```



#map of cpue by state
```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(maps) #to pull in the map data for geom_polygon
library(colorspace) #package with color palettes to choose from
library(viridis) #more color palettes

# lake info from LAGOS, https://portal.edirepository.org/nis/mapbrowse?packageid=edi.854.1 (want 'lake_information' file)
lagos_lake_info <- read_csv("LAGOS_lake_information_20Oct2023.csv")
nhd_lat_longs <- lagos_lake_info %>%
  select(lake_nhdid, lake_lat_decdeg, lake_lon_decdeg) #removes unnecessary columns

wi_cpue_lat_long <- wi %>%
  mutate(lake_nhdid = str_replace(nhdhr_id, "nhdhr_", "")) %>%  #remove nhdhr_ prefix from id
  left_join(nhd_lat_longs, by = "lake_nhdid")
mn_cpue_lat_long <- mn %>% 
  mutate(lake_nhdid = str_replace(nhdhr.id, "nhdhr_", "")) %>%    #remove nhdhr_ prefix from id
  left_join(nhd_lat_longs, by = "lake_nhdid")
mi_cpue_lat_long <- mi %>% 
  mutate(lake_nhdid = str_replace(nhdhr.id, "nhdhr_", "")) %>%    #remove nhdhr_ prefix from id
  left_join(nhd_lat_longs, by = "lake_nhdid")
  
#checked for lakes that didn't get lat longs with following code placed after the left join above 
  # %>% filter(is.na(lake_lat_decdeg))%>%group_by(lake_id)%>%summarise(total = n())
#  84600, 2621900, 2674100, 2691500 check out these nhd ids later

#basic map for reference code
Midwest <- map_data("state", region = c("minnesota", "wisconsin", "north dakota", "south dakota", "iowa", "illinois", "indiana", "michigan"))
midwest_map <- ggplot(data = Midwest)+
  geom_polygon(aes(x = long, y = lat, group = group, fill = region), color = "white")+
  coord_fixed(1.3)+
  guides(fill = FALSE)
midwest_map

#wisconsin map
WI <- map_data("state", region = "wisconsin") #get basic WI map
wi_map <- ggplot(data = WI)+
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  coord_fixed(1.3)+
  geom_point(data = wi_cpue_lat_long, aes(x = lake_lon_decdeg, y = lake_lat_decdeg, color = cpue ), alpha = .25)+
  facet_wrap(species_1~sampling_method)+
  scale_color_viridis(option = "H")
wi_map

#Minnesota map
MN <- map_data("state", region = "minnesota") #get basic WI map
mn_map <- ggplot(data = MN)+
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  coord_fixed(1.3)+
  geom_point(data = mn_cpue_lat_long, aes(x = lake_lon_decdeg, y = lake_lat_decdeg, color = cpue), alpha = .25)+
  facet_wrap(species_1~sampling_method)+
  scale_color_viridis(option = "H")
mn_map

#Michigan map
MI <- map_data("state", region = "michigan") #get basic WI map
mi_map <- ggplot(data = MI)+
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  coord_fixed(1.3)+
  geom_point(data = mi_cpue_lat_long, aes(x = lake_lon_decdeg, y = lake_lat_decdeg, color = cpue), alpha = .25)+
  facet_wrap(species_1~sampling_method)+
  scale_color_viridis(option = "H")
mi_map

#all states together - still needs work
MN_MI_WI <- map_data("state", region = c("minnesota", "wisconsin", "michigan"))
MN_MI_WI_map <- ggplot(data = MN_MI_WI)+
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  coord_fixed(1.3)+
  geom_point(data = cross_state_wae, aes(x = lake_lon_decdeg, y = lake_lat_decdeg, color = log(cpue)), alpha = .5)+ #can repeat this line for each state if needed
  facet_wrap(.~sampling_method)+
  scale_color_viridis(option = "H")+
  labs(title = "Filtered Walleye CPUE data by gear", x = "longitude", y = "latitude")
MN_MI_WI_map

#time with walleye?
ggplot(data = MN)+
  geom_polygon(aes(x = long, y = lat, group = group), fill = "white", color = "black")+
  coord_fixed(1.3)+
  geom_point(data = mn_cpue_lat_long %>% 
               filter(species_1 == "walleye" & sampling_method == "Standard gill net sets"), aes(x = lake_lon_decdeg, y = lake_lat_decdeg, color = cpue), alpha = .25)+
  facet_wrap(~year(date_1))+
  scale_color_viridis(option = "H")
```


#comparing acorss states
```{r}
#tallies of surveys
cross_state %>% 
  filter(state == "Minnesota") %>% 
  mutate(sampling_method = case_when(str_detect(sampling_method, "Standard gill") ~ "gill net",
                                     TRUE ~ "trap net")) %>% 
  ggplot() +
  geom_bar(aes(sampling_method)) +
  facet_wrap(~species_1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cross_state %>% 
  filter(state == "Wisconsin") %>% 
  ggplot() +
  geom_bar(aes(sampling_method)) +
  facet_wrap(~species_1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

cross_state %>% 
  filter(state == "Michigan") %>% 
  ggplot() +
  geom_bar(aes(sampling_method)) +
  facet_wrap(~species_1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  


#histogram of cpues
cross_state %>% 
  filter(cpue < 100) %>% 
  ggplot() +
  geom_histogram(aes(cpue), alpha = .6) +
  facet_grid(species_1~state, scales = "free")

cross_state %>% 
  filter(state == "Minnesota") %>% 
  ggplot()+
  geom_histogram(aes(cpue), alpha = .6) +
  facet_grid(~species_1, scales = "free")

cross_state %>% 
  filter(state == "Wisconsin") %>% 
  ggplot()+
  geom_histogram(aes(cpue), alpha = .6) +
  facet_grid(~species_1, scales = "free")

cross_state %>% 
  filter(state == "Michigan") %>% 
  ggplot()+
  geom_histogram(aes(cpue), alpha = .6) +
  facet_grid(~species_1, scales = "free")

#histogram of timing

#not sure why this plot is showing up funky? - Michigan?
cross_state %>% 
  mutate(julian_day = yday(date_1)) %>% 
  ggplot() +
  geom_histogram(aes(julian_day, fill = state), alpha = .4, binwidth = 5) +
  facet_wrap(~state, scales = "free")

cross_state %>% 
  mutate(julian_day = yday(date_1)) %>% 
  ggplot() +
  geom_density(aes(julian_day, fill = state), alpha = .4)

cross_state %>% 
  mutate(julian_day = yday(date_1)) %>% 
  ggplot() +
  geom_histogram(aes(julian_day, fill = state), binwidth = 5) +
  facet_wrap(~state)
```

#gears across states?
```{r}
cross_state <- read_csv("mn_wi_mi_filtered_cpue_11NOV23.csv")

cross_state %>% 
  group_by(species_1, state, sampling_method) %>% 
  count() %>% 
  print(n = nrow(.))
```

#plot for presentation
```{r}
plot <- mn_data %>% 
  filter(sampling_method_abbrev == "GN") %>% 
  filter(species.1 %in% c("walleye", "northern_pike", "largemouth_bass", "black_crappie")) %>% 
  collect() %>% 
  group_by(effort_ident, species.1) %>% 
  summarise(lake_id = unique(lake_id), sampling_method = unique(sampling_method), area_group = unique(area_group), date.1 = unique(date_clean), effort = unique(total_effort_1.1), n = n()) %>% 
  mutate(cpue = n/effort)

plot %>% 
  ungroup() %>% 
  filter(effort <= 20) %>% 
ggplot() +
  geom_smooth(aes(x = effort, y = cpue, color = area_group)) +
  facet_wrap(~ species.1, scales = "free")

plot %>% 
  ungroup() %>% 
  filter(effort <= 20) %>% 
ggplot() +
  geom_smooth(aes(x = effort, y = cpue)) +
  facet_wrap(~ species.1, scales = "free")
```


#notes for workshop- wisconsin
```{r}
#reading in data 
wi_inland_cpue <- read_csv("Data_and_Scripts/wi_data/wi_inland_cpue_19Mar2021.csv")
glimpse(wi_inland_cpue)

wi_inland_effort <- read_csv("Data_and_Scripts/wi_data/wi_inland_effort_19Mar2021.csv")
glimpse(wi_inland_effort)
inland_effort_keep <- wi_inland_effort %>% 
  group_by(gear,
         distance.shocked,
         weather.description,
         adverse.condition,
         volts,
         amps,
         pulse.rate.percent,
         duty.cycle,
         current.type,
         number.of.dippers,
         dipnet.bar.mesh.size,
         type.of.pass,
         deployment.depth,
         water.clarity,
         water.clarity.feet.amt,
         probe.no,
         probe.no.amt,
         run.type,
         average.droppers.per.probe) %>% 
  summarise(records = n()) 
write_csv(inland_effort_keep, "wi_effort_file_notes.csv")

wi_inland_fishobservations <- read_csv("Data_and_Scripts/wi_data/wi_inland_fishobservations_19Mar2021-001.csv")
glimpse(wi_inland_fishobservations) 
fishobs_notes <- wi_inland_fishobservations %>% 
  group_by(visit.type,
         net.number,
         mark.given,
         mark.found,
         second.mark.found,
         tag.number.given,
         second.tag.number.given,
         tag.number.found,
         second.tag.number.found,
         primary.survey.purpose,
         secondary.survey.purpose) %>% 
  summarise(n = n())
write_csv(fishobs_notes, "wi_fish_obs_notes.csv")

wi_inland_lenage <- read_csv("Data_and_Scripts/wi_data/wi_inland_lenage_19Mar2021.csv")
glimpse(wi_inland_lenage)

wi_win_cpe <- read_csv("Data_and_Scripts/wi_data/wi_winnebago_cpue_23Apr2021.csv")
glimpse(wi_win_cpe)

wi_win_fish <- read_csv("Data_and_Scripts/wi_data/wi_winnebago_fishobservations_23April2021.csv")
glimpse(wi_win_fish)

wi_cisco <- read_csv("Data_and_Scripts/wi_data/wi_cisco.csv")
glimpse(wi_cisco)
wi_cisco_notes <- wi_cisco %>% 
  group_by(Maxdpth,
         Mndepth,
         Stratind,
         Shore,
         Alk,
         Conduct,
         TSI,
         Cisco,
         C_Lst_yr,
         C_Source,
         VGN_Date,
         VGN_NN,
         VGN_Cisco,
         VGN_CiscoNN,
         VGN_CiscoRA,
         Overall_CiscoNN,
         Overall_CiscoRA,
         Overall_SmeltNN,
         Overall_LWhiteNN,
         TDO3,
         TDO3_Date,
         Slimysc,
         Othr_cld,
         Comments) %>% 
  summarize(n = n())
```

#cisco - wisco and michginan
```{r}
#wisconsin
wi_data <- open_dataset("Data_and_Scripts/wi_file_arrow")

wi_data %>% 
  distinct(lake_id, date_1, sampling_method) %>% 
  group_by(sampling_method) %>% 
  summarise(n = n()) %>% 
  collect() %>% 
  print(n = nrow(.))

wi_data %>% 
  filter(sampling_method == "vertical_gill_net") %>% 
  distinct(lake_id, date_1) %>% 
  collect() %>% 
  group_by(month(date_1)) %>% 
  count()

wi_data %>% 
  group_by(species_1) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))

cisco_surveys <- wi_data %>% 
  filter(sampling_method == "vertical_gill_net") %>% 
  distinct(lake_name_1, lake_id, date_1) %>% 
  collect()

#Michigan 
mi_data <- open_dataset("Data_and_Scripts/mi_file_arrow")

mi_data %>% 
  distinct(lake_id, date.1, sampling_method) %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  collect()

mi_data %>% 
  filter(species.1 == "cisco") %>%
  distinct(lake_id, date.1, sampling_method) %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  collect()
#cisco are only being caught in 

#minnesota
mn_data <- open_dataset("Data_and_Scripts/mn_file_arrow")

mn_data %>% 
  group_by(species.1) %>% 
  count() %>% 
  collect() %>% 
  arrange(species.1) %>% 
  print(n = nrow(.))

mn_data %>% 
  filter(species.1 == "cisco") %>%
  distinct(lake_id, date_clean, sampling_method) %>% 
  group_by(sampling_method) %>% 
  count() %>% 
  collect() %>% 
  print(n = nrow(.))
```

