---
title: "Hansen_Lab_QC"
author: "Denver Link"
date: "2024-04-08"
output: html_document
editor_options: 
  chunk_output_type: console
---

#library 
```{r}
library(tidyverse)
library(arrow)
```

#data
```{r}
mw_data <- open_dataset("hive", partitioning = c("state"))
glimpse(mw_data)
```

1. How many fish are in the database?
```{r}

```

2. How many lakes are within each state?
```{r}

```

3. How many sunfish were saught in each year in Mille Lacs?
```{r}

```

4. Do the units of length of each species/state combo make sense?
```{r}

```

5a. Which lakes captured the largest walleye, largemouth bass, crappie, and bluegill in each state?
5b. What year did that happen?
```{r}

```

6. Does every state have every species of fish observed?
```{r}

```

7a. How many surveys in each state caught NO fish?
7b. In states that record sub-effort, how many surveys had at least one net/run catch NO fish?
```{r}

```

8. When a state conducts a targeted survey, do they only capture that species of fish? Is it different between states?
```{r}

```

9. Based on the notes/flag column, which surveys should be NOT kept for CPUE and why?
```{r}

```

10a. Generally, when does each state conduct sampling?
10b. Do they depoly all gears at the same time? If not, how do the timings differ within a state for each gear?
```{r}

```


*Identify outliers in length, weight, length-weight ratio, or other columns*
```{r}

```

