---
title: "IN_Filtering"
author: "Holly Kundel"
date: '2022-06-29'
output: html_document
---
Load in required packages
```{r}
library(readr)
library(readxl)
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(forcats)
library(gghighlight)
library(stringr)
library(purrr)

options(scipen = 999) #ensures that survey ID numbers aren't rounded with scientific notation

```

Indiana is slightly different from other states because the data is divided into glacial lakes and reservoirs.

Load in Raw Data from Computer (in future try to read in data from explainer file in Drive)
Data Gretchen worked with
```{r, warning=FALSE}
in.fish.community.data <- read_excel("Data_and_Scripts/Data/Data_Summaries/IN/in_glacial_lakes2010_2020_fishdata.xlsx")
in.effort.community.data <- read_excel("Data_and_Scripts/Data/Data_Summaries/IN/in_glacial_lakes_2010_2020_effort.xlsx")
```

Add in lake ID Number for IN Glacial Lakes 9/7/22
```{r}

```

Add Gear Code to Effort Data to Match
Gretchen's code
```{r}
gear.translate <- data.frame("Gear"=unique(in.effort.community.data$Gear))%>% 
  mutate(gear_code = case_when(Gear=="Standard Trap Net" ~ "TN",
                             Gear=="Experimental Gill Net"~"GN",
                             Gear=="Nighttime Electrofishing"~"EF"))

in.effort.community.data <- in.effort.community.data %>%
  right_join(gear.translate) %>%  #should add something like , by = c("gear.name"= "Gear"
  rename(effort='Total Effort by Gear Type (Units; trap/gill nets = overnight lifts, total nighttime electrofishing = hours)',
         gear.name=Gear, Gear=gear_code )

#summarize catch by species and gear and lake/county/year
fish.catch <- in.fish.community.data %>%
  group_by(Lake, County, Year, Species, Gear) %>%
  summarise(total.catch=length(Length), mean.length=mean(Length)) %>%
  filter(Gear!="X"&Gear!="#REF!")
#a few missing data points for gear that we will ignore (5 fish out of 69k)

#identify surveyed years
surveyed.years <- unique(select(in.effort.community.data, Lake, County,Year, Gear, effort))

#expand fish and link to actual effort (expand will include unsurveyed years)
fish.long <- ungroup(fish.catch) %>%
  expand( Lake, County, Year, Species, Gear) %>%
  right_join(surveyed.years) #Joining, by = c("Lake", "County", "Year", "Gear")

#link to effort and #create temporary id for now
fish.catch.effort <- fish.catch %>% right_join(fish.long) %>% #Joining, by = c("Lake", "County", "Year", "Species", "Gear")
  replace_na(list(total.catch=0)) %>%
  mutate(cpue=total.catch/effort) %>%
  mutate(temp.id=paste(County, Lake, sep="_"))
```
My quick summary plot
```{r}
IN_fish_catch_effort_lk <- as.data.frame(fish.catch.effort)%>%
  filter(Species %in% c("Black Crappie", "Bluegill", "Largemouth Bass", "Northern Pike", "Smallmouth Bass", "Walleye", "Yellow Perch"))

IN_Lk_CATCH <- ggplot(data = IN_fish_catch_effort_lk, aes (x = Gear, y= total.catch, fill = Species))+
  geom_bar(stat = "identity")+
  facet_wrap(.~Species, scales = "free")+
  theme_bw()+
  labs(y = "Total Catch", title = "IN Species Total Catch by Gear: Glacial Lakes", caption = "*note differences in y-axis scales")


IN_Lk_CATCH

```

Summary Stats for All Species (Gretchen's Code)
```{r}
lake.count <- fish.catch.effort %>%
  filter(total.catch!=0) %>%
  group_by( Species) %>%
  summarise(N.lakes=length(unique(temp.id))) %>%
  arrange(-N.lakes) %>%
  mutate(percent.lakes=N.lakes/length(unique(fish.catch.effort$temp.id))) %>%
  mutate(Species=factor(Species, levels=arrange(lake.count,-N.lakes)$Species))

IN_all_Species <- ggplot(lake.count, aes(Species, percent.lakes))+
  geom_bar(stat="identity", fill="lavender", colour="navyblue")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(y = "Proportion of lakes\n(out of 132)", title = "In All Species")

IN_all_Species
```
Species Richness (Gretchen's Code)
```{r}
species.count <- fish.catch.effort %>% 
  filter(total.catch!=0) %>%
  group_by(temp.id) %>% 
  summarise(N.Species=length(unique(Species))) %>%
  arrange(-N.Species) %>%
  mutate(percent.species=N.Species/length(unique(fish.catch.effort$Species))) 

species.count <- species.count%>%
  mutate(temp.id=factor(temp.id, levels=arrange(species.count,-N.Species)$temp.id))

IN_Sp_Richness <- ggplot(species.count, aes(temp.id, N.Species))+
  geom_bar(stat="identity", fill="lightblue", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=90, hjust=1, vjust=.5))+
  labs(y = "Number of species\n(out of 83)", title = "IN Species Richness", x = "Lake")

IN_Sp_Richness
```
More Gretchen Code
```{r}
#extract species of interest
target.species=c("Bluegill", "Largemouth Bass","Northern Pike", "Smallmouth Bass", "Yellow Perch", "Black Crappie", "Walleye", "Cisco")

lake.count=lake.count %>%
  mutate(key.species=case_when(Species%in%target.species~1, TRUE~0))


IN_All_sp_highlight <- ggplot(lake.count, aes(Species, percent.lakes))+
  geom_bar(stat="identity", colour="black", fill="turquoise")+
  theme_classic()+theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = "none")+
  ylab("Proportion of lakes\n(out of 132)")+
  gghighlight(key.species==1)

IN_All_sp_highlight
```
Check Species by gear (Gretchen's code)
```{r}
lake.count.gear <- fish.catch.effort %>%
  filter(total.catch!=0) %>%
  group_by( Species, Gear) %>%
  summarise(N.lakes=length(unique(temp.id))) %>%
  arrange(-N.lakes, Gear) %>% 
  mutate(percent.lakes=N.lakes/length(unique(fish.catch.effort$temp.id))) %>%
  mutate(Species=factor(Species, levels=arrange(lake.count,-N.lakes)$Species))

lake.count.gear=lake.count.gear %>%
  mutate(key.species=case_when(Species%in%target.species~1, TRUE~0))


In_Sp_by_gear_plot <- ggplot(lake.count.gear, aes(Species, percent.lakes, group=Gear, fill=Gear))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  theme_classic()+theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.8,.8))+
  ylab("Proportion of lakes\n(out of 132)")+
  gghighlight(key.species==1, use_direct_label = FALSE, use_group_by = FALSE)+facet_wrap(~Gear, ncol=1)

In_Sp_by_gear_plot
```
Just target Species (Gretchen's Code)
```{r}
IN_target_sp_gears <- lake.count.gear %>%
  filter(key.species==1) %>%
  ggplot( aes(Gear, percent.lakes, group=Gear, fill=Gear))+geom_bar(stat="identity", colour="black", position="dodge")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.8,.2))+
  ylab("Proportion of lakes\n(out of 132)")+
  facet_wrap(~Species)

IN_target_sp_gears
```
Mean Catch by Gear (Gretchen)
```{r}
#mean catch by gear
mean.cpue <- fish.catch.effort %>%
  filter(total.catch!=0) %>%
  group_by( Species, Gear) %>%
  summarise(mean.non.zero.cpue=mean(total.catch/effort), lci.non.zero.cpue=quantile(total.catch/effort, .0275), uci.non.zero.cpue=quantile(total.catch/effort, .975, na.rm=T),mean.length=mean(mean.length), lci.length=quantile(mean.length, .0275, na.rm=T),uci.length=quantile(mean.length, .975, na.rm=T) )

mean.cpue <- mean.cpue %>%
  mutate(key.species=case_when(Species%in%target.species~1, TRUE~0))

#just target species

IN_Mean_CPUE_by_gear <- mean.cpue %>% 
  filter(key.species==1) %>%
  ggplot( aes(Species, mean.non.zero.cpue, group=Gear, colour=Gear))+
  geom_pointrange(aes(ymax=uci.non.zero.cpue, ymin=lci.non.zero.cpue))+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.8,.8))+
  ylab("Mean non-zero CPUE")+
  facet_wrap(~Gear, scales="free_y")

IN_Mean_CPUE_by_gear
```

```{r}
IN_Violin_CPUE_by_gear <- fish.catch.effort %>%
  filter(Species%in%target.species) %>%
  ggplot( aes(Species, total.catch/effort, group=Species, colour=Gear, fill=Gear))+
  geom_violin()+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.8,.8))+
  ylab("Non-Zero CPUE")+
  facet_wrap(~Gear, scales="free_y")+
  scale_y_log10()

IN_Violin_CPUE_by_gear 
```
Mean length and CPUE (Gretchen)
```{r}
IN_mean_length_CPUE <- fish.catch.effort %>%
  filter(Species%in%target.species)%>%
  ggplot( aes( total.catch/effort, mean.length, group=Gear, colour=Gear, shape=Gear))+
  geom_point(size=2, alpha=.7, position=position_jitter(width=.05))+
  theme_classic()+theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.8,.2))+
  xlab("Mean non-zero CPUE")+facet_wrap(~Species, scales="free")+
  ylab("Mean length")+
  scale_x_log10()

IN_mean_length_CPUE
```
End Gretchen's code looking at the community data

IN Reservoir Fish Data
Read in reservoir data
```{r}
IN_Resr_Fish <- read_excel("Data_and_Scripts/Data/Data_Summaries/IN/in_reservoir_fish_community_fish_data.xlsx")

IN_Resr_Effort <- read_excel("Data_and_Scripts/Data/Data_Summaries/IN/in_reservoir_fish_community_effort.xlsx")

```

Get list of all possible lakes, years, species, and gears
- Since the fish catch data doesn't have year info, will need to combine with effort
- then expand list
- join expanded list to catch (adding 0s where none are caught)
- then join with effort to remove survey gears that weren't used
```{r}
IN_Resr_List <- IN_Resr_Fish %>%
  rename(Survey_ID = 'Survey ID', Lake_Name = 'Lake Name')%>%
  group_by(Survey_ID, Species, Gear)%>%
  summarise(Total_Catch = n())%>%
  full_join(IN_Resr_Effort, by = "Survey_ID")%>%
  mutate(Sample_Start = ymd(Sample_Start_Date))%>%
  mutate(Year = year(Sample_Start))%>%
  mutate(Species = str_to_lower(Species))%>% #will help R grup species, uppercase and lowercase names used
  mutate(Gear_Abbrev = case_when(Gear=="Gill net std exp" ~ "GN",
                                 Gear == "Gill net STB exp" ~ "GN",
                                 Gear=="Trap net std"~"TN",
                                 Gear=="Night EF"~"EF"))%>% #will help keep gears straight, 7 entries for GN are slightly different (std vs STB)
  mutate(Gear_Abbrev = as.factor(Gear_Abbrev))%>% 
  filter(Survey_ID != 2, Survey_ID != 373, Survey_ID != 273) #2 and 373 don't have any effort data, 273 lacks fish data

# List of all possible Species, gears. Only one survey per lake so no need for years or lakes here
IN_Resr_List_Long <- IN_Resr_List %>% 
  expand(Survey_ID, Gear_Abbrev, Species) #this is too many years given they only did one survey per lake regardless of year but can filter out later, , Year=2016:2019. Not including year because there was only one survey per lake. 

#Combine catch, and rename columns for easier use
IN_Resr_Fish_Usable <- IN_Resr_Fish %>%
  rename(Survey_ID = 'Survey ID', Lake_Name = 'Lake Name', Length_in = `Length (in)`, Weight_lb = `weight (lb)`, Gear_Number = `Gear Number`)%>%
  mutate(Gear_Abbrev = case_when(Gear=="Gill net std exp" ~ "GN",
                                 Gear == "Gill net STB exp" ~ "GN",
                                 Gear=="Trap net std"~"TN",
                                 Gear=="Night EF"~"EF"))%>%
  mutate(Species = str_to_lower(Species))%>%
  group_by(Survey_ID, Species, Gear_Abbrev)%>%
  summarise(Total_Catch = n())

#Pull in catch data and add 0s if a species was not caught
IN_Resr_Total_Catch <- IN_Resr_Fish_Usable %>%
  right_join(IN_Resr_List_Long, by = c("Survey_ID", "Gear_Abbrev", "Species"))%>%
  mutate(Total_Catch = replace_na(Total_Catch, 0))

# Split effort up by gear (so it isn't one line per survey, it is one line per gear)
IN_Resr_Effort_by_Gear <- IN_Resr_Effort%>%
  rename(EF = 9, GN = 10, TN = 11)%>%
  pivot_longer(c(EF, GN, TN), names_to = "Gear_Abbrev", values_to = "Effort")%>%
  mutate(Effort_Units = case_when(Gear_Abbrev == "GN" ~ "Number of nets",
                                  Gear_Abbrev == "TN" ~ "Number of nets",
                                  Gear_Abbrev == "EF" ~ "Number of hours"))


# Combine catch and effort to calculate CPUE
IN_Resr_CPUE <- IN_Resr_Total_Catch %>%
  right_join(IN_Resr_Effort_by_Gear, by = c("Survey_ID", "Gear_Abbrev"))%>%
  filter(Survey_ID != 273) %>% #no catch data
  mutate(CPUE = Total_Catch/Effort)

In_Resr_CPUE_just_target_sp <- IN_Resr_CPUE %>%
  filter(Species %in% c("black crappie", "bluegill", "largemouth bass", "northern pike", "smallmouth bass", "walleye", "yellow perch")) #no cisco
```

Summary Stats and Plots
```{r}
IN_Resr_CATCH <- ggplot(data = In_Resr_CPUE_just_target_sp, aes (x = Gear_Abbrev, y= Total_Catch, fill = Species))+
  geom_bar(stat = "identity")+
  facet_wrap(.~Species, scales = "free")+
  theme_bw()+
  labs(y = "Total Catch", title = "IN Species Total Catch by Gear: Reservoirs", caption = "*note differences in y-axis scales")

IN_Resr_CATCH



IN_resr_number <- IN_Resr_CPUE %>%
  group_by(Lake_Name)%>%
  summarise(Total = n()) #63 reservoirs 

IN_resr_count <- IN_Resr_CPUE %>%
  filter(Total_Catch > 0) %>%
  group_by(Species) %>%
  summarise(N.resr=length(unique(Lake_Name))) %>%
  arrange(-N.resr) %>%
  mutate(Percent_lakes = N.resr/length(unique(IN_Resr_CPUE$Lake_Name))) %>%
  mutate(Species= as.factor(Species))%>%
  mutate(Species = fct_reorder(Species, Percent_lakes, .desc = TRUE)) #69 species
        

IN_all_Species_Reser <- ggplot(IN_resr_count, aes(Species, Percent_lakes))+
  geom_bar(stat="identity", fill="cyan3", colour="navyblue")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(y = "Proportion of reservoirs\n(out of 63)", title = "In All Species: Reservoirs")

IN_all_Species_Reser
```
Species Richness 
```{r}
species_count <- IN_Resr_CPUE %>% 
  filter(Total_Catch > 0) %>%
  group_by(Lake_Name) %>% 
  summarise(N.Species.res=length(unique(Species))) %>%
  arrange(-N.Species.res) %>%
  mutate(percent.species=N.Species.res/length(unique(fish.catch.effort$Species))) 

species_count2 <- species_count%>%
  mutate(Lake_Name= as.factor(Lake_Name))%>%
  mutate(Lake_Name = fct_reorder(Lake_Name, N.Species.res, .desc = TRUE))

IN_Sp_Richness_Resr <- ggplot(species_count2, aes(Lake_Name, N.Species.res))+
  geom_bar(stat="identity", fill="lightgreen", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=90, hjust=1, vjust=.5))+
  labs(y = "Number of species\n(out of 69)", title = "IN Species Richness Reservoirs", x = "Reservoir")

IN_Sp_Richness_Resr
```
Species Richness, highlight target species
```{r}
#extract species of interest
target_species <- c("black crappie", "bluegill", "largemouth bass", "northern pike", "smallmouth bass", "walleye", "yellow perch")

IN_resr_count_target <- IN_resr_count %>%
  mutate(target_species=case_when(Species%in%target_species~1, TRUE~0))


IN_All_sp_highlight <- ggplot(IN_resr_count_target, aes(Species, Percent_lakes))+
  geom_bar(stat="identity", colour="black", fill="orchid")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = "none")+
  labs(y = "Proportion of lakes\n(out of 63)", title = "IN All Species: Reservoirs")+
  gghighlight(target_species==1)


IN_All_sp_highlight
```

Species by Gear
```{r}
target_species_res <- c("black crappie", "bluegill", "largemouth bass", "northern pike", "smallmouth bass", "walleye", "yellow perch") #pasted from above

lake.count.gearR <- fish.catch.effort %>%
  filter(total.catch!=0) %>%
  group_by(Species, Gear) %>%
  summarise(N.lakes=length(unique(temp.id))) %>%
  arrange(-N.lakes, Gear) %>% 
  mutate(percent.res=N.lakes/length(unique(fish.catch.effort$temp.id))) %>%
  mutate(Species=factor(Species, levels=arrange(lake.count,-N.lakes)$Species))

lake.count.gearR_target=lake.count.gearR %>%
  mutate(Species = str_to_lower(Species))%>%
  mutate(target_species_res=case_when(Species %in% target_species_res~1, TRUE~0))%>%
  mutate(Gear = as.factor(Gear))


In_Sp_by_gear_plot <- ggplot(lake.count.gearR_target, aes(x = Species, y = percent.res, fill=Gear))+
  geom_bar(stat="identity", color="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(y="Proportion of reservoirs\n(out of 63)", title = "IN Reservoirs")+
  gghighlight(target_species_res==1, use_direct_label = FALSE, use_group_by = FALSE)+
  facet_wrap(.~Gear, ncol = 1)

In_Sp_by_gear_plot #the highlighted bars don't look right..? 
```


Just target Species by Gear
```{r}
IN_target_sp_gears_Resr <- lake.count.gearR_target %>%
  filter(target_species_res==1) 

IN_target_sp_gears_Resr_plot <- ggplot(data = IN_target_sp_gears_Resr, aes(Gear, percent.res, group=Gear, fill=Gear))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.8,.1))+
  ylab("Proportion of reservoirs\n(out of 63)")+
  facet_wrap(~Species)

IN_target_sp_gears_Resr_plot
```
Mean Catch by Gear
```{r}
Mean_CPUE_Resr <- IN_Resr_CPUE %>%
  filter(CPUE != 0) %>%
  group_by(Species, Gear_Abbrev) %>%
  summarise(Mean_non_0_CPUE = mean(CPUE), lci.non.zero.cpue=quantile(CPUE, .0275), uci.non.zero.cpue=quantile(CPUE, .975, na.rm=T))

Mean_CPUE_Resr2 <- Mean_CPUE_Resr %>%
  mutate(target_species_res=case_when(Species %in% target_species_res ~ 1, TRUE~0))%>%
  filter(target_species_res==1)

#just target species

IN_Mean_CPUE_by_gear_Resr <- ggplot(data = Mean_CPUE_Resr2, aes(Species, Mean_non_0_CPUE, group=Gear_Abbrev, color=Gear_Abbrev))+
  geom_pointrange(aes(ymax=uci.non.zero.cpue, ymin=lci.non.zero.cpue))+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = c(.95,.8))+
  facet_wrap(~Gear_Abbrev, scales="free_y")+
  labs(y = "Mean non-zero CPUE", color = "Gear")

IN_Mean_CPUE_by_gear_Resr
```
In the future, it may be nice to combine the lakes and reservoirs datasets to get the full picture.


IN lake and reservoir lists for temp
```{r}
IN_temp_list_lakes <-in.effort.community.data %>%
  mutate(Month = "June")%>%
  select(-Gear, -5)%>%
  group_by(Lake, County, Year, Month)%>%
  summarise(total = n())%>%
  select(-total)

write_csv(IN_temp_list_lakes, "IN_lake_list.csv")


#### Adding IN Lake IDs 9-7-22
IN_Lake_List <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/IN/2020FINFO_Lake_List.xlsx") #has IN lake IDs

IN_temp_list_lakes_with_ID <- IN_temp_list_lakes %>%
  mutate(County = str_to_title(County))%>%
  left_join(IN_Lake_List, by = c("Lake" = "Name", "County" = "County"))%>%
  select(-6, -7, -8, -9, -10, -11, -12)%>%
  filter(is.na(Id))#going to export as csv and manually fix :( 28 won't match

write_csv(IN_temp_list_lakes_with_ID, "IN_lk_temp_list_fix.csv")

IN_temp_list_resr <- IN_Resr_Effort %>%
  select(1:4)

write_csv(IN_temp_list_resr, "IN_reservoir_list.csv")
```

#Additional data activities for IN

IN Sp Map
```{r}
IN_Lake_Info <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/IN/IN_Lake_List2.csv") #manually fixed lake list
IN_Lat_Long <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/IN/2020FINFO_Lake_List.xlsx")

IN_Lat_Long_Basic <- IN_Lat_Long %>%
  rename(Lake_ID = Id)%>%
  select(Lake_ID, Latitude, Longitude)

IN_GL_CPUE <- fish.catch.effort %>%
  mutate(County = str_to_title(County))%>%
  left_join(IN_Lake_Info, by = c("Lake","County"))%>%
  rename(Lake_ID = Id, Total_Catch = total.catch)%>%
  ungroup()%>%
  select(Lake_ID, Species, Total_Catch)%>%
  left_join(IN_Lat_Long_Basic, by = "Lake_ID") #missing several lat longs
  
  
IN_Resr_CPUE2 <- IN_Resr_CPUE %>%
  rename(Lake_ID = Lake_ID_Number)%>%
  mutate(Species = str_to_title(Species))%>%
  ungroup()%>%
  select(Lake_ID, Species, Total_Catch, Latitude, Longitude)

IN_SP_Map <- IN_GL_CPUE %>%
  bind_rows(IN_Resr_CPUE2)%>%
  filter(Species %in% c("Black Crappie", "Bluegill", "Cisco", "Channel Catfish", "Largemouth Bass", "Northern Pike", "Smallmouth Bass", "Walleye", "White Crappie", "Yellow Perch"))%>%
   pivot_wider(names_from = Species, values_from = Total_Catch, values_fn = sum)
  
```

