---
title: "MI_Filtering"
author: "Holly Kundel"
date: '2022-06-13'
output: html_document
---
Load in required packages
```{r}
library(readr)
library(readxl)
library(dplyr)
library(lubridate)
library(tidyr)
library(googledrive)
library(ggplot2)
library(forcats)

options(scipen = 999) #ensures that survey ID numbers aren't rounded with scientific notation

```

Load in Raw Data from Drive Directly
- takes to long to go through all files. May incorporate this later.

This does NOT work, ignore for now
```{r}
Data_Folder <- drive_ls(path = "https://drive.google.com/drive/u/0/folders/1eHXdJvNMq46tBKURFTeViKWGdSBhWsL_")

MI_Folder <- drive_ls(path = "https://drive.google.com/drive/u/0/folders/1EYwXj3gRK5yavynOolig69C-5TaLhiOv")

MI_try1 <- drive_read_string(file = as_id("1E8M5VV2PII0ZKa_Yj7XtREOM_-VkH2scQJ2gBH7pVBQ"), type = "csv")
```

Load in Raw Data from my Computer
```{r}
MI_Effort_Raw <- read_excel("Data_and_Scripts/Data/Data_Summaries/MI/mi_snt_effort_data_mar2021.xlsx")

MI_Catch_Raw <-read_excel("Data_and_Scripts/Data/Data_Summaries/MI/mi_snt_catch_data_mar2021.xlsx")
```

Which gears catch which fish?
```{r}
MI_Catch <- ggplot(data = MI_Catch_Raw, aes(x = GEAR, fill = SPECIES))+
  geom_histogram(stat = "count")+
  facet_wrap(.~SPECIES)
  
MI_Catch

MI_Gear_Use <- ggplot(data = MI_Effort_Raw, aes(x = GEAR, fill = GEAR))+
  geom_histogram(stat = "count")

MI_Gear_Use
```
Note: we will likely drop "GLGNET" we guess this is "Great Lakes Gill Net" because it isn't hardly used, and therefore doesn't catch our species of interest often

Combine catch and effort to calculate CPUE
- First expand catch data to get list of all possible years, species, and gears
- Join above list with catch data, add 0s if no catch
- Join with effort data, only keep years/gears that were actually used and keep 0s where applicable. 


List of all possible years, species, and gears
```{r}
MI_all_potential_surveys <- MI_Catch_Raw %>%
  expand(NEW_KEY, SURVEY_YEAR, SPECIES, GEAR)
```

Adding 0s to the catch data
```{r}
MI_Catch_long <- MI_all_potential_surveys %>%
  left_join(MI_Catch_Raw, by = c("NEW_KEY", "SURVEY_YEAR", "SPECIES", "GEAR"))%>%
  mutate(FISH_COUNT =replace_na(FISH_COUNT, 0))%>%
  select(NEW_KEY, SURVEY_YEAR, SPECIES, GEAR, FISH_COUNT)
```

Join Catch and Effort Data
```{r}
MI_Catch_Effort <-MI_Catch_long %>%
  right_join(MI_Effort_Raw, by =  c("NEW_KEY", "SURVEY_YEAR", "GEAR"))%>%
  mutate(CPUE = FISH_COUNT/EFFORT)
```

Plotting summaries
```{r}
MI_Lakes_Num <- MI_Catch_Effort %>%
  group_by(NEW_KEY)%>%
  summarise(Total_Sp_Gear_Combos = n()) #464 Total Lakes


MI_Catch_Effort_Lakes <- MI_Catch_Effort %>%
  filter(FISH_COUNT > 0)%>%
  group_by(SPECIES, GEAR)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(MI_Catch_Effort$NEW_KEY)))

MI_Lake_prop <- ggplot(MI_Catch_Effort_Lakes, aes(x= GEAR, y = Prop_Lakes, fill = GEAR))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~SPECIES)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "MI fish catch and gear use", y="Proportion of lakes\n(out of 464)")
MI_Lake_prop
```

```{r}
MI_Catch_Effort_Species <- MI_Catch_Effort %>%
  filter(FISH_COUNT > 0)%>%
  group_by(SPECIES)%>%
  summarise(Num_Lks = n_distinct(NEW_KEY))%>%
  mutate(Prop_Lakes_SP = Num_Lks/length(unique(MI_Catch_Effort$NEW_KEY)))%>%
  mutate(Prop_Lakes_SP = round(Prop_Lakes_SP, digits = 2))%>%
  mutate(SPECIES= as.factor(SPECIES))%>%
  mutate(SPECIES = fct_reorder(SPECIES, Prop_Lakes_SP))

MI_Lake_prop <- ggplot(MI_Catch_Effort_Species, aes(x= SPECIES, y = Prop_Lakes_SP, fill = SPECIES))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  geom_text(aes(label = Prop_Lakes_SP), nudge_y = 0.05)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "MI fish catch by species", y="Proportion of lakes\n(out of 464)")
MI_Lake_prop
```
Mean non-zero CPUE
```{r}
MI_mean_CPUE_no_0 <- MI_Catch_Effort%>%
  filter(CPUE > 0)

MI_mean_CPUE_by_gear <- ggplot(MI_mean_CPUE_no_0, aes(x = SPECIES, y = CPUE, group = SPECIES, color = GEAR, fill = GEAR))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = GEAR), alpha = .25)+
  theme_classic()+
  facet_wrap(.~GEAR)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "MI CPUE by Gear", y = "non-zero CPUE (log scale)")
  
MI_mean_CPUE_by_gear

MI_mean_CPUE_by_gear2 <- ggplot(MI_mean_CPUE_no_0, aes(x = SPECIES, y = CPUE, group = SPECIES, fill = SPECIES))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = SPECIES), alpha = .25)+
  theme_classic()+
  facet_wrap(.~GEAR)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "MI CPUE by Gear", y = "non-zero CPUE (log scale)")
  
MI_mean_CPUE_by_gear2


```
Check to ensure that newly provided fish length data is already included in above data sets. 
```{r}
MI_fish_lengths <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/MI/mi_snt_catch_inchgrp_clean_0622.xls")
# more species than our 7 of interest. Lists survey, but not specific gear used. 
```


# Additional MI data activities

Get list of lakes/survey dates for temp data
```{r}
MI_Temp_List <- MI_Effort_Raw %>%
  select(Survey_Number, NEW_KEY, COUNTY, LAT_DD, LONG_DD, Water_Body_Name, SAMPLE_START_DATE)%>%
  group_by(Survey_Number, NEW_KEY, COUNTY, LAT_DD, LONG_DD, Water_Body_Name, SAMPLE_START_DATE)%>%
  summarise(Total = n())%>%
  select(-Total)

write_csv(MI_Temp_List, "MI_Lake_List.csv")
```
MI SP Map
```{r}
MI_SP_Map <- MI_Catch_Effort%>%
  rename(Lake_ID = NEW_KEY, Latitude = LAT_DD, Longitude = LONG_DD, Total_Catch = FISH_COUNT, Species = SPECIES)%>%
  select(Lake_ID, Species, Total_Catch, Latitude, Longitude)%>%
  pivot_wider(names_from = Species, values_from = Total_Catch, values_fn = sum)%>%
  rename("Black Crappie" = BCR, Bluegill = BLG, Cisco = CIS, "Largemouth Bass" = LMB, "Northern Pike" = NOP, "Smallmouth Bass" = SMB, Walleye = WAE, "Yellow Perch" = YEP)%>%
  mutate("Channel Catfish" = 0)%>%
  mutate("White Crappie" = 0)%>%
select(1,4,5,6, 12, 7, 8, 10, 13, 2,3)
```

