---
title: "MN_Filtering"
author: "Holly Kundel"
date: "5/13/2022"
output: github_document
---
Load in required packages
```{r}
library(readr)
library(readxl)
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(gghighlight)
library(colorspace)

options(scipen = 999) #ensures that survey ID numbers aren't rounded with scientific notation

```

##Load in raw  data

Data from Corey (received 5/3/22)
- has total catch and effort by species, not individual fish
```{r, warning=FALSE}
all_gn_tn <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/5.3.22_update/ALL-GNTN-SURVEYS.csv")

all_gde_gsh <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/5.3.22_update/ALL-GDEGSH-SURVEYS.csv")
```

Combine the two sheets from Corey into one, and create Effort data
```{r}

all_gn_tn_gde_gsh <- bind_rows(all_gn_tn, all_gde_gsh)

perch_target_gn_check <- all_gn_tn_gde_gsh %>%
  filter(GEAR %in% c("GN", "GDE", "GSH"))%>%
  filter(COMMON_NAME == "yellow perch")%>%
  filter(SURVEYTYPE == "Targeted Survey" | SURVEYTYPE == "Special Assessment") #693

All_Effort <- all_gn_tn_gde_gsh %>%
  select(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT)%>%
  group_by(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT)%>%
  summarise(NUMBER_OF_SP = n())%>%
  select(-NUMBER_OF_SP)%>%
  filter(SURVEYTYPE %in% c("Population Assessment", "Re-Survey", "Standard Survey", "Special Assessment", "Targeted Survey", "Large Lake Survey", "Research Survey", "Initial Survey"))

TN_Effort <- All_Effort %>%
  filter(GEAR == "TN")

GN_GDE_GSH_Effort <- All_Effort %>%
  filter(GEAR %in% c("GN", "GDE", "GSH"))

cols(all_gn_tn)

species_total <- all_gn_tn_gde_gsh %>%
  group_by(COMMON_NAME)%>%
  summarise(Total = n())
```

Load in Electrofishing Data
- From Corey Geving, via email 8/26/22

```{r}
# note, as of 12/22/22 these data have not been added to MN state summaries

EF_Catch_LMB_SMB <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/ELECTROFISHING_CATCH_LMB_SMB.csv")

EF_Effort_LMB_SMB <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/ELECTROFISHING_EFFORT.csv") #some non bass too


# get into cpue for purposes of quick summaries
EF_survey_catch_total <- EF_Catch_LMB_SMB %>%
  group_by(SURVEY_ID, SP)%>%
  summarize(TOTAL_CATCH = n())%>%
  mutate(COMMON_NAME = case_when(SP == "LMB" ~ "largemouth bass",
                                 SP == "SMB" ~ "smallmouth bass",
                                 TRUE ~ "NA"))

EF_LMB_SMB <- EF_Effort_LMB_SMB %>%
  left_join(EF_survey_catch_total, by = "SURVEY_ID")%>%
  drop_na(SP)%>%
  mutate(GEAR = "EF")%>%
  rename(LKNAME = WATER_BODY_NAME, DOW = DOW_KITTLE, SURVEYTYPE =SURVEY_TYPE_NAME, SURVEYDATE = SURVEY_ID_DATE, YEAR = SURVEY_YEAR)%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  select(SURVEY_ID, LKNAME, DOW, SURVEYTYPE, SURVEYDATE, COMMON_NAME, GEAR, CPUE, EFFORT, TOTAL_CATCH, YEAR )

write_csv(EF_LMB_SMB, "EF_LMB_SMB_14Feb23.csv")
```


MN Summaries
```{r}
all_mn <- all_gn_tn_gde_gsh %>%
  select(-FISHERIES_WATERBODY_ID)%>%
  bind_rows(EF_LMB_SMB)%>%
  filter(SURVEY_ID != 4421012345764001)%>% #this one survey has the columns shifted
  mutate(TARGET_SPECIES = case_when(COMMON_NAME == "bluegill" ~ "1",
                                    COMMON_NAME == "black crappie" ~ "1",
                                    COMMON_NAME == "largemouth bass" ~ "1",
                                    COMMON_NAME == "northern pike" ~ "1",
                                    COMMON_NAME == "smallmouth bass" ~ "1",
                                    COMMON_NAME == "walleye" ~ "1",
                                    COMMON_NAME == "yellow perch" ~ "1",
                                    TRUE ~ "0"))%>%
  mutate(GEAR_CATEGORY = case_when(GEAR == "GN" ~"GN",
                                   GEAR == "GDE" ~ "GN",
                                   GEAR == "GSH" ~ "GN",
                                   GEAR == "TN" ~ "TN",
                                   GEAR == "EF" ~ "EF"))

all_mn_sp <- all_mn %>% 
  group_by(COMMON_NAME)%>%
  summarise(Total = n()) #129 species (some aren't fish)
```



Plotting summaries
```{r}
MN_TOP_6 <- all_mn %>%
  filter(COMMON_NAME %in% c("bluegill", "black crappie", "largemouth bass", "northern pike", "smallmouth bass", "walleye", "yellow perch"))

MN_LK_Num <- MN_TOP_6 %>%
  group_by(DOW)%>%
  summarise(Total_gear_sp = n()) #2751 Total Lakes
MN_Lk_number <- length(MN_LK_Num)

MN_Catch_Effort_Lakes <- MN_TOP_6 %>%
  filter(TOTAL_CATCH > 0)%>%
  group_by(COMMON_NAME, GEAR)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(all_mn$DOW)))

MN_Lake_prop <- ggplot(MN_Catch_Effort_Lakes, aes(x= GEAR, y = Prop_Lakes, fill = GEAR))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~COMMON_NAME)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "MN fish catch and gear use", y="Proportion of lakes\n(out of 2,751)")
MN_Lake_prop
```

```{r}
MN_Gears_Grouped <- MN_TOP_6 %>%
  mutate(GEAR_CATEGORY = case_when(GEAR == "GN" ~"GN",
                                   GEAR == "GDE" ~ "GN",
                                   GEAR == "GSH" ~ "GN",
                                   GEAR == "TN" ~ "TN",
                                   GEAR == "EF" ~ "EF"))

MN_LK_Num2 <- MN_Gears_Grouped %>%
  group_by(DOW)%>%
  summarise(Total_gear_sp = n()) #2751 Total Lakes


MN_Catch_Effort_Lakes2 <- MN_Gears_Grouped %>%
  filter(TOTAL_CATCH > 0)%>%
  group_by(COMMON_NAME, GEAR_CATEGORY)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(all_mn$DOW)))

MN_Lake_prop <- ggplot(MN_Catch_Effort_Lakes2, aes(x= GEAR_CATEGORY, y = Prop_Lakes, fill = GEAR_CATEGORY))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~COMMON_NAME)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "MN fish catch and gear use", y="Proportion of lakes\n(out of 2,751)")
MN_Lake_prop
```
```{r}
MN_Catch_Effort_Species <- all_mn %>%
  filter(TOTAL_CATCH > 0)%>%
  group_by(COMMON_NAME)%>%
  summarise(Num_Lks = n_distinct(DOW))%>%
  mutate(Prop_Lakes_SP = Num_Lks/length(unique(all_mn$DOW)))%>%
  mutate(Prop_Lakes_SP = round(Prop_Lakes_SP, digits = 2))%>%
  mutate(SPECIES= as.factor(COMMON_NAME))%>%
  mutate(SPECIES = fct_reorder(SPECIES, Prop_Lakes_SP))%>%
  mutate(TARGET_SPECIES = case_when(COMMON_NAME == "bluegill" ~ "1",
                                    COMMON_NAME == "black crappie" ~ "1",
                                    COMMON_NAME == "largemouth bass" ~ "1",
                                    COMMON_NAME == "northern pike" ~ "1",
                                    COMMON_NAME == "smallmouth bass" ~ "1",
                                    COMMON_NAME == "walleye" ~ "1",
                                    COMMON_NAME == "yellow perch" ~ "1",
                                    TRUE ~ "0"))%>%
  filter(Prop_Lakes_SP >= 0.05)

all_mn_lk <- all_mn %>%
  group_by(DOW)%>%
  summarise(Total_gear_sp = n()) #2909 Total Lakes

MN_Lake_prop <- ggplot(MN_Catch_Effort_Species, aes(x= SPECIES, y = Prop_Lakes_SP, fill = SPECIES))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  geom_text(aes(label = Prop_Lakes_SP), nudge_y = 0.05, size = 2.5)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = "none")+
  labs(title = "Proportion of MN lakes with each species", y="Proportion of lakes\n(out of 2,909)")+
  gghighlight(TARGET_SPECIES == 1)

MN_Lake_prop
```

```{r}
MN_mean_CPUE_no_0 <- all_mn %>%
  filter(CPUE > 0)%>%
  filter(TARGET_SPECIES == 1)

MN_mean_CPUE_by_gear <- ggplot(MN_mean_CPUE_no_0, aes(x = COMMON_NAME, y = CPUE, group = COMMON_NAME, fill = COMMON_NAME))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = COMMON_NAME), alpha = .05)+
  theme_classic()+
  facet_wrap(.~GEAR_CATEGORY)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "MN CPUE by Gear", y = "non-zero CPUE (log scale)")
  
MN_mean_CPUE_by_gear
```


```{r}
MN_mean_CPUE_no_0_TN_GN <- all_mn %>%
  filter(CPUE > 0)%>%
  filter(TARGET_SPECIES == 1)%>%
  filter(GEAR_CATEGORY %in% c("GN", "TN"))

MN_mean_CPUE_by_gear_TN_GN <- ggplot(MN_mean_CPUE_no_0_TN_GN, aes(x = COMMON_NAME, y = CPUE, group = COMMON_NAME, fill = COMMON_NAME))+
  geom_jitter(aes(color = COMMON_NAME), alpha = .01)+
  geom_boxplot(color = "black")+
  theme_bw()+
  facet_wrap(.~GEAR_CATEGORY)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "MN CPUE by Gear", y = "non-zero CPUE (log scale)", x = "Species", fill = "Species", color = "Species")+
  scale_fill_discrete_qualitative(palette = "Set 2")+
  scale_color_discrete_qualitative(palette = "Set 2")
  
MN_mean_CPUE_by_gear_TN_GN
```



```{r}
MN_CPUE_no_0 <- all_mn %>%
  filter(CPUE > 0)

MN_mean_CPUE_by_gear <- ggplot(MN_CPUE_no_0, aes(x = COMMON_NAME, y = CPUE, group = COMMON_NAME, fill = COMMON_NAME))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = COMMON_NAME), alpha = .05)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = "none")+
  scale_y_log10()+
  labs(title = "MN CPUE by Gear", y = "non-zero CPUE (log scale)")
  
MN_mean_CPUE_by_gear
```

 Single species plots
```{r}
BLG_Gears <- MN_TOP_6 %>%
  filter(COMMON_NAME == "bluegill")%>%
  filter(CPUE > 0)

BLG_CPUE_by_gear <- ggplot(BLG_Gears, aes(x = COMMON_NAME, y = CPUE, group = GEAR_CATEGORY, fill = GEAR_CATEGORY))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = GEAR_CATEGORY), alpha = .05)+
  theme_classic()+
  facet_wrap(.~GEAR_CATEGORY)+
  theme(axis.title.x=element_blank())+
  scale_y_log10()+
  labs(title = "Bluegill CPUE by Gear", y = "non-zero CPUE (log scale)", fill = "Gear Type", color= "Gear Type", x = "Bluegill")
  
BLG_CPUE_by_gear

BLC_Gears <- MN_TOP_6 %>%
  filter(COMMON_NAME == "black crappie")%>%
  filter(CPUE > 0)

BLC_CPUE_by_gear <- ggplot(BLC_Gears, aes(x = COMMON_NAME, y = CPUE, group = GEAR_CATEGORY, fill = GEAR_CATEGORY))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = GEAR_CATEGORY), alpha = .05)+
  theme_classic()+
  facet_wrap(.~GEAR_CATEGORY)+
  theme(axis.title.x=element_blank())+
  scale_y_log10()+
  labs(title = "Black Crappie CPUE by Gear", y = "non-zero CPUE (log scale)", fill = "Gear Type", color= "Gear Type", x = "Black Crappie")
  
BLC_CPUE_by_gear

LMB_Gears <- MN_TOP_6 %>%
  filter(COMMON_NAME == "black crappie")%>%
  filter(CPUE > 0)%>%
  filter(GEAR_CATEGORY %in% c("GN", "TN"))

LMB_CPUE_by_gear <- ggplot(BLC_Gears, aes(x = COMMON_NAME, y = CPUE, group = GEAR_CATEGORY, fill = GEAR_CATEGORY))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = GEAR_CATEGORY), alpha = .05)+
  theme_classic()+
  facet_wrap(.~GEAR_CATEGORY)+
  theme(axis.title.x=element_blank())+
  scale_y_log10()+
  labs(title = "Largemouth Bass CPUE by Gear", y = "non-zero CPUE (log scale)", fill = "Gear Type", color= "Gear Type", x = "Largemouth Bass")
  
LMB_CPUE_by_gear
```



# Single Species Filtering

This was used for Ty's PGA work, I don't think individual scripts for each species makes sense though


BLUEGILL
- Only want trap net surveys, standard and targeted okay
- Targeted surveys need to be filtered for time of year/water temp

```{r}
BLG_MN <- all_gn_tn_gde_gsh %>%
  select(SURVEY_ID, GEAR, YEAR, COMMON_NAME, TOTAL_CATCH)%>%
  filter(COMMON_NAME == "bluegill")%>%
  filter(GEAR == "TN")%>%
  right_join(TN_Effort, by = c("SURVEY_ID", "GEAR", "YEAR"))%>%
  mutate(SURVEY_CATEGORY = case_when(SURVEYTYPE == "Standard Survey" ~ "Standard",
                                     SURVEYTYPE == "Population Assessment" ~ "Standard",
                                     SURVEYTYPE == "Re-Survey" ~ "Standard",
                                     SURVEYTYPE == "Initial Survey" ~ "Standard",
                                     SURVEYTYPE == "Large Lake Survey" ~ "Standard",
                                     SURVEYTYPE == "Targeted Survey" ~ "Targeted",
                                     SURVEYTYPE == "Special Assessment" ~ "Targeted",
                                     SURVEYTYPE == "Research Survey" ~ "Targeted")) %>%
  mutate(TOTAL_CATCH = replace_na(TOTAL_CATCH, 0))%>%
  mutate(COMMON_NAME = replace_na(COMMON_NAME, "bluegill"))%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  mutate(SURVEY_DATE = mdy(SURVEYDATE))%>%
  mutate(SURVEY_MONTH = month(SURVEY_DATE))

#write_csv(BLG_MN, "BLG_MN.csv")  #need to filter targeted surveys for time of year/water temp
  
  
  
```

YELLOW PERCH
- Only want gill net data, want GN, GDE, and GSH
- Maybe (not included rn) targeted fall TN surveys too..?
```{r}
YEP_MN <- all_gn_tn_gde_gsh %>%
  select(SURVEY_ID, GEAR, YEAR, COMMON_NAME, TOTAL_CATCH)%>%
  filter(COMMON_NAME == "yellow perch")%>%
  filter(GEAR %in% c("GN", "GDE", "GSH"))%>%
  right_join(GN_GDE_GSH_Effort, by = c("SURVEY_ID", "GEAR", "YEAR"))%>%
  mutate(SURVEY_CATEGORY = case_when(SURVEYTYPE == "Standard Survey" ~ "Standard",
                                     SURVEYTYPE == "Population Assessment" ~ "Standard",
                                     SURVEYTYPE == "Re-Survey" ~ "Standard",
                                     SURVEYTYPE == "Initial Survey" ~ "Standard",
                                     SURVEYTYPE == "Large Lake Survey" ~ "Standard",
                                     SURVEYTYPE == "Targeted Survey" ~ "Targeted",
                                     SURVEYTYPE == "Special Assessment" ~ "Targeted",
                                     SURVEYTYPE == "Research Survey" ~ "Targeted")) %>%
  mutate(TOTAL_CATCH = replace_na(TOTAL_CATCH, 0))%>%
  mutate(COMMON_NAME = replace_na(COMMON_NAME, "yellow perch"))%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  mutate(SURVEY_DATE = mdy(SURVEYDATE))%>%
  mutate(SURVEY_MONTH = month(SURVEY_DATE))

#write_csv(YEP_MN, "YEP_MN.csv")
```

CISCO
- Only want gill net data, unsure if there is enough data...
```{r}
TLC_MN <- all_gn_tn_gde_gsh %>%
  select(SURVEY_ID, GEAR, YEAR, COMMON_NAME, TOTAL_CATCH)%>%
  filter(COMMON_NAME == "tullibee (cisco)")%>%
  filter(GEAR %in% c("GN", "GDE", "GSH"))%>%
  right_join(GN_GDE_GSH_Effort, by = c("SURVEY_ID", "GEAR", "YEAR"))%>%
  mutate(SURVEY_CATEGORY = case_when(SURVEYTYPE == "Standard Survey" ~ "Standard",
                                     SURVEYTYPE == "Population Assessment" ~ "Standard",
                                     SURVEYTYPE == "Re-Survey" ~ "Standard",
                                     SURVEYTYPE == "Initial Survey" ~ "Standard",
                                     SURVEYTYPE == "Large Lake Survey" ~ "Standard",
                                     SURVEYTYPE == "Targeted Survey" ~ "Targeted",
                                     SURVEYTYPE == "Special Assessment" ~ "Targeted",
                                     SURVEYTYPE == "Research Survey" ~ "Targeted")) %>%
  mutate(TOTAL_CATCH = replace_na(TOTAL_CATCH, 0))%>%
  select(-COMMON_NAME)%>%
  mutate(COMMON_NAME = "cisco")%>% #making this easier to work with
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  mutate(SURVEY_DATE = mdy(SURVEYDATE))%>%
  mutate(SURVEY_MONTH = month(SURVEY_DATE))%>%
  select(1:3, 11, 4:10, 12:14) #reordering to match YEP and BLG

write_csv(TLC_MN, "TLC_MN.csv")
```

Combine BLG, YEP, and TLC data
```{r}
BLG_YEP_TLC_MN <- bind_rows(BLG_MN, YEP_MN, TLC_MN)
write_csv(BLG_YEP_TLC_MN, "BLG_YEP_TLC_MN.csv")
```

BLACK CRAPPIE
- Only want trap net surveys, standard and targeted okay
- Targeted surveys need to be filtered for time of year/water temp

```{r}
BLC_MN <- all_gn_tn_gde_gsh %>%
  select(SURVEY_ID, GEAR, YEAR, COMMON_NAME, TOTAL_CATCH)%>%
  filter(COMMON_NAME == "black crappie")%>%
  filter(GEAR == "TN")%>%
  right_join(TN_Effort, by = c("SURVEY_ID", "GEAR", "YEAR"))%>%
  mutate(SURVEY_CATEGORY = case_when(SURVEYTYPE == "Standard Survey" ~ "Standard",
                                     SURVEYTYPE == "Population Assessment" ~ "Standard",
                                     SURVEYTYPE == "Re-Survey" ~ "Standard",
                                     SURVEYTYPE == "Initial Survey" ~ "Standard",
                                     SURVEYTYPE == "Large Lake Survey" ~ "Standard",
                                     SURVEYTYPE == "Targeted Survey" ~ "Targeted",
                                     SURVEYTYPE == "Special Assessment" ~ "Targeted",
                                     SURVEYTYPE == "Research Survey" ~ "Targeted")) %>%
  mutate(TOTAL_CATCH = replace_na(TOTAL_CATCH, 0))%>%
  mutate(COMMON_NAME = replace_na(COMMON_NAME, "black crappie"))%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  mutate(SURVEY_DATE = mdy(SURVEYDATE))%>%
  mutate(SURVEY_MONTH = month(SURVEY_DATE))

#write_csv(BLG_MN, "BLG_MN.csv")
```

NORTHERN PIKE
- Only want GN data
```{r}

```

LARGEMOUTH BASS
- Using GN for now, but really we want EF data
```{r}

```

SMALLMOUTH BASS
- Using GN for now, but really we want EF data
```{r}

```

WALLEYE
- Only want GN data
```{r}

```

End single species filtering

_____________________________________________________________________________________________________________________________________________________
First draft of MN Filtering workflow
- Later hope to incorporate Paul's code which pulls the files directly from Google Drive
```{r}
MN_Fish_Catch <- all_gn_tn_gde_gsh %>%
  filter(COMMON_NAME %in% c("black crappie", "bluegill", "largemouth bass", 
                            "northern pike", "smallmouth bass", "walleye", "yellow perch"))%>%
  filter(SURVEYTYPE %in% c("Standard Survey", "Population Assessment", "Re-Survey", 
                           "Initial Survey", "Large Lake Survey", "Targeted Survey", "Special Assessment", "Research Survey"))%>%
  mutate(gear_check = case_when(COMMON_NAME == "black crappie" & GEAR == "TN" ~ 1,
                                COMMON_NAME == "black crappie" & GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" ~ 0,
                                COMMON_NAME == "bluegill" & GEAR == "TN" ~ 1,
                                COMMON_NAME == "bluegill" & GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" ~ 0,
                                COMMON_NAME == "largemouth bass"& GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" |GEAR == "TN" ~ 0,
                                COMMON_NAME == "northern pike" & GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" ~ 1,
                                COMMON_NAME == "northern pike" & GEAR == "TN" ~ 0,
                                COMMON_NAME == "smallmouth bass"& GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" |GEAR == "TN" ~ 0,
                                COMMON_NAME == "walleye" & GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" ~ 1,
                                COMMON_NAME == "walleye" & GEAR == "TN" ~ 0,
                                COMMON_NAME == "yellow perch" & GEAR == "GN" | GEAR == "GDE"| GEAR == "GSH" ~ 1,
                                COMMON_NAME == "yellow perch" & GEAR == "TN" ~ 0))
# check targeted surveys
#check temp and time of year
```



-----------------------------------------------------------------------------------------------------------------------------------------------------
older data, but this has individual fish listed for each line, instead of catch totals by species for each net

GN (note GDE = deep set GN and GSH = shallow set GN)
```{r, warning = FALSE, message=FALSE}
GN1 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region1_GN_raw_fish.csv")
GN2 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region2_GN_raw_fish.csv")
GN3 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region3_GN_raw_fish.csv")
GN4 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region4_GN_raw_fish.csv")

# note open excel sheet, and format the "SURVEY_ID" column as "number" with no decimal places, and re-save sheet
GND <- read_xlsx("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/All_Regions_GDE_raw_fish.xlsx")
GNS <- read_xlsx("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/All_Regions_GSH_raw_fish.xlsx")
```

TN
```{r, warning = FALSE, message=FALSE}
TN1 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region1_TN_raw_fish.csv")
TN2 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region2_TN_raw_fish.csv")
TN3 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region3_TN_raw_fish.csv")
TN4 <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/Region4_TN_raw_fish.csv")
```

GN summary
```{r}
GN_region1 <- GN1 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first (cuts out 448,000+ fish)
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>% #select identifying columns and group by species
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>% #sum up catch by species
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>% #set up date parsing with lubridate pkg
  mutate(YEAR = year(SAMPLE_DATE))%>% #separate out year
  mutate(MONTH = month(SAMPLE_DATE))%>% #separate out month
  mutate(GEAR_TYPE = "GN") #create gear type column so when combined with TN we can distinguish them

GN_region2 <- GN2 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first (cuts out 448,000+ fish)
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "GN")

GN_region3 <- GN3 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first (cuts out 448,000+ fish)
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "GN")

GN_region4 <- GN4 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first (cuts out 448,000+ fish)
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "GN")

GN_deep <- GND %>%
  mutate(ID_NBR = as.character(ID_NBR))%>%
  mutate(ID_NBR = str_pad(ID_NBR, 8, side = "left", pad ="0"))%>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first (cuts out 448,000+ fish)
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = ymd(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "GDE")

GN_shallow <- GNS %>%
  mutate(ID_NBR = as.character(ID_NBR))%>%
  mutate(ID_NBR = str_pad(ID_NBR, 8, side = "left", pad ="0"))%>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first (cuts out 448,000+ fish)
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = ymd(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "GSH")

GN_All_MN <- bind_rows(GN_region1, GN_region2, GN_region3, GN_region4, GN_deep, GN_shallow) #122,016
```

Trap Net summary
```{r}
TN_region1 <- TN1 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first 
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "TN")

TN_region2 <- TN2 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first 
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "TN")

TN_region3 <- TN3%>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first 
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "TN")

TN_region4 <- TN4 %>%
  filter(SP %in% c("BLC", "BLG", "LMB", "NOP", "SMB", "WAE", "YEP"))%>% #condense by keeping species of interest first 
  group_by(SURVEY_ID, ID_NBR, SAMPLE_DATE, SP)%>%
  summarise(TOTAL_CATCH = sum(FISH_COUNT))%>%
  mutate(SAMPLE_DATE = mdy(SAMPLE_DATE))%>%
  mutate(YEAR = year(SAMPLE_DATE))%>%
  mutate(MONTH = month(SAMPLE_DATE))%>%
  mutate(GEAR_TYPE = "TN")

TN_All_MN <- bind_rows(TN_region1, TN_region2, TN_region3, TN_region4)
```

Get list of lakes for Chris to pull in temp data for
```{r}
MN_Temp_List <- all_gn_tn_gde_gsh %>%
  select(SURVEY_ID, LKNAME, DOW, SURVEYDATE)%>%
  group_by(SURVEY_ID, LKNAME, DOW, SURVEYDATE)%>%
  summarise(Total = n())%>%
  select(SURVEY_ID, LKNAME, DOW, SURVEYDATE)

write_csv(MN_Temp_List, "MN_Lake_List.csv")
```
MN SP Map
```{r}
mn_lake_list_info <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/MN/mn_lake_list_info.csv")

MN_Lat_Longs <- mn_lake_list_info %>%
  select(DOW, LAKE_CENTER_LAT_DD5, LAKE_CENTER_LONG_DD5)%>%
  rename(Lake_ID = DOW, Latitude = LAKE_CENTER_LAT_DD5, Longitude = LAKE_CENTER_LONG_DD5)

MN_SP_MAP <- all_mn %>%
  rename(Lake_ID = DOW, Species = COMMON_NAME, Total_Catch = TOTAL_CATCH)%>%
  select(Lake_ID, Species, Total_Catch)%>%
  mutate(Species = str_to_title(Species))%>%
  filter(Species %in% c("Black Crappie", "Bluegill", "Tullibee (Cisco)", "Channel Catfish", "Largemouth Bass", "Northern Pike", "Smallmouth Bass", "Walleye", "White Crappie", "Yellow Perch"))%>%
  pivot_wider(names_from = Species, values_from = Total_Catch, values_fn = sum)%>%
  rename(Cisco = "Tullibee (Cisco)")%>%
  select(1,5,4,8,9,7,6,10,3,11,2)%>%
  left_join(MN_Lat_Longs, by = "Lake_ID")
```

