---
title: "MN_Fish_Data_for_Chris_18Apr2023"
author: "Holly Kundel"
date: '`r Sys.Date()`'
output: html_document
---

Load in required packages
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(data.table)
library(ggplot2)
```



Read in raw MN Data
```{r, warning=FALSE}
# Load in gillnet and trapnet data (direct from source or current versions in directory)
# direct from source
# GN_TN <- read_csv("E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_fish_effort_03May2022.csv")
# 
# #Load in shallow and deep gillnets, same kind of net, different gear code
# 
# GDE_GSH <- read_csv("E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_gde_gsh_fish_effort_03May2022.csv")


# as an RData file (this will need to be updated if source files change)
# save(GN_TN, GDE_GSH, file = "Data_and_Scripts/Data/Input/DNRcatchfiles.RData")
# 
# load("Data_and_Scripts/Data/Input/DNRcatchfiles.RData")

```

Combine GN, TN, GDE, GSH into one df
```{r}
all_GN_TN <- bind_rows(GN_TN, GDE_GSH)
```


Determine Number of unique surveys/gears and number of total species
```{r}
Effort_GN_TN <-  all_GN_TN %>%
  select( SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT)%>%
  group_by( SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT)%>%
  summarise(NUMBER_OF_SP = n())%>%
  select(-NUMBER_OF_SP)%>%
  filter(SURVEYTYPE %in% c("Population Assessment",
                           "Re-Survey",
                           "Standard Survey",
                           "Special Assessment",
                           "Targeted Survey",
                           "Large Lake Survey",
                           "Research Survey",
                           "Initial Survey"))%>% #only keep representative survey types
  mutate(SURVEY_ID_GEAR = str_c(SURVEY_ID, GEAR, sep = "_"))

#21,207 unique survey/gears (i.e. a survey may be listed more than once if more than one gear was used, each row is a unique survey and gear)

Num_species <- all_GN_TN %>%
  filter(SURVEYTYPE %in% c("Population Assessment",
                           "Re-Survey",
                           "Standard Survey",
                           "Special Assessment",
                           "Targeted Survey",
                           "Large Lake Survey",
                           "Research Survey",
                           "Initial Survey"))%>% #only keep representative survey types
  group_by(COMMON_NAME)%>%
  summarise(Total_Number_observed = sum(TOTAL_CATCH)) #129 total species, note not all of them are fish

Cisco_name_check <- Num_species%>%
  filter(str_detect(Num_species$COMMON_NAME, "cisco"))
```


Add instances of zero catch

* quick note about lake IDs (DOW), there are some streams/rivers listed and they have letters in the DOW ID and vary in length
  * creating a new column `WATERBODY_TYPE` to address this in case it matters for analysis
```{r}
MN_Catch_0 <- all_GN_TN %>%
  filter(SURVEYTYPE %in% c("Population Assessment",
                           "Re-Survey",
                           "Standard Survey",
                           "Special Assessment",
                           "Targeted Survey",
                           "Large Lake Survey",
                           "Research Survey",
                           "Initial Survey"))%>% #only keep representative survey types
  mutate(WATERBODY_TYPE = if_else(str_detect(DOW, "[:alpha:]"), paste("river/stream"), paste("lake")))%>%
  select(-CPUE)%>% #better to calculate CPUE ourselves
  pivot_wider(id_cols = c(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT, WATERBODY_TYPE),
              names_from = COMMON_NAME, values_from = TOTAL_CATCH, values_fill = 0) %>% #creates 0s, 21,207 surveys/gears matches check above
  select("SURVEY_ID",    #select columns to keep and
         "SURVEYDATE",
         "SURVEYTYPE",
         "DOW",
         "LKNAME",
         "YEAR",
         "GEAR",
         "EFFORT",
         "WATERBODY_TYPE",
         "black crappie",    #select species of interest
         "bluegill",
         "largemouth bass",
         "northern pike",
         "smallmouth bass",
         "walleye",
         "yellow perch",                 
         "tullibee (cisco)", #2364 instances    # note that cisco species are listed three ways in the data
         "cisco species",  # 163 instances
         "shortjaw cisco", # 1 instance
         "white sucker",
         "black bullhead",
         "brown bullhead",
         "yellow bullhead")%>%
  pivot_longer(cols = !c(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT, WATERBODY_TYPE),
               names_to = "COMMON_NAME", values_to = "TOTAL_CATCH")%>%
  mutate(GEAR_CATEGORY = case_when(GEAR == "GN" ~"GN",
                                   GEAR == "GDE" ~ "GN",
                                   GEAR == "GSH" ~ "GN",
                                   GEAR == "TN" ~ "TN",
                                   GEAR == "EF" ~ "EF"))%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  mutate(CPUE = round(CPUE, digits = 2)) #14 species times 21207 survey/gears = 296,898 rows 

#write_csv(MN_Catch_0, "MN_Fish_Data_21JUN2023.csv")

setDT(MN_Catch_0)


#tabling and summarizing for data checking
MN_Catch_0[ , .(n_surveyswithspecieslisted = .N)  , COMMON_NAME ] # number of surveys for each species (here we've included all zeros)
MN_Catch_0[ TOTAL_CATCH >0 , .(nsurveyswithspecies = .N) , COMMON_NAME ] # number of surveys where each species was caught (here we've nixed all zeros) this works b/c we don't have species dups with any "survey" (date,lake,gear, surveytype)
MN_Catch_0[ , .(n_indiv_obs = sum(TOTAL_CATCH)) , COMMON_NAME ][order(-n_indiv_obs)] # how many fish observations are in here for each species? #after our operation this still matches the Num_species file
MN_Catch_0[ , .N , .(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT, WATERBODY_TYPE,YEAR)  ][ , hist(YEAR) , ] # count of surveys per year
MN_Catch_0[ , .N , .(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT, WATERBODY_TYPE,YEAR)  ][ , .(nsurveys = .N, firstyear = min(YEAR), lastyear = max(YEAR)) , .(DOW, LKNAME) ] # summary of surveys by lake

MN_Catch_0[ , SURVEYDATE := as.IDate(SURVEYDATE, format = "%m/%d/%Y"), ]

MN_Catch_0[ TOTAL_CATCH > 0 , length(unique(DOW)) , COMMON_NAME ] # how many unique lakes was each species found in?

MN_Catch_0[ , CPUE := TOTAL_CATCH/EFFORT ,]

ggplot(MN_Catch_0, aes(SURVEYDATE, TOTAL_CATCH))+
  geom_point()+
  facet_wrap(~COMMON_NAME)

ggplot(MN_Catch_0, aes(SURVEYDATE, CPUE))+
  geom_point()+
  facet_wrap(~COMMON_NAME)+
  geom_smooth()

# now with non-zero ones

ggplot(MN_Catch_0[TOTAL_CATCH>0], aes(SURVEYDATE, CPUE))+
  geom_point()+
  facet_wrap(~COMMON_NAME)+
  geom_smooth()+
  scale_y_log10()

#compare CPUE by gears
ggplot(MN_Catch_0[TOTAL_CATCH>0], aes(SURVEYDATE, CPUE, color = GEAR_CATEGORY))+
  geom_point(alpha = .1)+
  facet_wrap(~COMMON_NAME)+
  geom_smooth()+
  scale_y_log10()










```

