---
title: "MN_Data_for_Chris"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---
This data is for Chris' work on his PGA modeling in MN

Load in required packages
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
```


Data Chris is currently using for reference
```{r}
Chris_MN_fish_3.23.22 <- read_csv("~/CASC/CASC_Fish_Abundance/Practice_MN_Fish_3.25.22.csv")
```

Columns:
- DOW
- COMMON_NAME
- SURVEYDATE
- TOTAL_CATCH
- EFFORT
- CPUE
- GEAR

Read in raw MN Data
```{r, warning=FALSE}
GN_TN <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/5.3.22_update/ALL-GNTN-SURVEYS.csv")

GDE_GSH <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/5.3.22_update/ALL-GDEGSH-SURVEYS.csv")

Electrofishing_catch <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/ELECTROFISHING_CATCH_LMB_SMB.csv")

Electrofishing_effort <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/ELECTROFISHING_EFFORT.csv")
```

Combine GN, TN, GDE, GSH into one df
```{r}
all_GN_TN <- bind_rows(GN_TN, GDE_GSH)

# create separate df of just surveys aka they survey effort
Effort_GN_TN <-  all_GN_TN %>%
  select(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT)%>%
  group_by(SURVEY_ID, SURVEYDATE, SURVEYTYPE, DOW, LKNAME, YEAR, GEAR, EFFORT)%>%
  summarise(NUMBER_OF_SP = n())%>%
  select(-NUMBER_OF_SP)%>%
  filter(SURVEYTYPE %in% c("Population Assessment", "Re-Survey", "Standard Survey", "Special Assessment", "Targeted Survey", "Large Lake Survey", "Research Survey", "Initial Survey"))%>%
  mutate(SURVEY_ID_GEAR = str_c(SURVEY_ID, GEAR, sep = "_"))
```

Add instances of zero catch to all GN (GDE and GSH included) and TN didn't work
```{r}
Wide_catch <- all_GN_TN %>%
  select(SURVEY_ID, GEAR, COMMON_NAME, TOTAL_CATCH)%>%
  mutate(ID = str_c(SURVEY_ID, GEAR, COMMON_NAME))%>%
  pivot_wider(id_cols = ID, names_from = COMMON_NAME, values_from = TOTAL_CATCH, values_fill = 0)

#mutate(SURVEY_ID_GEAR = str_c(SURVEY_ID, GEAR, sep = "_"))%>% select(SURVEY_ID_GEAR, COMMON_NAME, TOTAL_CATCH) %>%
```

```{r}
Expanded_Catch <- all_GN_TN %>%
  select(SURVEY_ID, GEAR, COMMON_NAME, TOTAL_CATCH) %>%
  mutate(SURVEY_ID_GEAR = str_c(SURVEY_ID, GEAR, sep = "_"))%>%
  select(SURVEY_ID_GEAR, COMMON_NAME, TOTAL_CATCH) %>%
  expand(SURVEY_ID_GEAR, COMMON_NAME)#note some dates appear in the COMMON_NAME column..?

Catch <- all_GN_TN %>%
  select(SURVEY_ID, GEAR, COMMON_NAME, TOTAL_CATCH) %>%
  mutate(SURVEY_ID_GEAR = str_c(SURVEY_ID, GEAR, sep = "_"))%>%
  select(SURVEY_ID_GEAR, COMMON_NAME, TOTAL_CATCH)

Catch_w_zeros <- Expanded_Catch %>%
  left_join(Catch, by = c("SURVEY_ID_GEAR", "COMMON_NAME"))%>%
  mutate(TOTAL_CATCH = replace_na(TOTAL_CATCH, 0))

CPUE_GN_TN_all_sp <- Catch_w_zeros %>%
  left_join(Effort_GN_TN, by = "SURVEY_ID_GEAR")%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  filter(COMMON_NAME != "7/12/2010")

CPUE_GN_TN_target_sp <- CPUE_GN_TN_all_sp %>%
  filter(COMMON_NAME %in% c("bluegill", "black crappie", "largemouth bass", "northern pike", "smallmouth bass", "walleye", "yellow perch"))%>%
  rename(GEAR_SPECIFIC = GEAR)%>%
  mutate(GEAR = case_when(GEAR_SPECIFIC == "GN" ~"GN",
                                   GEAR_SPECIFIC == "GDE" ~ "GN",
                                   GEAR_SPECIFIC == "GSH" ~ "GN",
                                   GEAR_SPECIFIC == "TN" ~ "TN"))%>%
  select(DOW, COMMON_NAME, SURVEYDATE, TOTAL_CATCH, EFFORT, CPUE, GEAR, GEAR_SPECIFIC)
  
```

Add covariates
```{r}
mn_lake_list_info <- read_csv("Data_and_Scripts/Data/Data_Summaries/MN/mn_lake_list_info.csv")

MN_lake_info <- mn_lake_list_info %>%
  select(DOW, MAX_DEPTH_FEET, LAKE_AREA_GIS_ACRES, LAKE_CENTER_UTM_EASTING, LAKE_CENTER_UTM_NORTHING, LAKE_CENTER_LAT_DD5, LAKE_CENTER_LONG_DD5)

MN_Fish_GN_TN <- CPUE_GN_TN_target_sp %>%
  left_join(MN_lake_info, by = "DOW")

write_csv(MN_Fish_GN_TN, "MN_Fish_GN_TN_Jan.12.23.csv")
```



Below is a horrible failure of using the expand function
```{r}
MN_all_potential_surveys <- all_GN_TN %>%
  expand(DOW, YEAR, GEAR, COMMON_NAME) #329,472,000 total
  
# If we decide to try more species, add those here
MN_expand_target_sp <- MN_all_potential_surveys %>%
  filter(COMMON_NAME %in% c("bluegill", "black crappie", "largemouth bass", "northern pike", "smallmouth bass", "walleye", "yellow perch")) %>% #17,740,800
  filter(GEAR %in% c("GN", "GDE", "GSH", "TN")) #3,548,160 was putting sp name in gear category..?

Catch_GN_TN <- all_GN_TN %>%
  select(1:3, 6,7,10,12)

Catch_long_GN_TN <- MN_expand_target_sp %>%
  left_join(Catch_GN_TN, by = c("DOW", "YEAR", "GEAR"))%>%
  mutate(TOTAL_CATCH = replace_na(TOTAL_CATCH, 0))%>%
  pivot_longer(cols = c(COMMON_NAME.x, COMMON_NAME.y), names_to = "Name", values_to = "COMMON_NAME")%>%
  drop_na(COMMON_NAME)%>%
  select(-Name)%>%
  distinct(.keep_all = TRUE)%>%
  select(DOW, YEAR, GEAR, TOTAL_CATCH, COMMON_NAME)%>%
  group_by(DOW, YEAR, GEAR, COMMON_NAME)%>%
  summarise(TOTAL_CATCH = sum(TOTAL_CATCH))


GN_TN_CPUE <- Effort_GN_TN %>%
  left_join(Catch_long_GN_TN, by = c("DOW", "YEAR", "GEAR"))
```

```{r}
EF_catch_total <- Electrofishing_catch %>%
  group_by(SURVEY_ID, SP)%>%
  summarize(TOTAL_CATCH = n())%>%
  mutate(COMMON_NAME = case_when(SP == "LMB" ~ "largemouth bass",
                                 SP == "SMB" ~ "smallmouth bass",
                                 TRUE ~ "NA"))

EF_LMB_SMB <- Electrofishing_effort %>%
  left_join(EF_catch_total, by = "SURVEY_ID")%>%
  drop_na(SP)%>%
  mutate(GEAR = "EF")%>%
  rename(LKNAME = WATER_BODY_NAME, DOW = DOW_KITTLE, SURVEYTYPE =SURVEY_TYPE_NAME, SURVEYDATE = SURVEY_ID_DATE, YEAR = SURVEY_YEAR)%>%
  mutate(CPUE = TOTAL_CATCH/EFFORT)%>%
  select(SURVEY_ID, LKNAME, DOW, SURVEYTYPE, SURVEYDATE, COMMON_NAME, GEAR, CPUE, EFFORT, TOTAL_CATCH, YEAR )%>%
  left_join(MN_lake_info, by = "DOW")


write_csv(EF_LMB_SMB, "MN_EF_LMB_SMB_Jan.12.23.csv")
```

Combine GN, TN, and EF
* change gear to GN, TN, EF, and have more specific under new column "gear_specific"
```{r}
# Note survey ID 4421012345764001 is entered wrong, remove it
```


checking MN DoWs that don't have an NHD ID 1/18/23
```{r}
noNHDID <- read_csv("Data_and_Scripts/Data/noNHDID.csv")

no_NHD <- noNHDID %>%
  mutate(DOW_8 = str_sub(DOW, -8))%>%
  mutate(DOW_08 = str_replace(DOW_8, "_", "0"))%>%
  mutate(child_DOW = str_detect(DOW_08, "[:digit:][:digit:][:digit:][:digit:][:digit:][:digit:][:digit:][1-9]"))%>%
  filter(child_DOW == TRUE)

write_csv(no_NHD, "no_NHD_short.csv")
```

