---
title: "MN_Flat_File_March23"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---

# Code from
2/2/23
First attempt to use the CASC Data Explainer File to rename columns and create units columns where needed. Uses MN data for test case
## Load libraries
```{r}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(arrow)
library(data.table)
library(googledrive)
library(janitor)
library(tidyr)
library(data.table)
library(tidyr)

```

#Testing Mike's WI code on MN data
```{r}
# find files on Google Drive Desktop
MN_files_list <- list.files(path = "G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files

MN_files_list #check that file names look correct

n <- length(MN_files_list)

for(i in 1:n) {
  #i = 1
  assign(gsub(".csv","", MN_files_list[i]),
         data.table(read_csv_arrow(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/",
                                          MN_files_list[i]))))
  
  #consider moving renaming into here!
  
  
  #this does those two steps in one package
# note we want to review a sorted list of column names to check misspelling etc.
cde %>% # call data explainer file
  filter(`new_file_name`== gsub(".csv","", MN_files_list)[i])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.))) %>% 
  transpose(keep.names = "newname") %>% 
  rename("oldname" = V1) %>% 
  assign("names", ., envir = .GlobalEnv)

#see if any column names will not have a match! 
# IF any pop FALSE, force stop and revist of data explainer ()
# - e.g., named something "total catch" when actual column name was "total_catch"
print(
cbind(colnames(get(gsub(".csv","", MN_files_list)[i])),
      colnames(get(gsub(".csv","", MN_files_list)[i])) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ]
)
)

# break the loop if the current file has column names not in the data explainer
if (all(cbind(colnames(get(gsub(".csv","", MN_files_list)[i])),  colnames(get(gsub(".csv","", MN_files_list)[i])) %in% names[ !str_detect(newname,"unique_row_key"), oldname, ])[,2]) == FALSE ) break

#now rename that file's colnames
setnames(get(gsub(".csv","", MN_files_list)[i]), colnames(get(gsub(".csv","", MN_files_list)[i])), names[!str_detect(newname,"unique_row_key")] [match(names(get(gsub(".csv","", MN_files_list)[i])),names[!str_detect(newname,"unique_row_key"),oldname]), newname] )

#confirm import of files:  
print(paste(gsub(".csv","", MN_files_list)[i] ,"added to workspace" ))  

}  

# gsub(".csv","", MN_files_list[i])

```
```




#_________________________________________________________________________________________________________________________________________________
Work below this line was pre 3/14/2023

## CASC Data Explainer
### Read in
```{r}
# change E (Mike's drive) to G (Holly's drive)
cde_raw <- data.table(read_excel("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/CASC_Data_Explainer_for_R_Data_Manipulation.xlsx")) #set as DT on the fly

```

### Prep 

* Data Explainer prep is now done in 'CASC_Data_Explainer_HLK.Rmd'

## State data
### Read in

We eventually want to be able to read this in from google drive directly, right now uses google drive desktops

```{r, warning = FALSE, message = FALSE}  
MN_files_list <- list.files(path = "G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data", pattern = ".+\\.csv") #only grab csv files

n <- length(MN_files_list) #may want to make it more than just n so it is different for each state??? e.g. "n_mn" ?

# Code from Mike
for(i in 1:n) {
  assign(gsub(".csv","", MN_files_list[i]),
         read_csv(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/",
                               MN_files_list[i]))) #don't use 'read.csv2, only get one column
} 

# get the list of data sets as data frames in R so they can be looped later

# plus by adding this, you won't have to remove the ".csv" every time you reference the list
MN_file_names <- str_replace(MN_files_list, ".csv", "")   #str_replace and gsub both work the same :) 

```

Process to fix file read in error, no need to run code in this chunk
```{r, echo=FALSE}

#This doesn't work because 'escape_double = TRUE'

for(i in 1:n) {
  assign(gsub(".csv","", MN_files_list[i]),
         data.table(read_csv_arrow(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/",
                               MN_files_list[i])))) #don't use 'read.csv2, only get one column
} 


#This doesn't work, only loads mn-gde_gsh_fish_effort okay, others don't work

for(i in 1:n) {
  assign(gsub(".csv","", MN_files_list[i]),
         data.table(read_csv_arrow(escape_double = FALSE,(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/",
                               MN_files_list[i]))))) #don't use 'read.csv2, only get one column
} 

# problem file is 'mn_gde_gsh...'

### fix mn_gde_gsh, was in an improper csv format
#fixed by reading into r with 'read_csv' from 'readr' pkg, and then using 'write_csv'

mn_gde_gsh <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/mn_gde_gsh_fish_effort_03May2022.csv")

#write_csv(mn_gde_gsh, "mn_gde_gs_fish_effort_03May2022_2.csv")
```


Get new column names from data explainer for all files (OUT of Date, see below)
```{r, echo=FALSE}
# THIS METHOD IS OUT OF DATE, SEE BELOW

#for a state file, display column names - 

MN_file_names <- gsub(".csv","", MN_files_list)

#oldnames
old_names <- cde %>% # call data explainer file
  filter(`new_file_name`== MN_file_names[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  make_clean_names() %>% 
  assign("oldnames", ., envir = .GlobalEnv)


# select(14:25) #keep only columns of relevance (remove our file tracking and urls)

# display our new names for that set - (esp. need to ID no matches)

cde %>% # call data explainer file
  filter(`new_file_name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  colnames() %>% 
  assign("newnames", ., envir = .GlobalEnv)

cbind(oldnames,newnames)

cde %>% # call data explainer file
  filter(`new_file_name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.))) %>% 
  transpose(keep.names = "rn")

colnames(get(gsub(".csv","", WI_files_list)[1]))cde %>% # call data explainer file
  filter(`new_file_name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  colnames() %>% 
  assign("newnames", ., envir = .GlobalEnv)

cbind(oldnames,newnames)
```

New Version of getting new column names from data explainer
```{r}
#code from Mike's WI flat-file work
# note we want to review a sorted list of column names to check misspelling etc.
column_names <- cde %>% # call data explainer file
  filter(`new_file_name`== MN_file_names[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.))) %>% 
  transpose(keep.names = "newname") %>% 
  rename("oldname" = V1) %>% 
  assign("names", ., envir = .GlobalEnv) # left with two columns of the original file names 'oldname' and the new data explainer names 'newname'


#see if any column names will not have a match! 
# IF any pop FALSE, force stop and revist the data explainer 
# - e.g., named something "total catch" when actual column name was "total_catch"

cbind(colnames(get(MN_file_names[1])),
      colnames(get(MN_file_names[1])) %in% names[ , oldname, ]) #"FALSE" means that our typed name is not a correct column name

# Once all columns are "TRUE" you can move on

names <- names[ match(colnames(get(MN_file_names[1])),
       names[!str_detect(newname,"unique_row_key"),oldname]), ] #check that there are no NAs

### pull in new code from Mike WI work file
```


