---
title: "SD_Filtering"
author: "Holly Kundel"
date: '2022-07-01'
output: html_document
---
Load in required packages
```{r}
library(readr)
library(readxl)
library(dplyr)
library(lubridate)
library(tidyr)
#library(googledrive)
library(ggplot2)
library(forcats)
library(colorspace) #for more color palettes

options(scipen = 999) #ensures that survey ID numbers aren't rounded with scientific notation

```

Load in Raw Data from my Computer ()
```{r, warning = FALSE}
SD_Fish <- read_excel("Data_and_Scripts/Data/Data_Summaries/SD/sd_lake_survey_fish_data_totalfish.xlsx")

SD_Effort <- read_excel("Data_and_Scripts/Data/Data_Summaries/SD/sd_lake_survey_fish_data_surveys.xlsx")

SD_NOP_Data <-read_excel("Data_and_Scripts/Data/Data_Summaries/SD/SD_NOP_DATA.xlsx", sheet = "LAKE_SURVEYS") #received separately 11/3/22

SD_Gears <- SD_Effort %>%
  group_by(Method)%>%
  summarise(total = n())

write_csv(SD_Gears, "SD_Gears.csv")

SD_Species <- SD_Fish %>%
  group_by(SpeciesName)%>%
  summarise(Total = n())
```

Data process:
- Since the fish catch data doesn't have year info, will need to combine with effort
- then expand list
- join expanded list to catch (adding 0s where none are caught)
- then join with effort to remove survey gears that weren't used

Create list of all possible lakes, years, species, and gears using `expand` so we can easily add 0s to catch data
- This list is NOT to calculate CPUE

Abandoning Attempt 1 and trying a new approach
```{r}
SD_Lake_List <- SD_Effort %>%
  full_join(SD_Fish, by = "SurveyID")

#check for missing catch data
SD_missing_catch <- SD_Lake_List %>%
  filter(is.na(FishCount))%>%
  distinct(SurveyID) #5894

#check for missing Effort
SD_missing_Effort <- SD_Lake_List %>%
  filter(is.na(Effort)) #0

#expand list
SD_Lake_List_Long <- SD_Lake_List %>%
  mutate(SurveyDate = date(SurveyDate))%>%
  mutate(Year = year(SurveyDate))%>%
  drop_na(FishCount)%>%
  expand(StateID, Year, SpeciesName, Method)

# add year to effort data
SD_Effort_Year <- SD_Effort %>%
  mutate(SurveyDate = date(SurveyDate))%>%
  mutate(Year = year(SurveyDate))

# combine expanded list to catch data to add 0s
SD_Fish_Long <- SD_Lake_List_Long %>%
  right_join(SD_Effort_Year, by = c("StateID","Year", "Method"))

```

Attempt 2
- because catch data has no tie to lake, year, or gear it is very hard to work with. It will likely be easiest to add the lake, gear, and year of the survey to the catch data via the survey ID.
```{r}
# create list of Lakes, Gears, and Years for each survey
SD_Survey_ID_list <- SD_Effort %>%
  mutate(SurveyDate = date(SurveyDate))%>%
  mutate(Year = year(SurveyDate))%>%
  select(SurveyID, SurveyDate, Year, StateID, Method, MethodCode)

#add info to the catch data
SD_Fish_plus <- SD_Fish %>%
  left_join(SD_Survey_ID_list, by = "SurveyID")
  
SD_Fish_Long <- SD_Fish_plus %>%
  expand()
```



```{r}
SD_Gear_Use <- ggplot(data = SD_Effort, aes(x = Method, fill = Method))+
  geom_histogram(stat = "count")

SD_Gear_Use
```
```{r}
SD_Surverys_Lk_yr <- SD_Effort_Year %>%
  group_by(StateID, Year)%>%
  summarise(Surveys_Per_Lk_Yr = n()) #2801

SD_Surverys_Lk_yr2 <- SD_Effort_Year %>%
  group_by(StateID, Method, Year)%>%
  summarise(Surveys_Per_method_Yr = n()) #5123

SD_Surveys_per_lake_per_yr <- ggplot(SD_Surverys_Lk_yr, aes(x = Surveys_Per_Lk_Yr))+
  geom_histogram(stat = "count")
SD_Surveys_per_lake_per_yr
```

```{r}
SD_FMA_ <- SD_Effort %>%
  group_by(FMA)%>%
  summarise(Total= n())

SD_Areas_Pierre <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_Areas.xlsx", sheet = "Pierre")

SD_Areas_Webster <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_Areas.xlsx", sheet = "Webster")

SD_Areas_Sioux_Falls <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_Areas.xlsx", sheet = "Sioux_Falls")

SD_Areas_Rapid_City <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_Areas.xlsx", sheet = "Rapid_City")

SD_Areas_Chamberlain <- read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_Areas.xlsx", sheet = "Chamberlain")
```

```{r}
Pierre <- SD_Areas_Pierre %>%
  select(Waterbodies, County)%>%
  left_join(SD_State_ID, by = c("Waterbodies" = "Waterbody", "County" = "County"))%>%
  distinct(Waterbodies, County, StateID)%>%
  rename(Waterbody = Waterbodies)%>%
  mutate(Area = "Pierre")

Webster <- SD_Areas_Webster%>%
  mutate(Area = "Webster")

Sioux_Falls <- SD_Areas_Sioux_Falls %>%
  select(Lake, County)%>%
  rename(Waterbody = Lake)%>%
  left_join(SD_State_ID, by = c("Waterbody", "County"))%>%
  distinct(Waterbody,County, StateID)%>%
  mutate(Area = "Sioux Falls")

Rapid_City <- SD_Areas_Rapid_City %>%
  rename(Waterbody = Waterbodies)%>%
  left_join(SD_State_ID, by = c("Waterbody", "County"))%>%
  distinct(Waterbody,County, StateID)%>%
  mutate(Area = "Rapid City")

Chamberlain <- SD_Areas_Chamberlain %>%
  rename(Waterbody = Waterbodies)%>%
  left_join(SD_State_ID, by = "Waterbody")%>%
  distinct(Waterbody, County, StateID)%>%
  select(Waterbody, County, StateID)%>%
  mutate(Area = "Chamberlain")

```

```{r}
SD_State_ID <- SD_Effort %>%
  select(StateID, Waterbody, County)

SD_Areas_IDs <- bind_rows(Pierre, Webster, Sioux_Falls, Rapid_City, Chamberlain)
write_csv(SD_Areas_IDs, "SD_Areas_IDs.csv")
```

SD lake list for temp
```{r}
SD_Temp_list <- SD_Effort %>%
  select(SurveyDate, StateID, Waterbody, County)%>%
  group_by(SurveyDate, StateID, Waterbody, County)%>%
  summarise(Total = n())%>%
  select(-Total)

write_csv(SD_Temp_list, "SD_Lake_List.csv")
```

Missing lake IDs were manually entered using information from the effort data and managed lakes list from SD
```{r}
SD_Area_Lake_List_Complete <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_Areas_IDs_Corrected.csv")
```

Overcomplicated it, trying a new method
```{r}
SD_Years <- tibble(Year = 1983:2021)

SD_Sp <- tibble(Species = c("Bluegill", "Black Crappie", "Largemouth Bass", "Northern Pike", "Smallmouth Bass", "Walleye", "Yellow Perch"))%>%
  mutate(Year = 1983)

SD_Expand <- SD_Years %>% 
  full_join(SD_Sp, by = "Year") %>%
  expand(Year, Species)%>%
  drop_na(Species) #not sure if I need this anymore

SD_Effort_Expand <- SD_Effort_Year %>%
  select(-Station)%>%
  full_join(SD_Expand, by = "Year")
  
#add all years to each lake, then re-join to effort and caatch
SD_Area_Lake_List_Complete_Years <- SD_Area_Lake_List_Complete %>% 
  bind_rows(SD_Expand)%>%
  select(-Comments)%>%
  expand(Waterbody, County, StateID, Area, Year, Species) #48 Gb vector... too big
```

```{r}
SD_CPUE <- SD_Effort_Year %>%
  full_join(SD_Fish_plus, by = "SurveyID")%>%
  mutate(SurveyDate = case_when(is.na(SurveyDate.x) ~ SurveyDate.y,
                                is.na(SurveyDate.y) ~ SurveyDate.x,
                                TRUE ~ SurveyDate.x))%>%
  mutate(StateID = case_when(is.na(StateID.x) ~ StateID.y,
                                is.na(StateID.y) ~ StateID.x,
                                TRUE ~ StateID.x))%>%
  mutate(Method = case_when(is.na(Method.x) ~ Method.y,
                                is.na(Method.y) ~ Method.x,
                                TRUE ~ Method.x))%>%
  mutate(MethodCode = case_when(is.na(MethodCode.x) ~ MethodCode.y,
                                is.na(MethodCode.y) ~ MethodCode.x,
                                TRUE ~ MethodCode.x))%>%
  select(SurveyDate, StateID, Waterbody, FMA, County, Method, MethodCode, Station, Effort, Year, SpeciesName, FishCount)%>%
  drop_na(SpeciesName)
```

Join effort to catch because each station has catch listed wiith unique survey IDs,
combine stations of same gear effort and combine their catch
```{r}
 #will add zeros for all species in next step

SD_CPUE_Grouped <- SD_CPUE %>%
  group_by(SurveyDate, Year, StateID, Waterbody, Method, MethodCode, SpeciesName)%>%
  summarise(Effort_Total = sum(Effort), Total_Fish = sum(FishCount))

SD_Gears <- SD_CPUE %>%
  group_by(Method, MethodCode)%>%
  summarise(Total = n())

SD_Long <- SD_CPUE_Grouped %>%
  mutate(Gear_Category = case_when(MethodCode == "9" ~ "EF",
                                   MethodCode == "10" ~ "EF",
                                   MethodCode == "11" ~ "EF",
                                   MethodCode == "16" ~ "EF",
                                   MethodCode == "21" ~ "TN",
                                   MethodCode == "23" ~ "TN",
                                   MethodCode == "24" ~ "TN",
                                   MethodCode == "30" ~ "GN",
                                   MethodCode == "31" ~ "GN",
                                   MethodCode == "35" ~ "GN",
                                   MethodCode == "99" ~ "GN",
                                   MethodCode == "111" ~ "EF",
                                   MethodCode == "112" ~ "EF",
                                   MethodCode == "113" ~ "EF",
                                   MethodCode == "114" ~ "EF",
                                   MethodCode == "116" ~ "EF",
                                   MethodCode == "212" ~ "TN",
                                   MethodCode == "213" ~ "TN",
                                   MethodCode == "310" ~ "GN",
                                   MethodCode == "311" ~ "GN",
                                   MethodCode == "312" ~ "GN",
                                   MethodCode == "314" ~ "GN",
                                   MethodCode == "315" ~ "GN",
                                   MethodCode == "316" ~ "GN",
                                   MethodCode == "317" ~ "GN",
                                   MethodCode == "550" ~ "GN",
                                   MethodCode == "551" ~ "TN",
                                   MethodCode == "552" ~ "GN"))


#expand year, StateID, me4thod, species
SD_Long_2 <- SD_Long %>%
  expand(Year, StateID, Gear_Category, SpeciesName)
```
Taking a break from Gretchen's method of adding zeros, will need to revisit this... Making summaries so we can reach back out to agency, we tend to drop zeros when making these any way..

Plotting summaries
```{r}
SD_Lakes_Num <- SD_Long %>%
  group_by(StateID)%>%
  summarise(Total_Sp_Gear_Combos = n()) #365 Total Lakes


SD_Catch_Effort_Lakes <- SD_Long %>%
  filter(Total_Fish > 0)%>%
  group_by(SpeciesName, Gear_Category)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(SD_Long$StateID)))

SD_Lake_prop <- ggplot(SD_Catch_Effort_Lakes, aes(x= Gear_Category, y = Prop_Lakes, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~SpeciesName)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD fish catch and gear use", y="Proportion of lakes\n(out of 365)")
SD_Lake_prop
```

```{r}
SD_Catch_Effort_Lakes3 <- SD_Long %>%
  filter(Total_Fish > 0)%>%
  group_by(SpeciesName, Method, Gear_Category)%>%
  summarise(Number_of_Fish = sum(Total_Fish))

SD_CE_WAE <- SD_Catch_Effort_Lakes3%>%
  filter(SpeciesName == "Walleye")%>%
  mutate(Method = as.factor(Method))%>%
  mutate(Gear_Category = as.factor(Gear_Category))%>%
  mutate(Method = fct_reorder(Method, Number_of_Fish, .desc = TRUE))

SD_Gear_WAE <- ggplot(data = SD_CE_WAE, aes(x= fct_reorder(Method, Number_of_Fish, .desc = TRUE), y = Number_of_Fish, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD Walleye Catch by Gear", y="Total Number of Fish Caught in Each Gear", x = "Detailed Gear Name", fill= "Gear Category")+
  scale_fill_discrete_sequential(palette = "Plasma", nmax = 5, order = c(1,2,4))

SD_Gear_WAE
```

```{r}
SD_CE_BLC <- SD_Catch_Effort_Lakes3%>%
  filter(SpeciesName == "Black Crappie")%>%
  mutate(Method = as.factor(Method))%>%
  mutate(Gear_Category = as.factor(Gear_Category))%>%
  mutate(Method = fct_reorder(Method, Number_of_Fish, .desc = TRUE))

SD_Gear_BLC <- ggplot(data = SD_CE_BLC, aes(x= fct_reorder(Method, Number_of_Fish, .desc = TRUE), y = Number_of_Fish, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD Black Crappie Catch by Gear", y="Total Number of Fish Caught in Each Gear", x = "Detailed Gear Name", fill= "Gear Category")+
  scale_fill_discrete_sequential(palette = "Plasma", nmax = 5, order = c(1,2,4))

SD_Gear_BLC
```
```{r}
SD_CE_BLG <- SD_Catch_Effort_Lakes3%>%
  filter(SpeciesName == "Bluegill")%>%
  mutate(Method = as.factor(Method))%>%
  mutate(Gear_Category = as.factor(Gear_Category))%>%
  mutate(Method = fct_reorder(Method, Number_of_Fish, .desc = TRUE))

SD_Gear_BLG <- ggplot(data = SD_CE_BLG, aes(x= fct_reorder(Method, Number_of_Fish, .desc = TRUE), y = Number_of_Fish, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD Bluegill Catch by Gear", y="Total Number of Fish Caught in Each Gear", x = "Detailed Gear Name", fill= "Gear Category")+
  scale_fill_discrete_sequential(palette = "Plasma", nmax = 5, order = c(1,2,4))

SD_Gear_BLG
```

```{r}
SD_CE_LMB <- SD_Catch_Effort_Lakes3%>%
  filter(SpeciesName == "Largemouth Bass")%>%
  mutate(Method = as.factor(Method))%>%
  mutate(Gear_Category = as.factor(Gear_Category))%>%
  mutate(Method = fct_reorder(Method, Number_of_Fish, .desc = TRUE))

SD_Gear_LMB <- ggplot(data = SD_CE_LMB, aes(x= fct_reorder(Method, Number_of_Fish, .desc = TRUE), y = Number_of_Fish, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD Largemouth Bass Catch by Gear", y="Total Number of Fish Caught in Each Gear", x = "Detailed Gear Name", fill= "Gear Category")+
  scale_fill_discrete_sequential(palette = "Plasma", nmax = 5, order = c(1,2,4))

SD_Gear_LMB
```

```{r}
SD_CE_SMB <- SD_Catch_Effort_Lakes3%>%
  filter(SpeciesName == "Smallmouth Bass")%>%
  mutate(Method = as.factor(Method))%>%
  mutate(Gear_Category = as.factor(Gear_Category))%>%
  mutate(Method = fct_reorder(Method, Number_of_Fish, .desc = TRUE))

SD_Gear_SMB <- ggplot(data = SD_CE_SMB, aes(x= fct_reorder(Method, Number_of_Fish, .desc = TRUE), y = Number_of_Fish, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD Smallmouth Bass Catch by Gear", y="Total Number of Fish Caught in Each Gear", x = "Detailed Gear Name", fill= "Gear Category")+
  scale_fill_discrete_sequential(palette = "Plasma", nmax = 5, order = c(1,2,4))

SD_Gear_SMB
```

```{r}
SD_CE_YEP <- SD_Catch_Effort_Lakes3%>%
  filter(SpeciesName == "Yellow Perch")%>%
  mutate(Method = as.factor(Method))%>%
  mutate(Gear_Category = as.factor(Gear_Category))%>%
  mutate(Method = fct_reorder(Method, Number_of_Fish, .desc = TRUE))

SD_Gear_YEP <- ggplot(data = SD_CE_YEP, aes(x= fct_reorder(Method, Number_of_Fish, .desc = TRUE), y = Number_of_Fish, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black")+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD Yellow Perch Catch by Gear", y="Total Number of Fish Caught in Each Gear", x = "Detailed Gear Name", fill= "Gear Category")+
  scale_fill_discrete_sequential(palette = "Plasma", nmax = 5, order = c(1,2,4))

SD_Gear_YEP
```

```{r}
SD_Catch_Effort_Species <- SD_Long %>%
  filter(Total_Fish > 0)%>%
  group_by(SpeciesName)%>%
  summarise(Num_Lks = n_distinct(StateID))%>%
  mutate(Prop_Lakes_SP = Num_Lks/length(unique(SD_Long$StateID)))%>%
  mutate(Prop_Lakes_SP = round(Prop_Lakes_SP, digits = 2))%>%
  mutate(SpeciesName= as.factor(SpeciesName))%>%
  mutate(SpeciesName = fct_reorder(SpeciesName, Prop_Lakes_SP))

SD_Lake_prop <- ggplot(SD_Catch_Effort_Species, aes(x= SpeciesName, y = Prop_Lakes_SP, fill = SpeciesName))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  geom_text(aes(label = Prop_Lakes_SP), nudge_y = 0.05)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD fish catch by species", y="Proportion of lakes\n(out of 365)")
SD_Lake_prop
```

```{r}
SD_CPUE <- SD_Long %>%
  mutate(CPUE = Total_Fish/Effort_Total)%>%
  rename(Species = SpeciesName)

SD_mean_CPUE_by_gear <- ggplot(SD_CPUE, aes(x = Species, y = CPUE, group = Species, fill = Species))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = Species), alpha = .05)+
  theme_classic()+
  facet_wrap(.~Gear_Category)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "SD CPUE by Gear", y = "non-zero CPUE (log scale)")
  
SD_mean_CPUE_by_gear
```


Additional Summaries August 22
```{r}
SD_Catch_Effort_Lakes2 <- SD_Long %>%
  filter(Total_Fish > 0)%>%
  group_by(SpeciesName, Method)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(SD_Long$StateID)))

SD_Lake_prop <- ggplot(SD_Catch_Effort_Lakes, aes(x= Gear_Category, y = Prop_Lakes, fill = Gear_Category))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~SpeciesName)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "SD fish catch and gear use", y="Proportion of lakes\n(out of 365)")
SD_Lake_prop
```

# Additional SD Data Activities

SD Species Map
```{r}
SD_NOP_catch <-read_excel("~/CASC/CASC_Fish_Abundance/Data_Summaries/SD/SD_NOP_DATA.xlsx", sheet = "NOP_AGE_GROWTH")

SD_SP_Map <- SD_Long %>%
  select(StateID, SpeciesName, Total_Fish)%>%
  rename(Lake_ID = StateID, Species=SpeciesName, Total_Catch = Total_Fish)
  
  
  
  rename(Lake_ID = NEW_KEY, Latitude = LAT_DD, Longitude = LONG_DD, Total_Catch = FISH_COUNT, Species = SPECIES)%>%
  select(Lake_ID, Species, Total_Catch, Latitude, Longitude)%>%
  pivot_wider(names_from = Species, values_from = Total_Catch, values_fn = sum)%>%
  rename("Black Crappie" = BCR, Bluegill = BLG, Cisco = CIS, "Largemouth Bass" = LMB, "Northern Pike" = NOP, "Smallmouth Bass" = SMB, Walleye = WAE, "Yellow Perch" = YEP)%>%
  mutate("Channel Catfish" = 0)%>%
  mutate("White Crappie" = 0)%>%
select(1,4,5,6, 12, 7, 8, 10, 13, 2,3)
```

