---
title: "WI_Filtering"
author: "Holly Kundel"
date: '2022-07-29'
output: html_document
---
Load in Packages
```{r}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(forcats)
library(arrow)

options(scipen = 999) #ensures that survey ID numbers aren't rounded with scientific notation
```


Load in Data AND remove streams and creeks, etc.
```{r}
WI_Fish_Data <- read_csv_arrow("Data_and_Scripts/Data/Data_Summaries/WI/wdnr_inland_fish_data.csv") #this is a very large file, so using arrow to load


# create new df so we don't alter the original raw data, will also select the columns we need.
WI_Fish_Dat_lakes <- WI_Fish_Data %>%
  select(wbic, waterbody.name)%>% #select columns of interest
  group_by(wbic, waterbody.name)%>% #group all fish from the same date
  summarise(Total_Fish = n()) #3859 lakes with fish data

#read in lat longs for wbics
WI_wbic_lat_longs <- read_csv("Data_and_Scripts/Data/Data_Summaries/WI/wi_lake_wbic_lat_long_PNF.csv")

# use 'left_join' to add the lat/longs to the lakes where we have fish data
WI_Lakes <- WI_Fish_Dat_lakes %>%
  left_join(WI_wbic_lat_longs, by = c("wbic" = "lake.id"))%>%
  rename(Latitude = lat, Longitude = long)%>%
  mutate(Lake_ID_Type = "WBIC")%>% #useful when combining with other states data
  mutate(State = "WI")%>% #will help distinguish other states when we combine
  rename(Lake_ID = wbic)%>% #creating a common column name to use with all states
  rename(Waterbody_Name = waterbody.name)%>% #rename to common column name
  select(Lake_ID, Lake_ID_Type, Waterbody_Name, State, Latitude, Longitude)%>% #re-order and keep columns of interest
  mutate(Lake_ID = as.character(Lake_ID)) #the other lake IDs are characters, so changing data type to match others

#attempt to remove creeks, streams, and rivers, not perfect, but non-lakes should be dropped when added to fish data
WI_Lakes3 <- WI_Lakes %>%
  mutate(River = str_detect(Waterbody_Name, "river")) %>%
  mutate(Creek = str_detect(Waterbody_Name, "creek"))%>%
  mutate(Canal = str_detect(Waterbody_Name, "canal"))%>%
  mutate(Stream = str_detect(Waterbody_Name, "stream"))%>%
  mutate(Brook = str_detect(Waterbody_Name, "brook"))%>% #added 7/29
  mutate(Flowage = str_detect(Waterbody_Name, "flowage"))%>% #added 7/29
  mutate(Channel = str_detect(Waterbody_Name, "channel"))%>% #added 7/29
  mutate(Ditch = str_detect(Waterbody_Name, "ditch"))%>% #added 7/29
  mutate(Lotic = case_when(River == TRUE ~ 1,
                           Creek == TRUE ~ 1,
                           Canal == TRUE ~ 1,
                           Stream == TRUE ~ 1,
                           Brook == TRUE ~ 1,
                           Flowage == TRUE ~ 1,
                           Channel == TRUE ~ 1,
                           Ditch == TRUE ~ 1,
                           TRUE ~ 0)) %>%
  filter(Lotic == 0) # Now 2048 after removing 7/29 additions, was 2195 instead of 3859

```

```
Winnebago_trawl_check <- WI_Fish_Data %>%
  select(1:15)%>%
  filter(wbic == 131100)%>%
  group_by(gear)%>%
  summarise(total = n()) #WI Fish Data only has electofishing and fyke netting for Lake Winnebago, so the trawl data will need to be added in separately
```

Do the above but with the effort data
```{r}
WI_Effort_Data <- read_csv("Data_and_Scripts/Data/Data_Summaries/WI/wdnr_inland_effort_data.csv")

# create new df so we don't alter the original raw data, will also select the columns we need.
WI_Effort_Dat_lakes <- WI_Effort_Data %>%
  select(wbic, waterbody.name, sample.date)

# create list of WI Gears for Jenna
WI_Gears <- WI_Effort_Data %>%
  group_by(gear)%>%
  summarise(Total=n())

#write_csv(WI_Gears, "WI_Gears")

#read in lat longs for wbics
WI_wbic_lat_longs <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/WI/wi_lake_wbic_lat_long_PNF.csv")

# use 'left_join' to add the lat/longs to the lakes where we have fish data
WI_Effort_lat_longs <- WI_Effort_Dat_lakes %>%
  left_join(WI_wbic_lat_longs, by = c("wbic" = "lake.id"))%>%
  rename(Latitude = lat, Longitude = long, Waterbody_Name = waterbody.name)%>%
  select(-state)

#attempt to remove creeks, streams, and rivers
WI_Effort_Lakes <- WI_Effort_lat_longs %>%
  mutate(River = str_detect(Waterbody_Name, "river")) %>%
  mutate(Creek = str_detect(Waterbody_Name, "creek"))%>%
  mutate(Canal = str_detect(Waterbody_Name, "canal"))%>%
  mutate(Stream = str_detect(Waterbody_Name, "stream"))%>%
  mutate(Brook = str_detect(Waterbody_Name, "brook"))%>% #added 7/29
  mutate(Flowage = str_detect(Waterbody_Name, "flowage"))%>% #added 7/29
  mutate(Channel = str_detect(Waterbody_Name, "channel"))%>% #added 7/29
  mutate(Ditch = str_detect(Waterbody_Name, "ditch"))%>% #added 7/29
  mutate(Lotic = case_when(River == TRUE ~ 1,
                           Creek == TRUE ~ 1,
                           Canal == TRUE ~ 1,
                           Stream == TRUE ~ 1,
                           Brook == TRUE ~ 1,
                           Flowage == TRUE ~ 1,
                           Channel == TRUE ~ 1,
                           Ditch == TRUE ~ 1,
                           TRUE ~ 0)) %>%
  filter(Lotic == 0)%>% #113,902 down to 59,238
  select(1:5)

write_csv(WI_Effort_Lakes, "WI_Lake_List.csv")
```

WI Data Summaries
- These summaries were run with the CPUE file since the fish file (where every line is an individual fish is too big), must find a way to make that usable, and then we won't use this file going forward


```{r}
WI_inland_cpue_data <- read_csv("~/CASC/CASC_Fish_Abundance/Data_Summaries/WI/wdnr_inland_cpue_data.csv")

#5853 with lotic

WI_inland_cpue_lakes <- WI_inland_cpue_data %>%
  rename(Waterbody_Name = waterbody.name)%>%
  mutate(River = str_detect(Waterbody_Name, "river")) %>%
  mutate(Creek = str_detect(Waterbody_Name, "creek"))%>%
  mutate(Canal = str_detect(Waterbody_Name, "canal"))%>%
  mutate(Stream = str_detect(Waterbody_Name, "stream"))%>%
  mutate(Brook = str_detect(Waterbody_Name, "brook"))%>% #added 7/29
  mutate(Flowage = str_detect(Waterbody_Name, "flowage"))%>% #added 7/29
  mutate(Channel = str_detect(Waterbody_Name, "channel"))%>% #added 7/29
  mutate(Ditch = str_detect(Waterbody_Name, "ditch"))%>% #added 7/29
  mutate(Lotic = case_when(River == TRUE ~ 1,
                           Creek == TRUE ~ 1,
                           Canal == TRUE ~ 1,
                           Stream == TRUE ~ 1,
                           Brook == TRUE ~ 1,
                           Flowage == TRUE ~ 1,
                           Channel == TRUE ~ 1,
                           Ditch == TRUE ~ 1,
                           TRUE ~ 0)) %>%
  filter(Lotic == 0)%>%
  select(1:11)%>%
  mutate(species = str_replace_all(species, "_", " "))%>%
  mutate(TARGET_SPECIES = case_when(species == "bluegill" ~ "1",           #target species is to flag our species of interest
                                    species == "black crappie" ~ "1",
                                    species == "largemouth bass" ~ "1",
                                    species == "northern pike" ~ "1",
                                    species == "smallmouth bass" ~ "1",
                                    species == "walleye" ~ "1",
                                    species == "yellow perch" ~ "1",
                                    TRUE ~ "0"))%>%
  mutate(Gear = case_when(cpe_units == "N/miles_shocked" ~ "EF",
                          cpe_units == "N/net_night" ~ "Netting",
                          TRUE ~ "NA"))%>%
  filter(Gear != "NA")
  
#1731 after applying my filters
#129 species, several hybrid sunfish

```

```{r}
WI_Catch_Effort_Lakes <- WI_inland_cpue_lakes %>%
  filter(total_catch > 0)%>%
  group_by(species, Gear)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(WI_inland_cpue_lakes$wbic)))

WI_Lake_prop <- ggplot(WI_Catch_Effort_Lakes, aes(x= Gear, y = Prop_Lakes, fill = Gear))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~species)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "WI fish catch and gear use", y="Proportion of lakes\n(out of 1,731)")

WI_Lake_prop
```
```{r}
WI_Catch_Effort_Lakes2 <- WI_inland_cpue_lakes %>%
  filter(total_catch > 0)%>%
  filter(TARGET_SPECIES == 1)%>%
  group_by(species, Gear)%>%
  summarise(Num_Lakes = n())%>%
  mutate(Prop_Lakes = Num_Lakes/length(unique(WI_inland_cpue_lakes$wbic)))

WI_Lake_prop2 <- ggplot(WI_Catch_Effort_Lakes2, aes(x= Gear, y = Prop_Lakes, fill = Gear))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  facet_wrap(.~species)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  labs(title = "WI fish catch and gear use", y="Proportion of lakes\n(out of 1,731)")

WI_Lake_prop2
```
```{r}
WI_Catch_Effort_Species <- WI_inland_cpue_lakes %>%
  filter(total_catch > 0)%>%
  group_by(species)%>%
  summarise(Num_Lks = n_distinct(wbic))%>%
  mutate(Prop_Lakes_SP = Num_Lks/length(unique(WI_inland_cpue_lakes$wbic)))%>%
  mutate(Prop_Lakes_SP = round(Prop_Lakes_SP, digits = 2))%>%
  mutate(SPECIES= as.factor(species))%>%
  mutate(SPECIES = fct_reorder(SPECIES, Prop_Lakes_SP))%>%
  mutate(TARGET_SPECIES = case_when(species == "bluegill" ~ "1",
                                    species == "black crappie" ~ "1",
                                    species == "largemouth bass" ~ "1",
                                    species == "northern pike" ~ "1",
                                    species == "smallmouth bass" ~ "1",
                                    species == "walleye" ~ "1",
                                    species == "yellow perch" ~ "1",
                                    TRUE ~ "0"))

all_wi_lk <- WI_inland_cpue_lakes%>%
  group_by(wbic)%>%
  summarise(Total_gear_sp = n()) #1731 Total Lakes

WI_Lake_prop <- ggplot(WI_Catch_Effort_Species, aes(x= SPECIES, y = Prop_Lakes_SP, fill = SPECIES))+
  geom_bar(stat="identity", colour="black", position="dodge")+
  geom_text(aes(label = Prop_Lakes_SP), nudge_y = 0.05, size = 2.5)+
  theme_classic()+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1), legend.position = "none")+
  labs(title = "Proportion of WI lakes with each species", y="Proportion of lakes\n(out of 1,731)")+
  gghighlight(TARGET_SPECIES == 1)

WI_Lake_prop
```
```{r}
WI_CPE_no_0_target <- WI_inland_cpue_lakes %>%
  filter(cpe > 0)%>%
  filter(TARGET_SPECIES == 1)

WI_mean_CPUE_by_gear <- ggplot(WI_CPE_no_0_target, aes(x = species, y = cpe, group = species, fill = species))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = species), alpha = .05)+
  theme_classic()+
  facet_wrap(.~Gear)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "WI CPUE by Gear", y = "non-zero CPUE (log scale)")
  
WI_mean_CPUE_by_gear
```

```{r}
WI_CPE_no_0 <- WI_inland_cpue_lakes %>%
  filter(cpe > 0)

WI_mean_CPUE_by_gear2 <- ggplot(WI_CPE_no_0, aes(x = species, y = cpe, group = species, fill = species))+
  geom_boxplot(color = "black")+
  geom_jitter(aes(color = species), alpha = .05)+
  theme_classic()+
  facet_wrap(.~Gear)+
  theme(axis.text.x=element_text(size=9, angle=45, hjust=1, vjust=1))+
  scale_y_log10()+
  labs(title = "WI CPUE by Gear", y = "non-zero CPUE (log scale)")
  
WI_mean_CPUE_by_gear2
```


# Additional WI Data Activities

WI Lake List for Temp UPDATE LATER WHEN REFINED BY FISH DATA
```{r}
WI_Lake_List_Temp <- WI_Lakes3%>%
  select(1,3,5,6)

write_csv(WI_Lake_List_Temp, "WI_Lake_List_draft.csv")
```

WI Species Map
```{r}
WI_Lat_Longs <- WI_wbic_lat_longs %>%
  select(2:4)%>%
  rename(Lake_ID = lake.id, Latitude = lat, Longitude = long)

WI_SP_MAP <- WI_inland_cpue_lakes%>%
  rename(Lake_ID = wbic, Species = species, Total_Catch = total_catch)%>%
  select(Lake_ID, Species, Total_Catch)%>%
  mutate(Species = str_to_title(Species))%>%
  mutate(Species = case_when(str_detect(Species, "cisco") ~ "Cisco",
                             str_detect(Species, "Cisco") ~ "Cisco",
                             TRUE ~ Species))%>%
  filter(Species %in% c("Black Crappie", "Bluegill", "Cisco", "Channel Catfish", "Largemouth Bass", "Northern Pike", "Smallmouth Bass", "Walleye", "White Crappie", "Yellow Perch"))%>%
  pivot_wider(names_from = Species, values_from = Total_Catch, values_fn = sum)%>%
  mutate("Channel Catfish"= 0)%>%
  mutate("White Crappie" = 0)%>%
  select(1,9,7,8,10,4,6,5,2,11,3)%>%
  left_join(WI_Lat_Longs, by = "Lake_ID")

```

