---
title: "WI_file cleanup"
author: "Mike Verhoeven"
date: "`r Sys.Date()`"
output: html_document
---


## State data
### Read in
From GDrive ?, not doing this for now, having issues with direct read from 
GDrive. Current startegy requires connect to GDrive for desktop. This code
is hard to WI data, can we create a more flexible/adaptable approach?

```{r}

# This code is still broken-- it is supposed to draw a file straight form the web to workspace
# import_id <- drive_get("https://drive.google.com/file/d/1huPEj93VZ01-q4YGcIVvHBR2YhbGMPUu/view?usp=share_link")
#  import_id %>%
#  drive_read_string() %>%
#    read_csv_arrow()




WI_files_list <- list.files(path = "E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/WI_raw_disaggregated_data", pattern = ".+\\.csv") #grabs only.csv files
WI_files_list

n <- length(WI_files_list)

for(i in 1:n) {
  assign(gsub(".csv","", WI_files_list[i]),
         data.table(read_csv_arrow(paste0("E:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/WI_raw_disaggregated_data/",
                                          WI_files_list[i]))))
  
  #consider moving renaming into here!
  
  
}  

# gsub(".csv","", WI_files_list[i])

```

Get new column names from data explainer for all files
```{r}
#for a state file, display column names - 

# gsub(".csv","", WI_files_list)

# #oldnames
# cde %>% # call data explainer file
#   filter(`new_file_name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
#   select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
#   make_clean_names() %>% 
#   assign("oldnames", ., envir = .GlobalEnv)


# select(14:25) #keep only columns of relevance (remove our file tracking and urls)

# display our new names for that set - (esp. need to ID no matches)

# cde %>% # call data explainer file
#   filter(`new_file_name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
#   select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
#   colnames() %>% 
#   assign("newnames", ., envir = .GlobalEnv)
# 
# cbind(oldnames,newnames)

#this does those two steps in one package
# note we want to review a sorted list of column names to check misspelling etc.
cde %>% # call data explainer file
  filter(`new_file_name`== gsub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.))) %>% 
  transpose(keep.names = "newname") %>% 
  rename("oldname" = V1) %>% 
  assign("names", ., envir = .GlobalEnv)

#see if nay column names will not have a match! 
# IF any pop FALSE, force stop and revist of data explainer ()
# - e.g., named something "total catch" when actual column name was "total_catch"

cbind(colnames(get(gsub(".csv","", WI_files_list)[1])),
      colnames(get(gsub(".csv","", WI_files_list)[1])) %in% names[ , oldname, ]
)

names <- names[ match(colnames(get(gsub(".csv","", WI_files_list)[1])),
       names[!str_detect(newname,"unique_row_key"),oldname]), ] 

get(gsub(".csv","", WI_files_list)[1]) %>% 
rename_with( ., function(x) {assign(x, value = names[oldname == x, newname])} )

function(x) {names[!str_detect(newname,"unique_row_key") & oldname == x, newname]}


colnames(get(gsub(".csv","", WI_files_list)[1])) <- #rename the columns
 #unique row key was for the data mgr, need to exclude it here





# note for files that have items in the "garbage bin" column, we NEED a different process where we remove garbage columns after 'filter' but before dropping na columns

 


#newnames
CASC_Data_explainer %>% # call data explainer file, I abbreviate it as CDE
  filter(`New File Name`== sub(".csv","", WI_files_list)[1])%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  clean_names() %>%  # clean up new names from data explainer
  colnames() #print only CDE column names





colnames(all_gde_gsh) <- (colnames(gde_gsh_col_id))
```


```

# Rename using a named vector and `all_of()`
lookup <- c(pl = "Petal.Length", sl = "Sepal.Length")
rename(iris, all_of(lookup))


# Using data.table
library(data.table)

# rename all columns for old to new
# Rename columns from list
setnames(my_dataframe, old = c('c1','c2','c3','c4','c5'), 
         new = c('id','pages','name','chapters','price'))

