---
title: "CASC_Data_Explainer_Work"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---

2/2/23
First attempt to use the CASC Data Explainer File to rename columns and create units columns where needed

```{r}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(arrow)
library(data.table)
library(tidyr)
library(tibble)
```

Read in CASC Data Explainer
```{r}
#CASC_Data_explainer_02Feb23 <- read_csv_arrow("Data_and_Scripts/Data/Data_Explainer/CASC_Data_Explainer_for_R_Data_Manipulation_02Feb23.csv")
  
CASC_Data_explainer <- read_excel("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/CASC_Data_Explainer_for_R_Data_Manipulation.xlsx")
```

Read in MN Data (starting with 2 files)
```{r}
mn_gde_gsh_fish_effort_03May2022 <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/mn_gde_gsh_fish_effort_03May2022.csv")

mn_fish_effort_03May2022 <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/mn_fish_effort_03May2022.csv")
```

Get new column names from data explainer
```{r}
CDE_cols <- colnames(CASC_Data_explainer_02Feb23) #all possible column names, including garbage column

gde_gsh_col_id <- CASC_Data_explainer_02Feb23 %>% # call data explainer file
  filter(`New File Name`== "mn_gde_gsh_fish_effort_03May2022")%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  select(14:25) #keep only columns of relevance (remove our file tracking and urls)
```

The attempts below did NOT work :(
```{r}
# note for files that have items in the "garbage bin" column, we may have a different process where we remove garbage columns after 'filter' but before dropping na columns

old_cols <- (gde_gsh_col_id[1,]) #pulls exisitng column names that are in row 1

new_cols <- data.frame(colnames(gde_gsh_col_id))#pulls column names that have a match in our file of interest 

all_gde_gsh_col <- mn_gde_gsh_fish_effort_03May2022 %>%
  rename(new_cols = old_cols)

renamed_columns <- gde_gsh_col_id %>%
  add_row()

sapply(all_gde_gsh, rename)


#almost works
colnames(all_gde_gsh) <- new_cols
```

#### THIS WORKS


Thanks to this: https://statisticsglobe.com/assign-column-names-based-on-existing-row-in-r

```{R}
gde_gsh_col_id <- CASC_Data_explainer %>% # call data explainer file
  filter(`New File Name`== "mn_gde_gsh_fish_effort_03May2022")%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  select(14:25) #keep only columns of relevance (remove our file tracking and urls)


colnames(all_gde_gsh) <- (colnames(gde_gsh_col_id))
```


Repeat above process with a file that has the a column in the "garbage bin" column
```{r}
mn_effort_col_id <- CASC_Data_explainer %>% # call data explainer file
  filter(`New File Name`== "mn_fish_effort_03May2022")%>% #keep only the row relevant to this file
  select_if(~ !any(is.na(.)))%>% #drop columns that contain NA (because that means df doesn't contain that info)
  


#colnames(all_gde_gsh) <- (colnames(gde_gsh_col_id))
```

Now rename columns with a loop
```{r}
MN_files_list <- list.files(path = "G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data", pattern = "mn_")
n <- length(MN_files_list)

for(i in 1:n) {
  assign(paste0("MN", i),
         read.csv2(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/",
                          MN_files_list[i])))
  
}

# Code from Mike
for(i in 1:n) {
  assign(gsub(".csv","", MN_files_list[i]),
         read_csv(paste0("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/",
                               MN_files_list[i]))) #don't use 'read.csv2, only get one column
  }  

#MN[i] <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/"[i])
```
Could maybe loop for renaming columns with code like:
MN_col_id[i] <- CASC_Data_explainer %>%
    filter('New File Name' == "i")%>%
    select_if(~ !any(is.na(.)))%>%
    select(-1:13)
  
  colnames(MN[i]) <- colnames(MN_col_id[i])
  
  But that didn't work lol

### Combine all MN files into one flat file
```{r}
MN_df <- as.data.frame(gsub(".csv","", MN_files_list)) %>% rename(file_name = 1)
CDE_cols <- as.data.frame(colnames(CASC_Data_explainer))#drop 1:16
CDE_keep_cols <- CDE_cols [17:79,]  #could also do %>%select(!Jan2023_inventoried:Targeted_or_Standard)



# 1 mn aged fish
mn_aged_fish <- read_csv("G:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/MN_raw_disaggregated_data/mn_aged_fish.csv")

MN_df_1_id_gb <- CASC_Data_explainer %>%   # get list of old column names and new column names
  filter(`New File Name`== MN_df[1,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))


GB1 <- as.data.frame(str_split_fixed(MN_df_1_id_gb$GarbageBin, ",", n = Inf))%>% # get list of columns in "garbage bin"
  pivot_longer(cols = 1:5, names_to = "v", values_to = "gb")%>%
  mutate(gb = str_trim(gb))

#vars <- GB1$gb
MN_1 <- mn_aged_fish %>% #drop garbage bin columns. SHOULD WORK BUT DOESN"T, only does first garbage column
  select(-any_of(GB1$gb))%>%
  select(contains("unit"))

MN_id_1 <- CASC_Data_explainer %>%   # get list of old column names and new column names
  filter(`New File Name`== MN_df[1,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols)) %>%
  select(-GarbageBin)

#MN1 <- mn_aged_fish %>% select(-ASSIGNED_AGE_SOURCE_NAME, -MAT, -CPUE, -AREA, -SRVY_DT)
col_order_1 <-as.character( MN_id_1[1,])

MN_1_ro <- MN_1[, col_order_1]
colnames(MN_1) <- (colnames(MN_id_1)) #rename columns (have to manually type df name in first 'colnames()' argument)

# 2 mn_aged_fish_1
MN_df_2_id <- CASC_Data_explainer %>%
  filter(`New File Name`== MN_df[2,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))

colnames(mn_aged_fish_1) <- (colnames(MN_df_2_id))

# 3 mn_aged_fish_2
MN_df_3_id <- CASC_Data_explainer %>%
  filter(`New File Name`== MN_df[3,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))

colnames(mn_aged_fish_2) <- (colnames(MN_df_3_id))  # I wish this worked colnames(MN_df[3,]) <- (colnames(MN_df_3_id))

# 4 mn_ef_lmb_smb_catch_26Aug2022
MN_df_4_id <- CASC_Data_explainer %>%
  filter(`New File Name`== MN_df[4,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))

colnames(mn_ef_lmb_smb_catch_26Aug2022) <- (colnames(MN_df_4_id))

# 5 mn_ef_lmb_smb_effort_26Aug2022
MN_df_5_id <- CASC_Data_explainer %>%
  filter(`New File Name`== MN_df[5,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))

colnames(mn_ef_lmb_smb_effort_26Aug2022) <- (colnames(MN_df_5_id))

#	6 mn_fish_effort_03May2022
MN_df_6_id <- CASC_Data_explainer %>%
  filter(`New File Name`== MN_df[6,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))

colnames(mn_fish_effort_03May2022) <- (colnames(MN_df_6_id))

# 7 mn_gde_gsh_fish_effort_03May2022
MN_df_7_id <- CASC_Data_explainer %>%
  filter(`New File Name`== MN_df[7,])%>%
  select_if(~ !any(is.na(.)))%>%
  select(any_of(CDE_keep_cols))

colnames(mn_gde_gsh_fish_effort_03May2022) <- (colnames(MN_df_7_id))

```

Now join or bind the above files into one
```{r}
# make blank df to full join with above, so all will have the same columns
Col_Keep_tibble <- tibble_row(Year = NA, Date = NA, End_Date = NA, start_time = NA, end_time = NA, `Time Zone` = NA, County = NA, `Lake ID` = NA,
                              Secondary_Lake_ID = NA, `Lake Name` = NA, `Lake Type` = NA, `Survey ID` = NA, SampleID = NA, survey_type = NA, `Site ID` = NA, 
                              lat_unspec = NA, lon_unspec = NA, lat_start = NA, lon_start = NA, lat_end = NA, lon_end = NA, location_notes = NA, `Sampling Method` = NA, 
                              `Sampling Method Abbrev` = NA, `Target Species` = NA, Secondary_Target_Species = NA, Species = NA, `Total Count` = NA, `Individual Count` = NA, 
                              Length  = NA, `Length Unit` = NA, Weight = NA, `Weight Unit` = NA, Age = NA, `Aging Structure` = NA, Young_of_Year = NA, 
                              `Backcalculated Age` = NA, Sex = NA, Reproductive_Condition = NA, Distance = NA, `Distance Units` = NA, 
                              `Sample Time` = NA, `Sample Time Units` = NA, `Total Effort 1` = NA, 
                              `Total Effort 2` = NA, `Total Effort 3` = NA, Total_Effort_4 = NA, Total_Effort_5 = NA, Effort_Units = NA, CPUE = NA, CPUE_Units = NA, 
                              c_lakesize = NA, c_lakesize_units = NA, Water_Temp = NA, Water_Temp_Units = NA, Sample_ID = NA, Notes = NA, GarbageBin = NA, 
                              Gear_Data = NA, water_clarity = NA, water_clarity_units = NA, mark_recap_data = NA, Aging_data = NA)

names(mn_aged_fish) %in% names(Col_Keep_tibble)
MN_1 <- mn_aged_fish %>%
  bind_cols()
```









```{r}
library(dtplyr)
```







this stuff doesn't work, and contains examples from the internet that failed to help
```{r}
#gde_gsh_col_id[1,i] <- colnames(gde_gsh_col_id)[i]


#all_gde_gsh_col <- mn_gde_gsh_fish_effort_03May2022 %>% rename(all_of(old_cols))


# Rename using a named vector and `all_of()`
lookup <- c(pl = "Petal.Length", sl = "Sepal.Length")
rename(iris, all_of(lookup))


# Using data.table
library(data.table)

# rename all columns for old to new
# Rename columns from list
setnames(my_dataframe, old = c('c1','c2','c3','c4','c5'), 
         new = c('id','pages','name','chapters','price'))
```
