---
title: "Fish_PA_Map"
author: "Holly Kundel"
date: "`r Sys.Date()`"
output: html_document
---

Load in required packages
```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(readxl)
library(stringr)
library(arrow)
library(lubridate)
library(colorspace)
```


Map to indicate presence/absence of various species
- esp. interested in differences between warm and cold water sp
- future direction: compare older data (1990- 2010) to more current data (2010 - present)

Ask Mike if we have lat longs for AR


First read in and run CASC Data Explainer Code

Read in catch and location data for each state

MN
```{r}
MN_GN_TN <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_fish_effort_03May2022.csv")

MN_1 <- MN_GN_TN %>%
  select(DOW, COMMON_NAME)%>%
  mutate(Species = ifelse(str_detect(COMMON_NAME, "cisco"), paste("cisco species"), paste(COMMON_NAME)))%>%
  rename(Lake_ID = DOW)%>%
  select(-COMMON_NAME)

MN_GDE_GSH <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_gde_gsh_fish_effort_03May2022.csv")

MN_2 <- MN_GDE_GSH %>%
  select(DOW, COMMON_NAME)%>%
  mutate(Species = ifelse(str_detect(COMMON_NAME, "cisco"), paste("cisco species"), paste(COMMON_NAME)))%>%
  rename(Lake_ID = DOW)%>%
  select(-COMMON_NAME)

MN_EF_bass <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_ef_lmb_smb_catch_26Aug2022.csv")

MN_3 <- MN_EF_bass %>%
  select(ID_NBR, SP)%>%
  mutate(Species = case_when(SP == "LMB" ~ "largemouth bass",
                             SP == "SMB" ~ "smallmouth bass",
                             TRUE ~ NA)) %>%
  rename(Lake_ID = ID_NBR)%>%
  select(-SP)

# combine into one, and group by lake
MN_species <- bind_rows(MN_1, MN_2, MN_3)%>%
  group_by(Lake_ID, Species)%>%
  summarise(Obs = n()) #37151 lakes

#condense to a few species of interest 
MN_fish_PA <- MN_species%>%
  filter(Species %in% c("largemouth bass",
                        "bluegill",
                        "walleye",
                        "cisco species",
                        "yellow perch",
                        "channel catfish",
                        "smallmouth bass",
                        "northern pike",
                        "black crappie"))

#add lat longs
MN_lake_list <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_lake_list.csv")

MN_lat_long <- MN_lake_list %>% 
  select(DOW_NBR_PRIMARY, DOW, LAKE_CENTER_LAT_DD5, LAKE_CENTER_LONG_DD5)%>%
  rename(long = LAKE_CENTER_LONG_DD5, lat = LAKE_CENTER_LAT_DD5)
  
MN_map_dat <- MN_fish_PA %>%
  left_join(MN_lat_long, by = c("Lake_ID" = "DOW"))%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Lake_ID = str_pad(Lake_ID, 8, side = "left", pad = "0"))
  
#write_csv(MN_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/MN_map_dat.csv")

MN_Lake_num <- MN_species %>%
  group_by(Lake_ID)%>%
  summarise(Total = n()) #3099
  
  # 79+134+364+43+402+464+2138+3099 = 6723
```


WI
```{r}
WI_fish <- read_csv_arrow("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/wi_raw_disaggregated_data/wi_inland_fishobservations_19Mar2021.csv")

WI_1 <- WI_fish %>%
  select(wbic, species, waterbody.name)%>%
  mutate(Species = ifelse(str_detect(species, "cisco"), paste("cisco species"), paste(species)))%>%
  rename(Lake_ID = wbic)%>%
  select(Lake_ID, Species, waterbody.name)

WI_CPUE <- read_csv_arrow("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/wi_raw_disaggregated_data/wi_inland_cpue_19Mar2021.csv")

WI_2 <- WI_CPUE %>%
  filter(total_catch > 0)%>% #filter out instances of zero catch
  select(wbic, species, waterbody.name)%>%
  mutate(Species = ifelse(str_detect(species, "cisco"), paste("cisco species"), paste(species)))%>%
  rename(Lake_ID = wbic)%>%
  select(Lake_ID, Species, waterbody.name)

WI_lakes <- bind_rows(WI_1, WI_2)%>%
  mutate(River = str_detect(waterbody.name, "river")) %>% #remove non-lakes
  mutate(Creek = str_detect(waterbody.name, "creek"))%>%
  mutate(Canal = str_detect(waterbody.name, "canal"))%>%
  mutate(Stream = str_detect(waterbody.name, "stream"))%>%
  mutate(Brook = str_detect(waterbody.name, "brook"))%>% 
  mutate(Flowage = str_detect(waterbody.name, "flowage"))%>% 
  mutate(Channel = str_detect(waterbody.name, "channel"))%>% 
  mutate(Ditch = str_detect(waterbody.name, "ditch"))%>% 
  mutate(Lotic = case_when(River == TRUE ~ 1,
                           Creek == TRUE ~ 1,
                           Canal == TRUE ~ 1,
                           Stream == TRUE ~ 1,
                           Brook == TRUE ~ 1,
                           Flowage == TRUE ~ 1,
                           Channel == TRUE ~ 1,
                           Ditch == TRUE ~ 1,
                           TRUE ~ 0)) %>%
  filter(Lotic == 0)
  
WI_Species <- WI_lakes %>%
  distinct()%>%
  select(Lake_ID, Species, waterbody.name)%>%
  group_by(Lake_ID, Species)%>%
  summarise(Obs = n()) #16356 waterbodies total, 31,593 lakes

WI_fish_PA <- WI_Species%>%
  mutate(Species = str_replace_all(Species, "_", " "))%>%
  filter(Species %in% c("largemouth bass",
                        "bluegill",
                        "walleye",
                        "cisco species",
                        "yellow perch",
                        "channel catfish",
                        "smallmouth bass",
                        "northern pike",
                        "black crappie"))%>%
  distinct()

# add lake lat longs

WI_lat_long <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/wi_lake_wbic_lat_long.csv")

WI_map_dat <- WI_fish_PA %>%
  left_join(WI_lat_long, by = c("Lake_ID" = "WBIC"))%>%
  rename(lat = Latitude, long = Longitude)%>%
  select(-Obs)%>%
  mutate(Lake_ID = as.character(Lake_ID))

write_csv(WI_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/WI_map_dat.csv") 

WI_lake_num <- WI_lakes %>%
  group_by(Lake_ID)%>%
  summarise(Total = n()) #2138 lakes

WI_3 <- WI_CPUE %>%
  select(wbic, species, waterbody.name)%>%
  filter(str_detect(species, "cisco"))%>%
  group_by(species)%>%
  summarise(Total = n())

WI_4 <- WI_map_dat %>%
  group_by(Species)%>%
  summarise(Total = n())
```


MI
```{r}
MI_Fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MI_Data/mi_raw_disaggregated_data/mi_statustrends_catch_16Mar2021.csv")

MI_Effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MI_Data/mi_raw_disaggregated_data/mi_statustrends_effort_16Mar2021.csv")

MI_map_dat <- MI_Fish %>%
  select(NEW_KEY, SPECIES, LAT_DD, LONG_DD)%>%
  group_by(NEW_KEY, SPECIES)%>%
  mutate(Species = case_when(SPECIES == "BCR" ~ "black crappie",
                             SPECIES== "BLG" ~ "bluegill",
                             SPECIES == "CIS" ~ "cisco species",
                             SPECIES == "LMB" ~ "largemouth bass",
                             SPECIES == "NOP" ~ "northern pike",
                             SPECIES == "SMB" ~ "smallmouth bass",
                             SPECIES == "WAE" ~ "walleye",
                             SPECIES == "YEP" ~ "yellow perch"))



MI_map_dat2 <- MI_map_dat %>%
  distinct()%>%
  ungroup()%>%
  rename(Lake_ID = NEW_KEY, lat = LAT_DD, long = LONG_DD)%>%
  select(Lake_ID, Species, long, lat)

MI_lakes <- MI_Effort %>%
  group_by(NEW_KEY)%>%
  summarise(total = n()) #464 lakes
  


#write_csv(MI_map_dat2, "Data_and_Scripts/Data/Rough_State_PA_May2023/MI_map_dat.csv")
```

SD
```{r, warning=FALSE}
SD <-  read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_catch_4Oct2021.csv") #catch

SD_Effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_effort_4Oct2021.csv")

SD_NOP <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_NOP_catch_effort_03Nov22.csv")

SD_NOP_effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_NOP_survey_effort_03Nov22.csv")

SD_lake_survey_ID <- SD_Effort %>%
  select(SurveyID, StateID)

SD_lakes <- SD_Effort %>%
  group_by(StateID)%>%
  summarise(total = n()) #402

SD_NOP_use <- SD_NOP %>%
  left_join(SD_NOP_effort, by = "SurveyID *")%>%
  rename(SpeciesName = 4, SurveyID = 3, StateID = 13)%>%
  select(SpeciesName, SurveyID, FishCount, StateID)

SD_Fish <- SD %>%
  left_join(SD_lake_survey_ID, by = "SurveyID")%>%
  bind_rows(SD_NOP_use)

SD_lat_long <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/ManagedFisheries.csv")

SDlatlong <- SD_lat_long %>%
  select(`StateID,C,75`,`Latitude,N,19,8`, `Longitude,N,19,8`)

SD_map_dat <- SD_Fish %>%
  left_join(SDlatlong, by = c("StateID"="StateID,C,75"))%>%
  rename(Lake_ID = StateID, Species = SpeciesName, lat = `Latitude,N,19,8`, long = `Longitude,N,19,8`)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = str_to_lower(Species))%>%
  distinct()


write_csv(SD_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/SD_map_dat.csv")
```

IA
```{r}
IA_CPUE <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IA_Data/ia_raw_disaggregated_data/cleanedIACPUEs.csv")

IA_lake_num <- IA_CPUE %>%
  group_by(LakeId)%>%
  summarise(total = n()) #443

IA_map_dat <- IA_CPUE %>%
  select(LakeId, CommonName, Latitude, Longitude)%>%
  distinct()%>%
  rename(Species = CommonName, lat = Latitude, long = Longitude, Lake_ID = LakeId)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = str_to_lower(Species))

#write_csv(IA_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/IA_map_dat.csv")
```

IL
```{r}
IL_fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IL_Data/il_raw_disaggregated_data/il_catch_age_effort_17Jan22.csv")

IL_fish <- IL_fish %>%
  select(LakeNumber, Species, Latitude, Longitude)%>%
  distinct()

IL_map_dat <- IL_fish %>%
  rename(Lake_ID = LakeNumber, lat = Latitude, long = Longitude)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = str_to_lower(Species))%>%
  mutate(Lake_ID = as.character(Lake_ID))

IL_num_lakes <- IL_fish %>%
  group_by(LakeNumber)%>%
  summarise(Total = n()) #364

#write_csv(IL_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/IL_map_dat.csv")
```

IN
need better lat long match for IN glacial lakes
```{r}
IN_lat_long <- read_excel("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/2020FINFO_Lake_List.xlsx")

IN_GL_fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/in_raw_disaggregated_data/in_glaciallakes_fish_6Oct2021.csv")

IN_1 <- IN_GL_fish %>%
  select(Lake, County, Species)%>%
  mutate(Species = str_to_lower(Species))%>%
  left_join(IN_lat_long, by = c("Lake" = "Name", "County" = "County"))%>%
  distinct()%>%
  select(Id, Species, Latitude, Longitude)%>%
  rename(Lake_ID = Id, lat = Latitude, long = Longitude)

IN_GL <- IN_GL_fish %>%
  group_by(Lake, County)%>%
  summarise(Total = n()) #134 glacial lakes

IN_R_fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/in_raw_disaggregated_data/in_reservoir_fish_community_fishdata_16Aug2022.csv")

IN_R_effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/in_raw_disaggregated_data/in_reservoir_fish_community_effort_16Aug2022.csv")

IN_2 <- IN_R_fish %>%
  left_join(IN_R_effort, by = c("Survey ID" = "Survey_ID"))%>%
  select(Lake_ID_Number, Species, Latitude, Longitude)%>%
  mutate(Species = str_to_lower(Species))%>%
  rename(Lake_ID = Lake_ID_Number, lat = Latitude, long = Longitude)%>%
  distinct()

IN_R <- IN_R_effort %>%
  group_by(Lake_ID_Number)%>%
  summarise(Total = n()) #79 res

IN_map_dat <- bind_rows(IN_1, IN_2)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = ifelse(str_detect(Species, "cisco"), paste("cisco species"), paste(Species)))

write_csv(IN_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/IN_map_dat.csv") 
```
3099+2048+464+402+443+364+134+79
[1] 7033 (total lakes)

Arkansas
Kansas, Nebraska, and Ontario WAE data from Lyndsie Wszola (7480 total)
```{r}
AR <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/AR_Data/ar_reservoir_age_8Oct2021.csv")
AR2 <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/AR_Data/ar_reservoir_age2_8Oct2021.csv")
AR2_rename <- AR2 %>%
  rename(Waterbody = waterbody)

AR_all <- bind_rows(AR, AR2_rename)%>%
  group_by(Waterbody, Location)%>%
  summarise(Total = n()) #3 lakes...

AR3 <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/AR_Data/ar_lat_longs_8May2023.csv")

NE <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/NE_KS_ON Growth Data - Lyndsie Wszola/NE_Standard_fish_19Jan2023.csv")

NE_lake_num <- NE %>%
  group_by(Waterbody_ID)%>%
  summarise(Total = n()) #225

ON <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/NE_KS_ON Growth Data - Lyndsie Wszola/ON_Standard_fish_19Jan2023.csv")

ON_lake_num <- ON %>%
  group_by(Waterbody_ID)%>%
  summarise(Total = n()) #21

KS <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/NE_KS_ON Growth Data - Lyndsie Wszola/KS_Standard_fish_19Jan2023.csv")

KS_lake_num <- KS %>%
  group_by(Waterbody_ID)%>%
  summarise(Total = n()) #200
```


Combine all states together
- total number of lakes: 
```{r}
Map_dat <- bind_rows(MN_map_dat, WI_map_dat, MI_map_dat2, SD_map_dat, IL_map_dat, IA_map_dat, IN_map_dat)

write_csv(Map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/All_states_map_dat.csv")

Map_dat_filtered_sp <- Map_dat %>%
  filter(Species %in% c("largemouth bass",
                        "bluegill",
                        "walleye",
                        "cisco species",
                        "yellow perch",
                        "smallmouth bass",
                        "northern pike",
                        "black crappie"))%>%
  filter(long > -104 & long < -69)%>%
  filter(lat > 33 & lat < 50)%>%
  filter(Lake_ID != "{C47CE045-4D52-4279-B982-E47E1409874F}")%>%
  filter(Lake_ID != "4232")%>%
  filter(Lake_ID != "SHP84")%>%
  filter(Lake_ID!= "118")

#removed channel catfish because we do not have data on that species from WI or MI
```

#Make the map
load in more libraries
```{r}
library(ggplot2) #for plotting
library(maps) #to pull in the map data for geom_polygon
library(colorspace)
```

Base state maps
```{r}
Midwest <- map_data("state", region = c("minnesota", "wisconsin", "north dakota", "south dakota", "iowa", "illinois", "indiana", "michigan"))

Midwest_noND <- map_data("state", region = c("minnesota", "wisconsin", "south dakota", "iowa", "illinois", "indiana", "michigan"))
midwest_map <- ggplot(data = Midwest)+
  geom_polygon(aes(x = long, y = lat, group = group, fill = region), color = "white")+
  coord_fixed(1.3)+
  guides(fill = FALSE)
```

```{r}
All_Lakes_map_noND <- ggplot()+
  geom_polygon(data = Midwest_noND, aes(x = long, y = lat, group = group,), color = "black", fill = "azure1")+
  coord_fixed(1.3)+
  guides(fill = FALSE)+
  geom_point(data = Map_dat_filtered_sp, aes(x = long, y = lat), color = "blue", alpha=0.1)+
  theme_void()

All_Lakes_map_noND

ggsave(file = "All_lakes_fish_map_May23.jpeg", 
       plot = All_Lakes_map_noND,
       device = "jpeg",
       dpi = "print")
```

```{r}
All_Lakes_SPECIES <- ggplot()+
  geom_polygon(data = Midwest_noND, aes(x = long, y = lat, group = group,), color = "black", fill = "azure1")+
  coord_fixed(1.3)+
  guides(fill = FALSE)+
  geom_point(data = Map_dat_filtered_sp, aes(x = long, y = lat, color = Species), alpha=0.2)+
  facet_wrap(.~Species, nrow = 2, ncol = 4)+
  theme_classic()+
  theme(legend.position = "none",
        axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank())+
  scale_color_discrete_qualitative(palette = "Dark 3")

All_Lakes_SPECIES

ggsave(file = "All_lakes_SPECIES_map_May23_2.jpeg", 
       plot = All_Lakes_SPECIES,
       device = "jpeg",
       dpi = "print",
       width = 12,
       height = 6)
```



```{r}
WI <- map_data("state", region = "wisconsin")

WI_Lakes_SPECIES <- ggplot()+
  geom_polygon(data = WI, aes(x = long, y = lat, group = group,), color = "black", fill = "azure1")+
  coord_fixed(1.3)+
  guides(fill = FALSE)+
  geom_point(data = WI_map_dat, aes(x = long, y = lat, color = Species), alpha=0.1)+
  facet_wrap(.~Species)+
  theme_classic()+
  theme(legend.position = "none")

WI_Lakes_SPECIES

```



Repeat to include Year
MN (1968 - 2022)
```{r}
MN_GN_TN <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_fish_effort_03May2022.csv")

MN_1 <- MN_GN_TN %>%
  select(DOW, COMMON_NAME, YEAR)%>%
  mutate(Species = ifelse(str_detect(COMMON_NAME, "cisco"), paste("cisco species"), paste(COMMON_NAME)))%>%
  rename(Lake_ID = DOW, Year = YEAR)%>%
  select(-COMMON_NAME)

MN_GDE_GSH <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_gde_gsh_fish_effort_03May2022.csv")

MN_2 <- MN_GDE_GSH %>%
  select(DOW, COMMON_NAME, YEAR)%>%
  mutate(Species = ifelse(str_detect(COMMON_NAME, "cisco"), paste("cisco species"), paste(COMMON_NAME)))%>%
  rename(Lake_ID = DOW, Year = YEAR)%>%
  select(-COMMON_NAME)

MN_EF_bass <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_raw_disaggregated_data/mn_ef_lmb_smb_catch_26Aug2022.csv")

MN_3 <- MN_EF_bass %>%
  mutate(SRVY_DT = mdy(SRVY_DT))%>%
  mutate(Year = year(SRVY_DT))%>%
  select(ID_NBR, SP, Year)%>%
  mutate(Species = case_when(SP == "LMB" ~ "largemouth bass",
                             SP == "SMB" ~ "smallmouth bass",
                             TRUE ~ NA)) %>%
  rename(Lake_ID = ID_NBR)%>%
  select(Lake_ID, Year, Species)

# combine into one, and group by lake
MN_species <- bind_rows(MN_1, MN_2, MN_3)%>%
  group_by(Lake_ID, Species, Year)%>%
  summarise(Obs = n()) 

#condense to a few species of interest 
MN_fish_PA <- MN_species%>%
  filter(Species %in% c("largemouth bass",
                        "bluegill",
                        "walleye",
                        "cisco species",
                        "yellow perch",
                        "channel catfish",
                        "smallmouth bass",
                        "northern pike",
                        "black crappie"))%>%
  distinct()

#add lat longs
MN_lake_list <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MN_Data/mn_lake_list.csv")

MN_lat_long <- MN_lake_list %>% 
  select(DOW_NBR_PRIMARY, DOW, LAKE_CENTER_LAT_DD5, LAKE_CENTER_LONG_DD5)%>%
  rename(long = LAKE_CENTER_LONG_DD5, lat = LAKE_CENTER_LAT_DD5)
  
MN_map_dat_y <- MN_fish_PA %>%
  left_join(MN_lat_long, by = c("Lake_ID" = "DOW"))%>%
  select(Lake_ID, Species, Year, long, lat)
  
#write_csv(MN_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/MN_map_dat.csv")

MN_Lake_num <- MN_species %>%
  group_by(Lake_ID)%>%
  summarise(Total = n()) #3099
  
  # 79+134+364+43+402+464+2138+3099 = 6723
```


WI
```{r}
WI_fish <- read_csv_arrow("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/wi_raw_disaggregated_data/wi_inland_fishobservations_19Mar2021.csv")

WI_1 <- WI_fish %>%
  select(wbic, species, waterbody.name, survey.year)%>%
  mutate(Species = ifelse(str_detect(species, "cisco"), paste("cisco species"), paste(species)))%>%
  rename(Lake_ID = wbic, Year = survey.year)%>%
  select(Lake_ID, Species, Year, waterbody.name)

WI_CPUE <- read_csv_arrow("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/wi_raw_disaggregated_data/wi_inland_cpue_19Mar2021.csv")

WI_2 <- WI_CPUE %>%
  filter(total_catch > 0)%>% #filter out instances of zero catch
  select(wbic, species, waterbody.name, survey.year)%>%
  mutate(Species = ifelse(str_detect(species, "cisco"), paste("cisco species"), paste(species)))%>%
  rename(Lake_ID = wbic, Year = survey.year)%>%
  select(Lake_ID, Species, Year, waterbody.name)

WI_lakes <- bind_rows(WI_1, WI_2)%>%
  mutate(River = str_detect(waterbody.name, "river")) %>% #remove non-lakes
  mutate(Creek = str_detect(waterbody.name, "creek"))%>%
  mutate(Canal = str_detect(waterbody.name, "canal"))%>%
  mutate(Stream = str_detect(waterbody.name, "stream"))%>%
  mutate(Brook = str_detect(waterbody.name, "brook"))%>% 
  mutate(Flowage = str_detect(waterbody.name, "flowage"))%>% 
  mutate(Channel = str_detect(waterbody.name, "channel"))%>% 
  mutate(Ditch = str_detect(waterbody.name, "ditch"))%>% 
  mutate(Lotic = case_when(River == TRUE ~ 1,
                           Creek == TRUE ~ 1,
                           Canal == TRUE ~ 1,
                           Stream == TRUE ~ 1,
                           Brook == TRUE ~ 1,
                           Flowage == TRUE ~ 1,
                           Channel == TRUE ~ 1,
                           Ditch == TRUE ~ 1,
                           TRUE ~ 0)) %>%
  filter(Lotic == 0)
  
WI_Species <- WI_lakes %>%
  distinct()%>%
  select(Lake_ID, Species, Year)%>%
  group_by(Lake_ID, Species, Year)%>%
  summarise(Obs = n()) 

WI_fish_PA <- WI_Species%>%
  mutate(Species = str_replace_all(Species, "_", " "))%>%
  filter(Species %in% c("largemouth bass",
                        "bluegill",
                        "walleye",
                        "cisco species",
                        "yellow perch",
                        "channel catfish",
                        "smallmouth bass",
                        "northern pike",
                        "black crappie"))%>%
  distinct()

# add lake lat longs

WI_lat_long <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/WI_Data/wi_lake_wbic_lat_long.csv")

WI_map_dat_y <- WI_fish_PA %>%
  left_join(WI_lat_long, by = c("Lake_ID" = "WBIC"))%>%
  rename(lat = Latitude, long = Longitude)%>%
  mutate(Lake_ID = as.character(Lake_ID))%>%
  select(Lake_ID, Species, Year, long, lat)

write_csv(WI_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/WI_map_dat.csv") 

WI_lake_num <- WI_lakes %>%
  group_by(Lake_ID)%>%
  summarise(Total = n()) #2138 lakes

WI_3 <- WI_CPUE %>%
  select(wbic, species, waterbody.name)%>%
  filter(str_detect(species, "cisco"))%>%
  group_by(species)%>%
  summarise(Total = n())

WI_4 <- WI_map_dat %>%
  group_by(Species)%>%
  summarise(Total = n())
```


MI
```{r}
MI_Fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MI_Data/mi_raw_disaggregated_data/mi_statustrends_catch_16Mar2021.csv")

MI_Effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/MI_Data/mi_raw_disaggregated_data/mi_statustrends_effort_16Mar2021.csv")

MI_map_dat <- MI_Fish %>%
  select(NEW_KEY, SPECIES, LAT_DD, LONG_DD, SURVEY_YEAR)%>%
  group_by(NEW_KEY, SPECIES, SURVEY_YEAR)%>%
  mutate(Species = case_when(SPECIES == "BCR" ~ "black crappie",
                             SPECIES== "BLG" ~ "bluegill",
                             SPECIES == "CIS" ~ "cisco species",
                             SPECIES == "LMB" ~ "largemouth bass",
                             SPECIES == "NOP" ~ "northern pike",
                             SPECIES == "SMB" ~ "smallmouth bass",
                             SPECIES == "WAE" ~ "walleye",
                             SPECIES == "YEP" ~ "yellow perch"))



MI_map_dat_y <- MI_map_dat %>%
  distinct()%>%
  ungroup()%>%
  rename(Lake_ID = NEW_KEY, lat = LAT_DD, long = LONG_DD, Year = SURVEY_YEAR)%>%
  select(Lake_ID, Species, Year, long, lat)

MI_lakes <- MI_Effort %>%
  group_by(NEW_KEY)%>%
  summarise(total = n()) #464 lakes
  


#write_csv(MI_map_dat2, "Data_and_Scripts/Data/Rough_State_PA_May2023/MI_map_dat.csv")
```

SD
```{r, warning=FALSE}
SD <-  read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_catch_4Oct2021.csv") #catch

SD_Effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_effort_4Oct2021.csv")

SD_NOP <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_NOP_catch_effort_03Nov22.csv")

SD_NOP_effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/sd_raw_disaggregated_data/sd_NOP_survey_effort_03Nov22.csv")

SD_lake_survey_ID <- SD_Effort %>%
  select(SurveyID, StateID)

SD_lakes <- SD_Effort %>%
  group_by(StateID)%>%
  summarise(total = n()) #402

SD_NOP_use <- SD_NOP %>%
  left_join(SD_NOP_effort, by = "SurveyID *")%>%
  rename(SpeciesName = 4, SurveyID = 3, StateID = 13)%>%
  select(SpeciesName, SurveyID, FishCount, StateID)

SD_Fish <- SD %>%
  left_join(SD_lake_survey_ID, by = "SurveyID")%>%
  bind_rows(SD_NOP_use)

SD_lat_long <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/SD_Data/ManagedFisheries.csv")

SDlatlong <- SD_lat_long %>%
  select(`StateID,C,75`,`Latitude,N,19,8`, `Longitude,N,19,8`)

SD_map_dat <- SD_Fish %>%
  left_join(SDlatlong, by = c("StateID"="StateID,C,75"))%>%
  rename(Lake_ID = StateID, Species = SpeciesName, lat = `Latitude,N,19,8`, long = `Longitude,N,19,8`)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = str_to_lower(Species))%>%
  distinct()


write_csv(SD_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/SD_map_dat.csv")
```

IA
```{r}
IA_CPUE <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IA_Data/ia_raw_disaggregated_data/cleanedIACPUEs.csv")

IA_lake_num <- IA_CPUE %>%
  group_by(LakeId)%>%
  summarise(total = n()) #443

IA_map_dat <- IA_CPUE %>%
  select(LakeId, CommonName, Latitude, Longitude)%>%
  distinct()%>%
  rename(Species = CommonName, lat = Latitude, long = Longitude, Lake_ID = LakeId)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = str_to_lower(Species))

#write_csv(IA_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/IA_map_dat.csv")
```

IL
```{r}
IL_fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IL_Data/il_raw_disaggregated_data/il_catch_age_effort_17Jan22.csv")

IL_fish <- IL_fish %>%
  select(LakeNumber, Species, Latitude, Longitude)%>%
  distinct()

IL_map_dat <- IL_fish %>%
  rename(Lake_ID = LakeNumber, lat = Latitude, long = Longitude)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = str_to_lower(Species))%>%
  mutate(Lake_ID = as.character(Lake_ID))

IL_num_lakes <- IL_fish %>%
  group_by(LakeNumber)%>%
  summarise(Total = n()) #364

#write_csv(IL_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/IL_map_dat.csv")
```

IN
need better lat long match for IN glacial lakes
```{r}
IN_lat_long <- read_excel("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/2020FINFO_Lake_List.xlsx")

IN_GL_fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/in_raw_disaggregated_data/in_glaciallakes_fish_6Oct2021.csv")

IN_1 <- IN_GL_fish %>%
  select(Lake, County, Species)%>%
  mutate(Species = str_to_lower(Species))%>%
  left_join(IN_lat_long, by = c("Lake" = "Name", "County" = "County"))%>%
  distinct()%>%
  select(Id, Species, Latitude, Longitude)%>%
  rename(Lake_ID = Id, lat = Latitude, long = Longitude)

IN_GL <- IN_GL_fish %>%
  group_by(Lake, County)%>%
  summarise(Total = n()) #134 glacial lakes

IN_R_fish <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/in_raw_disaggregated_data/in_reservoir_fish_community_fishdata_16Aug2022.csv")

IN_R_effort <- read_csv("D:/Shared drives/Hansen Lab/RESEARCH PROJECTS/Fish Survey Data/IN_Data/in_raw_disaggregated_data/in_reservoir_fish_community_effort_16Aug2022.csv")

IN_2 <- IN_R_fish %>%
  left_join(IN_R_effort, by = c("Survey ID" = "Survey_ID"))%>%
  select(Lake_ID_Number, Species, Latitude, Longitude)%>%
  mutate(Species = str_to_lower(Species))%>%
  rename(Lake_ID = Lake_ID_Number, lat = Latitude, long = Longitude)%>%
  distinct()

IN_R <- IN_R_effort %>%
  group_by(Lake_ID_Number)%>%
  summarise(Total = n()) #79 res

IN_map_dat <- bind_rows(IN_1, IN_2)%>%
  select(Lake_ID, Species, long, lat)%>%
  mutate(Species = ifelse(str_detect(Species, "cisco"), paste("cisco species"), paste(Species)))

write_csv(IN_map_dat, "Data_and_Scripts/Data/Rough_State_PA_May2023/IN_map_dat.csv") 
```

Cisco map
```{r}
Cisco_dat <- bind_rows(MN_map_dat_y, WI_map_dat_y, MI_map_dat_y)%>%
  filter(Species == "cisco species")%>%
  mutate(Decades = case_when(Year > 1940 & Year < 1960 ~ "Pre 1960",
                             Year >= 1960 & Year < 1970 ~ "1960s",
                             Year >= 1970 & Year < 1980 ~ "1970s",
                             Year >=1980 & Year < 1990 ~ "1980s",
                             Year >= 1990 & Year < 2000 ~ "1990s",
                             Year >= 2000 & Year < 2010 ~ "2000s",
                             Year >= 2010 & Year < 2020 ~ "2010s",
                             Year>= 2020 ~ "2020 - Present"))

Midwest_cisco_states<- map_data("state", region = c("minnesota", "wisconsin", "michigan"))

Cisco_map <- ggplot()+
  geom_polygon(data = Midwest_cisco_states, aes(x = long, y = lat, group = group,), color = "black", fill = "gray25")+
  coord_fixed(1.3)+
  guides(fill = FALSE)+
  geom_point(data = Cisco_dat, aes(x = long, y = lat, color = Decades), alpha=0.5, size = 2)+
  theme_void()+
  scale_color_discrete_sequential(palette = "YlOrRd")+
  facet_wrap(.~Decades)

Cisco_map

MN_cisco <- MN_map_dat_y %>%
  filter(Species == "cisco species")%>%
  distinct()

```















Base state maps
```{r}
Midwest <- map_data("state", region = c("minnesota", "wisconsin", "north dakota", "south dakota", "iowa", "illinois", "indiana", "michigan"))

Midwest_noND <- map_data("state", region = c("minnesota", "wisconsin", "south dakota", "iowa", "illinois", "indiana", "michigan"))
midwest_map <- ggplot(data = Midwest)+
  geom_polygon(aes(x = long, y = lat, group = group, fill = region), color = "white")+
  coord_fixed(1.3)+
  guides(fill = FALSE)
```

```{r}
All_Lakes_map_noND <- ggplot()+
  geom_polygon(data = Midwest_noND, aes(x = long, y = lat, group = group,), color = "black", fill = "azure1")+
  coord_fixed(1.3)+
  guides(fill = FALSE)+
  geom_point(data = All_State_Lat_Longs_edit, aes(x = Longitude, y = Latitude), color = "blue", alpha=0.1)+
  theme_void()

All_Lakes_map_noND
```


